### 2024-01-20 bio_dynamics/bio274_hw1.Rmd
bio274 hw1 q1 through 3

---
title: "BIOL 274 Homework 1"
author: "Ruby Krasnow"
date: "2024-01-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1
>Assume that once a drug is administered to a patient, it leaves the body following exponential decay. After administration of a dose, the concentration of the drug in the body decreases by 50% in 30 hours. Approximately how long does it take for the drug to decrease to 1% of its initial value (round to the nearest hour)?

We can represent the concentration of drug in the body at time $t$ as $x(t)=x_0e^{kt}$. Since the drug leaves the body according to an exponential decay model, that means $k$ should be negative.

We are given $\frac{x}{x_0}=0.5$ at $t=30$ hours. Plugging into our equation, we get
$$0.5=e^{30k}$$
$$\ln(0.5)=\ln(e^{30k})$$
$$\ln(0.5)=30k$$
$$k=\frac{\ln(0.5)}{30}$$
```{r solve k}
k = log(0.5)/30
k
```

<!-- So $k$ = `r k`. -->
Now, we want to know how long it will take for $x$ to be 1% of $x_0$, or at what time $t$ we will have $\frac{x(t)}{x(0)}=0.01$. 
$$0.01=e^{kt}$$
$$t=\frac{\ln{(0.01)}}{k}$$

```{r solve t}
t = log(0.01)/k
round(t, 0)
```

It will take approximately 199 hours for the drug to decrease to 1% of its initial value.

## Question 2
> Consider the population model: $\frac{dP}{dt}=0.3 \left( 1-\frac{P}{200} \right) \left(\frac{P}{50}-1 \right)P$.

>a. For what values of P is the population increasing?
b. For what values of P is the population decreasing?

If the population is increasing, that means the rate of change $\frac{dP}{dt}$ is positive, while a decreasing population means that $\frac{dP}{dt}$ is negative. First we need to find the equilibrium points, or the values of $P$ for which $\frac{dP}{dt}=0.3 \left( 1-\frac{P}{200} \right) \left(\frac{P}{50}-1 \right)P=0$.

We can see that this will clearly occur if $P=0$; i.e., the population has no organisms in it. The other equilibrium points occur when
$$1-\frac{P}{200}=0$$
$$1=\frac{P}{200}$$
$$P=200$$

and similarly when $1=\frac{P}{50}$ or $P=50$.
So we need to determine the sign of $\frac{dP}{dt}$ on the intervals $(0,50),(50,200),$ and $(200,\infty)$.
Let's evaluate some convenient values within these intervals: 25, 100, and 400.

```{r}
eqn <- function(P){
  0.3*(1-(P/200))*((P/50-1))*P
} 
eqn(25)
eqn(100)
eqn(400)
```

So the population is decreasing when $0<P<50$ or $P>200$ and the population is increasing when $50<P<200$. Let's confirm this with a quick plot:
```{r}
x=c(1:250)

ggplot()+
  geom_line(aes(x=x, y=0.3*(1-(x/200))*((x/50-1))*x))+
  geom_vline(xintercept = c(50, 200), linetype="dashed")+
  geom_hline(yintercept = 0, linetype="dashed", color="red")+
  theme_bw()+
  labs(x="P", y="dP/dt")
```

## Question 3
Find the general solution to each of the following DEs:

### a
$$ \frac{dy}{dt}=\frac{3t+1}{2y}$$
$$ 2y \, dy=(3t+1)dt$$
$$\int 2y \, dy=\int(3t+1)dt$$
$$y^2=\frac{3t^2}{2}+t+C$$
$$y=\sqrt{\frac{3t^2}{2}+t+C}$$

### b
$$\frac{dy}{dx}=x\sqrt[3]{y}$$

$$y^{-\frac{1}{3}} \, dy=x \, dx$$
$$\int y^{-\frac{1}{3}} \, dy= \int x \, dx$$
$$\frac{3}{2}y^{\frac{2}{3}}=\frac{x^2}{2}+C_1$$
$$y^{\frac{2}{3}}=\frac{x^2+C_2}{3}$$
$$y=\left(\frac{x^2+C_2}{3} \right)^\frac{3}{2}=\frac{(x^2+C_2)^{\frac{3}{2}}}{3\sqrt3}$$

### c
$$\frac{dp}{dt}=\frac{1}{3tp^2}$$
$$3p^2\,dp=\frac{1}{t}dt$$
$$\int 3p^2\,dp=\int \frac{1}{t}dt$$
$$p^3=\ln(t)+C$$
$$p=\sqrt[3]{\ln(t)+C}$$

### 2024-01-20 bio_dynamics/bio274_hw1.Rmd
finished hw1

@@
 $$k=\frac{\ln(0.5)}{30}$$
-```{r solve k}
+```{r solve_k_1}
 k = log(0.5)/30
@@
 
-```{r solve t}
+```{r solve_t_1}
 t = log(0.01)/k
@@
 
-### a
+### a)
 $$ \frac{dy}{dt}=\frac{3t+1}{2y}$$
@@
 
-### b
+### b)
 $$\frac{dy}{dx}=x\sqrt[3]{y}$$
@@
 
-### c
+### c)
 $$\frac{dp}{dt}=\frac{1}{3tp^2}$$
@@
 $$p=\sqrt[3]{\ln(t)+C}$$
+
+## Question 4
+>According to Newton’s law of cooling, the rate of decrease of temperature of a body is proportional to the difference between its temperature and that of its environment.
+If the temperature in your living room is 20°C, you remove a log from the fireplace at 100°C, and it takes 10 minutes to cool to 60°C, how long will it take to decrease the temperature of the log to 25°C?
+
+$$\frac{dT}{dt}=k(T_{body}-T_{env})=k(T_{body}-20)$$
+$$
+\frac{1}{(T_{body}-20)}dT=k\,dt
+$$
+$$\int\frac{1}{(T_{body}-20)}dT=kt+C$$
+
+$$\ln{(T_{body}-20)}=kt+C$$
+$$T_{body}=e^{kt+C}+20=Ce^{kt}+20$$
+
+Since $T_{body}=C+20$ at time $t=0$, we can write the equation as
+$$T_{body}(t)=(T_0-20)e^{kt}+20$$
+for the temperature of the body at time $t$.
+Now using the information we are given,
+$$60=(100-20)e^{10k}+20$$
+$$40=80e^{10k}$$
+$$\frac{1}{2}=e^{10k}$$
+$$\frac{\ln{(1/2)}}{10}=k$$
+
+```{r solve_k_2}
+k2 = log(1/2)/(10)
+k2
+```
+
+Now we want to solve for $t$ in the following equation:
+$$25=(60-20)e^{kt}+20$$
+$$5=40e^{kt}$$
+$$\frac{\ln{(1/8)}}{k}=t$$
+
+```{r solve_t_2}
+t2 = (log(1/8))/(k2)
+t2
+```
+
+*Quick sanity check*:
+If it took 10 minutes to cool from 100°C to 60°C and 30 minutes to cool from 60°C to 25°C, it should take 40 minutes to cool from 100°C to 25°C:
+```{r}
+# 25=(100-20)e^kt+20
+# 5=80e^kt
+# ln(1/16)=kt
+
+t3=(log(1/16))/(k2)
+t3
+```
+
+## Question 5
+Solve each of the following IVPs:
+
+### a)
+$$\frac{dx}{dt}=-xt,\, x(0)=1/\sqrt{\pi}$$
+
+$$\frac{1}{x}dx=-t\,dt$$
+$$\int \frac{1}{x}dx=\int-t\,dt$$
+$$\ln{(x)}=-\frac{t^2}{2}+C$$
+
+$$x(t)=Ce^{-t^2/2}$$
+is our general solution to the differential equation.
+Now we plug in the initial conditions:
+$$1/\sqrt{\pi}=Ce^{-(0^2/2)}=Ce^0=C$$
+
+$$x(t)=\frac{e^{-t^2/2}}{\pi}$$
+
+is the solution to the IVP.
+
+### b)
+$$\frac{dy}{dx}=xy^2+2y^2,\, y(0)=1$$
+
+$$\frac{dy}{dx}=y^2(x+2)$$
+$$y^{-2}dy=(x+2)dx$$
+
+$$-\frac{1}{y}=\frac{x^2}{2}+2x+C$$
+
+We can plug in our initial conditions now:
+$$-\frac{1}{1}=0+0+C$$
+so $C=-1$.
+$$-\frac{1}{y}=\frac{x^2}{2}+2x-1$$
+$$\frac{1}{y}=1-\frac{x^2}{2}-2x$$
+$$\frac{1}{y}=1-\frac{x^2}{2}-2x$$
+or
+$$\frac{2}{y}=2-x^2-4x$$
+
+$$y(x)=\frac{2}{2-x^2-4x}$$
+is the solution to the IVP.
+
+### c)
+
+$$ \frac{du}{dt}=\frac{\cos(t)}{9u^2}, \, u(0)=2$$
+
+$$ u^2du=\frac{\cos(t)}{9}dt$$
+$$ \int u^2du=\frac{1}{9}\int \cos(t)dt$$
+
+$$ \frac{u^3}{3}=\frac{1}{9}\sin(t)+C$$
+
+$$ u^3=\frac{\sin(t)}{3}+C$$
+
+Now we plug in the initial conditions:
+$$ 2^3=\frac{\sin(0)}{3}+C=0+C$$
+So $C=8$ and we have
+$$ u(t)=\sqrt[3]{\frac{\sin(t)}{3}+8}$$
+
+as the solution to the IVP.

### 2024-01-24 bio_dynamics/bio274_hw1.Rmd
revised hw1 based on class lecture

@@
 
+We can see that $\frac{dP}{dt}$ is indeed positive and negative over the expected intervals.
+
 ## Question 3
@@
 $$y^{\frac{2}{3}}=\frac{x^2+C_2}{3}$$
-$$y=\left(\frac{x^2+C_2}{3} \right)^\frac{3}{2}=\frac{(x^2+C_2)^{\frac{3}{2}}}{3\sqrt3}$$
+$$y(x)=\left(\frac{x^2+C_2}{3} \right)^\frac{3}{2}=\frac{(x^2+C_2)^{\frac{3}{2}}}{3\sqrt3}$$
+Note that $y(x)=0$ is also a solution to the ODE, because $\frac{dy}{dx}$ will equal 0 for all $x$ and $x\sqrt[3]{0}=0$.
 
@@
 $$
-$$\int\frac{1}{(T_{body}-20)}dT=kt+C$$
+$$\int\frac{1}{(T_{body}-20)}dT=\int k\,dt$$
+$$\ln{(T_{body}-20)}=kt+C_1$$
+$$T_{body}=e^{kt+C_1}+20=C_2e^{kt}+20$$
+where $C_2=e^{C_1}$.
 
-$$\ln{(T_{body}-20)}=kt+C$$
-$$T_{body}=e^{kt+C}+20=Ce^{kt}+20$$
-
-Since $T_{body}=C+20$ at time $t=0$, we can write the equation as
+Since $T_{body}=C_2+20$ at time $t=0$, we can write the equation as
 $$T_{body}(t)=(T_0-20)e^{kt}+20$$
@@
 $$\int \frac{1}{x}dx=\int-t\,dt$$
-$$\ln{(x)}=-\frac{t^2}{2}+C$$
+$$\ln{(x)}=-\frac{t^2}{2}+C_1$$
 
-$$x(t)=Ce^{-t^2/2}$$
-is our general solution to the differential equation.
+$$x(t)=C_2e^{-t^2/2}$$
+is our general solution to the differential equation, where $C_2=e^{C_1}$.
 Now we plug in the initial conditions:
-$$1/\sqrt{\pi}=Ce^{-(0^2/2)}=Ce^0=C$$
+$$1/\sqrt{\pi}=C_2e^{-(0^2/2)}=C_2e^0=C_2$$
 
-$$x(t)=\frac{e^{-t^2/2}}{\pi}$$
+$$x(t)=\frac{e^{-t^2/2}}{\sqrt{\pi}}$$
 
@@
 
-$$ \frac{u^3}{3}=\frac{1}{9}\sin(t)+C$$
+$$ \frac{u^3}{3}=\frac{1}{9}\sin(t)+C_1$$
 
-$$ u^3=\frac{\sin(t)}{3}+C$$
+$$ u^3=\frac{\sin(t)}{3}+C_2$$
 
+Where $C_2=3C_1$.
 Now we plug in the initial conditions:
-$$ 2^3=\frac{\sin(0)}{3}+C=0+C$$
-So $C=8$ and we have
+$$ 2^3=\frac{\sin(0)}{3}+C_2=0+C_2$$
+So $C_2=8$ and we have
 $$ u(t)=\sqrt[3]{\frac{\sin(t)}{3}+8}$$

### 2024-01-26 bio_dynamics/bio274_filing_cabinet.Rmd
started hw2

```{r}
x=c(100:5000)

fish1 <- tibble(x=x,
                orig=0.3*(1-(x/2500)),
                constant=0.3*(1-(x/2500))-100,
                third=0.3*(1-(x/2500))-x/3,
                prop=0.3*(1-(x/2500))-0.5*sqrt(x))
fish1 <- fish1 %>% pivot_longer(cols=2:5, names_to = "eqn", values_to = "dP_dt")

ggplot()+
 # geom_line(data=fish1, aes(x=x, y=dP_dt, color=eqn))+
  geom_line(data=fish1 %>% filter(eqn=="orig"), aes(x=x, y=dP_dt, color=eqn))+
  theme_bw()+
  labs(x="P", y="dP/dt", color="type")
```

### 2024-01-26 bio_dynamics/bio274_hw2.Rmd
started hw2

---
title: "BIOL 274 Homework 2"
author: "Ruby Krasnow"
date: "2024-01-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1
Suppose a species of fish in a particular lake has a population that is modeled by the Logistic
Growth model with $r = 0.3, K = 2500$, and $P_0 = 2500$. Adjust the model and write down an
IVP to represent each of the following situations.

>a. One hundred fish are harvested each year.
>b. One-third of the fish population is harvested annually.
>c. The number of fish harvested each year is proportional to the square root of the number
of fish in the lake.

We will assume that, in the absence of harvesting, the fish population at time t, $p(t)$, can be modeled by a discrete logistic growth equation with a time step of $\Delta t=1$ year. Given an initial population size of $P(0)=P_0 = 2500$, we have the following IVP:

\begin{equation}
    \begin{cases}
      \frac{dP}{dt}=0.3P \left( 1- \frac{P}{2500} \right) \\
      P(0)=2500
    \end{cases}       
\end{equation}

Although since we are working in discrete time, it might be more accurate to write the change in population as a difference equation, $P_{n+1}-P_n=\Delta P= 0.3P \left( 1- \frac{P}{2500} \right)$. For now we will continue using differential equation notation and apologize to Leibniz.

The initial conditions will remain the same for each of the three situations, but we can adjust the differential equation for each scenario as follows:

> **a.** One hundred fish are harvested each year:

$$\frac{dP}{dt}=0.3P \left( 1- \frac{P}{2500} \right)-100$$

> **b.** One-third of the fish population is harvested annually:

$$\frac{dP}{dt}=0.3P \left( 1- \frac{P}{2500} \right)-\frac{P}{3}$$

> **c.** The number of fish harvested each year is proportional to the square root of the number
of fish in the lake:

$$\frac{dP}{dt}=0.3P \left( 1- \frac{P}{2500} \right)-\alpha \sqrt{P}$$
Where $\alpha$ is a constant of proportionality.

Now we're going to have some fun and see how the population would change over a 100-year period if no harvesting occurred and under each of these three harvesting strategies! We will use the numerical ODE solver from the `deSolve` package.

```{r}

library(deSolve) # best package for solving ODEs in R

#First we define our state variables
state_fish <- c(Po = 2500, #original - no harvest
                Pc = 2500, #constant harvest (100 fish/year)
                P3 = 2500, #1/3 of the fish are harvested every year
                Ps = 2500) #fish are harvested prop. to the square root of P

#How many time steps (years) do we want the model to run for?
times_fish <- seq(0, 100, 1)

#Provide values for our parameters. This makes it easy to identify 
#how the population might change if we changed the carrying capacity 
#or had a different harvesting intensity in scenario C.
params_fish <- c(K=2500, #carrying capacity
                 a=0.5) #constant that scales the harvest intensity in scenario C
```

```{r}
#Define the model structure and the differential equations
fish_model <- function(t, state, parameters) {

  with(as.list(c(state, parameters)), { 
    dPo_dt <- 0.3*(1-(Po/K)) #original - no harvest
    dPc_dt <- 0.3*(1-(Pc/K))-100 #constant harvest (100 fish/year)
    dP3_dt <- 0.3*(1-(P3/K))-P3/3 #1/3 of the fish are harvested every year
    dPs_dt <- 0.3*(1-(Ps/K))-a*sqrt(Ps) #fish are harvested prop. to the square root of P

  #output
    return(list(c(dPo_dt, dPc_dt, dP3_dt, dPs_dt)))
  })
}

#Run the model
ode_output_fish <- ode(y = state_fish, t = times_fish, func = fish_model, parms = params_fish)

#Convert model output to a better format for visualization
fish_output <- ode_output_fish %>% 
  as.data.frame() %>% 
  pivot_longer(cols=2:5, names_to = "eqn", values_to = "P") %>% 
  mutate(names_full = case_when(
    eqn =="Po" ~ "No harvesting",
    eqn =="Pc" ~ "100 fish",
    eqn =="P3" ~ "1/3 of P",
    eqn =="Ps"~"0.5*sqrt(P)")
  )
```

Let's see what we got!
```{r}
ggplot()+
  geom_line(data=fish_output, aes(x=time, y=P, color=names_full))+
  theme_bw()+
  labs(x="Time (years)", y="Pop. size", color=NULL)
```

```{r}
#How many fish are left after 100 years?
fish_output %>% slice_tail(n=4)
```

Cool! So in the original formulation, the population doesn't change because our initial conditions had P equal to our carrying capacity, so the $\left( 1- \frac{P}{K} \right)$ term was equal to zero and therefore $\frac{dP}{dt}=0$.

When we harvest 100 fish per year, the population continues to decrease even beyond zero. To improve this model, we would need to add an if statement so that no harvesting occured once the population was at zero, since we cannot have negative fish.

```{r}
 fish_output %>% filter(eqn=="P3", P<1) %>% head()
```
Harvesting 1/3 of the population each year is not a sustainable management strategy, because we are basically at zero fish only 31 years after we started.

Harvest scenario 3 actually did pretty well, and the impact on the population would be even smaller if we changed the values of $\alpha$. If I was doing a whole project on these equations, it would be fun to test different values of $K$, $P_0$, and $\alpha$ and test the impacts on the population, dive into how the bifurcation points and stability of the solutions might change, and explore how we could use the model to identify an optimal harvest strategy.

## Question 2
Using Separation of Variables, find a solution to the IVP representing the situation in #1b.

Recall that the IVP for the situation in #1b is:
\begin{equation}
    \begin{cases}
      \frac{dP}{dt}=0.3P \left( 1- \frac{2500}{P} \right)-\frac{P}{3} \\
      P(0)=2500
    \end{cases}       
\end{equation}

$$\frac{dP}{dt}=\frac{3P}{10} \left( 1- \frac{P}{2500} \right)-\frac{P}{3}$$
$$dP\, \frac{1}{\frac{3P}{10} \left( 1- \frac{P}{2500} \right)-\frac{P}{3}}=dt$$
$$dP\, \frac{1}{\frac{3P}{10} - \frac{3P^2}{25000}-\frac{P}{3}}=dt$$
$$dP\, \frac{1}{P \left( \frac{3}{10} - \frac{3P}{25000}- \frac{1}{3} \right)} = dt $$
$$dP\, \frac{1}{P \left( \frac{9}{30} - \frac{3P}{25000}- \frac{10}{30} \right)} = dt $$
$$dP\, \frac{1}{P \left(- \frac{3P}{25000}- \frac{1}{30} \right)} = dt $$

Now we use partial fraction decomposition to rewrite the fraction, assuming there exist some $A, B$ such that
$$\frac{1}{P \left(- \frac{3P}{25000}- \frac{1}{30} \right)}=
\frac{A}{- \frac{3P}{25000}- \frac{1}{30}} + \frac{B}{P} $$

$$AP+B\left(- \frac{3P}{25000}- \frac{1}{30} \right)=1 $$
$$AP-\frac{3BP}{25000}-\frac{B}{30}=1 $$
$-\frac{B}{30}=1$, so $B=-30$.
$$P\left( A-\frac{3B}{25000} \right) = P\left( A+\frac{9}{2500} \right)=0 $$
$A=-\frac{9}{2500}$, so now
$$\frac{A}{- \frac{3P}{25000}- \frac{1}{30}} + \frac{B}{P}=$$
$$\frac{-9}{- \frac{3P}{10}- \frac{250}{3}} + \frac{-30}{P}$$
$$ \left( \frac{9}{ \frac{3P}{10}+ \frac{250}{3}} - \frac{30}{P} \right) dP= dt$$
$$ \left( \frac{9}{ \frac{9P+2500}{30}} - \frac{30}{P} \right) dP= dt$$
$$270 \int \frac{1}{ 9P+2500} dP-30\int  \frac{1}{P}\,dP= \int dt$$
$$30\ln{(9P+2500)}-30\ln(P)= t+C_1$$
$$\ln{(9P+2500)}-\ln(P)= \frac{t+C_1}{30}$$
$$\ln{\left(\frac{9P+2500}{P}\right)}= \frac{t+C_1}{30}$$
$$\frac{9P+2500}{P}= e^{\frac{t}{30}+\frac{C_1}{30}}=C_2e^{\frac{t}{30}}$$

$$9P+2500=PC_2e^{\frac{t}{30}}$$
$$9P-PC_2e^{\frac{t}{30}}=-2500$$
$$P\left (9-C_2e^{\frac{t}{30}}\right)=-2500$$

$$P=\frac{2500}{C_2e^{\frac{t}{30}}-9}$$

Now using our initial conditions,
$$2500=\frac{2500}{C_2-9}$$
$C_2-9=1$, so $C_2=10$

$$P=\frac{2500}{10e^{\frac{t}{30}}-9}$$

### 2024-01-27 bio_dynamics/bio274_hw2.Rmd
problem 2 hw2

@@
   with(as.list(c(state, parameters)), { 
-    dPo_dt <- 0.3*(1-(Po/K)) #original - no harvest
-    dPc_dt <- 0.3*(1-(Pc/K))-100 #constant harvest (100 fish/year)
-    dP3_dt <- 0.3*(1-(P3/K))-P3/3 #1/3 of the fish are harvested every year
-    dPs_dt <- 0.3*(1-(Ps/K))-a*sqrt(Ps) #fish are harvested prop. to the square root of P
+    dPo_dt <- 0.3*Po*(1-(Po/K)) #original - no harvest
+    dPc_dt <- 0.3*Pc*(1-(Pc/K))-100 #constant harvest (100 fish/year)
+    dP3_dt <- 0.3*P3*(1-(P3/K))-P3/3 #1/3 of the fish are harvested every year
+    dPs_dt <- 0.3*Ps*(1-(Ps/K))-a*sqrt(Ps) #fish are harvested prop. to the square root of P
 
@@
 
-When we harvest 100 fish per year, the population continues to decrease even beyond zero.
-To improve this model, we would need to add an if statement so that no harvesting occured once the population was at zero, since we cannot have negative fish.
-
-```{r}
- fish_output %>% filter(eqn=="P3", P<1) %>% head()
-```
-Harvesting 1/3 of the population each year is not a sustainable management strategy, because we are basically at zero fish only 31 years after we started.
+Harvesting 1/3 of the population each year is not a sustainable management strategy, because we have decreased the population to almost zero.
 
@@
 Recall that the IVP for the situation in #1b is:
-\begin{equation}
+\begin{equation}\label{IVP}
     \begin{cases}
@@
 
-$$P=\frac{2500}{10e^{\frac{t}{30}}-9}$$
+Our final solution is therefore:
+\begin{equation}\label{solution}P(t)=\frac{2500}{10e^{\frac{t}{30}}-9}\end{equation}
+Here is what it looks like:
+```{r}
+t=c(1:250)
+y=2500/(10*exp(t/30)-9)
+
+ggplot()+
+  geom_line(aes(x=t, y=y, color="P(t)"))+
+  theme_bw()+
+  labs(x="time", y="P")
+```
+
+Let's check and make sure it is actually a solution and satisfies our initial conditions:
+
+We want to show that the derivative with respect to $t$ of (\ref{solution}) is equivalent to $\frac{3P}{10} \left( 1- \frac{P}{2500} \right)-\frac{P}{3}$.
+
+Using the chain rule, we differentiate (\ref{solution}) to get:
+$$P'=2500(-1) \left( 10e^{\frac{t}{30}}-9 \right)^{-2} \left( 10e^{\frac{t}{30}} \right) \left(\frac{1}{30}\right)=\frac{-2500e^{\frac{t}{30}}}{3 \left( 10e^{\frac{t}{30}}-9 \right)^{2}}$$
+Since there are no $t$'s in the original $\frac{dP}{dt}$, we can use our solution to rewrite them as a function of $P$:
+$$P\left( 10e^{\frac{t}{30}}-9\right)=2500$$
+$$10e^{\frac{t}{30}}=\frac{2500}{P}+9$$
+$$e^{\frac{t}{30}}=\frac{250}{P}+\frac{9}{10}$$
+Now substituting this into our equation for $P'$, we get
+$$\frac{-2500 \left( \frac{250}{P}+\frac{9}{10} \right)}
+{3 \left( 10\left( \frac{250}{P}+\frac{9}{10} \right)-9 \right)^{2}} = 
+\frac{-2500 \left( \frac{250}{P}+\frac{9}{10} \right)}
+{3 \left( \frac{2500}{P}+9-9 \right)^{2}} = 
+\frac{-2500 \left( \frac{250}{P}+\frac{9}{10} \right)}
+{3 \left( \frac{2500^2}{P^2} \right)} = 
+\frac{ \frac{-250}{P}-\frac{9}{10}}
+{3 \left( \frac{2500}{P^2} \right)}$$
+Now multiply the numerator and denominator by $P$:
+
+$$\frac{-250-\frac{9P}{10}}
+{\frac{7500}{P}} = \left(-250-\frac{9P}{10}\right)\left( \frac{P}{7500}\right)=
+\frac{-250P}{7500}-\frac{9P^2}{75000}=P \left(- \frac{3P}{25000}- \frac{1}{30} \right)$$
+
+And see that our initial $\frac{dP}{dt}$,
+$$\frac{3P}{10} \left( 1- \frac{P}{2500} \right)-\frac{P}{3} = 
+\frac{3P}{10} - \frac{3P^2}{2500}-\frac{P}{3} =
+P \left(- \frac{3P}{25000}- \frac{1}{30} \right)$$
+
+So indeed, our solution satisfies the differential equation.
+
+And checking the initial conditions:
+$$P(0)=\frac{2500}{10e^{0}-9}=\frac{2500}{1}=2500$$
+We can also confirm by using R to calculate the derivative of our solution and plotting it alongside the original expression for $\frac{dP}{dt}$:
+
+```{r}
+dy=D(expression(2500/(10*exp(t/30)-9)), "t")
+
+ggplot()+
+  geom_line(aes(x=y, y=eval(dy), color="calculated"), linewidth=3, alpha=0.5)+
+  geom_line(aes(x=y, y= 0.3*y*(1-(y/2500))-y/3, color="original dP/dt"))+
+  scale_color_manual(values=c("lightblue", "black"))+
+  theme_bw()+
+  labs(x="P", y="dP/dt")
+```
+
+## Question 2
+Using Bernoulli's method, find a solution to the IVP representing the situation in #1b.
+
+$$\frac{dP}{dt} = \frac{3P}{10} - \frac{3P^2}{25000}-\frac{1P}{30}$$
+
+Bernoulli's method review:
+If the differential equation is in the form $y'(x)+p(x)y=g(x)y^n$,
+we can divide by $y^n$ and then use the substitution $z=y^{1-n}$.
+
+Here,
+$$P'(t)+\frac{1P}{30}=-\frac{3P^2}{25000}$$
+Divide by $P^2$:
+$$P'(P^{-2})+\frac{1}{30P}=-\frac{3}{25000}$$
+Use the substitution $z=P^{-1}$, $z'=-P^{-2}P'$:
+$$-z'+\frac{z}{30}=-\frac{3}{25000}$$
+$$z'-\frac{z}{30}=\frac{3}{25000}$$
+$$\mu=e^{\int \frac{-1}{30} dt}=e^{\frac{-t}{30}}$$
+
+$$\int \left( e^{\frac{-t}{30}}z \right)'dt=\int \frac{3e^{\frac{-t}{30}}}{25000} dt $$
+
+$$ e^{\frac{-t}{30}}z=\frac{-9e^{\frac{-t}{30}}}{2500}+C_1$$
+
+$$z=\frac{-9}{2500}+\frac{C_1}{e^{\frac{-t}{30}}}$$
+
+$$P=1/z= \frac{1}{\frac{-9}{2500}+C_1e^{\frac{t}{30}}}$$
+
+$$P(0)= 2500=\frac{1}{\frac{-9}{2500}+C_1}$$
+
+$$2500 \left( \frac{-9}{2500}+C_1 \right)=1$$
+
+$$-9+2500 C_1=1$$
+
+Then $2500 C_1=10$, so $C_1=\frac{1}{250}$
+
+$$P(t)= \frac{1} {\frac{-9}{2500}+\left( \frac{1}{250}\right) e^{\frac{t}{30}}}=
+\frac{2500} {10e^{\frac{t}{30}}-9}$$
+
+So we did get the same answer as when we used separation of variables.

### 2024-01-27 bio_dynamics/bio274_hw2.Rmd
q3 hw2

@@
 author: "Ruby Krasnow"
-date: "2024-01-26"
+date: "2024-01-27"
 output: pdf_document
@@
 
-## Question 2
+## Question 2 - for practice
 Using Bernoulli's method, find a solution to the IVP representing the situation in #1b.
@@
 So we did get the same answer as when we used separation of variables.
+
+## Question 3
+Figure 1.3.1 below is from your textbook. It shows an example of the logistic growth curve
+when $p(0) << N_0/a$. How would this curve change if $N_0 < aP(0)$?
+
+$$P(t)=\frac{N_0P_0}{aP_0+\left(N_0-aP_0\right)e^{-N_0t}}$$
+
+```{r}
+N0 = 0.1
+a=0.003
+p0_small=10
+p0_big=50
+
+x<-c(1:100)
+
+ggplot()+
+  geom_hline( linetype="dashed", aes(color="N0/a",yintercept=N0/a))+
+  geom_line(aes(x=x, y=(N0*p0_small)/(a*p0_small+(N0-a*p0_small)*exp(-N0*x)), color="p0 << N0/a"))+
+  geom_line(aes(x=x, y=(N0*p0_big)/(a*p0_big+(N0-a*p0_big)*exp(-N0*x)), color="p0 > N0/a"))+
+  theme_bw()+
+  labs(x="t", y="P", color=NULL)+
+  scale_color_manual(values=c("gray", "#0f85a0", "#dd4124"))+
+  ylim(0, NA)
+```
+
+If $N_0 < aP(0)$, that means that $N_0/a < P(0)$, or equivalently, $P(0)>N_0/a$.
+This is saying that our initial population size $P(0)=P_0$ is greater than our carrying capacity $N_0/a$, so the population size will decrease until it reaches the asymptote at $K=N_0/a$.
+
+## Question 4 & 5
+For problems 4 and 5, use Euler’s Method with the given step size, h, to approximate the solution
+to the given IVP over the time interval specified. Your answer should include a table of
+approximate values of the dependent variable. It should also include a sketch of the graph
+of the approximate solution.

### 2024-01-29 PDEs/math245_hw1.Rmd
PDE hw1

---
title: "MATH 245 Homework 1"
author: "Ruby Krasnow"
date: "2024-01-28"
output:
  pdf_document:
    extra_dependencies: ["derivative"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1
Show the function is a solution of the PDE:  
(a) $u_{xx}+u_{yy}=0$   
(i) $u(x,y)=e^x\sin(y)$  
$$u_x=e^x\sin(y), \, u_{xx}=e^x\sin(y), \, u_y=e^x\cos(y),\, u_{yy}=-e^x\sin(y)$$
Which means that $u_{xx}+u_{yy}=e^x\sin(y)+\left(-e^x\sin(y)\right)=0$ and so $u(x,y)=e^x\sin(y)$ is
a solution to the PDE.  
(ii) $u(x,y)=\log{\sqrt{x^2+y^2}}$  
Assuming this is meant to be $u(x,y)=\ln\sqrt{{{x^2+y^2}}}$,
$$u_x=\left({{x^2+y^2}}\right)^{-1/2}
\left(\frac{1}{2}\right)\left({{x^2+y^2}}\right)^{-1/2}
\left(2x\right)=x\left({{x^2+y^2}}\right)^{-1}$$

$$u_{xx}=x(-1)\left({{x^2+y^2}}\right)^{-2}\left(2x\right)+\left({{x^2+y^2}}\right)^{-1}$$

$$u_y=\left({{x^2+y^2}}\right)^{-1/2}
\left(\frac{1}{2}\right)\left({{x^2+y^2}}\right)^{-1/2}
\left(2y\right)=y\left({{x^2+y^2}}\right)^{-1}$$

$$u_{yy}=y(-1)\left({{x^2+y^2}}\right)^{-2}\left(2y\right)+\left({{x^2+y^2}}\right)^{-1}$$

$$u_{xx}+u_{yy}=\frac{-2x^2}{\left({{x^2+y^2}}\right)^{2}}+
\frac{-2y^2}{\left({{x^2+y^2}}\right)^{2}}+
\frac{2}{\left({{x^2+y^2}}\right)}=$$

$$\frac{-2\left({{x^2+y^2}}\right)}{\left({{x^2+y^2}}\right)^{2}}+ \frac{2}{\left({{x^2+y^2}}\right)}=0$$
Which means that $u(x,y)=\log{\sqrt{x^2+y^2}}$ is a solution to the PDE.

(b) $bu_x+au_y+u=0, \, u(x,y)=\exp{\left(\frac{-x}{b}\right)}f(ax-by)$ for arbitrary differentiable function $f$.
$$u_x=e^{\frac{-x}{b}} \left(\frac{-1}{b}\right)f(ax-by)+f'(ax-by)(a)e^{\frac{-x}{b}}$$
$$u_y=e^{\frac{-x}{b}}f'(ax-by)(-b)$$
Then,
$$bu_x+au_y+u=-e^{\frac{-x}{b}}f(ax-by)+(ab)\left(e^{\frac{-x}{b}}\right)f'(ax-by)+
e^{\frac{-x}{b}}f'(ax-by)(-ab)+e^{\frac{-x}{b}}f(ax-by)=0$$

So $u(x,y)=\exp{\left(\frac{-x}{b}\right)}f(ax-by)$ is a solution to the PDE for an arbitrary differentiable function $f$.

(c) $u_{xx}-\frac{1}{x}u_x-x^2u_{yy}=0, \, u(x,y)=f\left(2y+x^2\right)+g\left(2y-x^2\right)$ for arbitrary twice-differentiable functions $f$ and $g$.

$$u_x=\left(f'\left(2y+x^2\right)\right)(2x)+\left(g'\left(2y-x^2\right)\right)(-2x)$$
$$u_y=2 \left(f'\left(2y+x^2\right)\right) +2\left(g'\left(2y-x^2\right)\right)$$
$$u_{xx}=2f'+(f'')(4x^2)-2g'+(g'')(4x^2)$$
$$u_{yy}=4f''+4g''$$
$$u_{xx}-\frac{1}{x}u_x-x^2u_{yy}=$$ $$2f'+4x^2f''-2g'+4x^2g''-\frac{1}{x}\left(2xf'-2xg'\right)-x^2\left(4f''+4g''\right)=$$
$$2f'-2f'+4x^2f''-4x^2f''-2g'+2g'+4x^2g''-4x^2g''=0 $$

So $u(x,y)=f\left(2y+x^2\right)+g\left(2y-x^2\right)$ is a solution to the PDE for arbitrary twice-differentiable functions $f$ and $g$.

### Question 2
(a) 2nd-order linear homogenous  
(b) 4th-order linear inhomogenous  
(c) 2nd-order quasi-linear homogeneous  
(d) We can rewrite this as $u^{-5}u_{xx}+u^{-5}u_{yy}+u^{-4}f(x,y)=g(x,y)$ to see that it is a 2nd-order quasi-linear inhomogeneous PDE.

### Question 3
Use separation of variables to solve the following problems:  
#### (a)
$u_x+u=u_y, \, u(x,0)=4x^{-3x}$, use $u(x,y)=f(x)g(y)$

#### (b)
$x^2u_{xy}+9y^2u=0, \, u(x,0)=\exp{\left(\frac{1}{x}\right)}$, use $u(x,y)=f(x)g(y)$

#### (c)
$u_x^2+u_y^2=1$, use $u(x,y)=f(x)+g(y)$

#### (d)
$x^2u_x^2+y^2u_y^2=u^2$, use $u(x,y)=e^{f(x)}e^{g(y)}$

### Question 4
For each of the following IVPs, (i) find and plot the characteristic lines (curves), (ii) solve the IVP, and (iii) plot the solution of (a)-(b) for indicated time.  
#### (a)
$u_t+\left(1+x^2\right)u_x=0, \, u(0,x)=\arctan{(x)}, \, t=1,2,3$, and
what is $\lim\limits_{x \to \infty}u(t,x)$?

$$\frac{dx}{dt}=1+x^2$$
$$\frac{1}{1+x^2}dx=dt$$

$$\arctan{(x)}=t+C$$
So $x=\tan{(t+C)}$ are the characteristic curves of $u_t+\left(1+x^2\right)u_x=0$.  
On each of the curves, $u(x,t)$ is constant because $\frac{d}{dt}u(t,\tan{(t+C)}) = \pdv {u}{t}+\pdv{u}{x} \sec^2{(t+c)}$.  
And since $1+\tan^2\theta=\sec^2\theta$, this means $\frac{d}{dt}u(t,\tan{(t+C)}) = \pdv {u}{t}+\pdv{u}{x}\left( 1+\tan^2{(t+c)}\right)=u_t+\left(1+x^2\right)u_x$, which we know is 0.

Thus $u(t, \tan{(t+C)})=u(0, \tan{(0+C)})=u(0, C)$ is independent of $t$. Putting $x=\tan{(t+C)}$ and $C=\arctan{(x)}-t$, we have
$$u(t, x)=u(0,\arctan{(x)}-t)$$
It follows that $u(t,x)=f(\arctan{(x)}-t)$. And since we are given that $u(0,x)=\arctan{(x)}$, we have $f(\arctan{(x)}-0)=\arctan{(x)}$ so that $f(w)=w$ for any $w$, yielding our solution of 
$$u(t,x)=\arctan{(x)}-t$$

```{r}
C=1
df_4a <- tibble(x = seq(-2*pi, 2*pi, length.out = 1000),
                y=tan(x+C)) %>% 
  mutate(y = ifelse(abs(y) < 20, y, NA))

ggplot(df_4a, aes(x=x, y=y)) + 
  geom_line()+
  coord_cartesian(ylim=c(-10,10), xlim = c(-5,5))+
  theme_bw()+
  labs(x="t", y="x")+
  ggtitle("Characteristic curves for Problem 4a", subtitle = "with C=1")
```

Since the range of $\arctan$ is limited to $\left(\frac{-\pi}{2},\frac{\pi}{2}\right)$, as $\lim\limits_{t \to \infty}$ that means $u(t,x)=\arctan(x)-t$ will go to $-\infty$.

```{r}

x=seq(-10,10,length.out = 1000)
ggplot() + 
  geom_line(aes(x=x, y=atan(x)-1, color="1"))+
  geom_line(aes(x=x, y=atan(x)-2, color="2"))+
  geom_line(aes(x=x, y=atan(x)-3, color="3"))+
  theme_bw()+
  labs(x="x", y="u(x,t)", color="Value of t")+
  ggtitle("Solution curves for Problem 4a")
```

### 2024-01-29 PDEs/math245_hw1.Rmd
problem 4a

@@
 author: "Ruby Krasnow"
-date: "2024-01-28"
+date: "2024-01-29"
 output:
@@
 Use separation of variables to solve the following problems:
+
 #### (a)
@@
 For each of the following IVPs, (i) find and plot the characteristic lines (curves), (ii) solve the IVP, and (iii) plot the solution of (a)-(b) for indicated time.
+
 #### (a)
 $u_t+\left(1+x^2\right)u_x=0, \, u(0,x)=\arctan{(x)}, \, t=1,2,3$, and
-what is $\lim\limits_{x \to \infty}u(t,x)$?
+what is $\lim\limits_{t \to \infty}u(t,x)$?
 
@@
 So $x=\tan{(t+C)}$ are the characteristic curves of $u_t+\left(1+x^2\right)u_x=0$.  
-On each of the curves, $u(x,t)$ is constant because $\frac{d}{dt}u(t,\tan{(t+C)}) = \pdv {u}{t}+\pdv{u}{x} \sec^2{(t+c)}$.  
+On each of the curves, $u(x,t)$ is constant because $$\frac{d}{dt}u(t,\tan{(t+C)}) = \pdv {u}{t}+\pdv{u}{x} \sec^2{(t+c)}$$  
 And since $1+\tan^2\theta=\sec^2\theta$, this means $\frac{d}{dt}u(t,\tan{(t+C)}) = \pdv {u}{t}+\pdv{u}{x}\left( 1+\tan^2{(t+c)}\right)=u_t+\left(1+x^2\right)u_x$, which we know is 0.
 
-Thus $u(t, \tan{(t+C)})=u(0, \tan{(0+C)})=u(0, C)$ is independent of $t$. Putting $x=\tan{(t+C)}$ and $C=\arctan{(x)}-t$, we have
+Thus $u(t, \tan{(t+C)})=u(0, \tan{(0+C)})=u(0, C)$ is independent of $t$.  
+Putting $x=\tan{(t+C)}$ and $C=\arctan{(x)}-t$, we have
 $$u(t, x)=u(0,\arctan{(x)}-t)$$
-It follows that $u(t,x)=f(\arctan{(x)}-t)$. And since we are given that $u(0,x)=\arctan{(x)}$, we have $f(\arctan{(x)}-0)=\arctan{(x)}$ so that $f(w)=w$ for any $w$, yielding our solution of 
+It follows that $u(t,x)=f(\arctan{(x)}-t)$.  
+And since we are given that $u(0,x)=\arctan{(x)}$, we have $f(\arctan{(x)}-0)=\arctan{(x)}$ so that $f(w)=w$ for any $w$, yielding our solution of 
 $$u(t,x)=\arctan{(x)}-t$$
@@
 ```
+
+We can check our solution of $u(t,x)=\arctan{(x)}-t$ by differentiating, since $u_t=-1, u_x=\frac{1}{1+x^2}$, that means $u_t+\left(1+x^2\right)u_x=-1+1=0$.
+
+#### (b)
+$u_t-xu_x=0, \, u(0,x)=\frac{1}{1+x^2}, \, t=1,2,3$, and
+what is $\lim\limits_{t \to \infty}u(t,x)$?
+
+The directional derivative of $u$ in the direction of the vector $(1, -x)$ is zero. The curves in the $tx$ plane with $(1, -x)$ as a tangent vector have slopes $-x$:
+
+$$\frac{dx}{dt}=-x$$
+$$\frac{1}{x}dx=-dt$$
+Solving this ODE gives the equations for the characteristic lines:
+$$x=Ce^{-t}$$
+
+```{r, warning = FALSE}
+curves_4a <- tibble(C=c(1:5), seq=list(x)) %>% 
+  rowwise() %>% 
+  mutate(y=list(C*exp(-seq))) %>% 
+  unnest(c(seq,y))
+
+ggplot(data=curves_4a)+
+  geom_line(aes(x=seq, y=y, color=as.factor(C)))+
+  ylim(NA, 100)+
+  xlim(-5, NA)+
+  labs(x="t", y="x", color="C")+
+  theme_light()+
+  ggtitle("Characteristic curves for Problem 4b")
+```
+
+On each of the curves, $u(x,t)$ is constant because $$\frac{d}{dt}u(t,Ce^{-t}) = \pdv {u}{t}-Ce^{-t}\pdv{u}{x}=u_t-xu_x$$  
+and we know that $u_t-xu_x=0$.
+
+Set $\xi=xe^t$ and $\eta=x$. Then by the chain rule,
+$$u_x= \pdv{u}{x}=\pdv{u}{\eta}\pdv{\eta}{x}+\pdv{u}{\xi}\pdv{\xi}{x}=$$
+$$\pdv{u}{\eta}+\pdv{u}{\xi}e^t $$
+
+and similarly
+$$u_t= \pdv{u}{t}=\pdv{u}{\eta}\pdv{\eta}{t}+\pdv{u}{\xi}\pdv{\xi}{t}=\pdv{u}{\xi}xe^t $$
+
+Which means that 
+$$u_t-xu_x=\pdv{u}{\xi}xe^t -x\left(\pdv{u}{\eta}+\pdv{u}{\xi}e^t \right)=0$$
+Assuming $x\neq0$, this means
+$$u_\eta=0$$
+$$u=\int u_\eta \, d\eta \> + f(\xi)=0+ f(\xi)$$
+
+So we have $u(t,x)=f(\xi)=f(xe^t)$ and we know that $u(0,x)=\frac{1}{1+x^2}$, which means $f(xe^0)=f(x)=\frac{1}{1+x^2}$ and therefore $u(t,x)=\frac{1}{1+\left(xe^t\right)^2}$.
+
+We check by differentiating:
+$$u_t=-\left(1+\left(xe^t\right)^2\right)^{-2}\left(2xe^t\right)xe^t$$
+$$u_x=-\left(1+\left(xe^t\right)^2\right)^{-2}\left(2xe^t\right)e^t$$
+And we see that $u_t-xe_t$ indeed equals zero.
+Plotting the solution $u(t,x)=\frac{1}{1+\left(xe^t\right)^2}$ as a function of $x$ for $t=1,2,3$, we get:
+```{r, warning=FALSE}
+
+x=seq(-10,10,length.out = 1000)
+ggplot() + 
+  geom_line(aes(x=x, y=1/(1+(exp(1)*x)^2), color="1"))+
+  geom_line(aes(x=x, y=1/(1+(exp(2)*x)^2), color="2"))+
+  geom_line(aes(x=x, y=1/(1+(exp(3)*x)^2), color="3"))+
+  theme_bw()+
+  labs(x="x", y="u(x,t)", color="Value of t")+
+  xlim(-5,5)+
+  ggtitle("Solution curves for Problem 4b")
+```
+
+We can see that $u(t,x)$ will equal 1 at $x=0$ regardless of the value of $t$, but at all other values of $x$, $\lim\limits_{t \to \infty}u(t,x)=0$.

### 2024-01-29 PDEs/math245_hw1.Rmd
worked on hw1 PDEs

@@
 title: "MATH 245 Homework 1"
-author: "Ruby Krasnow"
+author: "Ruby Krasnow and Tommy Thach"
 date: "2024-01-29"
@@
 ### Question 2
-(a) 2nd-order linear homogenous  
-(b) 4th-order linear inhomogenous  
+(a) 2nd-order linear homogeneous  
+(b) 4th-order linear inhomogeneous  
 (c) 2nd-order quasi-linear homogeneous  
-(d) We can rewrite this as $u^{-5}u_{xx}+u^{-5}u_{yy}+u^{-4}f(x,y)=g(x,y)$ to see that it is a 2nd-order quasi-linear inhomogeneous PDE.
+(d) We can rewrite this as $u_{xx}+u_{yy}+f(x,y)u-g(x,y)u^5=0$, making it clear that this is a 2nd-order semi-linear homogeneous PDE.
 
@@
 
-```{r}
+```{r, echo=FALSE}
 C=1
@@
 
-```{r}
+```{r, echo=FALSE}
 
@@
 
-```{r, warning = FALSE}
-curves_4a <- tibble(C=c(1:5), seq=list(x)) %>% 
+```{r, warning = FALSE, echo=FALSE}
+curves_4b <- tibble(C=c(1:5), seq=list(x)) %>% 
   rowwise() %>% 
@@
 
-ggplot(data=curves_4a)+
+ggplot(data=curves_4b)+
   geom_line(aes(x=seq, y=y, color=as.factor(C)))+
-  ylim(NA, 100)+
-  xlim(-5, NA)+
+  coord_cartesian(ylim=c(NA, 100), xlim=c(-5, 9))+
   labs(x="t", y="x", color="C")+
@@
 Plotting the solution $u(t,x)=\frac{1}{1+\left(xe^t\right)^2}$ as a function of $x$ for $t=1,2,3$, we get:
-```{r, warning=FALSE}
+
+```{r, warning=FALSE, echo=FALSE}
 
@@
 
-We can see that $u(t,x)$ will equal 1 at $x=0$ regardless of the value of $t$, but at all other values of $x$, $\lim\limits_{t \to \infty}u(t,x)=0$.
+We can see that $u(t,x)$ will equal 1 at $x=0$ regardless of the value of $t$, but at all other values of $x$, $\lim\limits_{t \to \infty}u(t,x)=0$. This can be written as
+\begin{equation}
+\lim\limits_{t \to \infty}u(t,x)=
+    \begin{cases}
+        1 & \text{if } x=0\\
+        0 & \text{if } x\neq0
+    \end{cases}
+\end{equation}
+
+#### (c)
+$u_x+xu_y=\left(y-\frac{x^2}{2}\right)^2, \, u(0,y)=e^y$
+
+$\frac{dy}{dx}=x$
+Solving this ODE yields the equation for the characteristic curves: $y=\frac{x^2}{2}+C$
+
+```{r, warning = FALSE, echo=FALSE}
+curves_4c <- tibble(C=c(1:5), seq=list(x)) %>% 
+  rowwise() %>% 
+  mutate(y=list(0.5*(seq)^2+C)) %>% 
+  unnest(c(seq,y))
+
+ggplot(data=curves_4c)+
+  geom_line(aes(x=seq, y=y, color=as.factor(C)))+
+  coord_cartesian(ylim=c(NA, 45), xlim=c(-9.5, 9.5))+
+  labs(x="x", y="y", color="C")+
+  theme_light()+
+  ggtitle("Characteristic curves for Problem 4c")
+```
+
+Now we define our new coordinate system as $\xi=y-\frac{x^2}{2}, \, \eta=x$
+
+Then by the chain rule,
+$$u_x= \pdv{u}{x}=\pdv{u}{\eta}\pdv{\eta}{x}+\pdv{u}{\xi}\pdv{\xi}{x}=$$
+$$\pdv{u}{\eta}-x\pdv{u}{\xi}$$
+
+and similarly
+$$u_y= \pdv{u}{y}=\pdv{u}{\eta}\pdv{\eta}{y}+\pdv{u}{\xi}\pdv{\xi}{y}=$$
+$$\pdv{u}{\xi}$$
+
+$$u_x+xu_y=\pdv{u}{\eta}-x\pdv{u}{\xi}+x\pdv{u}{\xi}=\pdv{u}{\eta}$$
+
+And since we know $u_x+xu_y=\left(y-\frac{x^2}{2}\right)^2$, this means we have
+$$\pdv{u}{\eta}=\left(y-\frac{x^2}{2}\right)^2=\xi^2$$
+Now we integrate with respect to $\eta$ to find
+$$u=\eta\xi^2+F(\xi)$$
+
+At $u(0,y)$, this becomes $u=0y^2+F(y)$, or $u(0,y)=F(y)$.
+On the other hand, we are told that $u(0,y)=e^y$, which means that $F(y)=e^y$.
+$$u=\eta\xi^2+e^\xi$$
+Substitute back in $x$ and $y$:
+$$u(x,y)=x\left(y-\frac{x^2}{2}\right)^2+\exp{\left(y-\frac{x^2}{2}\right)}$$
+
+We can check by differentiating:
+$$u_x=-2x^2\left(y-\frac{x^2}{2}\right)+\left(y-\frac{x^2}{2}\right)^2+\exp{\left(y-\frac{x^2}{2}\right)}(-x)$$
+$$u_y=2x\left(y-\frac{x^2}{2}\right)+\exp{\left(y-\frac{x^2}{2}\right)}$$
+
+Which means that $u_x+xu_y=\left(y-\frac{x^2}{2}\right)^2$, as we wanted.
+
+#### (d)
+$2xu_x+(x+1)u_y=y$, for $x>0, u=2y$ on $x=1$
+
+$$\frac{dx}{dy}=\frac{x+1}{2x}=\frac{1}{2}+\frac{1}{2x}$$
+Solving this ODE yields the characteristic curves:
+$$y=\frac{x}{2}+\frac{\ln(x)}{2}+C$$
+$$C=2y-x-\ln(x)$$
+
+```{r, warning = FALSE, echo=FALSE}
+curves_4d <- tibble(C=c(1:5), seq=list(x)) %>% 
+  rowwise() %>% 
+  mutate(y=list((seq/2)+(log(seq)/2)+C)) %>% 
+  unnest(c(seq,y))
+
+ggplot(data=curves_4d)+
+  geom_line(aes(x=seq, y=y, color=as.factor(C)))+
+  coord_cartesian(ylim=c(-1, 12), xlim=c(-1, 9.5))+
+  labs(x="x", y="y", color="C")+
+  theme_light()+
+  ggtitle("Characteristic curves for Problem 4d")
+```
+
+Now we define our new coordinate system as $2y-x-\ln(x), \, \eta=x$
+
+Then by the chain rule,
+$$u_x= \pdv{u}{x}=\pdv{u}{\eta}\pdv{\eta}{x}+\pdv{u}{\xi}\pdv{\xi}{x}=$$
+
+$$\pdv{u}{\eta}+\pdv{u}{\xi}\left(-1-\frac{1}{x}\right) $$
+
+Similarly, $$u_y=\pdv{u}{y}=2\pdv{u}{\xi}$$
+
+Plugging into our initial PDE,
+$$2xu_x+(x+1)u_y=2x\pdv{u}{\eta}+\pdv{u}{\xi}2x\left(-1-\frac{1}{x}\right)+2(x+1)\pdv{u}{\xi}=2x\pdv{u}{\eta}$$
+
+And since we know $2xu_x+(x+1)u_y=y$, this means $y=2x\pdv{u}{\eta}$ or
+$$\frac{\eta}{2}+\frac{\ln(\eta)}{2}+\xi=2\eta \pdv{u}{\eta}$$

### 2024-01-30 PDEs/math245_hw1.Rmd
problem 3, colored text boxes

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-01-29"
+date: "2024-01-30"
 output:
   pdf_document:
-    extra_dependencies: ["derivative"]
+    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
 ---
@@
 $u_x+u=u_y, \, u(x,0)=4x^{-3x}$, use $u(x,y)=f(x)g(y)$
+$$u_x=f'g, \, u_y=fg'$$
+$$u_x+u-u_y=f'g+fg-fg'=f'g+(g-g')f=0$$
+$$f'g=-(g-g')f$$
+$$\frac{f'}{f}=\frac{g'-g}{g}=\lambda$$
+$$\frac{f'}{f}=\lambda, \,\frac{g'}{g}-1=\lambda $$
+Integrating the first ODE with respect to $x$,
+$$\ln(f)=\lambda x+C_1, \, f(x)=C_2e^{\lambda x}$$
+$$\frac{g'}{g}=\lambda+1 $$
+$$\ln(g)=\lambda y+y+C_3,\, g(y)=C_4e^{y(\lambda +1)}$$
+$$u(x,y)=f(x)g(y)=C_2e^{\lambda x}C_4e^{y(\lambda +1)}$$
+$$u(x,y)=Ce^{\lambda x+y(\lambda +1)}$$
+$$u(x,0)=4e^{-3x}=Ce^{\lambda x}$$
+
+Which means $C=4, \lambda=-3$ and our final solution is
+\begin{tcolorbox}[colback=white, title=Solution]
+$$u(x,y)=4e^{-3x-2y}$$
+\end{tcolorbox}
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+$$u_x+u=-12e^{-3x-2y}+4e^{-3x-2y}=-8e^{-3x-2y}=u_y$$
+\end{tcolorbox}
 
@@
 
+$$u_{xy}=f'g'$$
+$$x^2f'g'+9y^2fg=0$$
+
+$$x^2f'g'=-9y^2fg$$
+$$x^2\frac{f'}{f}=-9y^2\frac{g}{g'}=\lambda$$
+$$x^2\frac{f'}{f}=\lambda, \, -9y^2\frac{g}{g'}=\lambda$$
+$$x^2\frac{f'}{f}=\lambda \quad \longrightarrow \quad 
+\frac{f'}{f}=\lambda x^{-2} \quad \longrightarrow \quad
+\ln(f)=\frac{-\lambda}{x}+C_1 \quad \longrightarrow \quad 
+f(x)=C_2e^{\frac{-\lambda}{x}}$$
+
+$$-9y^2\frac{g}{g'}=\lambda \quad \longrightarrow \quad
+\frac{g}{g'}= \frac{\lambda}{-9y^2} \quad \longrightarrow \quad 
+\frac{g'}{g}= \frac{-9y^2}{\lambda} \quad \longrightarrow \quad 
+\ln(g)=\frac{-3y^3}{\lambda}+C_3 \quad \longrightarrow \quad 
+g(y)=C_4e^{\frac{-3}{\lambda}y^3}$$
+
+$$u(x,y)=f(x)g(y)=C_2e^{\frac{-\lambda}{x}}C_4e^{\frac{-3}{\lambda}y^3}=Ce^{\frac{-\lambda}{x}+\frac{-3}{\lambda}y^3}$$
+$$u(x,0)=e^{\frac{1}{x}}=Ce^{\frac{-\lambda}{x}}$$
+Which means $C=1, \lambda=-1$, and our final solution is
+\begin{tcolorbox}[colback=white, title=Solution]
+$$u(x,y)=e^{\frac{1}{x}+3y^3}$$
+\end{tcolorbox}
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+$$x^2u_{xy}+9y^2u=x^2e^{\frac{1}{x}+3y^3}\left(-x^{-2} \right) \left(9y^2 \right)+9y^2e^{\frac{1}{x}+y^3}=0$$
+\end{tcolorbox}
+
 #### (c)
 $u_x^2+u_y^2=1$, use $u(x,y)=f(x)+g(y)$
+$$u_x=f', u_y=g'$$
+$$(f')^2+(g')^2=1\quad \longrightarrow \quad (f')^2=1-(g')^2=\lambda^2$$
+$$(f')^2=\lambda^2\quad \longrightarrow \quad f'=\lambda\quad \longrightarrow \quad f(x)=\lambda x+C_1$$
+$$(g')^2=1-\lambda^2 \quad \longrightarrow \quad g'=\pm\sqrt{1-\lambda^2}\quad \longrightarrow \quad g(y)=\pm y\sqrt{1-\lambda^2} + C_2$$
+\begin{tcolorbox}[colback=white, title=Solution]
+$$u(x,y)=f(x)+g(y)=\lambda x \pm y\sqrt{1-\lambda^2}+C$$
+\end{tcolorbox}
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+$$u_x=\lambda, u_y=\pm \sqrt{1-\lambda^2},\quad u_x^2+u_y^2= \lambda ^2 + 1 - \lambda^2 = 1$$
+\end{tcolorbox}
 
@@
 $x^2u_x^2+y^2u_y^2=u^2$, use $u(x,y)=e^{f(x)}e^{g(y)}$
+
+$$u_x= f'e^{f+g}, u_y= g'e^{f+g}$$
+$$x^2u_x^2+y^2u_y^2-u^2=x^2(f')^2e^{2f+2g}+y^2(g')^2e^{2f+2g}-e^{2f+2g}=0$$
+$$x^2(f')^2+y^2(g')^2-1=0 \quad \longrightarrow \quad x^2(f')^2=1-y^2(g')^2=\lambda^2$$
+$$x^2(f')^2= \lambda^2 \quad \longrightarrow \quad f'=\frac{\lambda}{x} \quad \longrightarrow \quad f(x)=\lambda \ln(x)+C_1$$
+$$y^2(g')^2= 1-\lambda^2 \quad \longrightarrow \quad g'=\frac{\sqrt{1-\lambda^2}}{y} \quad \longrightarrow \quad g(y)=\sqrt{1-\lambda^2}\ln(y) +C_2$$
+
+$$u(x,y)=e^{f(x)}e^{g(y)}=C_3e^{\lambda \ln(x)}C_4e^{\sqrt{1-\lambda^2}\ln(y)}$$
+
+\begin{tcolorbox}[colback=white, title=Solution]
+$$u(x,y)=C x^\lambda y^{\sqrt{1-\lambda^2}}$$
+\end{tcolorbox}
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+$$u_x=\lambda C x^{\lambda-1} y^{\sqrt{1-\lambda^2}},u_y= \sqrt{1-\lambda^2} C x^\lambda y^{\sqrt{1-\lambda^2}-1}$$
+$$x^2u_x^2+y^2u_y^2=x^2\left(\lambda C x^{\lambda-1} y^{\sqrt{1-\lambda^2}} \right)^2+y^2 \left(\sqrt{1-\lambda^2} C x^\lambda y^{\sqrt{1-\lambda^2}-1}\right)^2 =$$
+
+$$x^2\lambda^2C^2 x^{2\lambda-2}y^{2\sqrt{1-\lambda^2}}+y^2(1-\lambda^2)C^2x^{2\lambda} y^{2\sqrt{1-\lambda^2}-2}=$$
+$$\lambda^2 C^2 x^{2\lambda}y^{2\sqrt{1-\lambda^2}}+(1-\lambda^2)C^2x^{2\lambda} y^{2\sqrt{1-\lambda^2}} = C^2x^{2\lambda} y^{2\sqrt{1-\lambda^2}}=u^2$$
+\end{tcolorbox}
 
@@
 And since we are given that $u(0,x)=\arctan{(x)}$, we have $f(\arctan{(x)}-0)=\arctan{(x)}$ so that $f(w)=w$ for any $w$, yielding our solution of
+
+\begin{tcolorbox}[colback=white, title=Solution]
 $$u(t,x)=\arctan{(x)}-t$$
+\end{tcolorbox}
 
@@
 
-We can check our solution of $u(t,x)=\arctan{(x)}-t$ by differentiating, since $u_t=-1, u_x=\frac{1}{1+x^2}$, that means $u_t+\left(1+x^2\right)u_x=-1+1=0$.
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+Since $u_t=-1, u_x=\frac{1}{1+x^2}$, that means $u_t+\left(1+x^2\right)u_x=-1+1=0$
+\end{tcolorbox}
 
@@
 
-So we have $u(t,x)=f(\xi)=f(xe^t)$ and we know that $u(0,x)=\frac{1}{1+x^2}$, which means $f(xe^0)=f(x)=\frac{1}{1+x^2}$ and therefore $u(t,x)=\frac{1}{1+\left(xe^t\right)^2}$.
+So we have $u(t,x)=f(\xi)=f(xe^t)$ and we know that $u(0,x)=\frac{1}{1+x^2}$, which means $f(xe^0)=f(x)=\frac{1}{1+x^2}$ and therefore
+\begin{tcolorbox}[colback=white, title=Solution]
+$$u(t,x)=\frac{1}{1+\left(xe^t\right)^2}$$
+\end{tcolorbox}
 
-We check by differentiating:
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
 $$u_t=-\left(1+\left(xe^t\right)^2\right)^{-2}\left(2xe^t\right)xe^t$$
@@
 And we see that $u_t-xe_t$ indeed equals zero.
+\end{tcolorbox}
+
 Plotting the solution $u(t,x)=\frac{1}{1+\left(xe^t\right)^2}$ as a function of $x$ for $t=1,2,3$, we get:
@@
 Substitute back in $x$ and $y$:
+\begin{tcolorbox}[colback=white, title=Solution]
 $$u(x,y)=x\left(y-\frac{x^2}{2}\right)^2+\exp{\left(y-\frac{x^2}{2}\right)}$$
+\end{tcolorbox}
 
-We can check by differentiating:
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
 $$u_x=-2x^2\left(y-\frac{x^2}{2}\right)+\left(y-\frac{x^2}{2}\right)^2+\exp{\left(y-\frac{x^2}{2}\right)}(-x)$$
 $$u_y=2x\left(y-\frac{x^2}{2}\right)+\exp{\left(y-\frac{x^2}{2}\right)}$$
-
 Which means that $u_x+xu_y=\left(y-\frac{x^2}{2}\right)^2$, as we wanted.
+\end{tcolorbox}
 
@@
 ```
-
-Now we define our new coordinate system as $2y-x-\ln(x), \, \eta=x$
-
-Then by the chain rule,
-$$u_x= \pdv{u}{x}=\pdv{u}{\eta}\pdv{\eta}{x}+\pdv{u}{\xi}\pdv{\xi}{x}=$$
-
-$$\pdv{u}{\eta}+\pdv{u}{\xi}\left(-1-\frac{1}{x}\right) $$
-
-Similarly, $$u_y=\pdv{u}{y}=2\pdv{u}{\xi}$$
-
-Plugging into our initial PDE,
-$$2xu_x+(x+1)u_y=2x\pdv{u}{\eta}+\pdv{u}{\xi}2x\left(-1-\frac{1}{x}\right)+2(x+1)\pdv{u}{\xi}=2x\pdv{u}{\eta}$$
-
-And since we know $2xu_x+(x+1)u_y=y$, this means $y=2x\pdv{u}{\eta}$ or
-$$\frac{\eta}{2}+\frac{\ln(\eta)}{2}+\xi=2\eta \pdv{u}{\eta}$$

### 2024-01-31 bio_dynamics/bio274_hw2.Rmd
bio274 hw2

@@
 author: "Ruby Krasnow"
-date: "2024-01-27"
-output: pdf_document
+date: "2024-01-31"
+output:
+  pdf_document:
+    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
 ---
@@
 
-Although since we are working in discrete time, it might be more accurate to write the change in population as a difference equation, $P_{n+1}-P_n=\Delta P= 0.3P \left( 1- \frac{P}{2500} \right)$.
+Note: Since we are working in discrete time, it might be more accurate to write the change in population as a difference equation, $P_{n+1}-P_n=\Delta P= 0.3P \left( 1- \frac{P}{2500} \right)$.
 For now we will continue using differential equation notation and apologize to Leibniz.
 
-The initial conditions will remain the same for each of the three situations, but we can adjust the differential equation for each scenario as follows:
+The initial conditions $P(0)=2500$ will remain the same for each of the three situations, but we can adjust the differential equation for each scenario as follows:
 
@@
 
-Now we're going to have some fun and see how the population would change over a 100-year period if no harvesting occurred and under each of these three harvesting strategies!
+Let's see how the population would change over a 100-year period if no harvesting occurred and under each of these three harvesting strategies!
 We will use the numerical ODE solver from the `deSolve` package.
@@
 Let's see what we got!
-```{r}
+```{r,   fig.height=3, fig.width=5, fig.width=5}
 ggplot()+
@@
 Here is what it looks like:
-```{r}
+```{r,   fig.height=3, fig.width=5  }
 t=c(1:250)
@@
 
-```{r}
+```{r,   fig.height=3, fig.width=5  }
 dy=D(expression(2500/(10*exp(t/30)-9)), "t")
@@
 
-```{r}
+```{r,   fig.height=3, fig.width=5  }
 N0 = 0.1
@@
 
-## Question 4 & 5
+## Euler's Method
 For problems 4 and 5, use Euler’s Method with the given step size, h, to approximate the solution
@@
 of the approximate solution.
+
+### Question 4
+$$\frac{dw}{dt}=(3-w)(w+1), \quad w(0)=4, \quad 0\leq t \leq 5, \quad h=1.0$$
+We start with a table that looks like this:
+```{r}
+data.frame(t=c(0:5), w=c(4, rep(NA, 5)))
+```
+
+**Manual computation**  
+At $t=1, \quad w=4+(3-4)(4+1)1 = 4+(-5)=-1$  
+But then for $t=2, \quad w=-1+(3--1)(-1+1)1= -1+0 = -1$  
+For $t=3, w=-1+(3--1)(-1+1)1= -1+0 = -1$  
+So once we hit $t=2$, the slope is always 0, making $y(t_n+1) = y(t_n) = -1$. The same thing will happen each time we compute $w$, so using Euler's method with a step size of h=1 has gotten us stuck!
+
+**Using a function**  
+Let's define a function to compute Euler's method and see if we can do better with smaller step sizes.
+
+Here we will write our own function, but note that the `deSolve` package includes a function `euler` that can do the same thing with many options for additional customization.
+There is also a simpler `euler` function in the package `demodelr` that takes fewer parameters.
+
+```{r}
+euler_fun <- function(slope_fun, x0, y0, h, N) {
+    x <- x0
+    y <- y0
+    num_steps <- N/h
+
+    for(i in 1:num_steps) {
+        y0 <- y0 + h * slope_fun(x0, y0)
+        x0 <- x0 + h
+        x <- c(x, x0)
+        y <- c(y, y0)
+    }
+
+    return(data.frame(x = x, y = y, h=h))
+}
+
+f <- function(x_n, y_n){
+  slope=(3-y_n)*(y_n+1)
+  slope
+}
+
+euler_fun(slope_fun=f, x0=0, y0=4, h=1, N=5) %>% rename(w=y, t=x)
+euler_fun(slope_fun=f, x0=0, y0=4, h=0.5, N=5) %>% rename(w=y, t=x)
+euler_fun(slope_fun=f, x0=0, y0=4, h=0.1, N=5)%>% rename(w=y, t=x)
+
+euler_output<- c(1, 0.5, 0.1) %>% 
+  map(\(x) euler_fun(f, x0=0, y0=4, h=x, N=5)) %>% 
+  bind_rows()
+
+```
+
+```{r,   fig.height=3, fig.width=5  }
+ggplot(data = euler_output)+
+  geom_point(aes(x=x, y=y, color=as.factor(h)))+
+  geom_line(aes(x=x, y=y, color=as.factor(h)), alpha=0.5)+
+  theme_light()+
+  labs(x="t", y="w", color="h")
+```
+
+Since this is a separable ODE and we are given initial conditions, we can also find the particular solution explicitly:
+$$\frac{dw}{dt}=(3-w)(w+1)=-w^2+2w+3$$
+
+$$\int \frac{dw}{w^2-2w-3}=\int -dt$$
+
+Using partial fraction decomposition for the left-hand side, we can rewrite the integral as:
+$$\frac{1}{4}\int \frac{1}{w-3}-\frac{1}{w+1}\>dw=\int -dt$$
+
+$$\frac{1}{4}\left( \ln(w-3)-\ln(w+1) \right)=-t+C_1$$
+$$ \ln\left(\frac{w-3}{w+1}\right) =-4t+C_2$$
+
+$$\frac{w-3}{w+1}=\frac{w+1-4}{w+1}=1+\frac{-4}{w+1}=C_3e^{-4t}$$
+$$\frac{-4}{w+1}=C_3e^{-4t}-1$$
+Plugging in our initial conditions, 
+$$\frac{-4}{5}=C_3-1 \quad \longrightarrow \quad C_3=\frac{1}{5}$$
+
+$$\frac{-4}{w+1}=\frac{1}{5}e^{-4t}-1$$
+\begin{tcolorbox}[colback=white, title=Solution]
+$$w(t)=\frac{4}{1-\frac{1}{5}e^{-4t}}-1$$
+\end{tcolorbox}
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution by differentiating]
+
+Direct differentiation of solution
+$$\frac{dw}{dt}= -4\left ( 1-\frac{1}{5}e^{-4t} \right)^{-2}\left(\frac{-1}{5}e^{-4t}\right)(-4) = \frac{16 \left(- \frac{1}{5} e^{-4t} \right)}{ \left(1 -\frac{1}{5}e^{-4t} \right)^2}$$
+Plugging solution into original ODE
+
+$$\frac{dw}{dt}=(3-w)(w+1)=\left(3-\frac{4}{1-\frac{1}{5}e^{-4t}}+1 \right) \left(\frac{4}{1-\frac{1}{5}e^{-4t}}-1+1 \right) =$$
+
+$$\left(4-\frac{4}{1-\frac{1}{5}e^{-4t}}\right)\left(\frac{4}{1-\frac{1}{5}e^{-4t}}\right) = $$
+
+$$\frac{16}{1-\frac{1}{5}e^{-4t}}-\frac{16}{\left(1-\frac{1}{5}e^{-4t}\right)^2} =
+\frac{16 \left( 1 - \frac{1}{5} e^{-4t} \right) -16}{ \left(1 -\frac{1}{5}e^{-4t} \right)^2}=$$
+$$\frac{16 \left(- \frac{1}{5} e^{-4t} \right)}{ \left(1 -\frac{1}{5}e^{-4t} \right)^2}$$
+
+\end{tcolorbox}
+
+So now we can add to our graph the actual solution:
+```{r}
+actual_eqn <- data.frame(t=seq(from=0, to=5, by=0.0001)) %>% mutate(
+w=4/((1-(1/5)*exp(-4*t)))-1)
+
+ggplot()+
+  geom_line(data=actual_eqn, aes(x=t, y=w, color="Actual solution"), linewidth=1.5)+
+  geom_point(data = euler_output, aes(x=x, y=y, color=as.factor(h)), size=1)+
+  geom_line(data = euler_output, aes(x=x, y=y, color=as.factor(h)), alpha=0.5, linewidth=1)+
+  theme_light()+
+  labs(x="t", y="w", color="h")+
+  scale_color_manual(values=c( "#e69b99","#89689d","#015b58", "black"))
+#Credit: hex codes are from Jake Lawlor's R package PNWcolors
+```
+
+We can also see that $\lim\limits_{t \to \infty}w(t)=3$ by looking at our expression for the actual solution, because $e^{-4t}$ will approach zero, so the denominator will approach $1-0$ and the overall solution will approach $\frac{4}{1}-1=3$.
+Clearly, the step size of $h=0.1$ very closely approximated the actual solution after around 30 time steps.
+
+### Question 5
+$$\frac{dy}{dt}=e^{2/y}, \quad y(0)=2, \quad 0\leq t \leq 2, \quad h=0.5$$
+
+**Manual computation**
+
+$(t_0=0, y_0=2) \quad \longrightarrow \quad$ slope $=e^{2/2}=e, y_1=2+\frac{e}{2} \approx$ `r exp(2+exp(1)/2)`
+```{r}
+y_1=round(2+exp(1)/2, 4)
+y_1
+```
+
+$(t_1=0.5, y_1 \approx 3.3591)\quad \longrightarrow \quad$ slope $=\exp{\left(\frac{2}{3.3591}\right)} \approx$ `r exp(2/3.3591)`
+```{r}
+y_2=round(3.3591+exp(2/3.3591)/2, 4)
+y_2
+```
+
+$(t_2=1, y_2 \approx 4.266)\quad \longrightarrow \quad$ slope $=\exp{\left(\frac{2}{4.266}\right)} \approx$ `r exp(2/4.266)`
+```{r}
+y_3=round(4.266+exp(2/4.266)/2, 4)
+y_3
+```
+
+$(t_3=1.5, y_3 \approx 5.0651)\quad \longrightarrow \quad$ slope $=\exp{\left(\frac{2}{5.0651}\right)} \approx$ `r exp(2/5.0651)`
+
+```{r}
+y_4=round(5.0651+exp(2/5.0651)/2, 4)
+y_4
+```
+
+$(t_4=2, y_4\approx 5.8072)$
+
+Using our function from earlier,
+```{r}
+new_f = function(x_n, y_n) {
+  slope = exp(2/y_n)
+  slope
+}
+
+euler_fun(slope_fun = new_f, h=0.5, N=2, x0=0, y0=2)
+```
+
+And we get the same result as our manual computation.
+
+```{r}
+library(PNWColors)
+
+euler_output_2<- c(1, 0.5, 0.1, 0.05) %>% 
+  map(\(x) euler_fun(new_f, x0=0, y0=2, h=x, N=2)) %>% 
+  bind_rows()
+
+ggplot()+
+  geom_point(data = euler_output_2, aes(x=x, y=y, color=as.factor(h)))+
+  geom_line(data = euler_output_2, aes(x=x, y=y, color=as.factor(h)), alpha=0.8)+
+  theme_light()+
+  labs(x="t", y="y", color="h")+
+  scale_color_manual(values=pnw_palette(name="Starfish",n=4,type="discrete"))
+```

### 2024-02-01 PDEs/math245_hw1.Rmd
hw2 for bio, hw1 for math

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-01-30"
+date: "2024-02-01"
 output:
@@
 
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+Since $u_t=-1, u_x=\frac{1}{1+x^2}$, that means $u_t+\left(1+x^2\right)u_x=-1+1=0$
+\end{tcolorbox}
+
 ```{r, echo=FALSE}
@@
 ```
-
-\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
-Since $u_t=-1, u_x=\frac{1}{1+x^2}$, that means $u_t+\left(1+x^2\right)u_x=-1+1=0$
-\end{tcolorbox}
 
@@
 #### (d)
-$2xu_x+(x+1)u_y=y$, for $x>0, u=2y$ on $x=1$
+$2x u_x+ (x+1)u_y=y$, for $x>0, u=2y$ on $x=1$
 
-$$\frac{dx}{dy}=\frac{x+1}{2x}=\frac{1}{2}+\frac{1}{2x}$$
+$$\frac{dx}{a}=\frac{dy}{b}=\frac{du}{c} \quad \longrightarrow \quad 
+\frac{dx}{2x}=\frac{dy}{x+1}=\frac{du}{y}$$
+
+$$\frac{dx}{2x}=\frac{dy}{x+1}$$
+$$\frac{x+1}{2x}\, dx=dy$$
+$$\int \frac{1}{2}+\frac{1}{2x}\, dx=\int dy$$
+
 Solving this ODE yields the characteristic curves:
-$$y=\frac{x}{2}+\frac{\ln(x)}{2}+C$$
-$$C=2y-x-\ln(x)$$
+$$y=\frac{x}{2}+\frac{\ln x}{2}-C$$
+$$C=\frac{x}{2}+\frac{\ln x}{2}-y$$
 
@@
   geom_line(aes(x=seq, y=y, color=as.factor(C)))+
-  coord_cartesian(ylim=c(-1, 12), xlim=c(-1, 9.5))+
+  coord_cartesian( xlim=c(-1, 9.5))+
   labs(x="x", y="y", color="C")+
@@
 ```
+
+Now define our new coordinate system as $xi=\frac{x}{2}+\frac{\ln x}{2}-y, \quad \eta=x$.  
+Then by the chain rule,
+$$u_x= \pdv{u}{x}=\pdv{u}{\eta}\pdv{\eta}{x}+\pdv{u}{\xi}\pdv{\xi}{x}=$$
+$$\pdv{u}{\xi}\left(\frac{1}{2}+ \frac{1}{2x}\right) +\pdv{u}{\eta}$$
+$$u_y= \pdv{u}{y}=\pdv{u}{\eta}\pdv{\eta}{y}+\pdv{u}{\xi}\pdv{\xi}{y}= -\pdv{u}{\xi}$$
+Plugging these into our original PDE,
+$$2x\left( \pdv{u}{\xi}\left(\frac{1}{2}+ \frac{1}{2x}\right) +\pdv{u}{\eta} \right)+\left(x+1\right)\left(-\pdv{u}{\xi}\right)=$$
+$$\pdv{u}{\xi}\left(x+1\right)+2x \pdv{u}{\eta}-\pdv{u}{\xi}\left(x+1\right)=2x \pdv{u}{\eta}$$
+So we must have $2x\pdv{u}{\eta}=y$. Now replace $\eta$ and $\xi$ by $x$ and $y$:
+$$2\eta\pdv{u}{\eta}=\frac{\eta}{2}+\frac{\ln\eta}{2}-\xi$$
+
+$$\pdv{u}{\eta}=\frac{1}{4}+\frac{\ln\eta}{4\eta}-\frac{\xi}{2n}$$
+
+Now integrate with respect to $\eta$:
+$$u=\frac{\eta}{4}+\frac{(\ln\eta)^2}{8}-\frac{\xi \ln\eta}{2}+F(\xi)$$
+$$u(x,y)=\frac{x}{4}+\frac{(\ln x)^2}{8}-\frac{\ln x}{2}\left(\frac{x}{2}+\frac{\ln x}{2}-y \right)+F\left(\frac{x}{2}+\frac{\ln x}{2}-y\right)$$
+Now we use the fact that $u(1,y)=2y$:
+$$u(1,y)=\frac{1}{4}+F\left(\frac{1}{2}-y\right)=2y$$
+$$F\left(\frac{1}{2}-y\right)=2y-\frac{1}{4}$$
+So $F(\alpha)=-2\alpha+\frac{3}{4}$ for arbitrary $\alpha$. Then
+$$u(x,y)=\frac{x}{4}+\frac{(\ln x)^2}{8}-\frac{\ln x}{2}\left(\frac{x}{2}+\frac{\ln x}{2}-y \right)-x-\ln x +2y+\frac{3}{4}$$
+Which can be simplified to
+$$u(x,y)=\frac{-3x}{4}-\frac{(\ln x)^2}{8}-\frac{x\ln x}{4}+\frac{y\ln x}{2}-\ln x+2y+\frac{3}{4}$$
+\begin{tcolorbox}[colback=white, title=Solution]
+$$u(x,y)=-\ln x\left(\frac{x}{4} +\frac{\ln x}{8}+ \frac{y}{2} +1 \right)-\frac{3x}{4}+2y+\frac{3}{4} $$
+\end{tcolorbox}
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+$$u_y=\frac{\ln x}{2}+2$$
+$$u_x=\frac{-1}{x} \left(\frac{x}{4} +\frac{\ln x}{8}- \frac{y}{2} +1 \right)+(-\ln x) \left(\frac{1}{4} +\frac{1}{8x}\right)-\frac{3}{4}$$
+Which means
+$$2xu_x+(x+1)u_y=\left( \frac{-x}{2}-\frac{\ln x}{4}+y-2-\frac {x \ln x}{2}-\frac{\ln x}{4}-\frac{3x}{2}\right)+ \left(\frac {x \ln x}{2}+\frac { \ln x}{2}+2x+2 \right)=$$
+$$\left(-\frac {x \ln x}{2}-\frac{\ln x}{2} -2x+-2+y\right)+ \left(\frac {x \ln x}{2}+\frac { \ln x}{2}+2x+2 \right)=y$$
+Which matches our original PDE.
+\end{tcolorbox}
+
+#### (e)
+$yu_x+xu_y=xy$, for $x,y \geq 0$ with $u(0,y)=\exp(-y^2)$
+
+$$\frac{dx}{a}=\frac{dy}{b}=\frac{du}{c} \quad \longrightarrow \quad 
+\frac{dx}{y}=\frac{dy}{x}=\frac{du}{xy}$$
+
+$$\frac{dx}{y}=\frac{dy}{x}\quad \longrightarrow \quad x\;dx=y\;dy\quad \longrightarrow \quad y^2=x^2+C$$
+
+$$y=\sqrt{x^2+C}, \quad C=y^2-x^2$$
+Where we kept the positive root for $y$ because it was given that $x,y \geq 0$.
+
+```{r, warning = FALSE, echo=FALSE}
+x=seq(-10,10,length.out = 1000)
+
+curves_4e <- tibble(C=c(1:5), seq=list(x)) %>% 
+  rowwise() %>% 
+  mutate(y=list(sqrt(seq^2+C))) %>% 
+  unnest(c(seq,y))
+
+ggplot(data=curves_4e)+
+  geom_line(aes(x=seq, y=y, color=as.factor(C)))+
+  labs(x="x", y="y", color="C")+
+  theme_light()+
+  ggtitle("Characteristic curves for Problem 4e")
+```
+
+$$\frac{dx}{y}=\frac{du}{xy} \quad\longrightarrow\quad du=x\;dx$$
+
+$$\frac{dy}{x}=\frac{du}{xy} \quad\longrightarrow\quad du=y\;dy$$
+
+$$2\,du= x\,dx+y\,dy=d\left(\frac{x^2+y^2}{2} \right)$$
+
+$$\int du= \frac{1}{4} \int d(x^2+y^2)$$
+
+$$u=\frac{x^2+y^2}{4}+f(C)\quad\longrightarrow\quad u=\frac{x^2+y^2}{4}+f(y^2-x^2)$$
+
+$$u(0,y)=\frac{y^2}{4}+f(y^2)=e^{-y^2}$$
+
+$$f(y^2)=e^{-y^2}-\frac{y^2}{4}$$
+
+$$f(\alpha)=e^{-\alpha}-\frac{\alpha}{4}$$
+
+$$u(x,y)=\frac{x^2+y^2}{4}+ e^{x^2-y^2}+\frac{x^2-y^2}{4}$$
+
+\begin{tcolorbox}[colback=white, title=Solution]
+$$u(x,y)=\frac{x^2}{2}+ e^{x^2-y^2}$$
+\end{tcolorbox}
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+$$u_x=x+2xe^{x^2-y^2}, \; u_y=-2ye^{x^2-y^2}$$
+$$yu_x+xu_y=xy+2xye^{x^2-y^2}-2xye^{x^2-y^2}=xy$$
+Which matches our original PDE.
+\end{tcolorbox}

### 2024-02-01 PDEs/math245_old.Rmd
hw2 for bio, hw1 for math

--- 2024-01-29 PDEs/math245_hw1.Rmd
@@
 ---
-title: "MATH 245 Homework 1"
-author: "Ruby Krasnow and Tommy Thach"
-date: "2024-01-29"
+title: "MATH 245 old and unused content"
+author: "Ruby Krasnow"
+date: "2024-01-31"
 output:
   pdf_document:
-    extra_dependencies: ["derivative"]
+    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
 ---
@@
 
-## Question 1
-Show the function is a solution of the PDE:  
-(a) $u_{xx}+u_{yy}=0$   
-(i) $u(x,y)=e^x\sin(y)$  
-$$u_x=e^x\sin(y), \, u_{xx}=e^x\sin(y), \, u_y=e^x\cos(y),\, u_{yy}=-e^x\sin(y)$$
-Which means that $u_{xx}+u_{yy}=e^x\sin(y)+\left(-e^x\sin(y)\right)=0$ and so $u(x,y)=e^x\sin(y)$ is
-a solution to the PDE.  
-(ii) $u(x,y)=\log{\sqrt{x^2+y^2}}$  
-Assuming this is meant to be $u(x,y)=\ln\sqrt{{{x^2+y^2}}}$,
-$$u_x=\left({{x^2+y^2}}\right)^{-1/2}
-\left(\frac{1}{2}\right)\left({{x^2+y^2}}\right)^{-1/2}
-\left(2x\right)=x\left({{x^2+y^2}}\right)^{-1}$$
-
-$$u_{xx}=x(-1)\left({{x^2+y^2}}\right)^{-2}\left(2x\right)+\left({{x^2+y^2}}\right)^{-1}$$
-
-$$u_y=\left({{x^2+y^2}}\right)^{-1/2}
-\left(\frac{1}{2}\right)\left({{x^2+y^2}}\right)^{-1/2}
-\left(2y\right)=y\left({{x^2+y^2}}\right)^{-1}$$
-
-$$u_{yy}=y(-1)\left({{x^2+y^2}}\right)^{-2}\left(2y\right)+\left({{x^2+y^2}}\right)^{-1}$$
-
-$$u_{xx}+u_{yy}=\frac{-2x^2}{\left({{x^2+y^2}}\right)^{2}}+
-\frac{-2y^2}{\left({{x^2+y^2}}\right)^{2}}+
-\frac{2}{\left({{x^2+y^2}}\right)}=$$
-
-$$\frac{-2\left({{x^2+y^2}}\right)}{\left({{x^2+y^2}}\right)^{2}}+ \frac{2}{\left({{x^2+y^2}}\right)}=0$$
-Which means that $u(x,y)=\log{\sqrt{x^2+y^2}}$ is a solution to the PDE.
-
-(b) $bu_x+au_y+u=0, \, u(x,y)=\exp{\left(\frac{-x}{b}\right)}f(ax-by)$ for arbitrary differentiable function $f$.
-$$u_x=e^{\frac{-x}{b}} \left(\frac{-1}{b}\right)f(ax-by)+f'(ax-by)(a)e^{\frac{-x}{b}}$$
-$$u_y=e^{\frac{-x}{b}}f'(ax-by)(-b)$$
-Then,
-$$bu_x+au_y+u=-e^{\frac{-x}{b}}f(ax-by)+(ab)\left(e^{\frac{-x}{b}}\right)f'(ax-by)+
-e^{\frac{-x}{b}}f'(ax-by)(-ab)+e^{\frac{-x}{b}}f(ax-by)=0$$
-
-So $u(x,y)=\exp{\left(\frac{-x}{b}\right)}f(ax-by)$ is a solution to the PDE for an arbitrary differentiable function $f$.
-
-(c) $u_{xx}-\frac{1}{x}u_x-x^2u_{yy}=0, \, u(x,y)=f\left(2y+x^2\right)+g\left(2y-x^2\right)$ for arbitrary twice-differentiable functions $f$ and $g$.
-
-$$u_x=\left(f'\left(2y+x^2\right)\right)(2x)+\left(g'\left(2y-x^2\right)\right)(-2x)$$
-$$u_y=2 \left(f'\left(2y+x^2\right)\right) +2\left(g'\left(2y-x^2\right)\right)$$
-$$u_{xx}=2f'+(f'')(4x^2)-2g'+(g'')(4x^2)$$
-$$u_{yy}=4f''+4g''$$
-$$u_{xx}-\frac{1}{x}u_x-x^2u_{yy}=$$ $$2f'+4x^2f''-2g'+4x^2g''-\frac{1}{x}\left(2xf'-2xg'\right)-x^2\left(4f''+4g''\right)=$$
-$$2f'-2f'+4x^2f''-4x^2f''-2g'+2g'+4x^2g''-4x^2g''=0 $$
-
-So $u(x,y)=f\left(2y+x^2\right)+g\left(2y-x^2\right)$ is a solution to the PDE for arbitrary twice-differentiable functions $f$ and $g$.
-
-### Question 2
-(a) 2nd-order linear homogeneous  
-(b) 4th-order linear inhomogeneous  
-(c) 2nd-order quasi-linear homogeneous  
-(d) We can rewrite this as $u_{xx}+u_{yy}+f(x,y)u-g(x,y)u^5=0$, making it clear that this is a 2nd-order semi-linear homogeneous PDE.
-
-### Question 3
-Use separation of variables to solve the following problems:
-
-#### (a)
-$u_x+u=u_y, \, u(x,0)=4x^{-3x}$, use $u(x,y)=f(x)g(y)$
-
-#### (b)
-$x^2u_{xy}+9y^2u=0, \, u(x,0)=\exp{\left(\frac{1}{x}\right)}$, use $u(x,y)=f(x)g(y)$
-
-#### (c)
-$u_x^2+u_y^2=1$, use $u(x,y)=f(x)+g(y)$
-
-#### (d)
-$x^2u_x^2+y^2u_y^2=u^2$, use $u(x,y)=e^{f(x)}e^{g(y)}$
-
-### Question 4
+### Homework 2 Question 4 - old version
 For each of the following IVPs, (i) find and plot the characteristic lines (curves), (ii) solve the IVP, and (iii) plot the solution of (a)-(b) for indicated time.

### 2024-02-01 bio_dynamics/bio274_hw2.Rmd
hw2 for bio, hw1 for math

@@
 author: "Ruby Krasnow"
-date: "2024-01-31"
+date: "2024-02-01"
 output:
@@
 library(tidyverse)
+library(PNWColors)
 ```
@@
   theme_bw()+
-  labs(x="Time (years)", y="Pop. size", color=NULL)
+  labs(x="Time (years)", y="Pop. size", color=NULL)+
+  scale_color_manual(values=pnw_palette(name="Sailboat",n=4,type="discrete"))
 ```
@@
   theme_bw()+
-  labs(x="time", y="P")
+  labs(x="time", y="P", color=NULL)
 ```
@@
 ggplot()+
-  geom_line(aes(x=y, y=eval(dy), color="calculated"), linewidth=3, alpha=0.5)+
-  geom_line(aes(x=y, y= 0.3*y*(1-(y/2500))-y/3, color="original dP/dt"))+
+  geom_line(aes(x=y, y=eval(dy), color="Calculated derivative"), linewidth=3, alpha=0.5)+
+  geom_line(aes(x=y, y= 0.3*y*(1-(y/2500))-y/3, color="Original dP/dt"))+
   scale_color_manual(values=c("lightblue", "black"))+
   theme_bw()+
-  labs(x="P", y="dP/dt")
+  labs(x="P", y="dP/dt", color=NULL)
 ```
 
-## Question 2 - for practice
+## Question 2 again - for practice with Bernoulli's method
 Using Bernoulli's method, find a solution to the IVP representing the situation in #1b.
@@
   labs(x="t", y="P", color=NULL)+
-  scale_color_manual(values=c("gray", "#0f85a0", "#dd4124"))+
+  scale_color_manual(values=c("gray",pnw_palette(name="Bay",n=2,type="discrete")))+
   ylim(0, NA)
@@
 ```
+
+The equations needed to use Euler's method are:
+$$t_{n+1}=t_n+h$$
+$$y_{n+1}=y_n+f(t_n, y_n)h$$
+where $f(t_n, y_n)$ is the slope we calculate by plugging $t_n, y_n$ into our original DE.
 
@@
   theme_light()+
-  labs(x="t", y="w", color="h")
+  labs(x="t", y="w", color="h")+
+  scale_color_manual(values=pnw_palette(name="Sailboat",n=3,type="discrete"))
 ```
@@
 ggplot()+
-  geom_line(data=actual_eqn, aes(x=t, y=w, color="Actual solution"), linewidth=1.5)+
+  geom_line(data=actual_eqn, aes(x=t, y=w, color="Actual solution"), linewidth=1)+
   geom_point(data = euler_output, aes(x=x, y=y, color=as.factor(h)), size=1)+
@@
   labs(x="t", y="w", color="h")+
-  scale_color_manual(values=c( "#e69b99","#89689d","#015b58", "black"))
-#Credit: hex codes are from Jake Lawlor's R package PNWcolors
+  scale_color_manual(values=c(pnw_palette(name="Sailboat",n=3,type="discrete"), "black"))
 ```
@@
 ```{r}
-library(PNWColors)
-
 euler_output_2<- c(1, 0.5, 0.1, 0.05) %>% 
@@
   labs(x="t", y="y", color="h")+
-  scale_color_manual(values=pnw_palette(name="Starfish",n=4,type="discrete"))
+  scale_color_manual(values=pnw_palette(name="Sailboat",n=4,type="discrete"))
 ```

### 2024-02-03 PDEs/math245_hw1.Rmd
bio hw 3

@@
 
-Set $\xi=xe^t$ and $\eta=x$. Then by the chain rule,
-$$u_x= \pdv{u}{x}=\pdv{u}{\eta}\pdv{\eta}{x}+\pdv{u}{\xi}\pdv{\xi}{x}=$$
-$$\pdv{u}{\eta}+\pdv{u}{\xi}e^t $$
+Set $\eta=xe^t$ and $\xi=x$. Then by the chain rule,
+$$u_x= \pdv{u}{x}=\pdv{u}{\xi}\pdv{\xi}{x}+\pdv{u}{\eta}\pdv{\eta}{x}=$$
+$$\pdv{u}{\xi}+\pdv{u}{\eta}e^t $$
 
 and similarly
-$$u_t= \pdv{u}{t}=\pdv{u}{\eta}\pdv{\eta}{t}+\pdv{u}{\xi}\pdv{\xi}{t}=\pdv{u}{\xi}xe^t $$
+$$u_t= \pdv{u}{t}=\pdv{u}{\xi}\pdv{\xi}{t}+\pdv{u}{\eta}\pdv{\eta}{t}=\pdv{u}{\eta}xe^t $$
 
 Which means that 
-$$u_t-xu_x=\pdv{u}{\xi}xe^t -x\left(\pdv{u}{\eta}+\pdv{u}{\xi}e^t \right)=0$$
+$$u_t-xu_x=\pdv{u}{\eta}xe^t -x\left(\pdv{u}{\xi}+\pdv{u}{\eta}e^t \right)=0$$
 Assuming $x\neq0$, this means
-$$u_\eta=0$$
-$$u=\int u_\eta \, d\eta \> + f(\xi)=0+ f(\xi)$$
+$$u_\xi=0$$
+$$u=\int u_\xi \, d\xi \> + f(\eta)=0+ f(\eta)$$
 
-So we have $u(t,x)=f(\xi)=f(xe^t)$ and we know that $u(0,x)=\frac{1}{1+x^2}$, which means $f(xe^0)=f(x)=\frac{1}{1+x^2}$ and therefore
+So we have $u(t,x)=f(\eta)=f(xe^t)$ and we know that $u(0,x)=\frac{1}{1+x^2}$, which means $f(xe^0)=f(x)=\frac{1}{1+x^2}$ and therefore
 \begin{tcolorbox}[colback=white, title=Solution]
@@
 
-Now we define our new coordinate system as $\xi=y-\frac{x^2}{2}, \, \eta=x$
+Now we define our new coordinate system as $\eta=y-\frac{x^2}{2}, \, \xi=x$
 
 Then by the chain rule,
-$$u_x= \pdv{u}{x}=\pdv{u}{\eta}\pdv{\eta}{x}+\pdv{u}{\xi}\pdv{\xi}{x}=$$
-$$\pdv{u}{\eta}-x\pdv{u}{\xi}$$
+$$u_x= \pdv{u}{x}=\pdv{u}{\xi}\pdv{\xi}{x}+\pdv{u}{\eta}\pdv{\eta}{x}=$$
+$$\pdv{u}{\xi}-x\pdv{u}{\eta}$$
 
 and similarly
-$$u_y= \pdv{u}{y}=\pdv{u}{\eta}\pdv{\eta}{y}+\pdv{u}{\xi}\pdv{\xi}{y}=$$
-$$\pdv{u}{\xi}$$
+$$u_y= \pdv{u}{y}=\pdv{u}{\xi}\pdv{\xi}{y}+\pdv{u}{\eta}\pdv{\eta}{y}=$$
+$$\pdv{u}{\eta}$$
 
-$$u_x+xu_y=\pdv{u}{\eta}-x\pdv{u}{\xi}+x\pdv{u}{\xi}=\pdv{u}{\eta}$$
+$$u_x+xu_y=\pdv{u}{\xi}-x\pdv{u}{\eta}+x\pdv{u}{\eta}=\pdv{u}{\xi}$$
 
 And since we know $u_x+xu_y=\left(y-\frac{x^2}{2}\right)^2$, this means we have
-$$\pdv{u}{\eta}=\left(y-\frac{x^2}{2}\right)^2=\xi^2$$
-Now we integrate with respect to $\eta$ to find
-$$u=\eta\xi^2+F(\xi)$$
+$$\pdv{u}{\xi}=\left(y-\frac{x^2}{2}\right)^2=\eta^2$$
+Now we integrate with respect to $\xi$ to find
+$$u=\xi\eta^2+F(\eta)$$
 
@@
 On the other hand, we are told that $u(0,y)=e^y$, which means that $F(y)=e^y$.
-$$u=\eta\xi^2+e^\xi$$
+$$u=\xi\eta^2+e^\eta$$
 Substitute back in $x$ and $y$:
@@
 
-Now define our new coordinate system as $xi=\frac{x}{2}+\frac{\ln x}{2}-y, \quad \eta=x$.  
+Now define our new coordinate system as $xi=\frac{x}{2}+\frac{\ln x}{2}-y, \quad \xi=x$.  
 Then by the chain rule,
-$$u_x= \pdv{u}{x}=\pdv{u}{\eta}\pdv{\eta}{x}+\pdv{u}{\xi}\pdv{\xi}{x}=$$
-$$\pdv{u}{\xi}\left(\frac{1}{2}+ \frac{1}{2x}\right) +\pdv{u}{\eta}$$
-$$u_y= \pdv{u}{y}=\pdv{u}{\eta}\pdv{\eta}{y}+\pdv{u}{\xi}\pdv{\xi}{y}= -\pdv{u}{\xi}$$
+$$u_x= \pdv{u}{x}=\pdv{u}{\xi}\pdv{\xi}{x}+\pdv{u}{\eta}\pdv{\eta}{x}=$$
+$$\pdv{u}{\eta}\left(\frac{1}{2}+ \frac{1}{2x}\right) +\pdv{u}{\xi}$$
+$$u_y= \pdv{u}{y}=\pdv{u}{\xi}\pdv{\xi}{y}+\pdv{u}{\eta}\pdv{\eta}{y}= -\pdv{u}{\eta}$$
 Plugging these into our original PDE,
-$$2x\left( \pdv{u}{\xi}\left(\frac{1}{2}+ \frac{1}{2x}\right) +\pdv{u}{\eta} \right)+\left(x+1\right)\left(-\pdv{u}{\xi}\right)=$$
-$$\pdv{u}{\xi}\left(x+1\right)+2x \pdv{u}{\eta}-\pdv{u}{\xi}\left(x+1\right)=2x \pdv{u}{\eta}$$
-So we must have $2x\pdv{u}{\eta}=y$. Now replace $\eta$ and $\xi$ by $x$ and $y$:
-$$2\eta\pdv{u}{\eta}=\frac{\eta}{2}+\frac{\ln\eta}{2}-\xi$$
+$$2x\left( \pdv{u}{\eta}\left(\frac{1}{2}+ \frac{1}{2x}\right) +\pdv{u}{\xi} \right)+\left(x+1\right)\left(-\pdv{u}{\eta}\right)=$$
+$$\pdv{u}{\eta}\left(x+1\right)+2x \pdv{u}{\xi}-\pdv{u}{\eta}\left(x+1\right)=2x \pdv{u}{\xi}$$
+So we must have $2x\pdv{u}{\xi}=y$. Now replace $\xi$ and $\eta$ by $x$ and $y$:
+$$2\xi\pdv{u}{\xi}=\frac{\xi}{2}+\frac{\ln\xi}{2}-\eta$$
 
-$$\pdv{u}{\eta}=\frac{1}{4}+\frac{\ln\eta}{4\eta}-\frac{\xi}{2n}$$
+$$\pdv{u}{\xi}=\frac{1}{4}+\frac{\ln\xi}{4\xi}-\frac{\eta}{2n}$$
 
-Now integrate with respect to $\eta$:
-$$u=\frac{\eta}{4}+\frac{(\ln\eta)^2}{8}-\frac{\xi \ln\eta}{2}+F(\xi)$$
+Now integrate with respect to $\xi$:
+$$u=\frac{\xi}{4}+\frac{(\ln\xi)^2}{8}-\frac{\eta \ln\xi}{2}+F(\eta)$$
 $$u(x,y)=\frac{x}{4}+\frac{(\ln x)^2}{8}-\frac{\ln x}{2}\left(\frac{x}{2}+\frac{\ln x}{2}-y \right)+F\left(\frac{x}{2}+\frac{\ln x}{2}-y\right)$$

### 2024-02-03 bio_dynamics/bio274_hw3.Rmd
bio hw 3

---
title: "BIOL 274 Homework 3"
author: "Ruby Krasnow"
date: "2024-02-03"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(PNWColors)
```

## Question 1
For each of the following DEs, state the order of the DE and whether it is linear or nonlinear. If it is linear, state whether the coefficients are constant or not:

## Question 2
A certain drug is being administered intravenously to a hospital patient who has had no prior drug treatments. Fluid containing 5 mg/cm^3^ of the drug enters the patient’s bloodstream at a rate of 100 cm^3^/hr. The drug is absorbed by tissues or otherwise leaves the bloodstream at a rate proportional to the amount present, with a rate constant of 0.4 (hr)^-1^.

(a) Assuming that the drug is always uniformly distributed throughout the bloodstream, write a differential equation for the amount of the drug that is present in the bloodstream at any time and state the initial condition.

Let $y(t)$ be the amount of the drug in the bloodstream at any time $t$. 
Then the rate of change of the drug, $\frac{dy}{dt}$, is equal to the rate of the drug entering the bloodstream minus the rate of the drug leaving the bloodstream:

$$\frac{dy}{dt}= \text{rate entering - rate leaving} = $$
$$\left(\frac{5 \text{ mg}}{\text{cm}^3}\right) \left( \frac{100 \text{ cm}^3}{\text{hr}}\right)-(y \text{ mg}) \left( \frac{0.4}{\text{hr}}   \right)=500-0.4y \;\text{mg hr}^{-1}$$

So $$\frac{dy}{dt}=500-0.4y$$

(b) Using the method of integrating factor find the solution to this initial value problem.

$$\frac{dy}{dt}=500-0.4y \quad\longrightarrow \quad \frac{dy}{dt}+0.4y=500$$
$$h(t)=e^{\int 0.4\,dt}=e^{0.4t} $$
$$y(t)= \frac{1}{h(t)}\int h(t)g(t)dt =$$
$$e^{-0.4t} \int 500e^{0.4t} = 500e^{-0.4t} \left( \frac{1}{0.4}e^{0.4t}+C_1\right)=1250+C_2e^{-0.4t}$$ where $C_2=500C_1$.

$$y(t)= 1250+C_2e^{-0.4t}$$

Check by differentiating:
$$y'(t)=-0.4\left(C_2e^{-0.4t}\right)$$
$$500-0.4y=500-0.4 \left(1250+C_2e^{-0.4t}\right)=$$
$$500-500-0.4\left(C_2e^{-0.4t}\right)=-0.4\left(C_2e^{-0.4t}\right)=y'(t)$$
So we have confirmed that out solution satisfies the original differential equation.

(c) How much of the drug is present in the bloodstream after a long time (i.e., as $t \rightarrow \infty$)?

$\lim\limits_{t \to \infty}e^{-0.4t}=0$, which means that $\lim\limits_{t \to \infty}y(t)=1250+0=1250$. Thus, 1250 mg of the drug is present in the bloodstream after a long time.

For problems 3 and 4, find the solution to each of the following IVPs:
## Question 3
$$ \frac{dy}{dx}= \frac{-y}{x}+2, \quad y(1)=3 $$

$$ \frac{dy}{dx}+ \frac{1}{x}y=2$$
$$h(x)=e^{\int \frac{1}{x}\,dx}=e^{\ln x}=x $$

$$y(x)= \frac{1}{h(x)}\int h(x)g(x)dx =$$
$$y(x)= \frac{1}{x}\int 2x\;dx =\frac{1}{x} \left(x^2+C\right) \quad\longrightarrow\quad y(x)=x+\frac{C}{x}$$

Plugging in our initial conditions, we get $y(1)=3=1+C$, which means $C=2$ and our final solution to the IVP is:
$$y(x)=x+\frac{2}{x} $$

Checking our solution by differentiating:
$$y'(x)=1- \frac{2}{x^2} $$
If we rewrite our original DE, we can see that it is equivalent:
$$ \frac{dy}{dx}= \frac{-y}{x}+2 = \frac{-x-\frac{2}{x}}{x}+2=-1- \frac{2}{x^2}+2=1- \frac{2}{x^2}$$

## Question 4
$$ \frac{dp}{dt}= \frac{1}{t+1}p+4t^2+4t, \quad p(1)=10$$
$$\frac{dp}{dt}-\frac{1}{t+1}p= 4t^2+4t$$

$$h(t)=e^{\int \frac{-1}{t+1}\,dx}=e^{-\ln (t+1)}= e^{\ln  \left(\frac{1}{t+1}\right)}=\frac{1}{t+1}$$

$$p(t)= \frac{1}{h(t)}\int h(t)g(t)dt =$$

$$(t+1) \int\frac{4t^2+4t}{t+1}dt= $$

$$(t+1) \int\frac{4t(t+1)}{t+1}dt = 4(t+1) \int t\;dt =$$

$$ 4(t+1) \left(\frac{t^2}{2} + C_1\right) = 2t^2(t+1)+ C_2(t+1) $$

Where $C_2=4C_1$. Therefore, our general solution is:
$$p(t)=(t+1)(2t^2+C_2) $$

Now plugging in our initial conditions $p(1)=10$, we get:
$$10=(2)(2+C_2) \quad\longrightarrow\quad 5=2+C_2 \quad\longrightarrow\quad C_2=3$$

So our final solution to the IVP is:
$$p(t)=(t+1)(2t^2+3) $$

Check by differentiating:
$$p'(t)=(t+1)(4t)+(2t^2+3)=4t^2+4t+2t^2+3 $$
And since 
$$ 2t^2+3 =\frac{(t+1)(2t^2+3)}{t+1} = \frac{p(t)}{t+1}$$,

This means our solution indeed satisfies the IVP.

## Question 5
A 400-gallon tank initially contains 200 gallons of water containing 2 parts per billion by weight of dioxin, an extremely potent carcinogen. Suppose water containing 5 parts per billion of dioxin flows into the top of the tank at a rate of 4 gallons per minute. The water in the tank is kept well mixed, and 2 gallons per minute are removed from the bottom of the tank. How much dioxin is in the tank when the tank is full?

```{r}
l_per_g <- 0.2641720524 # 1 L = 0.2641720524 gal
c <- 1/l_per_g
c
ppb5 <- 5*c

print(paste("1 part per billion = 1 µg per liter = 1 µg per", l_per_g,"gal = ", c, "µg/gal"))
print(paste("5 ppb =", ppb5, "µg/gal"))
```

<!-- $1 \text{ ppb} = \frac{1 \mu \text{g}} {\text{L}} = \frac{1 \mu\text{g}}{0.2641720524 \text{ gal}}= `r c` \mu \text{g/gal}$ -->

The volume of water at any time $t$ is
$$200\text{ gal}+\left( \frac{4 \text{ gal}}{\text{min}} -\frac{2 \text{ gal}}{\text{min}}\right)(t \text{ min})  = 200+2t$$

Let $y(t)$ be the amount of dioxin in the tank at any time $t$.

$$\frac{dy}{dt}= \text{rate entering - rate leaving} = $$

$$\left(\frac{18.927 \text{ µg}}{\text{gal}}\right) \left( \frac{4 \text{ gal}}{\text{min}}\right)-
\left(\frac{y \text{ µg}}{200+2t \text{ gal}}\right)\left( \frac{2 \text{ gal}}{\text{min}}\right)$$

$$\frac{dy}{dt}=4(18.927)- \frac{y}{100+t}= 75.708-\frac{y}{100+t}$$

$$\frac{dy}{dt}+\frac{1}{100+t}y= 75.708$$

$$h(t)=e^{\int \frac{1}{t+100}\,dx}=e^{\ln (t+100)}= t+100$$

$$y(t)= \frac{1}{h(t)}\int h(t)g(t)dt =$$

$$\frac{1}{t+100}\int 75.708(t+100)dt = \frac{75.708}{t+100} \left( \frac{t^2}{2}+100t+C_1\right)= \frac{75.708}{t+100}\left( \frac{t^2}{2}+100t+C_1\right) $$

$$y(t)=\frac{37.854t^2+7570.8t+C_2}{t+100}$$

$$y(0)=\frac{C_2}{100}$$
```{r}
200*2*c #200 gallons * 2 ppb * ~3.785 (µg/gal)/(1 ppb) = initial amount of dioxin, in µg = y_0
200*2*c*100 #C_2
```

$$y(t)=\frac{37.9t^2+7570.8t+151416.5}{t+100}$$

### 2024-02-03 bio_dynamics/bio274_hw3.Rmd
finished bio hw3

@@
 library(tidyverse)
-library(PNWColors)
 ```
@@
 For each of the following DEs, state the order of the DE and whether it is linear or nonlinear. If it is linear, state whether the coefficients are constant or not:
+
+(a) $\frac{dy}{dt}=t^2-y^2$. Rewriting as $\frac{dy}{dt}+y^2=t^2$, this is a first order nonlinear ODE.
+
+(b) $17y''=y'+t^2-3y$. Rewriting as  $17y''-y'+3y=t^2$, this is a second order linear ODE with constant coefficients.
+
+(c) $(\sin t)y= \frac{d^3y}{dt^3}- \frac{d^2y}{dt^2}+\cos t$. Rewriting as $\frac{d^3y}{dt^3}- \frac{d^2y}{dt^2}-(\sin t)y=-\cos t$, this is a third order linear ODE with non-constant coefficients.
+
+(d) $yy'=14e^t$ is a first order nonlinear ODE.
 
@@
 
+Since the patient has had no previous drug treatments, we can assume that there is no drug in the bloodstream at time $t=0$, i.e., we have an initial condition of $y(0)=0$.
+
 (b) Using the method of integrating factor find the solution to this initial value problem.
@@
 $$y(t)= 1250+C_2e^{-0.4t}$$
+Plugging in our initial condition,
+$0=1250+C_2e^0 \quad\longrightarrow\quad C_2=-1250$
+
+$$y(t)= 1250-1250e^{-0.4t}$$
 
 Check by differentiating:
-$$y'(t)=-0.4\left(C_2e^{-0.4t}\right)$$
-$$500-0.4y=500-0.4 \left(1250+C_2e^{-0.4t}\right)=$$
-$$500-500-0.4\left(C_2e^{-0.4t}\right)=-0.4\left(C_2e^{-0.4t}\right)=y'(t)$$
+$$y'(t)=-0.4\left(-1250e^{-0.4t}\right)=500e^{-0.4t}$$
+Now we rewrite our initial DE:
+$$500-0.4y=$$
+$$500-0.4 \left(1250-1250e^{-0.4t}\right)=$$
+$$500-500+500e^{-0.4t}=$$
+$$500e^{-0.4t}=y'(t)$$
 So we have confirmed that out solution satisfies the original differential equation.
@@
 
-For problems 3 and 4, find the solution to each of the following IVPs:
 ## Question 3
@@
 And since 
-$$ 2t^2+3 =\frac{(t+1)(2t^2+3)}{t+1} = \frac{p(t)}{t+1}$$,
-
-This means our solution indeed satisfies the IVP.
+$$ 2t^2+3 =\frac{(t+1)(2t^2+3)}{t+1} = \frac{p(t)}{t+1}$$
+we get $p'(t)=\frac{1}{t+1}p+4t^2+4t$, which exactly matches our original ODE.
 
@@
 
-```{r}
+```{r, echo=FALSE}
 l_per_g <- 0.2641720524 # 1 L = 0.2641720524 gal
 c <- 1/l_per_g
-c
 ppb5 <- 5*c
 
-print(paste("1 part per billion = 1 µg per liter = 1 µg per", l_per_g,"gal = ", c, "µg/gal"))
+print(paste("1 part per billion = 1 µg/L = 1 µg per", l_per_g,"gal = ", c, "µg/gal"))
 print(paste("5 ppb =", ppb5, "µg/gal"))
 ```
-
-<!-- $1 \text{ ppb} = \frac{1 \mu \text{g}} {\text{L}} = \frac{1 \mu\text{g}}{0.2641720524 \text{ gal}}= `r c` \mu \text{g/gal}$ -->
 
@@
 $$\left(\frac{18.927 \text{ µg}}{\text{gal}}\right) \left( \frac{4 \text{ gal}}{\text{min}}\right)-
-\left(\frac{y \text{ µg}}{200+2t \text{ gal}}\right)\left( \frac{2 \text{ gal}}{\text{min}}\right)$$
+\left(\frac{y(t) \text{ µg}}{200+2t \text{ gal}}\right)\left( \frac{2 \text{ gal}}{\text{min}}\right)$$
 
@@
 
-$$\frac{1}{t+100}\int 75.708(t+100)dt = \frac{75.708}{t+100} \left( \frac{t^2}{2}+100t+C_1\right)= \frac{75.708}{t+100}\left( \frac{t^2}{2}+100t+C_1\right) $$
+$$\frac{1}{t+100}\int 75.708(t+100)dt = \frac{75.708}{t+100} \left( \frac{t^2}{2}+100t+C_1\right)$$
 
@@
 $$y(0)=\frac{C_2}{100}$$
+
+The starting amount of dioxin, in µg $(y_0)$, is:
 ```{r}
-200*2*c #200 gallons * 2 ppb * ~3.785 (µg/gal)/(1 ppb) = initial amount of dioxin, in µg = y_0
-200*2*c*100 #C_2
+200*2*c #200 gallons * 2 ppb * 3.785 (µg/gal)/(1 ppb)
+
+```
+Which means that $C_2$ is 151416.5 and the solution to the IVP is
+
+$$y(t)=\frac{37.9t^2+7570.8t+151416.5}{t+100}$$
+
+Check our solution by differentiating:
+$$y'(t)=\frac{(t+100)[(37.9)(2t)+7570.8]-(37.9t^2+7570.8t+151416.5)}{(t+100)^2}=$$
+$$\frac{75.708t^2+7570.8t+7570.8t+757082-(37.9t^2+7570.8t+151416.5)}{(t+100)^2}= $$
+
+$$\frac{75.708t^2+15141.6t+757082-(37.9t^2+7570.8t+151416.5)}{(t+100)^2}$$
+
+Our original ODE was
+
+$$\frac{dy}{dt}= 75.708-\frac{y}{100+t} = $$
+
+$$75.708- \frac{37.9t^2+7570.8t+151416.5}{(t+100)^2}=$$
+
+$$75.708 \frac{(t+100)^2}{(t+100)^2}- \frac{37.9t^2+7570.8t+151416.5}{(t+100)^2}=$$
+
+$$\frac{75.708(t+100)^2 - (37.9t^2+7570.8t+151416.5)}{(t+100)^2}=$$
+
+$$\frac{75.708t^2+15141.6t+757082-(37.9t^2+7570.8t+151416.5)}{(t+100)^2}$$
+Which is exactly what we got by differentiating our solution.
+
+However, the question does not ask for the solution to the IVP, it asks how much dioxin is in the tank when the tank is full.
+The tank will be full at the time $t$ such that $200+2t=400$, which means the tank is full at $t=100$ minutes.
+Now we can plug $t=100$ into our solution:
+
+$$y(t)=\frac{37.9t^2+7570.8t+151416.5}{t+100}$$
+
+$$y(100)=\frac{379000+757080+151416.5}{200}=6437.5 \text{ µg} =6.44 \text{ mg}$$
+
+```{r}
+# Using unrounded values gives us the same answer, 6.44 mg
+((c*10*100*100)+(c*2*1000*100)+(c*4*10000))/200/1000
+
+# Divide by 400 gal and convert units to ppb
+((c*10*100*100)+(c*2*1000*100)+(c*4*10000))/200/(400*c)
+
 ```
 
-$$y(t)=\frac{37.9t^2+7570.8t+151416.5}{t+100}$$
+So, there are approximately 6.44 mg of dioxin in the tank when it is full (containing 400 gallons), corresponding to around 4.25 parts per billion.
+Given that the initial concentration was 2 parts per billion and the water entering the tank is 5 parts per billion, this seems like a reasonable value.

### 2024-02-07 kelp_bacteria.R
setting up citation mgmt

@@ -1,0 +1,21 @@
+#Kelp microbiome review
+
+library(metagear)
+library(RefManageR)
+library(PRISMAstatement)
+
+prisma(found = 5933,
+       found_other = 106,
+       no_dupes = 776, 
+       screened = 776, 
+       screen_exclusions = 13, 
+       full_text = 763,
+       full_text_exclusions = 17, 
+       qualitative = 746, 
+       quantitative = 319,
+       width = 800, height = 800)
+
+
+
+ReadZotero(user = "8204423", .params=list(q = "kelp", key = "QJbKC91HmGH0hwnxNXOZHOFH",
+                                          collection = "LMELYQTG", limit=5))

### 2024-02-07 kelp_bacteria.R
bib2df

@@ -1,8 +1,9 @@
 #Kelp microbiome review
 
 library(metagear)
-library(RefManageR)
+#library(RefManageR)
 library(PRISMAstatement)
+library(bib2df)
 
 prisma(found = 5933,
        found_other = 106,
@@ -15,7 +16,4 @@ prisma(found = 5933,
        quantitative = 319,
        width = 800, height = 800)
 
-
-
-ReadZotero(user = "8204423", .params=list(q = "kelp", key = "QJbKC91HmGH0hwnxNXOZHOFH",
-                                          collection = "LMELYQTG", limit=5))
+all_refs <- bib2df("retrieved_biome.bib")

### 2024-02-09 kelp_bacteria.R
biblio

@@ -1,19 +1,0 @@
-#Kelp microbiome review
-
-library(metagear)
-#library(RefManageR)
-library(PRISMAstatement)
-library(bib2df)
-
-prisma(found = 5933,
-       found_other = 106,
-       no_dupes = 776, 
-       screened = 776, 
-       screen_exclusions = 13, 
-       full_text = 763,
-       full_text_exclusions = 17, 
-       qualitative = 746, 
-       quantitative = 319,
-       width = 800, height = 800)
-
-all_refs <- bib2df("retrieved_biome.bib")

### 2024-02-09 kelp_sysrev/kelp_bacteria.R

@@ -1,0 +1,19 @@
+#Kelp microbiome review
+
+library(PRISMAstatement)
+library(tidyverse)
+
+prisma(found = 1448,
+       found_other = 106,
+       no_dupes = 1483, 
+       screened = 776, 
+       screen_exclusions = 13, 
+       full_text = 763,
+       full_text_exclusions = 17, 
+       qualitative = 746, 
+       quantitative = 319,
+       width = 800, height = 800)
+
+
+screened_refs <- read_csv()
+

### 2024-02-09 PDEs/math245_hw2.Rmd
math hw proiblem 2

---
title: "MATH 245 Homework 2"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-02-09"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1
Determine the region in which the given equation is hyperbolic, parabolic, elliptic, or
singular.

a) $$u_{xx}+y^2u_{yy}+u_x-u+x^2=0 $$

b) $$u_{xx}-yu_{yy}+xu_x+yu_y+u=0 $$

## Question 2
Using a factorization similar to the wave equation, solve the following IVP:
\begin{equation}
    \begin{cases}
      u_{xx}+2u_{xy}-3u_{yy}=0 & x\subset \mathbb{R},\; y> 0\\
      u(0,x)=\sin{x} & x\subset \mathbb{R}\\
      u_y(0,x) = x & x\subset \mathbb{R}
    \end{cases}       
\end{equation}

First, we can factor the equation as follows:
$$\left( \pdv{}{x}+3\pdv{}{y}\right) \left(\pdv{}{x}-\pdv{}{y}\right)u=0$$
or $$\left( \partial_{x}+3\partial_{y}\right) \left(\partial_{x}-\partial_{y}\right)u=0$$

Then set $\left(\pdv{}{x}-\pdv{}{y}\right)u=v$, giving us

$\left( \pdv{}{x}+3\pdv{}{y}\right)v= v_x+3v_t=0$

which we know has the solution $v(x,y)=f(3x-y)$, so
$$u_x-u_y=f(3x-y)$$

On (characteristic) lines with the slope $y=-x+c$, or $y+x=\text{constant}$, we must have $u_x-u_y=f(3x-y)=0$.
Set $\eta=x+y$ and $\xi=x$. 
Then by the chain rule,
$$u_x= u_{\eta}+u_{\xi}, \; u_y=u_{\eta}$$
And let's rewrite $y$ as $y=\eta-x=\eta-\xi$.

So
$$u_x-u_y=f(3x-y) \quad\longrightarrow\quad u_{\xi}=f(3\xi-\eta+\xi)$$

$$u_{\xi}=f(4\xi-\eta)$$
Now integrate with respect to $\xi$:
$$u(\eta, \xi)=F(4\xi-\eta)+G(\eta)$$
where $F$ is the antiderivative of $f$ with respect to $\xi$.

Now convert back to our original variables:
$$u(x,y)=F(3x-y)+G(x+y)$$

Using the fact that $u(0,x)=\sin{x}$,

 \begin{equation} \label{eq:1} u(0,x)=\sin{x}=F(3x)+G(x) \end{equation}
now replace $x$ with a new neutral variable, $\alpha$ and differentiate:
 $$\sin{\alpha}=F(3\alpha)+G(\alpha) $$
\begin{equation} \label{eq:2} \cos{\alpha}=3F'(3\alpha)+G'(\alpha) \end{equation}

But we can also differentiate $u(x,y)=F(3x-y)+G(x+y)$ with respect to $y$ to get
$$u_y(x,y)=-F'(3x-y)+G'(x+y)$$
but from our initial conditions, we know
$$u_y(0,x)=-F'(3x-0)+G'(x+0)=x$$

Let's replace $x$ by our neutral variable $\alpha$ and solve for $F'$:
$$F'(\alpha)=G'(3\alpha)-\alpha$$
Now plug this into \ref{eq:2}:
$$\cos{\alpha}=3G'(\alpha)-3\alpha+G'(\alpha)$$

$$G(\alpha)= \frac{1}{4} \int {\cos{\alpha}+3\alpha} = \frac{\sin{\alpha}}{4}+ \frac{3 \alpha^2}{8} $$
So that means \ref{eq:1} becomes:
$$\sin{\alpha}=F(3\alpha)+\frac{\sin{\alpha}}{4}+ \frac{3 \alpha^2}{8}$$

$$F(\alpha)=\frac{3\sin{(\frac{\alpha}{3})}}{4}- \frac{\alpha^2}{24}$$

Which means $u(x,y)=F(3x-y)+G(x+y)$ becomes

$$u(x,y)=\frac{3}{4}\sin{\left(x-\frac{y}{3}\right)}- \frac{(3x-y)^2}{24}+  \frac{\sin{(x+y)}}{4}+ \frac{3(x+y)^2}{8}=$$

\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=\frac{3}{4}\sin{\left(x-\frac{y}{3}\right)}+\frac{\sin{(x+y)}}{4}+ xy+  \frac{y^2}{3} $$
\end{tcolorbox}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]

 $$u_y=\frac{-1}{4}\cos{\left(x-\frac{y}{3}\right)}+\frac{\cos{(x+y)}}{4}+ x+  \frac{2y}{3} $$
  $$u_{yy}=\frac{-1}{12}\sin{\left(x-\frac{y}{3}\right)}-\frac{\sin{(x+y)}}{4}+ \frac{2}{3} $$

$$u_x=\frac{3}{4}\cos{\left(x-\frac{y}{3}\right)}+\frac{\cos{(x+y)}}{4}+y$$

$$u_{xx}=\frac{-3}{4} \sin{\left( x-\frac{y}{3} \right)} - \frac{\sin{(x+y)}}{4}$$

$$u_{xy}=\frac{1}{4}\sin{\left(x-\frac{y}{3}\right)}-\frac{\sin{(x+y)}}{4}+1$$

Check that $u_{xx}+2u_{xy}-3u_{yy}=0$
$$
\left(\frac{-3}{4}+ \frac{2}{4}+\frac{1}{4}\right)\sin{\left(x-\frac{y}{3}\right)}+
\left(\frac{-1}{4}+ \frac{-2}{4}+\frac{3}{4}\right)\sin{(x+y)}+(0+2-3)=0
$$

\end{tcolorbox}

## Question 3

### 2024-02-09 kelp_sysrev/kelp_bacteria.R

@@ -6,14 +6,12 @@ library(tidyverse)
 prisma(found = 1448,
        found_other = 106,
        no_dupes = 1483, 
-       screened = 776, 
-       screen_exclusions = 13, 
-       full_text = 763,
-       full_text_exclusions = 17, 
-       qualitative = 746, 
-       quantitative = 319,
+       screened = 1483, 
+       screen_exclusions = 1308, 
+       full_text = 175,
+       full_text_exclusions = 125, 
+       qualitative = 50, 
+       quantitative = 50,
        width = 800, height = 800)
 
-
-screened_refs <- read_csv()
 

### 2024-02-10 bio_dynamics/bio274_hw4.Rmd
bio274 HW4

---
title: "BIOL 274 Homework 4"
author: "Ruby Krasnow"
date: "2024-02-09"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1
Solve each of the following IVPs:

(a)
\begin{equation}
    \begin{cases}
      y'' + 6y' + 8y = 0\\
      y(0)=3\\
      y'(0)=-16
    \end{cases}       
\end{equation}

The characteristic equation is
$$\lambda^2+6\lambda+8=0 \quad\longrightarrow\quad (\lambda+4)(\lambda+2)=0 $$

Which means we have two distinct real roots, $\lambda_1=-4$ and $\lambda_1=-2$.

Thus, our general solution is 
$$y=C_1e^{-4x}+C_2e^{-2x}$$

We differentiate to find that
$$y'=-4C_1e^{-4x}-2C_2e^{-2x}$$

Now we plug in our initial conditions:
$$3=C_1+C_2$$
$$-16=-4C_1-2C_2$$

Multiply both sides of the first equation by 2 and add to the second equation:

$-10=-2C_1 \quad\longrightarrow\quad C_1=5$, so $C_2=3-5=-2$.
Which means the solution to the IVP is

\begin{tcolorbox}[colback=white, title=Solution]
$$y(x)= 5e^{-4x}-2e^{-2x}$$
\end{tcolorbox}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
$$y'(x)= -20e^{-4x}+4e^{-2x}$$

$$y''(x)= 80e^{-4x}-8e^{-2x}$$

$$y'' + 6y' + 8y = (80-120+40)e^{-4x}+(-8+24-16)e^{-2x}=0$$

\end{tcolorbox}

(b)
\begin{equation}
    \begin{cases}
      y'' + 6y' + 9y = 0\\
      y(0)=\frac{1}{3}\\
      y'(0)=\frac{1}{2}
    \end{cases}       
\end{equation}

The characteristic equation is
$$\lambda^2+6\lambda+9=0 \quad\longrightarrow\quad (\lambda+3)(\lambda+3)=0 $$
So we have a repeated real root, $\lambda_1=\lambda_2=-3$.
Thus, our general solution is 
$$y=C_1e^{-3x}+C_2xe^{-3x}$$
We differentiate to find that
$$y'=-3C_1e^{-3x}+C_2e^{-3x}-3C_2xe^{-3x}$$

Now we plug in our initial conditions:
$$\frac{1}{3}=C_1$$

$$\frac{1}{2}=-3C_1+C_2 \quad\longrightarrow\quad C_2=\frac{3}{2}$$
Which means the solution to the IVP is

\begin{tcolorbox}[colback=white, title=Solution]
$$y(x)= \frac{1}{3}e^{-3x}+\frac{3}{2}xe^{-3x}$$
\end{tcolorbox}
\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]

$$y'(x)= -e^{-3x}-\frac{9}{2}xe^{-3x}+\frac{3}{2}e^{-3x}=\frac{1}{2}e^{-3x}-\frac{9}{2}xe^{-3x}$$
$$y''(x)=-\frac{3}{2}e^{-3x}+\frac{27}{2}xe^{-3x}-\frac{9}{2}e^{-3x}=\frac{27}{2}xe^{-3x}-6e^{-3x}$$
Plugging these expressions into our original ODE, we see that
$$y'' + 6y' + 9y =
\left(\frac{27}{2} -\frac{54}{2}+\frac{27}{2} \right)xe^{-3x} + \left(-6 +3 +3\right)e^{-3x}=0 $$
\end{tcolorbox}

## Question 2
Consider the equation
$$ \frac{d^2y}{dt^2}+ \frac{k}{m}y=0 $$
for the motion of a single (undamped) harmonic oscillator.

(a)  
Consider the function $y(t)=\cos{(\beta t)}$. Under what conditions on $\beta$ is $y(t)$ a solution?

$\lambda^2+\frac{k}{m}=0$

$$\frac{-b\pm\sqrt{b^2-4ac}}{2a} $$

$$\frac{0\pm\sqrt{0-4\frac{k}{m}}}{2} $$

$$0\pm\sqrt{\frac{k}{m}}\:i $$
$y(t)=\cos{(\beta t)}$ is only a solution if $\beta=\sqrt{\frac{k}{m}}$.

(b)  
What initial conditions $(t=0)$ correspond to this solution?

(c)
In terms of $k$ and $m$, what is the period of this solution?

$y(t)=\cos{(\sqrt{\frac{k}{m}} t)}$
$y'(t)=-\sqrt{\frac{k}{m}}\sin{(\sqrt{\frac{k}{m}} t)}$
$y''(t)=-\frac{k}{m}\cos{(\sqrt{\frac{k}{m}} t)}$

Then $y''+\frac{k}{m}y=-\frac{k}{m}\cos{(\sqrt{\frac{k}{m}} t)}+\frac{k}{m}\cos{(\sqrt{\frac{k}{m}} t)}=0$

## Question 3  
Find the general solution for each of the following.  
(a)
\begin{equation}\label{eq:1} y''+3y'+2y=4t^2 -2t \end{equation}

The characteristic equation is
$$\lambda^2+3\lambda+2=0 \quad\longrightarrow\quad (\lambda+2)(\lambda+1)=0 $$

Which means we have two distinct real roots, $\lambda_1=-2$ and $\lambda_1=-1$.

Thus, our complementary function (solution) is 
$$y_c=C_1e^{-2t}+C_2e^{-t}$$

Let's assume our solution takes the form $y=At^2+Bt+C$.
Then $y'=2At+B$ and $y''=2A$, so the left-hand side of \ref{eq:1} becomes

$$2A+6At+3B+2At^2+2Bt+2C=4t^2-2t$$

We can split this up by coefficients of $t$ into
$$2At^2=4t^2$$
$$6At+2Bt=-2t $$
$$2A+3B+2C=0$$

Which means that $A=2,\; B=-7,\; C= \frac{17}{2}$ and so 
$$y_p=2t^2-7t+ \frac{17}{2}$$

Combine $y_p$ with $y_c$ to find our general solution:
\begin{tcolorbox}[colback=white, title=Solution]
$$y(t)=C_1e^{-2t}+C_2e^{-t}+2t^2-7t+ \frac{17}{2}$$
\end{tcolorbox}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]

$$y'(t)=-2C_1e^{-2t}-C_2e^{-t}+4t-7$$

$$y''(t)=4C_1e^{-2t}+C_2e^{-t}+4$$

$$y''+3y'+2y=\left(4C_1-6C_1+2C_1 \right)e^{-2t}+
\left(C_2-3C_1+2C_1 \right)e^{-t}+4+12t-21+4t^2-14t+17=4t^2-2t $$
\end{tcolorbox}

(b)
\begin{equation}\label{eq:2} y''-y'-2y=15e^2t \end{equation}

The characteristic equation is
$$\lambda^2-\lambda-2=0 \quad\longrightarrow\quad (\lambda-2)(\lambda+1)=0 $$

Which means we have two distinct real roots, $\lambda_1=2$ and $\lambda_1=-1$.

Thus, our complementary function (solution) is 
$$y_c=C_1e^{2t}+C_2e^{-t}$$

Let's first assume our solution takes the form $y=Ae^{2t}$.
Then $y'=2Ae^{2t}$ and $y''=4Ae^{2t}$, so the left-hand side of \ref{eq:2} becomes

$$4Ae^{2t}-2Ae^{2t}-2Ae^{2t}=0 \neq 15e^2t $$

Oh no! What happened? Well, $e^{2t}$ satisfies the associated homogenous DE and so is part of the complementary solution. Instead, we will try $y=Ate^{2t}$. Then $y'=2Ate^{2t}+Ae^{2t}$ and $y''=4Ate^{2t}+4Ae^{2t}$, so the left-hand side of \ref{eq:2} becomes

$$y''-y'-2y= \left(4A-2A-2A\right)te^{2t}+ \left(4A-A\right)e^{2t} $$

Which gives us $3Ae^{2t}=15e^{2t}$, meaning $A=5$ and we have
$$y_p=5te^{2t}$$

Combine $y_p$ with $y_c$ to find our general solution:
\begin{tcolorbox}[colback=white, title=Solution]
$$y(t)=C_1e^{2t}+C_2e^{-t}+5te^{2t}$$ or
$$y(t)=(C_1+5t)e^{2t}+C_2e^{-t}$$
\end{tcolorbox}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
$$y'(t)=(2C_1+10t+5)e^{2t}-C_2e^{-t}$$

$$y''(t)=(4C_1+20t+20)e^{2t}+C_2e^{-t}$$

$$y''-y'-2y= (4C_1+20t+20-2C_1-10t-5-2C_1-10t)e^{2t}+(C_2+C_2-2C_2)e^{-t}=15e^{2t}$$

\end{tcolorbox}

(c)
\begin{equation}\label{eq:3} y''-2y'-3y= \cos{(2t)}\end{equation}

The characteristic equation is
$$\lambda^2-2\lambda-3=0 \quad\longrightarrow\quad (\lambda-3)(\lambda+1)=0 $$

Which means we have two distinct real roots, $\lambda_1=3$ and $\lambda_1=-1$.

Thus, our complementary function (solution) is 
$$y_c=C_1e^{3t}+C_2e^{-t}$$

Let's assume our solution takes the form $y=A\sin{(2t)}+B\cos{(2t)}$. Then 
$$y'=2A\cos{(2t)}-2B\sin{(2t)} $$

$$y''=-4A\sin{(2t)}-4B\cos{(2t)} $$

So the left-hand side of \ref{eq:3} becomes
$$y''-2y'-3y=(-4A+4B-3A)\sin{(2t)}+(-4B-4A-3B)\cos{(2t)} $$
Compare the coefficients of the $\sin$ and $\cos$ terms to the original DE:
$$-7A+4B=0 $$
$$-4A-7B=1 $$

$A= \frac{4B}{7}$
$$-\frac{16B}{7}-\frac{49B}{7}=1 $$
Which gives us that $B=\frac{-7}{65}, \; A= \frac{-4}{65}$.
Thus, our particular solution is 
$$y_p=-\frac{4}{65} \sin{(2t)}-\frac{7}{65}\cos{(2t)}$$

Combine $y_p$ with $y_c$ to find our general solution:
\begin{tcolorbox}[colback=white, title=Solution]

$$y(t)=C_1e^{3t}+C_2e^{-t}-\frac{4}{65} \sin{(2t)}-\frac{7}{65}\cos{(2t)}$$

\end{tcolorbox}
\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]

$$y'(t)=3C_1e^{3t}-C_2e^{-t}+\frac{14}{65} \sin{(2t)}-\frac{8}{65}\cos{(2t)}$$

$$y''(t)=9C_1e^{3t}+C_2e^{-t}+\frac{16}{65} \sin{(2t)}+\frac{28}{65}\cos{(2t)}$$

$$
y''-2y'-3y=(9-6-3)C_1e^{3t}+(1+2-3)C_2e^{-t}+ \\
\left(\frac{16}{65}-\frac{28}{65}+\frac{12}{65}\right)\sin{(2t)}+ \\
\left(\frac{28}{65}+\frac{16}{65}+\frac{21}{65}\right)\cos{(2t)}=
$$
$$\cos{(2t)}$$
\end{tcolorbox}

(d)
\begin{equation}\label{eq:4} y''+4y= e^t+\sin{(2t)}\end{equation}

The characteristic equation is
$$\lambda^2+4=0 \quad\longrightarrow\quad \lambda=0\pm2i $$
Thus, our complementary function (solution) is 
$$y_c=C_1\sin{(2t)}+C_2\cos{(2t)}$$
Let's break up our right hand side to find $y_{p_1}$ and $y_{p_2}$.

For $y_{p_1}$, we assume our solution takes the form $y=Ae^t$. Then 
$Ae^t+4Ae^t=e^t$, so $A= \frac{1}{5}$ and we have 
$$y_{p_1}=\frac{e^t}{5}$$

For $y_{p_2}$, we cannot assume our solution takes the form $y=A\sin{(2t)}+B\cos{(2t)}$ because we had complex roots $\pm2i$ for our characteristic equation. Instead, we try

$$y=At\sin{(2t)}+Bt\cos{(2t)}$$

$$y'=2At\cos{(2t)}+A\sin{(2t)}+B\cos{(2t)}-2Bt\sin{(2t)}=$$
$$(2At+B)\cos{(2t)}+(-2Bt+A)\sin{(2t)}$$

$$y''= (-4At-2B)\sin{(2t)}+2A\cos{(2t)}+(-4Bt+2A)\cos{(2t)}-2B\sin{(2t)}=$$

$$ (-4At-4B)\sin{(2t)}+(-4Bt+4A)\cos{(2t)}$$

$$ (-4At-4B+4At)\sin{(2t)}+(-4Bt+4A+4Bt)\cos{(2t)}=\sin{(2t)}$$

$$ -4B\sin{(2t)}+4A\cos{(2t)}=\sin{(2t)}$$
$$A=0, B= -\frac{1}{4}$$
$$y_{p_2}=-t\frac{\cos{(2t)}}{4}$$
$$y_p=\frac{e^t}{5}-\frac{t\cos{(2t)}}{4} $$
Now adding this to our complementary function, we get

\begin{tcolorbox}[colback=white, title=Solution]

$$y(t)=C_1\sin{(2t)}+C_2\cos{(2t)}+\frac{e^t}{5}-\frac{t\cos{(2t)}}{4}$$
\end{tcolorbox}
\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]

$$y'(t)=2C_1\cos{(2t)}-2C_2\sin{(2t)}+\frac{e^t}{5}-\frac{\cos{(2t)}}{4}+\frac{t\sin{(2t)}}{2}$$

$$y''(t)=-4C_1\sin{(2t)}-4C_2\cos{(2t)}+\frac{e^t}{5}+\frac{\sin{(2t)}}{2}+\frac{\sin{(2t)}}{2}+t\cos{(2t)}$$

$$ y''+4y= \left(-4C_1+1+4C_1 \right)\sin{(2t)}+ \left(-4C_2+t+4C_2-t\right)\cos{(2t)}+\frac{e^t}{5}++\frac{4e^t}{5}= $$
$$ e^t+\sin{(2t)} $$
\end{tcolorbox}

### 2024-02-10 bio_dynamics/bio274_hw4.Rmd
bio hw 4 prob 2

@@
 author: "Ruby Krasnow"
-date: "2024-02-09"
+date: "2024-02-10"
 output:
@@
 
-$\lambda^2+\frac{k}{m}=0$
+$$\lambda^2+\frac{k}{m}=0$$
 
@@
 
+We can check this:
+$$y(t)=\cos{(\sqrt{\frac{k}{m}} t)}$$
+$$y'(t)=-\sqrt{\frac{k}{m}}\sin{(\sqrt{\frac{k}{m}} t)}$$
+$$y''(t)=-\frac{k}{m}\cos{(\sqrt{\frac{k}{m}} t)}$$
+$$y''+\frac{k}{m}y=-\frac{k}{m}\cos{(\sqrt{\frac{k}{m}} t)}+\frac{k}{m}\cos{(\sqrt{\frac{k}{m}} t)}=0$$
+
 (b)  
 What initial conditions $(t=0)$ correspond to this solution?
+
+$y(0)=\cos{(0)}=1$, so we must have the initial condition $y(0)=1$.
 
@@
 
-$y(t)=\cos{(\sqrt{\frac{k}{m}} t)}$
-$y'(t)=-\sqrt{\frac{k}{m}}\sin{(\sqrt{\frac{k}{m}} t)}$
-$y''(t)=-\frac{k}{m}\cos{(\sqrt{\frac{k}{m}} t)}$
-
-Then $y''+\frac{k}{m}y=-\frac{k}{m}\cos{(\sqrt{\frac{k}{m}} t)}+\frac{k}{m}\cos{(\sqrt{\frac{k}{m}} t)}=0$
+In the equation for simple harmonic oscillation $y(t) = A \cos(\omega t + \varphi)$, $\omega=\sqrt{\frac{k}{m}}$ represents the angular frequency, measured in radians per second.
+$\omega$ is equivalent to $\frac{2\pi}{T}$, where $T$ is the period of the oscillation (the time it takes to complete one oscillation and return to the starting position).
+This means we can find T by taking $\frac{2\pi}{\omega}=\frac{2\pi}{\sqrt{\frac{k}{m}}}$, or
+$$T=2\pi \sqrt{\frac{m}{k}}$$
 
@@
 
-$A= \frac{4B}{7}$
+$$A= \frac{4B}{7}$$
 $$-\frac{16B}{7}-\frac{49B}{7}=1 $$

### 2024-02-13 PDEs/math245_hw2.Rmd
math hw2 num 4

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-02-09"
+date: "2024-02-12"
 output:
@@
 
+$a=1, b=0, c=-y^2$, so we have $b^2-ac=0-(-y^2)=y^2$. This will be positive everywhere except for $y=0$, so the equation is hyperbolic where $y \neq 0$ and parabolic for $y=0$.
+
 b) $$u_{xx}-yu_{yy}+xu_x+yu_y+u=0 $$
+
+$a=1, b=0, c=-y$, so we have $b^2-ac=0-(-y)=y$. Thus, the equation will be hyperbolic where $y>0$, parabolic where $y=0$, and elliptic where $y<0$.
 
@@
 ## Question 3
+
+Solve the Neumann boundary value problem for the wave equation on half line:
+\begin{equation}
+    \begin{cases}
+      u_{tt}=c^2u_{xx}+f(t,x) & 0<x<\infty\\
+      u(0,x)=\phi{x} & 0<x<\infty\\
+      u_t(0,x) = \psi{x} & 0<x<\infty\\
+      u_x(t,0) = h(t) & t>0
+    \end{cases}       
+\end{equation}
+
+## Question 4
+
+Consider the 3D wave equation for $u(t,x,y,z)$:
+
+$$u_{tt}=c^2\Delta u \quad\quad (x,y,z) \in \mathbb{R}^3, \quad t>0$$
+
+Change the coordinates to spherical coordinates.
+Assume the solution is spherically symmetric, so that $u(t,x,y,z)=u(t,r)$ and does not depend on $\theta$ and $\phi$.
+Find the solution for $u(0,r)=0$ and
+
+\begin{equation} u_t(0,r) = \begin{cases}
+  1  & \lvert{r}\rvert \leq 1 \\
+  0 &  \lvert{r}\rvert > 1
+\end{cases}
+\end{equation}
+
+Hint: use the substitution $u(t,r)= \frac{1}{r} w(t,r)$.
+
+First, we need to derive the formula for the Laplacian in spherical coordinates.
+
+We know the equation for the Laplacian in polar coordinates is:
+
+$$\frac{\partial^2}{\partial r^2}+\frac{1}{r}\frac{\partial}{\partial r}+ \frac{1}{r^2}\frac{\partial^2}{\partial \theta^2}$$
+
+Now let's convert to spherical coordinates:
+
+$$r=\sqrt{x^2+y^2+z^2}=\sqrt{s^2+z^2}$$
+
+$$x=s \cos{\phi}$$
+$$y=s \sin{\phi}$$
+$$z=r \cos{\theta}$$
+$$s=r \sin{\theta}$$
+
+By the two-dimensional Laplacian, we have
+
+$$u_{zz}+u_{ss}=u_{rr}+\frac{1}{r}u_r+\frac{1}{r^2}u_{\theta\theta}$$ and
+
+$$u_{xx}+u_{yy}=u_{ss}+\frac{1}{s}u_s+\frac{1}{s^2}u_{\phi\phi}$$
+We add these two equations and cancel $u_s$ to get
+
+$$\Delta u = u_{rr}+\frac{1}{r}u_r+\frac{1}{r^2}u_{\theta\theta}+\frac{1}{s}u_s+\frac{1}{s^2}u_{\phi\phi}$$
+
+Now since $u$ doesn't depend on $\theta$ or $\phi$, we have
+$$\Delta u = u_{rr}+\frac{1}{r}u_r+\frac{1}{s}u_s=u_{rr}+\frac{1}{r}u_r+\frac{1}{r \sin{\theta}}u_s$$
+
+$$ \frac{\partial u}{\partial s} = \frac{\partial u}{\partial r} \frac{\partial r}{\partial s}+\frac{\partial u}{\partial\theta}\frac{\partial\theta}{\partial s}+\frac{\partial u}
+{\partial\phi}\frac{\partial\phi}{\partial s}=u_r\frac{1}{ \sin{\theta}}+0+0=u_r\frac{s}{ r}$$
+
+So with our change of variables, we have
+
+$$u_{tt}=c^2 \left(u_{rr} + \frac{2}{r}u_r \right)$$
+
+Now set $w=ru$, or $u= \frac{w}{r}$. Then 
+$$w_t=ru_t, \quad w_{tt}=ru_{tt}, \quad u_{tt}= \frac{w_{tt}}{r} $$
+$$w_t=ru_t, \quad w_{tt}=ru_{tt}, \quad u_{tt}= \frac{w_{tt}}{r} $$
+$$u_r = \frac{w_r}{r}-\frac{w}{r^2} $$
+
+$$u_{rr} = \frac{w_{rr}}{r}-\frac{2w_r}{r^2}+\frac{2w}{r^3} $$
+So $$u_{tt}=c^2 \left(u_{rr} + \frac{2}{r}u_r \right)$$ becomes
+
+$$\frac{w_{tt}}{r} =c^2 \left(\frac{w_{rr}}{r}-\frac{2w_r}{r^2}+\frac{2w}{r^3} + \frac{2}{r} \left(\frac{w_r}{r}-\frac{w}{r^2} \right) \right)$$
+which simplifies to
+$$w_{tt}=c^2 w_{rr}$$
+but this is just the wave equation, which we know has the solution
+ $$w(t,r)= \frac{\varphi(r+ct)+\varphi(r-ct)}{2}+\frac{1}{2c} \int_{r-ct}^{r+ct}{\psi(s)ds}$$
+ Since $\varphi=0$, 
+ $$w(t,r)= \frac{1}{2c} \int_{r-ct}^{r+ct}{\psi(s)ds}$$
+
+Now we have 4 cases:
+
+Case 1: $r-ct \geq -1, r+ct \leq 1$
+$$w(t,r)= \frac{1}{2c} \int_{r-ct}^{r+ct}{s\;ds}$$
+
+Case 2:$r-ct < -1, r+ct>1$
+$$w(t,r)= \frac{1}{2c} \int_{-1}^{1}{s\;ds}$$
+
+Case 3:$r-ct < -1, r+ct \leq 1$
+$$w(t,r)= \frac{1}{2c} \int_{-1}^{r+ct}{s\;ds}$$
+
+Case 4:$r-ct \geq -1, r+ct>1$
+$$w(t,r)= \frac{1}{2c} \int_{r-ct}^{1}{s\;ds}$$
+
+Since $u=\frac{w}{r}$, this means we have
+\begin{equation}u(t,r)=
+    \begin{cases} 
+      \frac{1}{2c} \int_{r-ct}^{r+ct}{s\;ds} & r-ct \geq -1, r+ct \leq 1\\
+      \\
+      \frac{1}{2c} \int_{-1}^{1}{s\;ds} & r-ct < -1, r+ct > 1\\
+      \\
+      \frac{1}{2c} \int_{-1}^{r+ct}{s\;ds} & r-ct < -1, r+ct \leq 1\\
+      \\
+      \frac{1}{2c} \int_{r-ct}^{1}{s\;ds} & r-ct \geq -1, r+ct > 1\\
+
+    \end{cases}       
+\end{equation}
+
+## Question 5
+Consider the following Dirichlet boundary value problem:
+\begin{equation} 
+\begin{cases}
+  u_{tt}+x(t,x)u_t=u_{xx} & 0<x<1 \\
+  u(0,x)=\phi(x) & 0<x<1 \\
+  u_t(0,x)=\psi(x) & 0<x<1 \\
+  u(t,0) = u(t,1)=0 & t \geq 0 \\
+\end{cases}
+\end{equation}
+
+Assume that $\lvert a(t,x)\rvert \leq m$ for some constant $m$ and all $0<x<1$ and $t\geq 0$. Let 
+$$E(t)= \frac{1}{2} \int_{0}^{1}{ \left(u_t(t,x)^2+u_x(t,x)^2\right)}\:dx $$
+
+(1) Show that $\frac{dE(t)}{dt} \leq 2mE(t)$ for $t \geq 0$.
+(2) Use part (a) and show that $\frac{d}{dt} \left(e^{-2mE(t)}\right) \leq0$ for 
+all $t \geq 0$. Hence, by integration from $[0,t]$, we get that
+
+$$E(t) \leq e^{2mt}E(0) \quad \text{for all } t \geq 0$$
+
+(3) If $\phi(x)=\psi(x)=0$ for all $0<x<1$, what does this say about $E(t)$ for $t \geq 0$ and hence about $u(t,x)$ for $t \geq 0$?
+
+(4) Use the previous part to prove uniqueness of the following problem:
+\begin{equation} 
+\begin{cases}
+  u_{tt}+a(t,x)u_t=u_{xx} & 0<x<1, t>0 \\
+  u(0,x)=\phi(x) & 0<x<1 \\
+  u_t(0,x)=\psi(x) & 0<x<1 \\
+  u(t,0) = f(t) & t \geq 0 \\
+  u(t,1) = g(t) & t \geq 0 \\
+\end{cases}
+\end{equation}
+
+## Problem 6
+
+Does the D’Alembert method work for the wave equation $u_{tt} = c(x)^2u_{xx}$? What about $u_{tt} = c(t)^2u_{xx}$? Why?
+
+## Problem 7 (The Poisson-Darboux Equation)   
+Solve the initial value problem
+
+\begin{equation} 
+\begin{cases}
+  u_{tt}-u_{xx}- \frac{2}{x}u_x=0 & -\infty <x<\infty, t>0 \\
+  u(0,x)=0 & -\infty <x<\infty \\
+  u_t(0,x)=g(x) & -\infty <x<\infty \\
+\end{cases}
+\end{equation}
+
+where $g(x)=-g(x)$ is an even function. Hint: set $w=xu$.
+
+## Problem 8   
+Solve the following characteristic initial value problem:
+\begin{equation} 
+\begin{cases}
+  y^3u_{xx}-yu_{yy}+u_y=0 & 0<x<4, \quad \lvert y \rvert \leq 2 \sqrt{2} \\
+  u(x,y)= f(x) & x+ \frac{y^2}{2}=4 \text{for} 2 \leq x \leq 4 \\
+  u(x,y)=g(x) & x- \frac{y^2}{2}=0 \text{for} 0 \leq x \leq 2 \\
+\end{cases}
+\end{equation}
+
+where $f(2)=g(2)$. Hint: Use the transformation $\eta=x-\frac{y^2}{2}$ and $\xi=x+\frac{y^2}{2}$ and express the PDE in the coordinates $(\xi, \eta)$.

### 2024-02-13 kelp_sysrev/kelp_bacteria.R

@@ -14,4 +14,6 @@ prisma(found = 1448,
        quantitative = 50,
        width = 800, height = 800)
 
+#files for full-text screening
+inital_screen <- read_csv("./kelp_sysrev/initial_screen.csv")
 

### 2024-02-13 PDEs/math245_hw2.Rmd
more PDEs

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-02-12"
+date: "2024-02-13"
 output:
@@
 ```
+
+## Done, feel OK: 1, 2
+## Done, very low confidence: 4, 6
+## Mostly done: 7
+## Not started: 3, 5, 8
 
@@
 
+Let's try the factorization method if $c=c(x)$:
+$$\left( \partial_{t}+c(x)\partial_{x}\right) \left(\partial_{t}-c(x)\partial_{x}\right)u=0$$
+Set $\xi=x+c(x)t$ and $\eta=x-c(x)t$.
+
+By the chain rule,
+
+$$\frac{\partial}{\partial x} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial x}+\frac{\partial}{\partial\eta}\frac{\partial\eta}{\partial x}$$
+
+$$=\partial_{\xi}(1+c')+\partial_{\eta}\,(1-c')$$
+
+$$\frac{\partial}{\partial t} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial t}+\frac{\partial }{\partial\eta}\frac{\partial\eta}{\partial t}$$
+$$=c\partial_{\xi}-c\partial_{\eta}$$
+When we plug these back into
+$$\left( \partial_{t}+c\partial_{x}\right) \left(\partial_{t}-c\partial_{x}\right)u=0$$
+we get
+
+$$ \left(c\partial_{\xi}-c\partial_{\eta}+c\partial_{\xi}(1+c')+c\partial_{\eta}\,(1-c')\right)
+\left(c\partial_{\xi}-c\partial_{\eta} -c\partial_{\xi}(1+c')-c\partial_{\eta}\,(1-c')\right)u=0$$
+
+When $c$ is constant, $c'=0$ and this simplifies to $-4c^2 u_{\xi\eta}=0$, allowing us to integrate to find $u(x,t)=F(x+ct)+G(x-ct)$.
+But we cannot do the same simplification when we have the $c'$ terms, meaning we cannot solve the wave equation by the same method when the wave speed is not constant.
+
+Similarly,
+
+Let's try the factorization method if $c=c(t)$:
+$$\left( \partial_{t}+c(t)\partial_{x}\right) \left(\partial_{t}-c(t)\partial_{x}\right)u=0$$
+Set $\xi=x+c(t)t$ and $\eta=x-c(t)t$.
+
+By the chain rule,
+
+$$\frac{\partial}{\partial x} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial x}+\frac{\partial}{\partial\eta}\frac{\partial\eta}{\partial x}=
+\partial_{\xi}+\partial_{\eta}$$
+
+$$\frac{\partial}{\partial t} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial t}+\frac{\partial }{\partial\eta}\frac{\partial\eta}{\partial t}=\partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)
+$$
+
+When we plug these back into
+$$\left( \partial_{t}+c\partial_{x}\right) \left(\partial_{t}-c\partial_{x}\right)u=0$$
+we get
+
+$$\left( \partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)+c\partial_{\xi}+c\partial_{\eta} \right)
+\left(\partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)-c\partial_{\xi}-c\partial_{\eta}\right)u=0$$
+Where again we cannot cancel terms to simplify our equation as we could with constant $c$.
+
 ## Problem 7 (The Poisson-Darboux Equation)   
@@
 
-where $g(x)=-g(x)$ is an even function. Hint: set $w=xu$.
+where $g(x)=g(-x)$ is an even function. Hint: set $w=xu$.
+
+Using the results from when we set $w=ru$ in Problem 4,
+$$u_t=\frac{w_t}{x}, u_{tt}=\frac{w_{tt}}{x} $$
+$$u_x= \frac{w_x}{x}- \frac{w}{x^2} $$
+
+$$u_{xx}=  \frac{w_{xx}}{x}-\frac{2w_x}{x^2}+\frac{2w}{x^3}$$
+So $u_{tt}-u_{xx}- \frac{2}{x}u_x=0$ becomes
+
+$$\frac{w_{tt}}{x} -\frac{w_{xx}}{x}+\frac{2w_x}{x^2}- \frac{2w}{x^3}- \frac{2}{x} \left(\frac{w_x}{x}- \frac{w}{x^2}\right)=0 $$
+Which simplifies to $w_{tt}-w_{xx}=0$, the wave equation with $c=1$.
+By the D'Alembert formula, since we have $\varphi=0$,
+$$w(t,x)=\frac{1}{2c} \int_{x-t}^{x+t}{\psi(s)ds}$$
+$w_t=xu_t$
+Since $y=x$ is an odd function and $u_t(0,x)=g(x)$ is an even function, $w_t=xg(x)=h(x)$ must be an odd function.
 
@@
   y^3u_{xx}-yu_{yy}+u_y=0 & 0<x<4, \quad \lvert y \rvert \leq 2 \sqrt{2} \\
-  u(x,y)= f(x) & x+ \frac{y^2}{2}=4 \text{for} 2 \leq x \leq 4 \\
-  u(x,y)=g(x) & x- \frac{y^2}{2}=0 \text{for} 0 \leq x \leq 2 \\
+  u(x,y)= f(x) & x+ \frac{y^2}{2}=4 \text{ for } 2 \leq x \leq 4 \\
+  u(x,y)=g(x) & x- \frac{y^2}{2}=0 \text{ for } 0 \leq x \leq 2 \\
 \end{cases}

### 2024-02-13 PDEs/math245_old.Rmd
math hw proiblem 2
math hw2 num 4
more PDEs

@@
 ```
+
+Then set $\xi=x+3y, \; \eta=x-y$, giving us
+
+$$\partial_x= \pdv{}{\xi} \pdv{\xi}{x}+\pdv{}{\eta} \pdv{\eta}{x}=\partial_{\xi}+\partial_{\eta}$$
+$$\partial_y= \pdv{}{\xi} \pdv{\xi}{y}+\pdv{}{\eta} \pdv{\eta}{y}= 3\partial_{\xi}-\partial_{\eta}$$
+Which means that 
+$$\partial_{x}+3 \partial_{y}= 10 \partial_{\xi}-2\partial_{\eta}$$
+
+$$\partial_{x}-\partial_{y}=-2\partial_{\xi}+2\partial_{\eta} $$
+
+Which means we can rewrite the original equation as 
+$$\left( \partial_{x}+3\partial_{y}\right) \left(\partial_{x}-\partial_{y}\right)u=0 \quad\longrightarrow\quad -8\partial$$
+\begin{equation}\label{eq:sphere1}
+x=r \sin{\theta} \cos{\phi}, \quad y=r \sin{\theta} \sin{\phi}, \quad z=r \cos{\theta}
+\end{equation}
+
+\begin{equation}\label{eq:sphere2}
+r=\sqrt{x^2+y^2+z^2}, \quad \theta=\arccos{(\frac{z}{r})}, \quad \phi=\arctan{(\frac{y}{x})} \end{equation}
+
+$$\theta=\arccos{(\frac{z}{r})}$$
+
+Using the chain rule:
+$$\Delta u= \frac{{\partial}^{2}f}{\partial x^{2}}+\frac{{\partial}^{2}f}{\partial y^{2}}+\frac{{\partial}^{2}f}{\partial z^{2}}$$
+
+Now we find the following values from (\ref{eq:sphere1}):
+$$\frac{\partial r}{\partial x}=\sin(\theta)\cos(\phi),
+\qquad\frac{\partial\theta}{\partial x}=\frac{1}{r}\cos(\theta)\cos(\phi),
+\qquad\frac{\partial\phi}{\partial x}=-\frac{1}{r}\frac{\sin(\phi)}{\sin(\theta)}$$
+$$\frac{\partial r}{\partial y}=\sin(\theta)\sin(\phi),\qquad\frac{ \partial\theta}{\partial y}=\frac{1}{r}\cos(\theta)\sin(\phi),\qquad\frac{\partial\phi}{\partial y}=\frac{1}{r}\frac{\cos(\phi)}{\sin(\theta)} $$
+
+$$\frac{\partial r}{\partial z}=\cos(\theta),\qquad\frac{\partial \theta}{\partial z}=-\frac{1}{r}\sin(\theta),\qquad\frac{\partial\phi}{\partial z}=0$$
+$$ \frac{\partial u}{\partial x} = \frac{\partial u}{\partial r} \frac{\partial r}{\partial x}+\frac{\partial u}{\partial\theta}\frac{\partial\theta}{\partial x}+\frac{\partial u}
+{\partial\phi}\frac{\partial\phi}{\partial x}=u_r\sin(\theta)\cos(\phi)$$
+
+$$\frac{\partial u_x}{\partial x} = \frac{\partial u_x}{\partial r} \frac{\partial r}{\partial x}+\frac{\partial u_x}{\partial\theta}\frac{\partial\theta}{\partial x}+\frac{\partial u_x}{\partial\phi}\frac{\partial\phi}{\partial x}= $$
+
+$$u_{rr}\sin^2(\theta)\cos^2(\phi)+ \frac{u_r}{r} \cos^2(\theta)\cos^2(\phi)- $$
+
+$$u(x,t)= \frac{\varphi(x+ct)+\varphi(x-ct)}{2}+\frac{1}{2c} \int_{x-ct}^{x+ct}{\psi(s)ds}$$
+Assume $c$ is not a constant, but instead a function of $x$.

### 2024-02-14 PDEs/math245_hw2.Rmd
PDE tears

@@
 
-## Done, feel OK: 1, 2
+## Done, feel OK: 1, 2, 5, 7
 ## Done, very low confidence: 4, 6
-## Mostly done: 7
-## Not started: 3, 5, 8
+## Not started: 3, 8
 
@@
 $$F'(\alpha)=G'(3\alpha)-\alpha$$
-Now plug this into \ref{eq:2}:
+Now plug this into (\ref{eq:2}):
 $$\cos{\alpha}=3G'(\alpha)-3\alpha+G'(\alpha)$$
@@
 $$G(\alpha)= \frac{1}{4} \int {\cos{\alpha}+3\alpha} = \frac{\sin{\alpha}}{4}+ \frac{3 \alpha^2}{8} $$
-So that means \ref{eq:1} becomes:
+So that means (\ref{eq:1}) becomes:
 $$\sin{\alpha}=F(3\alpha)+\frac{\sin{\alpha}}{4}+ \frac{3 \alpha^2}{8}$$
@@
 
-$$x=s \cos{\phi}$$
-$$y=s \sin{\phi}$$
-$$z=r \cos{\theta}$$
+$$x=s \cos{\phi}, \quad y=s \sin{\phi}, \quad z=r \cos{\theta}$$
 $$s=r \sin{\theta}$$
@@
 
-$$u_{zz}+u_{ss}=u_{rr}+\frac{1}{r}u_r+\frac{1}{r^2}u_{\theta\theta}$$ and
+$$u_{zz}+u_{ss}=u_{rr}+\frac{1}{r}u_r+\frac{1}{r^2}u_{\theta\theta},$$
 
@@
 $$u_{rr} = \frac{w_{rr}}{r}-\frac{2w_r}{r^2}+\frac{2w}{r^3} $$
-So $$u_{tt}=c^2 \left(u_{rr} + \frac{2}{r}u_r \right)$$ becomes
+So $u_{tt}=c^2 \left(u_{rr} + \frac{2}{r}u_r \right)$ becomes
 
-$$\frac{w_{tt}}{r} =c^2 \left(\frac{w_{rr}}{r}-\frac{2w_r}{r^2}+\frac{2w}{r^3} + \frac{2}{r} \left(\frac{w_r}{r}-\frac{w}{r^2} \right) \right)$$
+$$\frac{w_{tt}}{r} =c^2 \left(\frac{w_{rr}}{r}-\frac{2w_r}{r^2}+\frac{2w}{r^3} + \frac{2}{r} \left(\frac{w_r}{r}-\frac{w}{r^2} \right) \right),$$
 which simplifies to
-$$w_{tt}=c^2 w_{rr}$$
-but this is just the wave equation, which we know has the solution
+$$w_{tt}=c^2 w_{rr}.$$
+But this is just the wave equation, and we can use d'Alembert's formula to find the solution:
  $$w(t,r)= \frac{\varphi(r+ct)+\varphi(r-ct)}{2}+\frac{1}{2c} \int_{r-ct}^{r+ct}{\psi(s)ds}$$
@@
 
+$$= \frac{1}{4c} ((r+ct)^2-(r-ct)^2)$$
+$$= \frac{1}{4c}(r^2+2crt+c^2t^2-r^2+2crt-c^2t^2)= \frac{4crt}{4c}=rt  $$
+
 Case 2:$r-ct < -1, r+ct>1$
-$$w(t,r)= \frac{1}{2c} \int_{-1}^{1}{s\;ds}$$
+$$w(t,r)= \frac{1}{2c} \int_{-1}^{1}{s\;ds}=$$
+$$w(t,r)= \frac{1}{4c} (1-1)=0$$
 
@@
 
+$$= \frac{1}{4c} ((r+ct)^2-1)$$
+
 Case 4:$r-ct \geq -1, r+ct>1$
 $$w(t,r)= \frac{1}{2c} \int_{r-ct}^{1}{s\;ds}$$
+$$= \frac{1}{4c} (1-(r-ct)^2)$$
 
 Since $u=\frac{w}{r}$, this means we have
+\begin{tcolorbox}[colback=white, title=Solution]
 \begin{equation}u(t,r)=
     \begin{cases} 
-      \frac{1}{2c} \int_{r-ct}^{r+ct}{s\;ds} & r-ct \geq -1, r+ct \leq 1\\
-      \\
-      \frac{1}{2c} \int_{-1}^{1}{s\;ds} & r-ct < -1, r+ct > 1\\
-      \\
-      \frac{1}{2c} \int_{-1}^{r+ct}{s\;ds} & r-ct < -1, r+ct \leq 1\\
-      \\
-      \frac{1}{2c} \int_{r-ct}^{1}{s\;ds} & r-ct \geq -1, r+ct > 1\\
-
+      t & \text{ if } r-ct \geq -1, r+ct \leq 1\\
+      0 & \text{ if } r-ct < -1, r+ct > 1\\
+      \frac{1}{4rc} ((r+ct)^2-1) & \text{ if } r-ct < -1, r+ct \leq 1\\
+      \frac{1}{4rc} (1-(r-ct)^2) &\text{ if }  r-ct \geq -1, r+ct > 1\\
     \end{cases}       
 \end{equation}
+\end{tcolorbox}
 
@@
 
-(1) Show that $\frac{dE(t)}{dt} \leq 2mE(t)$ for $t \geq 0$.
-(2) Use part (a) and show that $\frac{d}{dt} \left(e^{-2mE(t)}\right) \leq0$ for 
+(1) Show that 
+\begin{equation}\label{eqn:E} 
+\frac{dE(t)}{dt} \leq 2mE(t)
+\end{equation}
+for $t \geq 0$.
+
+First differentiate $E(t)$:
+
+$$ \frac{dE}{dt}= \frac{d}{dt} \left[\frac{1}{2} \int_{0}^{1}{ \left(u_t^2+u_x^2\right)}\:dx \right]$$ 
+\begin{equation} \label{eqn:prob5a}
+=\frac{1}{2} \int_{0}^{1}{ \frac{\partial}{\partial t} \left(u_t^2+u_x^2\right)} \:dx
+\end{equation}
+
+\begin{equation} \label{eqn:prob5b}
+=\frac{1}{2} \int_{0}^{1}{ 2u_t u_{tt}+ 2u_x u_{xt}}\:dx
+\end{equation}
+
+\begin{equation}\label{eqn:prob5c}
+=\int_{0}^{1}{ u_t u_{tt}}\: dx + \int_{0}^{1}{2u_x u_{xt}}\:dx =: I+J.
+\end{equation}
+
+The equality (\ref{eqn:prob5a}) follows from differentiation under the integral sign, while (\ref{eqn:prob5b}) follows from the chain rule for partial derivatives.
+
+Consider the integral $J$ in (\ref{eqn:prob5c}).
+By integrating by parts, we can move one of the partials $\frac{\partial}{\partial x}$ to the other factor, at the cost of introducing a minus sign and a boundary term.
+Hence $J$ becomes
+
+\begin{equation} \label{eqn:J}
+J=\int_{0}^{1}{ u_x u_{xt}}\: dx =- \int_{0}^{1}{u_xx u_t}\:dx + 
+\left[ u_x u_t \right]_{x=0}^{x=1}.
+\end{equation}
+
+The boundary term vanishes, since $u(t,0) \equiv u(t,1)  \equiv 0$ for $t>0$ implies that $u_t$ is identically zero at $x=0,1$.
+So, substituting (\ref{eqn:J}) for $J$, equation (\ref{eqn:prob5c}) becomes
+ $$ I+J=\int_{0}^{1}{u_t u_{tt}} \;dx - \int_{0}^{1}{u_{xx} u_t} \;dx$$
+ $$= \int_{0}^{1} {u_t (u_{tt}-u_{xx})} \;dx$$
+\begin{equation} \label{eqn:J2}
+=\int_{0}^{1}{ u_t(-au_t) }\:dx
+\end{equation}
+
+\begin{equation} \label{eqn:J3}
+=\int_{0}^{1}{ (-a)u_t^2 }\: dx.
+\end{equation}
+
+Here, equality (\ref{eqn:J2}) just uses the PDE.
+Since $u_t^2 \geq 0$ and $-a \leq \lvert -a \rvert \leq m$, we see that $(-a)u_t^2 \leq mu_t^2$.
+Hence, the expression in (\ref{eqn:J3}) satisfies the inequality
+
+\begin{equation} \label{eqn:IJ}
+I+J= \int_{0}^{1}{ (-a)u_t^2 }\:dx \leq \int_{0}^{1}{ mu_t^2 }\:dx \leq m \int_{0}^{1}{ u_t^2 + u_x^2 }\:dx =2mE.
+\end{equation}
+
+Where we also used the fact that $u_x^2 \geq 0$ means that $m\int_{0}^{1}{ u_t^2 }\:dx \leq m\int_{0}^{1}{ (u_t^2 + u_x^2) }\:dx$.
+
+The desired inequality (\ref{eqn:E}) follows from (\ref{eqn:IJ}) and the fact that $\frac{dE}{dt}=I+J.$
+
+(2) Use part (1) and show that $\frac{d}{dt} \left( e^{-2mE(t)} \right) \leq 0$ for 
 all $t \geq 0$. Hence, by integration from $[0,t]$, we get that
 
-$$E(t) \leq e^{2mt}E(0) \quad \text{for all } t \geq 0$$
+\begin{equation}\label{eqn:leq}
+E(t) \leq e^{2mt} E(0) \text{ for all } t \geq 0.
+\end{equation}
+
+By the product rule, 
+$$ \frac{d}{dt} \left(e^{-2mt}E\right)=-2me^{-2mt}E+e^{-2mt}\frac{dE}{dt} $$
+$$=e^{-2mt}\left(\frac{dE}{dt}-2mE\right) $$
+$$\leq e^{-2mt} \cdot 0=0. $$
 
 (3) If $\phi(x)=\psi(x)=0$ for all $0<x<1$, what does this say about $E(t)$ for $t \geq 0$ and hence about $u(t,x)$ for $t \geq 0$?
+
+Since $u(0,x) \equiv \varphi(x) \equiv 0$ for $0<x<1$, we see that $u_x$ is identically zero at time $t=0$. Similarly, $u_t(0,x) \equiv \psi(x) \equiv 0$ for $0<x<1$. Thus at $t=0$,
+
+\begin{equation}\label{eqn:E0}
+E(0)= \frac{1}{2} \int_{0}^{1}{ \left(u_t(t,0)^2+u_x(t,0)^2\right)}\:dx =\frac{1}{2} \int_{0}^{1}{0^2+0^2 \:dx=0.}
+\end{equation}
+
+But (\ref{eqn:E0}) together with the inequality (\ref{eqn:leq}) implies that
+
+$$0 \leq E(t)  \leq e^{2mt}E(0) \leq 0$$
+for all $t \geq 0$.
+Hence the energy $E$ is identically zero.
+Since the integrand $u_t^2+u_x^2$ is nonnegative, this is only possible if $u_t \equiv u_x \equiv 0$ for $t>0, 0<x<1$, meaning that $u$ varies with neither time nor position.
+But this implies that $u$ must be identically zero everywhere.
 
@@
 \end{equation}
+
+Let $u$ and $v$ be two solutions, and define $w:=u-v$.
+Observe that $w$ satisfies the original boundary value problem along with the conditions specified in part (3), implying that $w \equiv 0$.
+Hence $u-v\equiv 0$ and so any solution $u$ must be unique.
+$\square$
 
@@
   u(0,x)=0 & -\infty <x<\infty \\
-  u_t(0,x)=g(x) & -\infty <x<\infty \\
+  u_t(0,x)=g(x) & -\infty <x< \infty \\
 \end{cases}
@@
 $$\frac{w_{tt}}{x} -\frac{w_{xx}}{x}+\frac{2w_x}{x^2}- \frac{2w}{x^3}- \frac{2}{x} \left(\frac{w_x}{x}- \frac{w}{x^2}\right)=0 $$
-Which simplifies to $w_{tt}-w_{xx}=0$, the wave equation with $c=1$.
-By the D'Alembert formula, since we have $\varphi=0$,
-$$w(t,x)=\frac{1}{2c} \int_{x-t}^{x+t}{\psi(s)ds}$$
-$w_t=xu_t$
-Since $y=x$ is an odd function and $u_t(0,x)=g(x)$ is an even function, $w_t=xg(x)=h(x)$ must be an odd function.
+Which simplifies to $w_{tt}-w_{xx}=0$, and so $w$ solves the wave equation when $-\infty <x<\infty, t>0$.
+
+Now observe that $w(0,x)=xu(0,x)=0$ and $w_t(0,x)=xu_t(0,x)=xg(x)$. Hence, $w$ solves the initial value problem:
+
+\begin{equation} \label{eqn:prob7}
+    \begin{cases}
+      w_{tt}-w_{xx} & -\infty <x<\infty, t>0 \\
+      w(0,x)=0 & -\infty <x<\infty, t>0 \\
+      w_t(0,x) = xg(x) & -\infty <x<\infty, t>0
+    \end{cases}       
+\end{equation}
+
+whre $g(x)$ is an even function.
+
+By d'Alembert's formula, the solution to (\ref{eqn:prob7}) is given by:
+$$w(t,x)=\frac{1}{2} \int_{x-t}^{x+t}{sg(s)ds}$$
+Hence since $w_t=xu_t$, 
+\begin{equation} \label{eqn:prob7a}
+w(t,x)=\frac{1}{2x} \int_{x-t}^{x+t}{sg(s)ds,}
+\end{equation}
+assuming $x \neq 0$.
+
+To handle the case when $x=0$, assume that $u$ is continuous in $x$, so that $u(t,0)=\lim_{x\to 0} u(t,x)$. Define
+
+$$I(y)=\int_{0}^{y}sg(s)\;ds. $$
+
+This allows us to rewrite (\ref{eqn:prob7a}) as
+
+$$u(t,x)=\frac{I(x+t)-I(x-t)}{2x}= \frac{I(x+t)-I(t-x)}{2x},$$
+Noting that the evenness of $g(s)$ means that $sg(s)$ is odd, and so $I(y)$ is even (one can check by substitution that $I(y)=I(-y)$). Hence, letting $x \rightarrow 0$, we see that
+
+\begin{equation}\label{eqn:prob7b}
+u(t,0)= \lim_{x\to 0} \frac{I(t+x)-I(t-x)}{2x}.
+\end{equation}
+
+But this is just the form of a derivative, namely the symmetric derivative of $I$ at $t$.
+By the Fundamental Theorem of Calculus, assuming $g$ is continuous, we know the limit (\ref{eqn:prob7b}) exists and is necessarily equal to $I'(t)=tg(t)$.
+
+Thus $u(t,x)$ is given for any $t \geq 0, x \in \mathbb{R}$ by
+
+\begin{equation} \label{eqn:prob7c} 
+u(t,x)=
+\begin{cases}
+      \frac{1}{2x} \int_{x-t}^{x+t}{sg(s)ds,} & x \neq 0, \\
+      tg(t), & x=0 \\
+    \end{cases} 
+\end{equation}
 

### 2024-02-16 PDEs/math245_hw2.Rmd
math hw prob 8

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-02-13"
+date: "2024-02-15"
 output:
@@
 ## Done, feel OK: 1, 2, 5, 7
-## Done, very low confidence: 4, 6
+## Done, very low confidence: 4
+## Redo: 6
 ## Not started: 3, 8
@@
 $$u_x-u_y=f(3x-y)$$
+Now we can reorder our equation as $$\left(\partial_{x}-\partial_{y}\right)
+\left( \partial_{x}+3\partial_{y}\right)u=0$$
+and set $w=\left( \partial_{x}+3\partial_{y}\right)u$
 
-On (characteristic) lines with the slope $y=-x+c$, or $y+x=\text{constant}$, we must have $u_x-u_y=f(3x-y)=0$.
-Set $\eta=x+y$ and $\xi=x$. 
-Then by the chain rule,
-$$u_x= u_{\eta}+u_{\xi}, \; u_y=u_{\eta}$$
-And let's rewrite $y$ as $y=\eta-x=\eta-\xi$.
+Then $$w_x-w_y=0 $$
 
-So
-$$u_x-u_y=f(3x-y) \quad\longrightarrow\quad u_{\xi}=f(3\xi-\eta+\xi)$$
+Which we know has the solution $w(x,y)=g(x+y)$. So $u_x+3u_y=g(x+y)$, which gives us a system of two equations:
 
-$$u_{\xi}=f(4\xi-\eta)$$
-Now integrate with respect to $\xi$:
-$$u(\eta, \xi)=F(4\xi-\eta)+G(\eta)$$
-where $F$ is the antiderivative of $f$ with respect to $\xi$.
+\begin{equation}
+\begin{cases}
+u_x-u_y=f(3x-y)
+u_x+3u_y=g(x+y)
+\end{cases}
+\end{equation}
 
-Now convert back to our original variables:
+Subtract the first equation from the second:
+$$4u_y=-f(3x-y)+g(x+y) $$
+
+Now we can integrate with respect to $y$ to find that:
 $$u(x,y)=F(3x-y)+G(x+y)$$
+where $F$ is the antiderivative of $-f$ with respect to $y$ and $G$ is the antiderivative of $g$ with respect to $y$.
 
@@
 where $f(2)=g(2)$. Hint: Use the transformation $\eta=x-\frac{y^2}{2}$ and $\xi=x+\frac{y^2}{2}$ and express the PDE in the coordinates $(\xi, \eta)$.
+
+Set $\eta=x-\frac{y^2}{2}$ and $\xi=x+\frac{y^2}{2}$. Then
+
+$$\frac{\partial u}{\partial x} = \frac{\partial u}{\partial \xi} \frac{\partial \xi}{\partial x}+\frac{\partial u}{\partial\eta}\frac{\partial\eta}{\partial x} = \frac{\partial u}{\partial \xi}+\frac{\partial u}{\partial\eta}$$
+
+$$\frac{\partial^2 u}{\partial x^2} = \frac{\partial }{\partial x} \left(\frac{\partial u}{\partial \xi}\right) +\frac{\partial }{\partial x} \left(\frac{\partial u}{\partial \eta}\right)$$
+
+$$=\frac{\partial^2 u}{\partial \xi^2} \frac{\partial \xi}{\partial x}+
+\frac{\partial^2 u}{\partial\eta \partial \xi}\frac{\partial \xi}{\partial x} +
+\frac{\partial^2 u}{\partial\eta \partial \xi}\frac{\partial\eta}{\partial x}+
+\frac{\partial^2 u}{\partial \eta^2}\frac{\partial \eta}{\partial x}  $$
+
+Which we can simplify to
+
+$$u_{xx}= u_{\xi \xi}+2 u_{\xi \eta}+u_{\eta \eta}$$
+
+Now we do the same for $y$:
+
+$$\frac{\partial u}{\partial y} = \frac{\partial u}{\partial \xi} \frac{\partial \xi}{\partial y}+\frac{\partial u}{\partial\eta}\frac{\partial\eta}{\partial y} = y\frac{\partial u}{\partial \xi}-y\frac{\partial u}{\partial\eta}$$
+
+But since $\eta=x-\frac{y^2}{2}$ and $\xi=x+\frac{y^2}{2}$, we can rewrite $y$ as $\sqrt{\xi-\eta}$.
+
+$$\frac{\partial u}{\partial y} = \sqrt{\xi-\eta} \left(\frac{\partial u}{\partial \xi}-\frac{\partial u}{\partial\eta}\right)$$
+
+Now using the product rule,
+$$\frac{\partial^2 u}{\partial y^2} = \sqrt{\xi-\eta}\; \frac{\partial}{\partial y}\left(\frac{\partial u}{\partial \xi}-\frac{\partial u}{\partial\eta}\right)+\left(\frac{\partial u}{\partial \xi}-\frac{\partial u}{\partial\eta}\right)$$
+
+Where the second term has been simplified because $\frac{\partial}{\partial y}$ of $\sqrt{\xi-\eta}$ is simply $\frac{\partial}{\partial y}y=1$.
+
+Now we use the chain rule again for the first term:
+
+Now using the product rule,
+$$\frac{\partial^2 u}{\partial y^2} = \sqrt{\xi-\eta}\left(\frac{\partial^2 u}{\partial \xi^2} \frac{\partial \xi}{\partial y}+
+\frac{\partial^2 u}{\partial\eta \partial \xi}\frac{\partial\eta}{\partial y} -
+\frac{\partial^2 u}{\partial \eta^2}\frac{\partial \eta}{\partial y}-
+\frac{\partial^2 u}{\partial \eta \partial \xi}\frac{\partial \xi}{\partial y}\right)
++
+\left(\frac{\partial u}{\partial \xi}-\frac{\partial u}{\partial\eta}\right)$$
+
+Which simplifies to
+
+$$u_{yy}= \left(\xi-\eta \right)\left(u_{\xi \xi}+u_{\eta \eta} -2 u_{\xi \eta}\right) +u_\xi-u_\eta$$
+
+This means $y^3u_{xx}-yu_{yy}+u_y=0$ becomes
+
+$$\left(\xi-\eta \right)^{\frac{3}{2}} \left(u_{\xi \xi}+2 u_{\xi \eta}+u_{\eta \eta}\right) - \left(\xi-\eta \right)^{\frac{1}{2}} \left[\left(\xi-\eta \right)\left(u_{\xi \xi}-2 u_{\xi \eta}+u_{\eta \eta} \right) +u_\xi-u_\eta \right]+\left(\xi-\eta \right)^{\frac{1}{2}} \left(u_\xi-u_\eta\right)=0  $$
+
+$$\left(\xi-\eta \right)^{\frac{3}{2}} \left(u_{\xi \xi}+2 u_{\xi \eta}+
+u_{\eta \eta}\right) - \left(\xi-\eta \right)^{\frac{3}{2}} \left(u_{\xi \xi}-2 u_{\xi \eta}+
+u_{\eta \eta}\right)-
+\left(\xi-\eta \right)^{\frac{1}{2}} \left(u_\xi-u_\eta\right)+
+\left(\xi-\eta \right)^{\frac{1}{2}} \left(u_\xi-u_\eta\right)=0  $$
+
+$$\left(\xi-\eta \right)^{\frac{3}{2}} \left(4\, u_{\xi \eta}\right)=0$$
+
+But by integrating with respect to both variables, we recognize $u_{\xi \eta}=0$ as having the solution
+
+$$u(\xi, \eta)=h_1(\xi)+h_2(\eta),$$
+
+or in our original coordinates,
+
+$$u(x,y)=h_1\left(x+\frac{y^2}{2}\right)+h_2\left(x-\frac{y^2}{2}\right) $$
+
+When $x=2$,
+
+$f(x)=h_1(4)+h_2(2x-4)$
+
+$z=2x-4, \quad x=\frac{z}{2} + 2$
+
+$h_2(z)=f\left(\frac{z}{2}+2\right)-h_1(4)$
+
+and
+$g(x)=h_1(2x)+h_2(0)$
+
+$h_1(z)=g\left(\frac{z}{2}\right)-h_2(0)$
+
+$h_1+h_2=g\left(\frac{z}{2}\right)-h_2(0)+f(\frac{z}{2} + 2)-h_1(4)$
+
+But $h_2(0)+h_1(4)$ is just $g(x)$ when $x=2$ (or equivalently, $f(x)$ when $x=2$),
+
+so after replacing $z$ by $x+\frac{y^2}{2}$ and $x-\frac{y^2}{2}$, we get
+
+\begin{tcolorbox}[colback=white, title=Solution]
+$$u(x,y)=g\left(\frac{x}{2}+\frac{y^2}{4}\right)+f\left(\frac{x}{2}+\frac{y^2}{4}+2\right)+f(2)$$
+\end{tcolorbox}

### 2024-02-16 PDEs/math245_old.Rmd
PDE tears
math hw prob 8

@@
 Assume $c$ is not a constant, but instead a function of $x$.
+
+Since $u=\frac{w}{r}$, this means we have
+\begin{equation}u(t,r)=
+    \begin{cases} 
+      \frac{1}{2cr} \int_{r-ct}^{r+ct}{s\;ds} & r-ct \geq -1, r+ct \leq 1\\
+      \\
+      \frac{1}{2cr} \int_{-1}^{1}{s\;ds} & r-ct < -1, r+ct > 1\\
+      \\
+      \frac{1}{2cr} \int_{-1}^{r+ct}{s\;ds} & r-ct < -1, r+ct \leq 1\\
+      \\
+      \frac{1}{2cr} \int_{r-ct}^{1}{s\;ds} & r-ct \geq -1, r+ct > 1\\
+
+    \end{cases}       
+\end{equation}
+
+On (characteristic) lines with the slope $y=-x+c$, or $y+x=\text{constant}$, we must have $u_x-u_y=f(3x-y)=0$.
+Set $\eta=x+y$ and $\xi=x$. 
+Then by the chain rule,
+$$u_x= u_{\eta}+u_{\xi}, \; u_y=u_{\eta}$$
+And let's rewrite $y$ as $y=\eta-x=\eta-\xi$.
+
+So
+$$u_x-u_y=f(3x-y) \quad\longrightarrow\quad u_{\xi}=f(3\xi-\eta+\xi)$$
+
+$$u_{\xi}=f(4\xi-\eta)$$
+Now integrate with respect to $\xi$:
+$$u(\eta, \xi)=F(4\xi-\eta)+G(\eta)$$
+where $F$ is the antiderivative of $f$ with respect to $\xi$.

### 2024-02-16 PDEs/math245_hw2.Rmd
finished PDE prob 8

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-02-15"
+date: "2024-02-16"
 output:
@@
 
-## Done, feel OK: 1, 2, 5, 7
-## Done, very low confidence: 4
+## Done, feel OK: 1, 2, 5, 7, 8
+## Done, low confidence: 4
 ## Redo: 6
-## Not started: 3, 8
+## Not started: 3
 
@@
 Using a factorization similar to the wave equation, solve the following IVP:
-\begin{equation}
+\begin{equation*}
     \begin{cases}
@@
     \end{cases}       
-\end{equation}
+\end{equation*}
 
@@
 
-\begin{equation}
+\begin{equation*}
 \begin{cases}
@@
 \end{cases}
-\end{equation}
+\end{equation*}
 
@@
 Solve the Neumann boundary value problem for the wave equation on half line:
-\begin{equation}
+\begin{equation*}
     \begin{cases}
@@
 
-\begin{equation} u_t(0,r) = \begin{cases}
+\begin{equation*} u_t(0,r) = \begin{cases}
   1  & \lvert{r}\rvert \leq 1 \\
@@
 \begin{tcolorbox}[colback=white, title=Solution]
-\begin{equation}u(t,r)=
+\begin{equation*}u(t,r)=
     \begin{cases} 
@@
     \end{cases}       
-\end{equation}
+\end{equation*}
 \end{tcolorbox}
@@
 Consider the following Dirichlet boundary value problem:
-\begin{equation} 
+\begin{equation*} 
 \begin{cases}
@@
 (4) Use the previous part to prove uniqueness of the following problem:
-\begin{equation} 
+\begin{equation*} 
 \begin{cases}
@@
 
-\begin{equation} 
+\begin{equation*} 
 \begin{cases}
@@
 Solve the following characteristic initial value problem:
-\begin{equation} 
+\begin{equation*} 
 \begin{cases}
@@
 
-When $x=2$,
+Now we use our given conditions. When  $2 \leq x \leq 4$, we have $g(x)=h_1(2x)+h_2(0)$, which means $h_1(z_1)=g\left(\frac{z_1}{2}\right)-h_2(0)$.
 
-$f(x)=h_1(4)+h_2(2x-4)$
+Similarly, when $0 \leq x \leq 2$, we have $f(x)=h_1(4)+h_2(2x-4)$, which means $h_2(z_2)=f\left(\frac{z_2}{2}+2\right)-h_1(4)$.
 
-$z=2x-4, \quad x=\frac{z}{2} + 2$
-
-$h_2(z)=f\left(\frac{z}{2}+2\right)-h_1(4)$
-
-and
-$g(x)=h_1(2x)+h_2(0)$
-
-$h_1(z)=g\left(\frac{z}{2}\right)-h_2(0)$
-
-$h_1+h_2=g\left(\frac{z}{2}\right)-h_2(0)+f(\frac{z}{2} + 2)-h_1(4)$
+Thus,
+$$h_1(z_1)+h_2(z_2)=g\left(\frac{z_1}{2}\right)-h_2(0)+f\left(\frac{z_2}{2} + 2\right)-h_1(4)$$
 
@@
 
-so after replacing $z$ by $x+\frac{y^2}{2}$ and $x-\frac{y^2}{2}$, we get
+so after replacing $z_1$ by $x+\frac{y^2}{2}$ and $z_2$ by $x-\frac{y^2}{2}$, we get
 
 \begin{tcolorbox}[colback=white, title=Solution]
-$$u(x,y)=g\left(\frac{x}{2}+\frac{y^2}{4}\right)+f\left(\frac{x}{2}+\frac{y^2}{4}+2\right)+f(2)$$
+$$u(x,y)=g\left(\frac{x}{2}+\frac{y^2}{4}\right)+f\left(\frac{x}{2}+\frac{y^2}{4}+2\right)-f(2)$$
 \end{tcolorbox}
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+$$u_y(x,y)=\frac{y}{2}g'\left(\frac{x}{2}+\frac{y^2}{4}\right)+\frac{y}{2}f'\left(\frac{x}{2}+\frac{y^2}{4}+2\right)$$
+
+$$u_{yy}(x,y)=\frac{y^2}{4}g''\left(\frac{x}{2}+\frac{y^2}{4}\right)+
+\frac{1}{2}g'\left(\frac{x}{2}+\frac{y^2}{4}\right)+
+\frac{y^2}{4}f''\left(\frac{x}{2}+\frac{y^2}{4}+2\right)+
+\frac{1}{2}f'\left(\frac{x}{2}+\frac{y^2}{4}+2\right)$$
+
+$$u_x(x,y)=\frac{1}{2}g'\left(\frac{x}{2}+\frac{y^2}{4}\right)+\frac{1}{2}f'\left(\frac{x}{2}+\frac{y^2}{4}+2\right)-f'(2)$$
+
+$$u_{xx}(x,y)=\frac{1}{4}g''\left(\frac{x}{2}+\frac{y^2}{4}\right)+\frac{1}{4}f''\left(\frac{x}{2}+\frac{y^2}{4}+2\right)-f''(2)$$
+
+This means $y^3u_{xx}-yu_{yy}+u_y=0$ becomes
+
+\begin{equation*}
+  \begin{aligned}
+  & \left[\frac{y^3}{4} g''\left(\frac{x}{2} +\frac{y^2}{4}\right) + \frac{y^3}{4} f''\left(\frac{x}{2}+\frac{y^2}{4}+2\right)\right]- \\
+  & \left[\frac{y^3}{4}g''\left(\frac{x}{2}+\frac{y^2}{4}\right)+
+\frac{y}{2} g' \left(\frac{x}{2}+\frac{y^2}{4} \right)+
+\frac{y^3}{4} f'' \left( \frac{x}{2}+ \frac{y^2}{4}+2 \right)+
+\frac{y}{2}f' \left( \frac{x}{2}+ \frac{y^2}{4}+2 \right) \right] + \\
+  & \left[ \frac{y}{2} g' \left( \frac{x}{2}+ \frac{y^2}{4} \right)+ \frac{y}{2}f' \left( \frac{x}{2}+ \frac{y^2}{4}+2 \right) \right]=0
+ \end{aligned}
+\end{equation*}
+
+\end{tcolorbox}

### 2024-02-17 PDEs/math245_hw2.Rmd
hw2 pde neumann

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-02-16"
+date: "2024-02-17"
 output:
@@
 ```
-
-## Done, feel OK: 1, 2, 5, 7, 8
-## Done, low confidence: 4
-## Redo: 6
-## Not started: 3
 
@@
       u_{tt}=c^2u_{xx}+f(t,x) & 0<x<\infty\\
-      u(0,x)=\phi{x} & 0<x<\infty\\
-      u_t(0,x) = \psi{x} & 0<x<\infty\\
+      u(0,x)=\phi(x) & 0<x<\infty\\
+      u_t(0,x) = \psi(x) & 0<x<\infty\\
       u_x(t,0) = h(t) & t>0
@@
 \end{equation*}
+
+The method we will use is to first convert to homogenous Neumann boundary conditions, and then extend the equation to the entire line using the reflection method.
+
+First, define $w(t,x)=u(t,x)-H(t)$, where $H_x(t)=h(t)$.
+
+Then our IBVP becomes:
+\begin{equation*}
+    \begin{cases}
+      \square w= \square u - H''(t)=f(t,x)-H''(t) \equiv F(t,x) & x \geq 0, t>0\\
+      w(0,x)=u(0,x)-H(0)=\phi(x) - H(0) \equiv \Phi(x) & x \geq 0\\
+      w_t(0,x) =  u_t(0,x)  - H_t(0)= \psi(x) - H_t(0) \equiv \Psi(x) & x \geq 0\\
+      w_x(0,x) = u_x(t,0) - H_x(0)= u_x(t,0)-h(t) \equiv 0 & t \geq 0
+    \end{cases}       
+\end{equation*}
+
+Now, since we have homogeneous Neumann boundary conditions, we will use even extensions of the initial data to the whole line, namely:
+
+\begin{equation} 
+\Phi_{\text{even}}=
+    \begin{cases} 
+      \Phi(x) & \text{for } x \geq 0\\
+      \Phi(-x) & \text{for } x \leq 0\\
+    \end{cases}  
+    ,\quad
+\Psi_{\text{even}}=
+    \begin{cases}
+      \Psi(x) & \text{for } x \geq 0\\
+      \Psi(-x) & \text{for } x \leq 0\\
+    \end{cases}
+    ,\quad
+F_{\text{even}}=
+    \begin{cases} 
+      F(t,x) & \text{for } x \geq 0\\
+      F(t,-x) & \text{for } x \leq 0\\
+      \end{cases}
+\end{equation}
+
+Then we assume $U(t,x)$ is an extension of $w(t,x)$ to the whole line.
+Since the initial condition and non-homogeneous parts are both even, the solution $U$ is even, so we automatically have $U_x(t,0)=0$.
+
+Consider
+\begin{equation} 
+    \begin{cases} 
+      \square U(t,x)=F_{\text{even}}(t,x) & \text{for } x \in \mathbb{R}, t>0\\
+      U(0,x)=\Phi_{\text{even}}(x) & \text{for } x \in \mathbb{R} \\
+      U_t(0,x)=\Psi_{\text{even}}(x) & \text{for } x \in \mathbb{R}
+    \end{cases} 
+\end{equation}
+
+The solution at point $(t_0, x_0)$ is
+
+\begin{equation*}
+U(t_0,x_0)= \frac{\Phi_{\text{even}}(x_0+ct_0)+\Phi_{\text{even}}(x_0-ct_0)}{2} +
+\frac{1}{2c} \int_{x_0-ct_0}^{x_0+ct_0}\Psi_{\text{even}}(x)dx + \\
+\frac{1}{2c} \int_{0}^{t_0}  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} F_{\text{even}}(t,x) dx\,dt
+\end{equation*}
+
+Assume $t_0>0, x_0>0 \Rightarrow x_0+ct_0 >0$. Now assume $x_0>ct_0$, so the domain of dependence is entirely to the right of the line $x=ct$. This means $w(t_0, x_0)=U(t_0,x_0)$ and our solution is
+\begin{equation*}
+w(t_0,x_0)= \frac{\Phi(x_0+ct_0)+\Phi(x_0-ct_0)}{2} +
+\frac{1}{2c} \int_{x_0-ct_0}^{x_0+ct_0}\Psi(x)dx + \\
+\frac{1}{2c} \int_{0}^{t_0}  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} F(t,x) dx\,dt 
+\end{equation*}
+
+\begin{equation*}
+\frac{\phi(x_0+ct_0)-H(0)+\phi(x_0-ct_0)-H(0)}{2} +
+\frac{1}{2c} \int_{x_0-ct_0}^{x_0+ct_0}\psi(x)-H'(0)dx + \\
+\frac{1}{2c} \int_{0}^{t_0}  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x)-H''(t) dx\,dt
+\end{equation*}
+
+\begin{align*}
+& w(t_0,x_0)=\frac{\phi(x_0+ct_0)+\phi(x_0-ct_0)}{2}- H(0) +
+\frac{1}{2c} \int_{x_0-ct_0}^{x_0+ct_0}\psi(x) dx - \frac{H'(0)}{2c}\left[(x_0+ct_0)-(x_0-ct_0)\right]+ \\
+& \frac{1}{2c} \int_{0}^{t_0}  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) dx\,dt -
+\frac{1}{2c} H''(t) \int_{0}^{t_0}  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} dx\,dt
+\end{align*}
+
+We can rewrite this as
+\begin{align*}
+& w(t_0,x_0)= \color{blue} \frac{\phi(x_0+ct_0)+\phi(x_0-ct_0)}{2} +
+\frac{1}{2c} \int_{x_0-ct_0}^{x_0+ct_0}\psi(x) dx +
+\frac{1}{2c} \int_{0}^{t_0}  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) dx\,dt \\
+\color{black}
+& -H(0)-H'(0)t_0- \frac{1}{2c}\int_{0}^{t_0} H''(t) 2c(t_0-t)dt.
+\end{align*}
+
+Call the section in blue "A".
+
+This means we have 
+\begin{equation}\label{eqn:w1} 
+w(t_0,x_0)=A-H(0)-H'(0)t_0- \int_{0}^{t_0} H''(t)(t_0-t)\,dt 
+\end{equation}
+
+Now we use integration by parts on the final integral in (\ref{eqn:w1}), setting
+$$u=(t_0-t), \quad du=dt, \quad dv=H''(t)dt, \quad v=H'(t)$$,
+
+so that 
+$$\int_{0}^{t_0} H''(t)(t_0-t)\,dt =H'(t)(t_0-t)\vert_{0}^{t_0}+\int_{0}^{t_0} H'(t)dt=-H'(0)t_0+H(t_0)-H(0) $$
+So (\ref{eqn:w1}) becomes
+
+$$w(t_0,x_0)=A-H(0)-H'(0)t_0+H'(0)t_0-H(t_0)+H(0)$$
+$$w(t_0,x_0)=A-H(t_0)$$
+
+Since $u(t,x)=u(t,x)+H(t)$, we get $u(t_0, x_0)=A$, or
+\begin{equation} \label{eqn:part1}
+u(t_0, x_0)=\frac{\phi(x_0+ct_0)+\phi(x_0-ct_0)}{2} +
+\frac{1}{2c} \int_{x_0-ct_0}^{x_0+ct_0}\psi(x) dx +
+\frac{1}{2c} \int_{0}^{t_0}  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) dx\,dt
+\end{equation}
+
+when $x_0>ct_0$.
+
+Now we must consider what happens when $x_0<ct_0$, such that $x_0-ct_0<0$. Then we have
+\begin{align*}
+&w(t_0,x_0)= \frac{\phi(x_0+ct_0)-H(0)+\phi(ct_0-x_0)-H(0)}{2} \;+ \\
+& \frac{1}{2c} \int_{x_0-ct_0}^{0} \psi(-x)-H'(0)\,dx\; + 
+\frac{1}{2c} \int_{0}^{x_0+ct_0} \psi(x) -H'(0)\,dx \;+ \\
+&\frac{1}{2c} \iint\limits_\Delta F_{\text{even}}(t,x) \, dx \, dt
+\end{align*}
+
+\begin{align*}
+&w(t_0,x_0)= \frac{\phi(x_0+ct_0)+\phi(ct_0-x_0)}{2}-H(0) +\\
+ & \color{violet} \frac{1}{2c} \int_{x_0-ct_0}^{0}\psi(-x)-H'(0)\,dx \;+ 
+\frac{1}{2c} \int_{0}^{x_0+ct_0}\psi(x) -H'(0)\,dx \; \color{black} + \\
+&\frac{1}{2c}\iint\limits_\Delta F_{\text{even}}(t,x)\,dx\,dt
+\end{align*}
+
+Violet section:
+\begin{align*}
+& \frac{1}{2c} \int_{x_0-ct_0}^{0}\psi(-x)\,dx \; + 
+\frac{1}{2c} \int_{0}^{x_0+ct_0}\psi(x)\,dx \; 
+-\frac{1}{2c} \left[ H'(0)(0-(x-ct_0))+H'(0)(x+ct_0-0) \right] = \\
+& \frac{1}{2c} \left[\int_{0}^{ct_0-x_0}\psi(x)\,dx +  \int_{0}^{x_0+ct_0}\psi(x)\,dx \right]-H'(0)t_0
+\end{align*}
+
+Now we investigate $\frac{1}{2c}\iint\limits_\Delta F_{\text{even}}(t,x)\,dx\,dt$.
+
+```{r, warning=FALSE, fig.height=3, fig.width=5, echo=FALSE}
+c=1
+t_0=5
+x_0=3
+x_1=seq(-20, 20, by=0.2)
+x_2=seq(-20, 3, by=0.2)
+x_3=seq(3, 20, by=0.2)
+rm(x)
+
+ggplot()+
+  geom_line(aes(x=x_2, y=t_0-(1/c)*(x_0-x_2)))+
+  geom_line(aes(x=x_3, y=t_0+(1/c)*(x_0-x_3)))+
+  lims(y=c(0, 10), x=c(-5, 10))+
+  geom_vline(xintercept=0, linetype="dashed")+
+  geom_hline(yintercept = t_0-x_0/c, linetype="dashed")+
+  geom_text(aes(x=-0.5, y=0.6, label="I"))+
+  geom_text(aes(x=3, y=0.6, label="II"))+
+  geom_text(aes(x=3, y=t_0-x_0/c+1, label="III"))+
+  annotate("text", label="'('* x[0]* ','* t[0] *')'", x=3, y=5.5, size=3, parse = TRUE)+
+  labs(x="x", y="t")+
+  theme_light()
+```
+
+<!-- \newcommand{\Int}{\int\limits} -->
+
+$$ \text{I}= \frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{x_0-c(t_0-t)}^{0} F(t,-x)\;dx\;dt , \quad
+\text{II}= \frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{x_0+c(t_0-t)} F(t,x)\;dx\;dt,$$
+$$\text{III}= \frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} F(t,x) \;dx\;dt $$
+
+Now we will evaluate each integral separately:
+
+$$ \text{I}= \frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{x_0-c(t_0-t)}^{0} f(t,-x)-H''(t)\;dx\;dt$$
+$$ = \frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{x_0-c(t_0-t)}^{0} f(t,-x)\;dx\;dt -\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}}H''(t) \int_{x_0-c(t_0-t)}^{0}dx\;dt$$
+$$ = -\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{-x_0+c(t_0-t)}^{0} f(t,x)\;dx\;dt +\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}}H''(t) (x_0-c(t_0-t))\;dt$$
+$$ = \frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{c(t_0-t)-x_0} f(t,x)\;dx\;dt +\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}}H''(t) (x_0-c(t_0-t))\;dt$$
+
+$$\text{II}= \frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{x_0+c(t_0-t)} f(t,x)\;dx\;dt-
+\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} H''(t)\int_{0}^{x_0+c(t_0-t)}dx\;dt$$
+
+$$= \frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{x_0+c(t_0-t)} f(t,x)\;dx\;dt-
+\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} H''(t)(x_0+c(t_0-t))\;dt$$
+
+$$\text{III}= \frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) \;dx\;dt -\frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} H''(t) \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} dx\;dt $$
+$$ = \frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) \;dx\;dt -\frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} H''(t) \;2c(t_0-t)\;dt $$
+$$ = \frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) \;dx\;dt -\int_{t_0-\frac{x_0}{c}}^{t_0} H''(t) \;(t_0-t)\;dt $$
+Now we use integration by parts on the final integral in the above expression, setting
+$$u=(t_0-t), \quad du=dt, \quad dv=H''(t)dt, \quad v=H'(t)$$
+
+so that 
+$$\int_{t_0-\frac{x_0}{c}}^{t_0}  H''(t)(t_0-t)\,dt =H'(t)(t_0-t)\vert_{t_0-\frac{x_0}{c}}^{t_0}+\int_{t_0-\frac{x_0}{c}}^{t_0}  H'(t)dt$$
+$$=-\frac{x_0}{c} H'\left(t_0-\frac{x_0}{c} \right)+H(t_0)-H\left(t_0-\frac{x_0}{c} \right)$$
+
+So $\text{III}$ is
+$$\frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) \;dx\;dt+\frac{x_0}{c} H'\left(t_0-\frac{x_0}{c} \right)-H(t_0)+H\left(t_0-\frac{x_0}{c} \right) $$
+
+And now we can find the sum of the three regions:
+
+\begin{align*}
+\text{I}+\text{II}+\text{III}&= \frac{1}{2c} \int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{c(t_0-t)-x_0} f(t,x)\;dx\;dt +\frac{1}{2c} \int_{0}^{t_0-\frac{x_0}{c}} H''(t) (x_0-c(t_0-t))\;dt + \\
+& \frac{1}{2c} \int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{x_0+c(t_0-t)} f(t,x)\;dx\;dt-
+\frac{1}{2c} \int_{0}^{t_0-\frac{x_0}{c}} H''(t)(x_0+c(t_0-t))\;dt + \\
+& \frac{1}{2c} \int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) \;dx\;dt+ \\
+& \frac{x_0}{c} H'\left(t_0-\frac{x_0}{c} \right) - H(t_0)+H\left(t_0-\frac{x_0}{c} \right)
+\end{align*}
+
+\begin{align*}
+& =\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{c(t_0-t)-x_0} f(t,x)\;dx\;dt + 
+\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{x_0+c(t_0-t)} f(t,x)\;dx\;dt + 
+\frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) \;dx\;dt+ \\
+& \color{blue} \frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}}H''(t) (x_0-c(t_0-t))\;dt
+-\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}}H''(t) (x_0+c(t_0-t))\;dt\,+\\
+& \frac{x_0}{c} H'\left(t_0-\frac{x_0}{c} \right)-H(t_0)+H\left(t_0-\frac{x_0}{c} \right)
+\end{align*}
+
+\begin{align*}
+\color{blue}
+\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} H''(t) (x_0-c(t_0-t))\;dt
+-\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} H''(t) (x_0+c(t_0-t))\;dt = \\
+\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} H''(t) (-c(t_0-t))\;dt
+-\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} H''(t)c(t_0-t)\;dt = \\
+-\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} H''(t) c(t_0-t)\;dt
+-\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} H''(t)c(t_0-t)\;dt = \\
+-\int_{0}^{t_0-\frac{x_0}{c}} H''(t) (t_0-t)\;dt
+\end{align*}
+
+Now we use integration by parts, so that 
+$$\int_{0}^{t_0-\frac{x_0}{c}}  H''(t)(t_0-t)\,dt =H'(t)(t_0-t)\vert_{0}^{t_0-\frac{x_0}{c}}+\int_{0}^{t_0-\frac{x_0}{c}}  H'(t)dt = \\
+\frac{x_0}{c}H'\left(t_0-\frac{x_0}{c}\right)+H\left(t_0-\frac{x_0}{c}\right)-H(0)$$
+
+\begin{align*}
+\text{I}& +\text{II}+\text{III} = \\
+& \frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{c(t_0-t)-x_0} f(t,x)\;dx\;dt + 
+\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{x_0+c(t_0-t)} f(t,x)\;dx\;dt + 
+\frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) \;dx\;dt \\
+& \color{blue}-\frac{x_0}{c}H'\left(t_0-\frac{x_0}{c}\right)-H\left(t_0-\frac{x_0}{c}\right)+H(0)+\\
+\color{black}
+& \frac{x_0}{c} H'\left(t_0-\frac{x_0}{c} \right)-H(t_0)+H\left(t_0-\frac{x_0}{c} \right)
+\end{align*}
+
+\begin{align*}
+\text{I}& +\text{II}+\text{III} = \\
+& \frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{c(t_0-t)-x_0} f(t,x)\;dx\;dt + 
+\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{x_0+c(t_0-t)} f(t,x)\;dx\;dt + 
+\frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) \;dx\;dt + \\
+& H(0) -H(t_0)
+\end{align*}
+
+\begin{align*}
+& w(t_0,x_0)= \frac{\phi(x_0+ct_0) +\phi(ct_0-x_0)}{2}-H(0) + \\
+& \frac{1}{2c} \left[\int_{0}^{ct_0-x_0}\psi(x)\,dx +  \int_{0}^{x_0+ct_0}\psi(x)\,dx \right]-H'(0)t_0 + \\
+& \frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{c(t_0-t)-x_0} f(t,x)\;dx\;dt + 
+\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{x_0+c(t_0-t)} f(t,x)\;dx\;dt + 
+\frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) \;dx\;dt + \\
+& H(0) -H(t_0)
+\end{align*}
+
+\begin{tcolorbox}[colback=white, title=Solution]
+
+So for $ct_0>x_0$, $u(t_0,x_0) =w(t_0,x_0)+H(t_0)$, which means
+\begin{align*}
+&u(t_0,x_0)= \frac{\phi(x_0+ct_0) +\phi(ct_0-x_0)}{2}+ \\
+& \frac{1}{2c} \left[\int_{0}^{ct_0-x_0}\psi(x)\,dx +  \int_{0}^{x_0+ct_0}\psi(x)\,dx \right]-H'(0)t_0 + \\
+& \frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{c(t_0-t)-x_0} f(t,x)\;dx\;dt + 
+\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{x_0+c(t_0-t)} f(t,x)\;dx\;dt + \\
+& \frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) \;dx\;dt \\
+\end{align*}
+
+While for $ct_0>x_0$, from (\ref{eqn:part1}), we have
+$$u(t_0, x_0)=\frac{\phi(x_0+ct_0)+\phi(x_0-ct_0)}{2} +
+\frac{1}{2c} \int_{x_0-ct_0}^{x_0+ct_0}\psi(x) dx +
+\frac{1}{2c} \int_{0}^{t_0}  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) dx\,dt$$
+
+\end{tcolorbox}
+
+```{r, warning=FALSE, fig.height=3, fig.width=5, echo=FALSE}
+c=1
+t_0=5
+x_0=3
+x_1=seq(-20, 20, by=0.2)
+x_2=seq(-20, x_0, by=0.2)
+x_3=seq(x_0, 20, by=0.2)
+x_4=seq(0, t_0-c*x_0, by=0.2)
+rm(x)
+
+ggplot()+
+  geom_line(aes(x=x_2, y=t_0-(1/c)*(x_0-x_2)))+
+  geom_line(aes(x=x_3, y=t_0+(1/c)*(x_0-x_3)))+
+  geom_polygon(aes(x=c(0, 0, x_0, x_0+c*t_0), y=c(0,t_0-x_0/c, t_0,0)), fill="blue", alpha=0.3)+
+  geom_polygon(aes(x=c(0, 0, t_0-c*x_0), y=c(0,t_0-x_0/c, 0 )), fill="red", alpha=0.5)+
+
+  lims(y=c(0, 10), x=c(-5, 10))+
+  geom_vline(xintercept=0, linetype="dashed")+
+  geom_hline(yintercept = t_0-x_0/c, linetype="dashed")+
+  geom_text(aes(x=-0.5, y=0.6, label="I"))+
+  geom_text(aes(x=3, y=0.6, label="II"))+
+  geom_text(aes(x=3, y=t_0-x_0/c+1, label="III"))+
+  annotate("text", label="'('* x[0]* ','* t[0] *')'", x=3, y=5.5, size=3, parse = TRUE)+
+  labs(x="x", y="t")+
+  theme_light()
+
+```
+
+When $x_0<ct_0$, we're integrating $f$ over the domain of dependence (the shaded blue region, including the red triangle) but then adding $f$ integrated over the red triangle.
+When $x_0<ct_0$, we just integrate $f$ over the blue triangle.
+
+```{r, warning=FALSE, echo=FALSE, fig.height=3, fig.width=5}
+x_0=6
+x_1=seq(-20, 20, by=0.2)
+x_2=seq(-20, x_0, by=0.2)
+x_3=seq(x_0, 20, by=0.2)
+
+ggplot()+
+  geom_line(aes(x=x_2, y=t_0-(1/c)*(x_0-x_2)))+
+  geom_line(aes(x=x_3, y=t_0+(1/c)*(x_0-x_3)))+
+  geom_polygon(aes(x=c(x_0-c*t_0,  x_0+c*t_0, x_0) , y=c(0,0, t_0)), fill="blue", alpha=0.3)+
+ lims(y=c(0, 10), x=c(-5, 15))+
+  geom_vline(xintercept=0, linetype="dashed")+
+  geom_hline(yintercept = t_0-x_0/c, linetype="dashed")+
+  annotate("text", label="'('* x[0]* ','* t[0] *')'", x=x_0, y=5.5, size=3, parse = TRUE)+
+  labs(x="x", y="t")+
+  theme_light()
+```
 
@@
 
-Let's try the factorization method if $c=c(x)$:
+Let's assume that we can factor this wave equation with $c=c(x)$ as we could with constant $c$:
 $$\left( \partial_{t}+c(x)\partial_{x}\right) \left(\partial_{t}-c(x)\partial_{x}\right)u=0$$
-Set $\xi=x+c(x)t$ and $\eta=x-c(x)t$.
+Now, let's distribute:
+$$\left( \partial_{t}+c(x)\partial_{x}\right) \left(u_{t}-c(x)u_{x}\right)=0$$
+$$u_{tt}-c(x)u_{xt} +c(x)u_{xt}-c(x)\left(c'(x)u_x+c(x)u_{xx}\right) =0$$
+$$u_{tt}-c(x)^2u_{xx}- c(x)c'(x)u_x=0$$
+But this is only equivalent to our original equation if $c'(x)=0$ for all $x$, i.e., if $c$ is constant.
 
-By the chain rule,
+We can repeat the process for $c=c(t)$:
+$$\left( \partial_{t}+c(t)\partial_{x}\right) \left(\partial_{t}-c(t)\partial_{x}\right)u=0$$
+$$\left( \partial_{t}+c(t)\partial_{x}\right) \left(u_{t}-c(t)u_{x}\right)=0$$
+$$u_{tt}-(c'(t)u_{x}+c(t)u_{xt}) +c(t)u_{xt}-c(t)^2u_{xx} =0$$
+$$u_{tt}-c(t)^2u_{xx}- c'(t)u_x=0$$
 
-$$\frac{\partial}{\partial x} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial x}+\frac{\partial}{\partial\eta}\frac{\partial\eta}{\partial x}$$
-
-$$=\partial_{\xi}(1+c')+\partial_{\eta}\,(1-c')$$
-
-$$\frac{\partial}{\partial t} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial t}+\frac{\partial }{\partial\eta}\frac{\partial\eta}{\partial t}$$
-$$=c\partial_{\xi}-c\partial_{\eta}$$
-When we plug these back into
-$$\left( \partial_{t}+c\partial_{x}\right) \left(\partial_{t}-c\partial_{x}\right)u=0$$
-we get
-
-$$ \left(c\partial_{\xi}-c\partial_{\eta}+c\partial_{\xi}(1+c')+c\partial_{\eta}\,(1-c')\right)
-\left(c\partial_{\xi}-c\partial_{\eta} -c\partial_{\xi}(1+c')-c\partial_{\eta}\,(1-c')\right)u=0$$
-
-When $c$ is constant, $c'=0$ and this simplifies to $-4c^2 u_{\xi\eta}=0$, allowing us to integrate to find $u(x,t)=F(x+ct)+G(x-ct)$.
-But we cannot do the same simplification when we have the $c'$ terms, meaning we cannot solve the wave equation by the same method when the wave speed is not constant.
-
-Similarly,
-
-Let's try the factorization method if $c=c(t)$:
-$$\left( \partial_{t}+c(t)\partial_{x}\right) \left(\partial_{t}-c(t)\partial_{x}\right)u=0$$
-Set $\xi=x+c(t)t$ and $\eta=x-c(t)t$.
-
-By the chain rule,
-
-$$\frac{\partial}{\partial x} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial x}+\frac{\partial}{\partial\eta}\frac{\partial\eta}{\partial x}=
-\partial_{\xi}+\partial_{\eta}$$
-
-$$\frac{\partial}{\partial t} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial t}+\frac{\partial }{\partial\eta}\frac{\partial\eta}{\partial t}=\partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)
-$$
-
-When we plug these back into
-$$\left( \partial_{t}+c\partial_{x}\right) \left(\partial_{t}-c\partial_{x}\right)u=0$$
-we get
-
-$$\left( \partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)+c\partial_{\xi}+c\partial_{\eta} \right)
-\left(\partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)-c\partial_{\xi}-c\partial_{\eta}\right)u=0$$
-Where again we cannot cancel terms to simplify our equation as we could with constant $c$.
+Which, as before, is only equivalent to our original equation if $c'(t)=0$ for all $t$, i.e., if $c$ is constant.
 

### 2024-02-17 PDEs/math245_old.Rmd
hw2 pde neumann

@@
 where $F$ is the antiderivative of $f$ with respect to $\xi$.
+
+## HW 2 Problem 6
+
+Does the D’Alembert method work for the wave equation $u_{tt} = c(x)^2u_{xx}$? What about $u_{tt} = c(t)^2u_{xx}$? Why?
+
+Let's try the factorization method if $c=c(x)$:
+$$\left( \partial_{t}+c(x)\partial_{x}\right) \left(\partial_{t}-c(x)\partial_{x}\right)u=0$$
+Set $\xi=x+c(x)t$ and $\eta=x-c(x)t$.
+
+By the chain rule,
+
+$$\frac{\partial}{\partial x} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial x}+\frac{\partial}{\partial\eta}\frac{\partial\eta}{\partial x}$$
+
+$$=\partial_{\xi}(1+c')+\partial_{\eta}\,(1-c')$$
+
+$$\frac{\partial}{\partial t} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial t}+\frac{\partial }{\partial\eta}\frac{\partial\eta}{\partial t}$$
+$$=c\partial_{\xi}-c\partial_{\eta}$$
+When we plug these back into
+$$\left( \partial_{t}+c\partial_{x}\right) \left(\partial_{t}-c\partial_{x}\right)u=0$$
+we get
+
+$$ \left(c\partial_{\xi}-c\partial_{\eta}+c\partial_{\xi}(1+c')+c\partial_{\eta}\,(1-c')\right)
+\left(c\partial_{\xi}-c\partial_{\eta} -c\partial_{\xi}(1+c')-c\partial_{\eta}\,(1-c')\right)u=0$$
+
+When $c$ is constant, $c'=0$ and this simplifies to $-4c^2 u_{\xi\eta}=0$, allowing us to integrate to find $u(x,t)=F(x+ct)+G(x-ct)$.
+But we cannot do the same simplification when we have the $c'$ terms, meaning we cannot solve the wave equation by the same method when the wave speed is not constant.
+
+Similarly,
+
+Let's try the factorization method if $c=c(t)$:
+$$\left( \partial_{t}+c(t)\partial_{x}\right) \left(\partial_{t}-c(t)\partial_{x}\right)u=0$$
+Set $\xi=x+c(t)t$ and $\eta=x-c(t)t$.
+
+By the chain rule,
+
+$$\frac{\partial}{\partial x} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial x}+\frac{\partial}{\partial\eta}\frac{\partial\eta}{\partial x}=
+\partial_{\xi}+\partial_{\eta}$$
+
+$$\frac{\partial}{\partial t} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial t}+\frac{\partial }{\partial\eta}\frac{\partial\eta}{\partial t}=\partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)
+$$
+
+When we plug these back into
+$$\left( \partial_{t}+c\partial_{x}\right) \left(\partial_{t}-c\partial_{x}\right)u=0$$
+we get
+
+$$\left( \partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)+c\partial_{\xi}+c\partial_{\eta} \right)
+\left(\partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)-c\partial_{\xi}-c\partial_{\eta}\right)u=0$$
+Where again we cannot cancel terms to simplify our equation as we could with constant $c$.
+
+<!-- paste0("(", "x[0]","','","t[0]", ")")) -->

### 2024-02-19 PDEs/math245_hw2.Rmd
pde hw revisions

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-02-17"
+date: "2024-02-18"
 output:
@@
 ## Question 1
-Determine the region in which the given equation is hyperbolic, parabolic, elliptic, or
-singular.
+Determine the region in which the given equation is hyperbolic, parabolic, elliptic, or singular.
 
-a) $$u_{xx}+y^2u_{yy}+u_x-u+x^2=0 $$
+a) $$u_{xx}-y^2u_{yy}+u_x-u+x^2=0 $$
 
@@
 $a=1, b=0, c=-y$, so we have $b^2-ac=0-(-y)=y$. Thus, the equation will be hyperbolic where $y>0$, parabolic where $y=0$, and elliptic where $y<0$.
+
+Both PDEs are nowhere singular, because the coefficient of $u_{xx}$ is never zero.
 
@@
 
-First, we can factor the equation as follows:
+We are looking for a solution $u(x,y)$ for the above IVP. First, we can factor the given PDE as follows:
 $$\left( \pdv{}{x}+3\pdv{}{y}\right) \left(\pdv{}{x}-\pdv{}{y}\right)u=0$$
@@
 
-$$u(x,y)=\frac{3}{4}\sin{\left(x-\frac{y}{3}\right)}- \frac{(3x-y)^2}{24}+  \frac{\sin{(x+y)}}{4}+ \frac{3(x+y)^2}{8}=$$
+\begin{tcolorbox}[colback=white, title=Solution]
 
-\begin{tcolorbox}[colback=white, title=Solution]
-$$u(x,y)=\frac{3}{4}\sin{\left(x-\frac{y}{3}\right)}+\frac{\sin{(x+y)}}{4}+ xy+  \frac{y^2}{3} $$
+$$u(x,y)=\frac{3}{4}\sin{\left(x-\frac{y}{3}\right)}- \frac{(3x-y)^2}{24}+  \frac{\sin{(x+y)}}{4}+ \frac{3(x+y)^2}{8}$$
 \end{tcolorbox}
@@
 \begin{equation} 
-\Phi_{\text{even}}=
+\Phi_{\text{even}}(x)=
     \begin{cases} 
@@
     ,\quad
-\Psi_{\text{even}}=
+\Psi_{\text{even}}(x)=
     \begin{cases}
@@
     ,\quad
-F_{\text{even}}=
+F_{\text{even}}(t,x)=
     \begin{cases} 
@@
 \begin{equation*}
-\frac{\phi(x_0+ct_0)-H(0)+\phi(x_0-ct_0)-H(0)}{2} +
+=\frac{\phi(x_0+ct_0)-H(0)+\phi(x_0-ct_0)-H(0)}{2} +
 \frac{1}{2c} \int_{x_0-ct_0}^{x_0+ct_0}\psi(x)-H'(0)dx + \\
@@
 & \frac{1}{2c} \int_{0}^{t_0}  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) dx\,dt -
-\frac{1}{2c} H''(t) \int_{0}^{t_0}  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} dx\,dt
+\frac{1}{2c} \int_{0}^{t_0} H''(t)  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} dx\,dt
 \end{align*}
@@
 Now we use integration by parts on the final integral in (\ref{eqn:w1}), setting
-$$u=(t_0-t), \quad du=dt, \quad dv=H''(t)dt, \quad v=H'(t)$$,
+$$u=(t_0-t), \quad du=-dt, \quad dv=H''(t)dt, \quad v=H'(t)$$,
 
@@
 \end{align*}
+
+Where $\Delta$ is the 2D  region in the diagram below.
 
@@
 Now we use integration by parts on the final integral in the above expression, setting
-$$u=(t_0-t), \quad du=dt, \quad dv=H''(t)dt, \quad v=H'(t)$$
+$$u=(t_0-t), \quad du=-dt, \quad dv=H''(t)dt, \quad v=H'(t)$$
 
@@
 
-\begin{tcolorbox}[colback=white, title=Solution]
-
-So for $ct_0>x_0$, $u(t_0,x_0) =w(t_0,x_0)+H(t_0)$, which means
-\begin{align*}
-&u(t_0,x_0)= \frac{\phi(x_0+ct_0) +\phi(ct_0-x_0)}{2}+ \\
-& \frac{1}{2c} \left[\int_{0}^{ct_0-x_0}\psi(x)\,dx +  \int_{0}^{x_0+ct_0}\psi(x)\,dx \right]-H'(0)t_0 + \\
-& \frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{c(t_0-t)-x_0} f(t,x)\;dx\;dt + 
-\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{x_0+c(t_0-t)} f(t,x)\;dx\;dt + \\
-& \frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) \;dx\;dt \\
-\end{align*}
-
-While for $ct_0>x_0$, from (\ref{eqn:part1}), we have
-$$u(t_0, x_0)=\frac{\phi(x_0+ct_0)+\phi(x_0-ct_0)}{2} +
-\frac{1}{2c} \int_{x_0-ct_0}^{x_0+ct_0}\psi(x) dx +
-\frac{1}{2c} \int_{0}^{t_0}  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) dx\,dt$$
-
-\end{tcolorbox}
-
 ```{r, warning=FALSE, fig.height=3, fig.width=5, echo=FALSE}
@@
 ```
+
+\begin{tcolorbox}[colback=white, title=Solution]
+
+So for $ct_0>x_0$, $u(t_0,x_0) =w(t_0,x_0)+H(t_0)$, which means
+\begin{align*}
+&u(t_0,x_0)= \frac{\phi(x_0+ct_0) +\phi(ct_0-x_0)}{2}+ \\
+& \frac{1}{2c} \left[\int_{0}^{ct_0-x_0}\psi(x)\,dx +  \int_{0}^{x_0+ct_0}\psi(x)\,dx \right]-H'(0)t_0 + \\
+& \frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{c(t_0-t)-x_0} f(t,x)\;dx\;dt + 
+\frac{1}{2c}\int_{0}^{t_0-\frac{x_0}{c}} \int_{0}^{x_0+c(t_0-t)} f(t,x)\;dx\;dt + \\
+& \frac{1}{2c}\int_{t_0-\frac{x_0}{c}}^{t_0} \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) \;dx\;dt \\
+\end{align*}
+
+While for $ct_0>x_0$, from (\ref{eqn:part1}), we have
+$$u(t_0, x_0)=\frac{\phi(x_0+ct_0)+\phi(x_0-ct_0)}{2} +
+\frac{1}{2c} \int_{x_0-ct_0}^{x_0+ct_0}\psi(x) dx +
+\frac{1}{2c} \int_{0}^{t_0}  \int_{x_0-c(t_0-t)}^{x_0+c(t_0-t)} f(t,x) dx\,dt$$
+
+\end{tcolorbox}
 
@@
 $$w_{tt}=c^2 w_{rr}.$$
-But this is just the wave equation, and we can use d'Alembert's formula to find the solution:
+But this is just the wave equation, giving us the following IVP:
+
+\begin{equation*} 
+\begin{cases}
+w_{tt}=c^2 w_{rr} \\
+u(0,r)=0 \\
+u_t(0,r) = \begin{cases}
+  1  & \lvert{r}\rvert \leq 1 \\
+  0 &  \lvert{r}\rvert > 1
+  \end{cases}
+\end{cases}
+\end{equation*}
+
+So can use d'Alembert's formula to find the solution:
  $$w(t,r)= \frac{\varphi(r+ct)+\varphi(r-ct)}{2}+\frac{1}{2c} \int_{r-ct}^{r+ct}{\psi(s)ds}$$
@@
 \begin{equation}\label{eqn:prob5c}
-=\int_{0}^{1}{ u_t u_{tt}}\: dx + \int_{0}^{1}{2u_x u_{xt}}\:dx =: I+J.
+=\int_{0}^{1}{ u_t u_{tt}}\: dx + \int_{0}^{1}{u_x u_{xt}}\:dx =: I+J.
 \end{equation}
@@
 \begin{equation} \label{eqn:J}
-J=\int_{0}^{1}{ u_x u_{xt}}\: dx =- \int_{0}^{1}{u_xx u_t}\:dx + 
+J=\int_{0}^{1}{ u_x u_{xt}}\: dx =- \int_{0}^{1}{u_{xx} u_t}\:dx + 
 \left[ u_x u_t \right]_{x=0}^{x=1}.
@@
 
-Now we use our given conditions. When  $2 \leq x \leq 4$, we have $g(x)=h_1(2x)+h_2(0)$, which means $h_1(z_1)=g\left(\frac{z_1}{2}\right)-h_2(0)$.
+Now we use our given conditions. When  $2 \leq x \leq 4$, we have $g(x)=h_1(2x)+h_2(0)$, which means $h_1(x)=g\left(\frac{x}{2}\right)-h_2(0)$.
 
-Similarly, when $0 \leq x \leq 2$, we have $f(x)=h_1(4)+h_2(2x-4)$, which means $h_2(z_2)=f\left(\frac{z_2}{2}+2\right)-h_1(4)$.
+Similarly, when $0 \leq x \leq 2$, we have $f(x)=h_1(4)+h_2(2x-4)$, which means $h_2(x)=f\left(\frac{x}{2}+2\right)-h_1(4)$.
 
 Thus,
-$$h_1(z_1)+h_2(z_2)=g\left(\frac{z_1}{2}\right)-h_2(0)+f\left(\frac{z_2}{2} + 2\right)-h_1(4)$$
+$$h_1(x)+h_2(x)=g\left(\frac{x}{2}\right)-h_2(0)+f\left(\frac{x}{2} + 2\right)-h_1(4)$$
 
@@
 
-so after replacing $z_1$ by $x+\frac{y^2}{2}$ and $z_2$ by $x-\frac{y^2}{2}$, we get
+so after replacing the $x$ in $h_1(x)$ by $x+\frac{y^2}{2}$ and the $x$ in $h_2(x)$ by $x-\frac{y^2}{2}$, we get
 
@@
 
-$$u_x(x,y)=\frac{1}{2}g'\left(\frac{x}{2}+\frac{y^2}{4}\right)+\frac{1}{2}f'\left(\frac{x}{2}+\frac{y^2}{4}+2\right)-f'(2)$$
+$$u_x(x,y)=\frac{1}{2}g'\left(\frac{x}{2}+\frac{y^2}{4}\right)+\frac{1}{2}f'\left(\frac{x}{2}+\frac{y^2}{4}+2\right)$$
 
-$$u_{xx}(x,y)=\frac{1}{4}g''\left(\frac{x}{2}+\frac{y^2}{4}\right)+\frac{1}{4}f''\left(\frac{x}{2}+\frac{y^2}{4}+2\right)-f''(2)$$
+$$u_{xx}(x,y)=\frac{1}{4}g''\left(\frac{x}{2}+\frac{y^2}{4}\right)+\frac{1}{4}f''\left(\frac{x}{2}+\frac{y^2}{4}+2\right)$$
 

### 2024-02-23 PDEs/math245_hw2.Rmd
started pde hw3

@@
 
-First, define $w(t,x)=u(t,x)-H(t)$, where $H_x(t)=h(t)$.
+First, define $w(t,x)=u(t,x)-H(t)$, where $H'(t)=h(t)$.
 
@@
       w(0,x)=u(0,x)-H(0)=\phi(x) - H(0) \equiv \Phi(x) & x \geq 0\\
-      w_t(0,x) =  u_t(0,x)  - H_t(0)= \psi(x) - H_t(0) \equiv \Psi(x) & x \geq 0\\
-      w_x(0,x) = u_x(t,0) - H_x(0)= u_x(t,0)-h(t) \equiv 0 & t \geq 0
+      w_t(0,x) =  u_t(0,x)  - H'(0)= \psi(x) - H'(0) \equiv \Psi(x) & x \geq 0\\
+      w_x(0,x) = u_x(t,0) - H'(0)= u_x(t,0)-h(t) \equiv 0 & t \geq 0
     \end{cases}       
@@
 \begin{cases}
-w_{tt}=c^2 w_{rr} \\
-u(0,r)=0 \\
+w_{tt}=c^2 w_{rr} & t>0, \, r \geq 0\\
+u(0,r)=0 & r \geq 0\\
 u_t(0,r) = \begin{cases}
@@
   0 &  \lvert{r}\rvert > 1
-  \end{cases}
+  \end{cases} & r \geq 0\\
 \end{cases}
@@
 
-So can use d'Alembert's formula to find the solution:
+Although $r$ is nonnegative, since $\varphi=0$ and there are absolute value bars in $\phi$, we can safely use d'Alembert's formula to find the solution:
  $$w(t,r)= \frac{\varphi(r+ct)+\varphi(r-ct)}{2}+\frac{1}{2c} \int_{r-ct}^{r+ct}{\psi(s)ds}$$

### 2024-02-23 PDEs/math245_hw3.Rmd
started pde hw3

---
title: "MATH 245 Homework 3"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-02-23"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1  (Parity of Solution)
Prove the following.

**(a)** If initial conditions of the wave equation on the whole line are even(odd), the solution is
even(odd).
\newline

  Consider the IVP
\begin{equation*}
    \begin{cases}
      u_{tt}=c^2u_{xx}\ & -\infty <x<\infty\\
      u(0,x)=\varphi(x) & -\infty<x<\infty\\
      u_t(0,x) = \psi(x) & -\infty<x<\infty\\
    \end{cases}       
\end{equation*}

where $\varphi(x)$ and $\psi(x)$ are even functions, i.e., $\varphi(-x)=\varphi(x)$ and $\psi(-x)=\psi(x)$. Then using D'Alembert's formula, we know the solution $u$ is

$$u(x,t)= \frac{\varphi(x+ct)+\varphi(x-ct)}{2}+\frac{1}{2c} \int_{x-ct}^{x+ct}{\psi(s)ds}$$
Now plug in $-x$:
$$u(-x,t)= \frac{\varphi(-x+ct)+\varphi(-x-ct)}{2}+\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(s)ds}$$
$$ = \frac{\varphi(x-ct)+\varphi(x+ct)}{2}+\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(-s)ds}$$
After setting $y=-s, dy=-ds$ in the integral of $\psi$:

$$ u(-x,t)= \frac{\varphi(x-ct)+\varphi(x+ct)}{2}+\frac{1}{2c} \int_{x-ct}^{x+ct}{\psi(y)dy}=u(x,t).$$
Therefore, even initial conditions imply that the solution of the wave equation is even.

Similarly, assume $\varphi(x)$ and $\psi(x)$ are odd functions, i.e., $\varphi(-x)=-\varphi(x)$ and $\psi(-x)=-\psi(x)$. Then using D'Alembert's formula, we know $u(-x, t)$ is

$$u(-x,t)= \frac{\varphi(-x+ct)+\varphi(-x-ct)}{2}+\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(s)ds}$$
$$= \frac{-\varphi(x-ct)-\varphi(x+ct)}{2}-\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(-s)ds}$$

After setting $y=-s, dy=-ds$ in the integral of $\psi$:
$$u(-x,t)= \frac{-\varphi(x-ct)-\varphi(x+ct)}{2}-\frac{1}{2c} \int_{x-ct}^{x+ct}{\psi(-y)dy}=-u(x,t).$$
Therefore, odd initial conditions imply that the solution of the wave equation is odd.

**(b)** If the initial condition of the heat equation on the whole line is even(odd), the solution is even(odd).

## Question 2  (Speed of Heat vs Wave)
Consider the traveling wave $u(x, t) = f(x-at)$ where $f$ is a given function of one variable.

(a) If it is a solution of the wave equation, show that the speed must be $a =\pm c$ (unless $f$ is a
linear function).

(b) If it is a solution of the diffusion equation, find $f$ and show that the speed $a$ is arbitrary.

## Question 4  (Maximum Principle)

Consider two solutions $u(t,x)$ and $v(t,x)$ of the diffusion equation in $\Omega_T = \{ 0 \leq x \leq l, 0 \leq t \leq \infty \}$

(a) Let $M(T)=$ the maximum of $u(t, x)$ in the closed rectangle $\Omega_T = \{ 0 \leq x \leq l, 0 \leq t \leq \infty \}$. Does $M(T)$ increase or decrease as a function of $T$? Explain.

(b) Let $m(T)=$ the minimum of $u(t, x)$ in the closed rectangle $\Omega_T = \{ 0 \leq x \leq l, 0 \leq t \leq \infty \}$. Does $m(T)$ increase or decrease as a function of $T$? Explain.

(c) **Comparison Principle**
If $u \leq v$ for $t=0,$ for $x=0, $ for $x=l$, then $u \leq v$ for $0 \leq t \leq \infty$ and $0 \leq x \leq l$.

## Question 5 (Diffusion Equation with Dissipation). 
Solve the following IVP for constant dissipation $b>0$.

\begin{equation}\label{eqn:dissipation}
\begin{cases}
u_t-ku_{xx}+bu=0 & -\infty \leq x \leq \infty, \; t>0 \\
u(0,x)=\psi(x), & -\infty \leq x \leq \infty

\end{cases}
\end{equation}

We will use the change of variables $u(t,x)=e^{-bt}v(t,x)$

## Question 6
Prove that there is no maximum principle for the wave equation.

### 2024-02-23 PDEs/math245_hw3.Rmd
pde problem 6

@@
 ```
+##Done: 1, 6
+##Not started: 2-5
 
 ## Question 1  (Parity of Solution)
-Prove the following.
 
-**(a)** If initial conditions of the wave equation on the whole line are even(odd), the solution is
+**(a)** Prove that if initial conditions of the wave equation on the whole line are even(odd), the solution is
 even(odd).
 \newline
-
+\newline
 Consider the IVP
@@
 $$u(-x,t)= \frac{\varphi(-x+ct)+\varphi(-x-ct)}{2}+\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(s)ds}$$
-$$ = \frac{\varphi(x-ct)+\varphi(x+ct)}{2}+\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(-s)ds}$$
+$$= \frac{\varphi(x-ct)+\varphi(x+ct)}{2}+\frac{1}{2c} \int_{-x-ct}^{-x+ct}{\psi(-s)ds}$$
 After setting $y=-s, dy=-ds$ in the integral of $\psi$:
@@
 
-**(b)** If the initial condition of the heat equation on the whole line is even(odd), the solution is even(odd).
+**(b)** Prove that if the initial condition of the heat equation on the whole line is even(odd), the solution is even(odd).
+
+  Consider the IVP
+\begin{equation*}
+    \begin{cases}
+      u_{t}=k u_{xx} & -\infty <x<\infty\\
+      u(0,x)=f(x)\\
+    \end{cases}       
+\end{equation*}
+
+Then we know
+
+$$u(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy $$
+Which means, when $x=-x$:
+$$u(-x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(-x-y)^2/4kt}f(y)\;dy $$
+Now use the change of variable $y=-s, dy=-ds$:
+$$u(-x,t)= \frac{-1}{\sqrt{4 \pi kt}} \int_{\infty}^{-\infty} e^{-(-x+s)^2/4kt}f(-s)\;ds $$
+
+$$= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-((-1)(x-s))^2/4kt}f(-s)\;ds $$
+But $e^{-((-1)(x-s))^2/4kt}$ is equivalent to $e^{-(x-s)^2/4kt}$, so we have
+
+$$u(-x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(-s)\;ds $$
+
+Now assume $f$ is odd. By definition we have $f(-s)=-f(s)$, so
+
+$$u(-x,t)=\frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(-s)\;ds = 
+\frac{-1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(s)\;ds = -u(x,t) $$
+
+And therefore our solution is odd when our initial $f(x)$ is an odd function. Similarly, if we assume $f(x)$ is even, by definition we have $f(-s)=f(s)$, so
+$$u(-x,t)=\frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(-s)\;ds = 
+\frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(s)\;ds = u(x,t), $$
+meaning our solution is also even.
 
@@
 
+## Question 3 (Heat Equation with Robin boundary)
+Consider the following problem with a Robin boundary condition:
+\begin{equation*}
+    \begin{cases}
+      u_{t}=k u_{xx} & 0 <x<\infty,0 <t<\infty \\
+      u(0,x)=x & t=0,0 <x<\infty \\
+      u_x(0,t)-2u(0,t)=0 & x=0
+    \end{cases}       
+\end{equation*}
+
 ## Question 4  (Maximum Principle)
 
-Consider two solutions $u(t,x)$ and $v(t,x)$ of the diffusion equation in $\Omega_T = \{ 0 \leq x \leq l, 0 \leq t \leq \infty \}$
+Consider two solutions $u(t,x)$ and $v(t,x)$ of the diffusion equation in $\{ 0 \leq x \leq l, 0 \leq t \leq \infty \}$
 
-(a) Let $M(T)=$ the maximum of $u(t, x)$ in the closed rectangle $\Omega_T = \{ 0 \leq x \leq l, 0 \leq t \leq \infty \}$. Does $M(T)$ increase or decrease as a function of $T$? Explain.
+(a) Let $M(T)=$ the maximum of $u(t, x)$ in the closed rectangle $\Omega_T = \{ 0 \leq x \leq l, 0 \leq t \leq T\}$. Does $M(T)$ increase or decrease as a function of $T$?
 
-(b) Let $m(T)=$ the minimum of $u(t, x)$ in the closed rectangle $\Omega_T = \{ 0 \leq x \leq l, 0 \leq t \leq \infty \}$. Does $m(T)$ increase or decrease as a function of $T$? Explain.
+The maximum principle says that if $u(x, t)$ satisfies the diffusion equation in a rectangle $\{ 0 \leq x \leq x, 0 \leq t \leq T \}$, then the maximum value of $u(x, t)$ is assumed either initially $(t = 0)$ or on the lateral sides $(x = 0$ or $x = l)$.
 
-(c) **Comparison Principle**
-If $u \leq v$ for $t=0,$ for $x=0, $ for $x=l$, then $u \leq v$ for $0 \leq t \leq \infty$ and $0 \leq x \leq l$.
+By the definition of maximum, if $u$ assumes a higher value on the lateral sides $x = 0$ or $x = l$ when $T$ is extended from some $T_1$ to a new $T$, $T_2 > T_1$, then the maximum of $u$ will increase to that new value; otherwise, the maximum of $u$ will remain the same.
+In other words, $M$ is non-decreasing (increases, but not strictly) as a function of $T$.
 
-## Question 5 (Diffusion Equation with Dissipation). 
+(b) Let $m(T)=$ the minimum of $u(t, x)$ in the closed rectangle $\Omega_T = \{ 0 \leq x \leq l, 0 \leq t \leq T \}$. Does $m(T)$ increase or decrease as a function of $T$?
+
+By the definition of minimum, if $u$ assumes a lower value on the lateral sides $x = 0$ or $x = l$ when $T$ is extended from some $T_1$ to a new $T$, $T_2 > T_1$, then the minimum of $u$ will decrease to that new value; otherwise, the minimum of $u$ will remain the same.
+In other words, $M$ is non-increasing (decreases, but not strictly) as a function of $T$.
+
+(c) Comparison Principle: Show that if $u \leq v$ for $t=0,x=0,$ and $x=l$, then $u \leq v$ for $0 \leq t \leq \infty$ and $0 \leq x \leq l$.
+
+## Question 5 (Diffusion Equation with Dissipation)
 Solve the following IVP for constant dissipation $b>0$.
@@
 ## Question 6
-Prove that there is no maximum principle for the wave equation.
+
+Suppose that there is a maximum principle for the wave equation and that $u(x,t)$ is a solution to the wave equation in the closed rectangle $\Omega_T = \{ 0 \leq x \leq \pi, \; 0 \leq t \leq \pi/c \}$.
+Then the maximum of $u$ is assumed either initially (at $t=0$) or on the lines $x=0$ or $x=\pi$.
+
+Consider the particular solution $u(x, t) = \sin x \sin ct$. This clearly satisfies the wave equation:
+$$u_x=\cos x \sin ct $$
+$$u_{xx}=-\sin x \sin ct $$
+$$u_t=\sin x (c \cos ct) $$
+$$u_{tt}=-\sin x (c^2 \sin ct)= c^2 u_{xx}$$
+
+According to our presumed maximum principle, the maximum of $u(x,t)$ would have to be one of the following:
+$$u(x,0)= \sin x \sin 0=0 $$
+$$u(0, t)=\sin 0 \sin ct = 0 $$
+$$u(\pi, t)=\sin \pi \sin ct = 0 $$
+
+But $u(x,t)$ has a higher value when $x=\pi/2$ and $t=\pi/2c$:  
+$$u\left(\frac{\pi}{2}, \frac{\pi}{2c}\right)=\sin \left(\frac{\pi}{2}\right) \sin \left(\frac{\pi}{2}\right)=1$$
+
+This contradicts the definition of maximum, so we have found a counterexample and therefore there is no maximum principle for the wave equation.

### 2024-02-23 PDEs/math245_hw3.Rmd
pde hw3 mostly done

@@
 ```
-##Done: 1, 6
-##Not started: 2-5
+##Done: 1, 2, 4, 5, 6.
+
+##Not Done: 3
 
@@
 $$u(-x,t)= \frac{-\varphi(x-ct)-\varphi(x+ct)}{2}-\frac{1}{2c} \int_{x-ct}^{x+ct}{\psi(-y)dy}=-u(x,t).$$
-Therefore, odd initial conditions imply that the solution of the wave equation is odd.
+Therefore, using the uniqueness of the wave equation, odd initial conditions imply that the solution of the wave equation is odd.
 
@@
 Which means, when $x=-x$:
+
 $$u(-x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(-x-y)^2/4kt}f(y)\;dy $$
 Now use the change of variable $y=-s, dy=-ds$:
+
 $$u(-x,t)= \frac{-1}{\sqrt{4 \pi kt}} \int_{\infty}^{-\infty} e^{-(-x+s)^2/4kt}f(-s)\;ds $$
@@
 
-Now assume $f$ is odd. By definition we have $f(-s)=-f(s)$, so
+Now assume $f$ is odd. By definition, we have $f(-s)=-f(s)$, so
 
@@
 \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-s)^2/4kt}f(s)\;ds = u(x,t), $$
-meaning our solution is also even.
+meaning (using the uniqueness of the heat equation), our solution is also even.
 
@@
 
-(a) If it is a solution of the wave equation, show that the speed must be $a =\pm c$ (unless $f$ is a
-linear function).
+(a) If it is a solution of the wave equation, show that the speed must be $a =\pm c$ (unless $f$ is a linear function).
+
+\begin{equation}\label{eqn:tderiv}
+u_t = -af'(x-at), \quad u_{tt}=a^2 f''(x-at)
+\end{equation}
+
+\begin{equation}\label{eqn:xderiv}
+u_x = f'(x-at), \quad u_{xx}=f''(x-at)
+\end{equation}
+
+If $u(x, t) = f(x-at)$ is a solution to the wave equation, then we must have $u_{tt}=c^2 u_{xx}$. By substituting (\ref{eqn:tderiv}) and (\ref{eqn:xderiv}) into the PDE, we get
+
+$$a^2 f''=c^2f'' ,$$
+which implies at $a^2=c^2$, or $a= \pm c$. However, if $f$ is a linear function, then the second derivative $f''=0$ and we cannot conclude that $a^2=c^2$.
 
@@
 
+If $u(x, t) = f(x-at)$ is a solution to the wave equation, then we must have $u_{t}=k u_{xx}$. By substituting (\ref{eqn:tderiv}) and (\ref{eqn:xderiv}) into the PDE, we get
+
+$$-af'=kf'' $$
+$$-af'=kf'' $$
+$$ \frac{-a}{k}= \frac{f''}{f'} $$
+Now if we set $\xi=x-at$, 
+$$\frac{-a}{k}=\frac{f''}{f'}= \frac{d}{d\xi} \ln(f') $$
+So integrating with respect to $\xi$,
+$$\frac{-a}{k}\xi+C= \ln (f')$$
+$$ f'(\xi) = e^{ \frac{-a}{k} \xi+C_1} = C_2 e^{\frac{-a}{k} \xi} $$
+Now if we integrate again with respect to $\xi$,
+$$f(\xi)=\frac{-k}{a}C_2 e^{\frac{-a}{k}\xi}+C_3 $$
+
+Set a new constant $C_4=\frac{-k}{a}C_3$,
+$$f(\xi)=C_4 e^{\frac{-a}{k}\xi}+C_3 $$
+
+If we change back to the original variables,
+$$f(x-at)=C_4 e^ \frac{-a(x-at)}{k}+C_3 $$
+
+Thus, we found the solution
+\begin{equation} \label{eqn:findf}
+u(x, t) = f(x-at)=A e^{\frac{-a}{k}(x-at)}+B,
+\end{equation}
+where $A$ and $B$ are constants.
+
+We can check that this satisfies the heat equation:
+$$u_t=A\frac{a^2}{k} e^ {\frac{-a}{k}(x-at)}$$
+
+$$u_x=A\frac{-a}{k} e^ {\frac{-a}{k}(x-at)}$$
+
+$$u_{xx}=A\frac{a^2}{k^2} e^ {\frac{-a}{k}(x-at)}$$
+$$u_t-ku_{xx}=A\frac{a^2}{k} e^ {\frac{-a}{k}(x-at)}-kA\frac{a^2}{k^2} e^ {\frac{-a}{k}(x-at)}=0$$
+However, (\ref{eqn:findf}) will be a solution regardless of the value of $a$; i.e., $a$ is arbitrary.
+
 ## Question 3 (Heat Equation with Robin boundary)
 Consider the following problem with a Robin boundary condition:
-\begin{equation*}
+\begin{equation*} (\ast)
     \begin{cases}
@@
 \end{equation*}
+
+The purpose of this exercise is to verify the solution formula for $(\ast)$. Let $f(x)=x$ for $x>0$ and let $f(x)=x+1-e^{2x}$ for $x<0$, and let
+
+$$v(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy. $$
+**(a)** What PDE and initial conditions does $v(x,t)$ satisfy for $-\infty < x < \infty$?
+
+$v(x,t)$ satisfies
+
+\begin{equation*}
+    \begin{cases}
+      u_{t}=k u_{xx} & 0 <x<\infty,0 <t<\infty \\
+      u(0,x)=x & t=0,0 <x<\infty \\
+    \end{cases}       
+\end{equation*}
+
+**(b)** Let $w=v_x-2v$. What PDE and initial condition does $w(x,y)$ satisfy for $-\infty < x < \infty$?
+
+**(c)** Show that $f'(x)-2f(x)$ is an odd function for $x \neq 0$.
+\begin{equation*}  f(x)=
+\begin{cases}
+   x & \text{ if } x >0 \\
+   x+1-e^{2x} & \text{ if } x<0\\
+\end{cases}
+\end{equation*}
+
+\begin{equation*}  f'(x)=
+\begin{cases}
+   1 & \text{ if } x >0 \\
+   1-2e^{2x} & \text{ if } x<0\\
+\end{cases}
+\end{equation*}
+
+\begin{equation*}  f'(x)-2f(x)=
+\begin{cases}
+   1-2x & \text{ if } x >0 \\
+   -1-2x & \text{ if } x<0\\
+\end{cases}
+\end{equation*}
+
+Which means that
+\begin{equation*}  f'(-x)-2f(-x)=
+\begin{cases}
+   1-2(-x) & \text{ if } (-x) >0 \\
+   -1-2(-x) & \text{ if } (-x) <0\\
+\end{cases}
+\end{equation*}
+
+But this is equivalent to
+\begin{equation*}
+f'(-x)-2f(-x)=
+\begin{cases}
+   1+2x & \text{ if } x <0 \\
+    -1+2x & \text{ if } x>0\\
+\end{cases}
+\end{equation*}
+
+which is exactly $-(f'(x)-2f(x))$. Therefore, $f'(x)-2f(x)$ is an odd function of $x$.
+
+**(d)** Use Exercise 2.4.11 to show that $w$ is an odd function of $x$.
+
+**(e)** Deduce that $v(x, t)$ satisfies $(\ast)$ for $x>0$. Assuming uniqueness, deduce that the solution of $(\ast)$ is given by
+$$v(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy. $$
 
@@
 
+Define $w(t,x)=u(t,x)-v(t,x)$ and note that $w$ solves the heat equation in $\{ 0 \leq x \leq l, 0 \leq t \leq T \}$:
+$$w_t-kw_{xx}=u_t-v_t-k(u_{xx}-v_{xx})=(u_t-ku_{xx})-(v_t-kv_{xx})=0-0=0 $$
+Since we are assuming $u \leq v$ for $t=0,x=0,$ and $x=l$, then $w \leq 0$ for $t=0,x=0,$ and $x=l$.
+But by the Maximum Principle, $w$ assumes its maximum value on one of those three lines.
+This means we can take $T$ to be arbitrarily large and still have $w \leq 0$, i.e., $w \leq 0$ on $\{ 0 \leq x \leq l, 0 \leq t \leq \infty \}$ and therefore $u \leq v$ for $0 \leq t \leq \infty$ and $0 \leq x \leq l$.
+
 ## Question 5 (Diffusion Equation with Dissipation)
@@
 
-We will use the change of variables $u(t,x)=e^{-bt}v(t,x)$
+We will use the change of variables $u(t,x)=e^{-bt}v(t,x)$, and differentiate to find that:
+
+$$u_t=e^{-bt}v_t-be^{-bt}v$$
+$$u_x=e^{-bt}v_x$$
+$$u_{xx}=e^{-bt}v_{xx}$$
+
+So $u_t-ku_{xx}+bu=0$ becomes
+
+$$e^{-bt}v_t-be^{-bt}v-k \left(e^{-bt}v_{xx}\right) + be^{-bt}v=0 $$
+After canceling the $be^{-bt}v$ terms and dividing by $e^{-bt}$ (which we can do because it cannot equal zero), we are left with
+
+$$v_t-kv_{xx}=0, \quad v(0,x)=u(0,x)=\psi(x) $$
+Hence, the general solution of $v$ is
+
+$$v(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy, $$
+and the general solution of $u$ is therefore
+
+$$u(x,t)= \frac{e^{-bt}}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy. $$
 

### 2024-02-27 PDEs/math245_hw3.Rmd
finished pde hw3

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-02-23"
+date: "2024-02-27"
 output:
@@
 ```
-##Done: 1, 2, 4, 5, 6.
-
-##Not Done: 3
 
@@
 
+\begin{equation} \label{eqn:heatIVP}
+    \begin{cases}
+      v_{t}=k v_{xx} & -\infty <x<\infty,0 <t<\infty \\
+      v(0,x)=f(x) & -\infty  <x<\infty \\
+    \end{cases}       
+\end{equation}
+
+where 
+\begin{equation} f(x)=
+    \begin{cases}
+      x & x>0 \\
+      x+1-e^{2x} & x<0 \\
+    \end{cases}       
+\end{equation}
+
+**(b)** Let $w=v_x-2v$. What PDE and initial condition does $w(x,y)$ satisfy for $-\infty < x < \infty$?
+
+Let $v(x,t)$ be a solution of $v_x=v_{tt}$.
+Now, we know that any partial derivatives of a solution to the heat equation are also solutions, and any linear combination of solutions is also a solution.
+Thus, if $v(x,t)$ is a solution to $v_x=v_{tt}$, so is $w=v_x-2v$.
+The initial conditions will be
+
 \begin{equation*}
-    \begin{cases}
-      u_{t}=k u_{xx} & 0 <x<\infty,0 <t<\infty \\
-      u(0,x)=x & t=0,0 <x<\infty \\
+w(x,0)=v_x(0,x)-2v(0,x)= f'(x)-2f(x) = \begin{cases}
+   1-2x & \text{ if } x >0 \\
+   -1-2x & \text{ if } x<0\\
 \end{cases}
@@
 
-**(b)** Let $w=v_x-2v$. What PDE and initial condition does $w(x,y)$ satisfy for $-\infty < x < \infty$?
+Thus, $w(x,t)$ satisfies the PDE $v_x=v_{tt}$ with initial conditions $$w(x,0)= \begin{cases}
+   1-2x & \text{ if } x >0 \\
+   -1-2x & \text{ if } x<0\\
+\end{cases}$$
 
@@
 
+By Problem 1 in this assignment, we know that given the diffusion equation on the whole line with the usual initial condition $u(x, 0)=\phi(x), \phi(x)$ odd, the solution $u(x, t)$ is also an odd function of $x$.
+
+To summarize, we know that
+\begin{itemize}
+  \item $f'(x)-2f(x)$ is an odd function of $x$
+  \item Given a diffusion equation with odd initial conditions, the solution is also odd
+  \item $w(x,t)$ is a solution to the diffusion equation with initial conditions $f'(x)-2f(x)$
+\end{itemize}
+
+Combining these three points, it is clear that $w(x,t)$ must be an odd function of $x$.
+\newline
+
 **(e)** Deduce that $v(x, t)$ satisfies $(\ast)$ for $x>0$. Assuming uniqueness, deduce that the solution of $(\ast)$ is given by
 $$v(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy. $$
+
+Since $w(x,t)$ is an odd function, it must automatically satisfy the boundary conditions $w(0,t)=0$. We can solve the problem on the half-line by restricting $x>0$. We know that the solution to 
+$$w_t=kw_{xx}, \quad w(x,0)=f'(x)-2f(x), \quad -\infty <x<\infty $$
+
+is given by $$w(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt} \left[f'(y)-2f(y) \right]\;dy. $$
+
+Now we can use our solution for $w$ to solve for $v$:
+$$v_x-2v= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt} \left[f'(y)-2f(y) \right]\;dy $$
+$$v_x-2v= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt} f'(y)\;dy - 
+\frac{2}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt} f(y)\;dy$$
+
+We can see that the first term on the right-hand side is the solution to $(vx)_t = k(vx)_{xx}, -\infty<x<\infty$ with the initial condition $v_x(x, 0) = f'(x)$, i.e., it equals $v_x$.
+Thus, we can subtract $v_x$ from both sides of the above equation and divide by $-2$ to find show that $(\ast)$ is indeed satisfied by
+$$v(x,t)= \frac{1}{\sqrt{4 \pi kt}} \int_{-\infty}^{\infty} e^{-(x-y)^2/4kt}f(y)\;dy, \quad x>0 $$
+
+Because any solution satisfying the heat equation with specified initial and boundary conditions must be the one and only solution, this must be the unique solution to $(\ast)$.
 

### 2024-02-28 bio_dynamics/bio274_hw5.Rmd
started bio hw 5

---
title: "BIOL 274 Homework 5"
author: "Ruby Krasnow"
date: "2024-02-27"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Define each of the following scalars, vectors, and matrices:
$$s=3, r=-2, q= \frac{1}{4} $$

\begin{equation*}
    v = \begin{bmatrix} -8 \\ 4 \\ -16 \end{bmatrix},\quad
    w = \begin{bmatrix} 2 \\ -1 \\ 3 \\ 4 \end{bmatrix},\quad
    A = \begin{bmatrix} 
    -1 & 3 & 2 \\ 
    2 & 0 & -1 \\ 
    3 & 5 & -4 \end{bmatrix},
\end{equation*}

\begin{equation*}
    B = \begin{bmatrix} 
    0 & -2 & 7 & 4 \\ 
    6 & 0 & -5 & 1 \\
    0 & 1 & 4 & -2 \end{bmatrix},\quad \text{and} \quad
    C = \begin{bmatrix} 
    11 & -3 & 1 & 0 \\
    -2 & 5 & 0 & -1 \\
    1 & 3 & 0 & 4 \end{bmatrix}
\end{equation*}

We will also define them in R, for use later in checking our manual computations
```{r}
#scalars
s=3
r=-2
q=(1/4)

#vectors
v=c(-8, 4, -16)
w=c(2,-1,3,4)

#matrices
A = matrix(c(-1, 2, 3, 3, 0, 5, 2, -1, -4), nrow=3, ncol=3)
B = matrix(c(0,6,0,-2,0,1,7,-5,4,4,1,-3), nrow=3)
C = matrix(c(11,-2,1,-3,5,3,1,0,0,0,-1,-4), nrow=3)
```

\newcommand{\negA}{
    \begin{bmatrix} 
    1 & -3 & -2 \\ 
    -2 & 0 & 1 \\ 
    -3 & -5 & 4 \end{bmatrix}}

\newcommand{\B}{
    \begin{bmatrix} 
    0 & -2 & 7 & 4 \\ 
    6 & 0 & -5 & 1 \\ 
    0 & 1 & 4 & -3 \end{bmatrix}}

\newcommand{\C}{
    \begin{bmatrix} 
    11 & -3 & 1 & 0 \\ 
    -2 & 5 & 0 & -1 \\ 
    1 & 3 & 0 & -4 \end{bmatrix}}

Use these to determine each of the following, if possible. If not possible, state why.

## 1. $sw$

\begin{equation*}
    sw = 3\begin{bmatrix} 2 \\ -1 \\ 3 \\ 4 \end{bmatrix} = 
    \begin{bmatrix} 3(2) \\ 3(-1) \\ 3(3) \\ 3(4) \end{bmatrix} =
    \begin{bmatrix} 6 \\ -3 \\ 9 \\ 12 \end{bmatrix}
\end{equation*}

Check our manual computation:
```{r}
as.vector(s*w)
```

## 2. $qv$

\begin{equation*}
    qv = \frac{1}{4} \begin{bmatrix} -8 \\ 4 \\ -16 \end{bmatrix} =
    \begin{bmatrix} -8/4 \\ 4/4 \\ -16/4 \end{bmatrix} =
     \begin{bmatrix} -2 \\ 1 \\ -4 \end{bmatrix}
\end{equation*}

Check our manual computation:
```{r}
as.vector(q*v)
```

##3. $v+w$

This is not possible, because $v \in \mathbb{R}^3$ while $w \in \mathbb{R}^4$, and in order to add two vectors, they must be in the same n-dimensional space $\mathbb{R}^n$, where $n$ is the same for both vectors. In other words, $v$ has three elements while $w$ has four. To add vectors, they must be the same size.

When we try to do this in R, we should get an error:
```{r, error=TRUE}
# the "error=TRUE" argument allows it to show errors without halting R
as.vector(v+w)
```

##4. $(s+r)w$

$$(s+r)w=(3-2)w=w=\begin{bmatrix} 2 \\ -1 \\ 3 \\ 4 \end{bmatrix}$$

Check our manual computation:
```{r}
as.vector((s+r) %*% w)
```

##5. $Aw$

This is not possible, because $A$ is a 3 by 3 matrix while $w$ is a vector in $\mathbb{R}^4$. In order for this to be possible, $w$ would need to have three elements, not four.

When we try to do this in R, we should get an error:
```{r, error=TRUE}
A %*% w
```

##6. $(-A)v$

$$ -A = \negA $$
$$ -Av = \negA \begin{bmatrix} -8 \\ 4 \\ -16 \end{bmatrix} =
\begin{bmatrix}
(1)(-8)+(-3)(4)+(-2)(-16) \\
(-2)(-8)+(0)(4)+(1)(-16) \\
(-3)(-8)+(-5)(4)+(4)(-16)
\end{bmatrix} = \begin{bmatrix} 12 \\ 0 \\ -60 \end{bmatrix} $$

Now we will check our manual computation:
```{r}
negA = -1*A #Define -A

negA %*% v #multiply A by v
```

##7. $Bw$

$$Bw=\B \begin{bmatrix} 2 \\ -1 \\ 3 \\ 4 \end{bmatrix} =
\begin{bmatrix}
(0)(2)+(-2)(-1)+(7)(3)+(4)(4) \\
(6)(2)+(0)(-1)+(-5)(3)+(1)(4) \\
(0)(2)+(1)(-1)+(4)(3)+(-3)(4)
\end{bmatrix} = \begin{bmatrix} 39 \\ 1 \\ -1 \end{bmatrix}$$

Now we will check our manual computation:
```{r}
B %*% w #multiply B by w
```

##8. $B+C$

$$B+C=\B+\C=
\begin{bmatrix}
0+11 & -2-3 & 7+1 & 4+0 \\
6-2 & 0+5 & -5+0 & 1-1 \\
0+1 & 1+3 & 4+0 & -3-4 \end{bmatrix} =
\begin{bmatrix}
11 & -5 & 8 & 4 \\
4 & 5 & -5 & 0 \\
1 & 4 & 4 & -7 \end{bmatrix}$$

Now we will check our manual computation:
```{r}
B + C #add B and C
```

##9. $C-B$

$$C-B=\C-\B=
\begin{bmatrix}
11-0 & -3-2 & 1-7 & 0-4 \\
-2-6 & 5-0 & 0-(-5) & -1-1 \\
1-0 & 3-1 & 0-4 & -4-(-3) \end{bmatrix} =
\begin{bmatrix}
11 & -5 & -6 & -4 \\
-8 & 5 & 5 & -2 \\
1 & 2 & -4 & -1 \end{bmatrix}$$

Now we will check our manual computation:
```{r}
C-B #add B and C
```

##10. $Cv$

This is not possible, because $C$ is a 3 by 4 matrix while $v$ is a vector in $\mathbb{R}^3$. In order for this to be possible, $v$ would need to have four elements, not three.

When we try to do this in R, we should get an error:
```{r, error=TRUE}
C %*% v
```

##11. $(s-r)B$

### 2024-02-28 bio_dynamics/bio274_hw5.Rmd
finished bio hw5

@@
 author: "Ruby Krasnow"
-date: "2024-02-27"
+date: "2024-02-28"
 output:
@@
 library(tidyverse)
+library(expm)
 ```
@@
 
+\newcommand{\A}{
+    \begin{bmatrix} 
+    -1 & 3 & 2 \\ 
+    2 & 0 & -1 \\ 
+    3 & 5 & -4 \end{bmatrix}}
+
 \newcommand{\negA}{
@@
     1 & 3 & 0 & -4 \end{bmatrix}}
+
+\renewcommand{\v}{\begin{bmatrix} -8 \\ 4 \\ -16 \end{bmatrix}}
+
+\newcommand{\w}{\begin{bmatrix} 2 \\ -1 \\ 3 \\ 4 \end{bmatrix}}
 
@@
 \begin{equation*}
-    qv = \frac{1}{4} \begin{bmatrix} -8 \\ 4 \\ -16 \end{bmatrix} =
+    qv = \frac{1}{4} \v =
     \begin{bmatrix} -8/4 \\ 4/4 \\ -16/4 \end{bmatrix} =
@@
 
-##3. $v+w$
+## 3. $v+w$
 
@@
 When we try to do this in R, we should get an error:
-```{r, error=TRUE}
-# the "error=TRUE" argument allows it to show errors without halting R
-as.vector(v+w)
+```{r, error=TRUE, results='hide'}
+v+w
 ```
 
-##4. $(s+r)w$
+## 4. $(s+r)w$
 
-$$(s+r)w=(3-2)w=w=\begin{bmatrix} 2 \\ -1 \\ 3 \\ 4 \end{bmatrix}$$
+$$(s+r)w=(3-2)w=w=\w$$
 
@@
 
-##5. $Aw$
+## 5. $Aw$
 
@@
 
-##6. $(-A)v$
+## 6. $(-A)v$
 
 $$ -A = \negA $$
-$$ -Av = \negA \begin{bmatrix} -8 \\ 4 \\ -16 \end{bmatrix} =
+$$ -Av = \negA \v =
 \begin{bmatrix}
@@
 
-negA %*% v #multiply A by v
+negA %*% v #multiply -A by v
 ```
 
-##7. $Bw$
+## 7. $Bw$
 
-$$Bw=\B \begin{bmatrix} 2 \\ -1 \\ 3 \\ 4 \end{bmatrix} =
+$$Bw=\B \w =
 \begin{bmatrix}
@@
 
-##8. $B+C$
+## 8. $B+C$
 
@@
 6-2 & 0+5 & -5+0 & 1-1 \\
-0+1 & 1+3 & 4+0 & -3-4 \end{bmatrix} =
-\begin{bmatrix}
+0+1 & 1+3 & 4+0 & -3-4 \end{bmatrix}$$
+
+$$=\begin{bmatrix}
 11 & -5 & 8 & 4 \\
@@
 
-##9. $C-B$
+## 9. $C-B$
 
@@
 \begin{bmatrix}
-11-0 & -3-2 & 1-7 & 0-4 \\
+11-0 & -3-(-2) & 1-7 & 0-4 \\
 -2-6 & 5-0 & 0-(-5) & -1-1 \\
-1-0 & 3-1 & 0-4 & -4-(-3) \end{bmatrix} =
-\begin{bmatrix}
-11 & -5 & -6 & -4 \\
+1-0 & 3-1 & 0-4 & -4-(-3) \end{bmatrix}$$
+
+$$=\begin{bmatrix}
+11 & -1 & -6 & -4 \\
 -8 & 5 & 5 & -2 \\
@@
 
-##10. $Cv$
+## 10. $Cv$
 
@@
 
-##11. $(s-r)B$
+## 11. $(s-r)B$
+
+$$s-r=3-(-2)=5$$
+
+$$5B= 5\B = \begin{bmatrix}
+5(0) & 5(-2) & 5(7) & 5(4) \\
+5(6) & 5(0) & 5(-5) & 5(1) \\
+5(0) & 5(1) & 5(4) & 5(-3) \end{bmatrix}=
+\begin{bmatrix}
+0 & -10 & 35 & 20 \\
+30 & 0 & -25 & 5 \\
+0 & 5 & 20 & -15\end{bmatrix}$$
+
+Now we will check our manual computation:
+```{r}
+(s-r)*B
+```
+
+## 12. $(-sB)v$
+
+This is not possible, because $-sB$ will be a matrix of the same size as $B, \;3 \times 4$, while $v$ is a vector in $\mathbb{R}^3$.
+In order for this to be possible, $v$ would need to have four elements, not three.
+
+When we try to do this in R, we should get an error:
+```{r, error=TRUE}
+(-s*B) %*% v
+```
+
+R tells us that the error occurs when we try and multiply $B$ with $v$ because it recognizes that $(-sB)v=-s(Bv)$.
+
+## 13. $(qA)v$
+
+We can rewrite this as follows:
+$$(qA)v=q(Av)= -q(-Av)=\frac{-1}{4} \left(-Av\right). $$
+And since we found $-Av$ in problem 6 above, we can just divide every element in that vector by $-4$ to find our solution:
+
+$$-q \v=
+\begin{bmatrix} -12/4 \\ 0 \\ 60/4 \end{bmatrix}=
+\begin{bmatrix} -3 \\ 0 \\ 15 \end{bmatrix}$$
+
+Now we will check our manual computation:
+```{r}
+(q*A) %*% v
+```
+
+## 14. $(rC)w$
+
+$$rC=-2 \C=
+\begin{bmatrix}
+-22 & 6 & -2 & 0 \\
+4 & -10 & 0 & 2 \\
+-2 & -6 & 0 & 8
+\end{bmatrix}$$
+
+Now multiplying this by $w$,
+
+$$\begin{bmatrix}
+-22 & 6 & -2 & 0 \\
+4 & -10 & 0 & 2 \\
+-2 & -6 & 0 & 8
+\end{bmatrix} \w =
+\begin{bmatrix}
+(-22)(2)+(6)(-1)+(-2)(3)+(0)(4) \\
+(4)(2)+(-10)(-1)+(0)(3)+(2)(4) \\
+(-2)(2)+(-6)(-1)+(0)(3)+(8)(4)
+\end{bmatrix} =
+\begin{bmatrix} -56 \\ 26 \\ 34 \end{bmatrix}$$
+
+Now we will check our manual computation:
+```{r}
+(r*C) %*% w
+```
+
+## 15. $BA$
+
+This is not possible because $B$ is a $3 \times 4$ matrix, while A is a $3 \times 3$ matrix, and in order to multiply two matrices, the second matrix must have the same number of rows as the first matrix has columns.
+
+When we try to do this in R, we should get an error:
+```{r, error=TRUE}
+B %*% A
+```
+
+## 16. $AB$
+
+This is valid matrix multiplication because $A$ is a $3 \times 3$ matrix and $B$ is a $3 \times 4$ matrix.
+
+$$A B = \A\B$$
+
+$$= \begin{bmatrix}
+(-1)(0)+(3)(6)+(2)(0) & (-1)(-2)+(3)(0)+(2)(1) & (-1)(7)+(3)(-5)+(2)(4) & (-1)(4)+(3)(1)+(2)(-3) \\
+(2)(0)+(0)(6)+(-1)(0) & (2)(-2)+(0)(0)+(-1)(1) & (2)(7)+(0)(-5)+(-1)(4) & (2)(4)+(0)(1)+(-1)(-3) \\
+(3)(0)+(5)(6)+(-4)(0) & (3)(-2)+(5)(0)+(-4)(1) & (3)(7)+(5)(-5)+(-4)(4) & (3)(4)+(5)(1)+(-4)(-3) \end{bmatrix}$$
+
+$$=\begin{bmatrix}
+18 & 4 & -14 & -7 \\
+0 & -5 & 10 & 11 \\
+30 & -10 & -20 & 29
+\end{bmatrix} $$
+
+Now we will check our manual computation:
+```{r}
+A %*% B
+```
+
+## 17. $CA$
+This is not possible because $C$ is a $3 \times 4$ matrix, while A is a $3 \times 3$ matrix, and in order to multiply two matrices, the second matrix must have the same number of rows as the first matrix has columns.
+
+When we try to do this in R, we should get an error:
+```{r, error=TRUE}
+C %*% A
+```
+
+## 18. $B^2$
+
+This is not possible because $B$ is not a square matrix.
+Computing $B^2$ is equivalent to $B(B)$, which would be trying to multiply a $3 \times 4$ matrix by another $3 \times 4$ matrix, and in order to multiply two matrices, the second matrix must have the same number of rows as the first matrix has columns.
+
+When we try to do this in R, we should get an error:
+```{r, error=TRUE}
+B %*% B
+```
+
+## 19. $A^2$
+
+$$A^2=AA=\A \A$$
+$$=\begin{bmatrix}
+(-1)(-1)+(3)(2)+(2)(3) & (-1)(3)+(3)(0)+(2)(5) & (-1)(2)+(3)(-1)+(2)(-4) \\
+(2)(-1)+(0)(2)+(-1)(3) & (2)(3)+(0)(0)+(-1)(5) & (2)(2)+(0)(-1)+(-1)(-4) \\
+(3)(-1)+(5)(2)+(-4)(3) & (3)(3)+(5)(0)+(-4)(5) & (3)(2)+(5)(-1)+(-4)(-4)
+\end{bmatrix}$$
+
+$$=\begin{bmatrix}
+13 & 7 & -13 \\
+-5 & 1 & 8 \\
+-5 & -11 & 17
+\end{bmatrix}$$
+
+Now we will check our manual computation:
+```{r, message=FALSE}
+A %*% A
+
+#or using a function from the package 'expm' 
+A %^% 2
+```
+## 20. $A^3$
+
+Since $A^3$ is equivalent to $A^2 A$, we can simply multiply the matrix we found in the previous part by $A$:
+
+$$A^2 A = \begin{bmatrix}
+13 & 7 & -13 \\
+-5 & 1 & 8 \\
+-5 & -11 & 17
+\end{bmatrix} \A$$
+
+$$=\begin{bmatrix}
+(13)(-1)+(7)(2)+(-13)(3) & (13)(3)+(7)(0)+(-13)(5) & (13)(2)+(7)(-1)+(-13)(-4) \\
+(-5)(-1)+(1)(2)+(8)(3) & (-5)(3)+(1)(0)+(8)(5) & (-5)(2)+(1)(-1)+(8)(-4) \\
+(-5)(-1)+(-11)(2)+(17)(3) & (-5)(3)+(-11)(0)+(17)(5) & (-5)(2)+(-11)(-1)+(17)(-4)
+\end{bmatrix}$$
+
+$$=\begin{bmatrix}
+-38 & -26 & 71 \\
+31 & 25 & -43 \\
+34 & 70 & -67
+\end{bmatrix}$$
+
+Now we will check our manual computation:
+```{r}
+A %*% A %*% A
+
+#or using a function from the package 'expm' 
+A %^% 3
+```

### 2024-03-05 PDEs/math245_hw3.Rmd
math hw3, kelp stuff

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-02-27"
+date: "2024-03-01"
 output:
@@
 
-We can see that the first term on the right-hand side is the solution to $(vx)_t = k(vx)_{xx}, -\infty<x<\infty$ with the initial condition $v_x(x, 0) = f'(x)$, i.e., it equals $v_x$.
+We can see that the first term on the right-hand side is the solution to $(v_x)_t = k(v_x)_{xx}, -\infty<x<\infty$ with the initial condition $v_x(x, 0) = f'(x)$, i.e., it equals $v_x$.
 Thus, we can subtract $v_x$ from both sides of the above equation and divide by $-2$ to find show that $(\ast)$ is indeed satisfied by

### 2024-03-05 kelp_sysrev/dataverse_files/1-SF_dada2_pipeline.R

@@ -1,0 +1,137 @@
+##### WORKSPACE SET UP #####
+## load packages
+library(dada2)
+library(readr)
+library(stringr)
+library(tidyverse)
+
+## set path on cluster where your reads are
+path = ""
+
+
+## list the files in the path
+list.files(path)
+
+## set sample names
+fnFs <- sort(list.files(path, pattern="_R1_001.fastq.gz", full.names = TRUE)) #change the pattern to match all your R1 files
+fnRs <- sort(list.files(path, pattern="_R2_001.fastq.gz", full.names = TRUE))
+
+sample.names <- sapply(strsplit(basename(fnFs), "_S"), `[`, 1) 
+sample.names <- sapply(strsplit(basename(fnRs), "_S"), `[`, 1) 
+#change the delimiter in quotes and the number at the end of this command to decide how to split up the file name, 
+#and which element to extract for a unique sample name
+
+head(sample.names)
+
+
+##### IDENFITY PRIMER SEQUENCE AND LENGTH #####
+FWD = nchar(c("GTGYCAGCMGCCGCGGTAA"))
+REV = nchar(c("GGACTACHVGGGTWTCTAAT"))
+
+##### INSPECT READ QUALITY PROFILES #####
+#plotQualityProfile(fnFs[1:9])
+#plotQualityProfile(fnFs[60:63])
+#plotQualityProfile(fnRs[1:9])
+#plotQualityProfile(fnRs[60:63])
+
+
+##### FILTER AND TRIM #####
+# Place filtered files in filtered/ subdirectory
+filtFs <- file.path(path, "filtered", paste0(sample.names, "_R1_filt.fastq.gz"))
+filtRs <- file.path(path, "filtered", paste0(sample.names, "_R2_filt.fastq.gz"))
+names(filtFs) <- sample.names
+names(filtRs) <- sample.names
+
+##### SANITY CHECK AND FILTERING #####
+
+## last sanity check that sample match 
+any(duplicated(c(fnFs, fnRs)))
+any(duplicated(c(filtFs, filtRs)))
+
+length(fnFs)
+length(fnRs)
+
+head(fnFs)
+head(fnRs)
+head(filtFs)
+head(filtRs)
+
+
+## filter out basic things, like N in sequencing and set minimal length for sequence
+filterAndTrim(fnFs, filtFs, fnRs, filtRs, 
+              truncLen=c(FWnumber, RVnumber), ## cut the sequences at this length. FW will usually be longer than RV
[…]

### 2024-03-05 kelp_sysrev/dataverse_files/2-SF_merging_and_taxonomy.R

@@ -1,0 +1,165 @@
+## https://github.com/benjjneb/dada2/issues/95
+## with lots of help from Bia Segovia and Geneveive Lajoie
+
+##### WORKSPACE SET UP #####
+## load packages
+library(tidyverse)
+library(phyloseq)
+library(dada2)
+library(data.table)
+library(zoo)
+
+
+## set working directory 
+setwd("/parfreylab/schenk/spring_freshet_output")
+
+
+##### READ IN DATA #####
+## load processed sequence files 
+st2021 = readRDS("seqtab_nochim_SF_2021.rds")
+st2022.lib1 = readRDS("seqtab_nochim_seqtab_SF_2022_lib1.RDS")
+st2022.lib2 = readRDS("seqtab_nochim_seqtab_SF_2022_lib2.RDS")
+
+
+##### MERGE DATASETS ####
+## merge both datatsets 
+st.all <- mergeSequenceTables(st2021, st2022.lib1, st2022.lib2, tryRC=TRUE)
+## remove bimeras
+seqtab.nochim <- removeBimeraDenovo(st.all, method="consensus", multithread=TRUE,verbose = FALSE)
+dim(seqtab.nochim)
+
+
+# Collapse ASVs that only differ by end base pair
+seqtab.nochim<-collapseNoMismatch(seqtab.nochim, verbose=T) # Check Issue 716 from dada2 github
+# https://github.com/benjjneb/dada2/issues/626
+dim(seqtab.nochim)
+
+## save seqtab file as RDS
+write_rds(seqtab.nochim, "seqtab_nochim_SF_2021_and_2022.rds")
+
+##### ASSIGN TAXONOMY #####
+## SILVA needs to be downloaded in the directory to run
+taxa <- assignTaxonomy(seqtab.nochim, "silva_nr_v138_train_set.fa", 
+                       multithread=TRUE, tryRC=TRUE)
+
+# Save taxonomy table
+write.table(data.frame("row_names"=rownames(taxa),taxa),"SF_taxonomy_SILVAv138_16s.txt", 
+            row.names=FALSE, quote=F, sep="\t")
+
+## add species information to taxonomy data
+taxa_sp <- addSpecies(taxa, "silva_species_assignment_v138.fa")
+
+#NA taxa are hard to separate later if they have no label. apply "Unassigned" label here now.
[…]

### 2024-03-05 kelp_sysrev/dataverse_files/3-SF_filtering_and_coveraged_based_rarefaction.R

@@ -1,0 +1,241 @@
+##### WORKSPACE SETUP #####
+library(phyloseq)
+library(tidyverse)
+library(dplyr)
+library(data.table)
+library(readr)
+library(devtools)
+library(metagMisc)
+library(iNEXT)
+library(vegan)
+
+## tell R where to get the files
+
+## read in files
+sf = readRDS("unfiltered_phyloseq_SF.rds")
+sf
+
+##### REMOVE UNWANTED TAXA FROM TAXONOMY FILE #####
+sf = subset_taxa(sf, Kingdom!="Eukaryota"&
+                   Kingdom!="Unassigned" &
+                   Order!="Chloroplast" &
+                   Family!="Mitochondria" &
+                   Family!="Chloroplast" &
+                   Species !="Chloroplast" &
+                   Genus!="Pseudomonas")
+
+
+##### FILTERING - REMOVE SAMPLES WITH LESS THAN 1000 READS #####
+## Remove samples with less than N reads 
+sample_sums(sf)
+plot(sort(sample_sums(sf)))
+
+## add number of total sequences in a sample (Read depth)
+# e.g. if one sample had 3 ASV with 1, 5, and 7 reads respectively, the Read_depth of that sample would be 13
+# this is the same as colSums for this dataset because of the orientation of the OTU table (rows=asv, columns = samples)
+# rowSums would sum the total number of sequences each ASV has. So if across 10 samples that ASV was detected 50 time, the rowSum would be 50
+sf@sam_data$read_depth_noofftargets = sample_sums(sf) 
+
+## check which samples have less than 1000 reads
+which(sf@sam_data$read_depth_noofftargets < 1000) 
+# this one gets rid of all the controls and nothing else
+# decided to remove samples with less than 1000 reads
+sf.pruned <- prune_samples(sample_sums(sf) >= 1000, sf)
+## write file to know which reds were lost here
+sf.below1000 <- prune_samples(sample_sums(sf) < 1000, sf)
+sf.below1000 = as.matrix(sf.below1000@sam_data)
+write_rds(sf.below1000, "SF_samples_less_than_1000.csv")
+
+##### FILTERING - REMOVE INDIVIDUAL ASVS WITH LESS THAN 100 READS #####
+## extract OTU dataframe from phyloseq object
[…]

### 2024-03-05 kelp_sysrev/dataverse_files/4-SF_alpha_diversity_field.R

@@ -1,0 +1,144 @@
+##### WORKSPACE SETUP #####
+library(phyloseq)
+library(tidyverse)
+library(data.table)
+library(vegan)
+library(car)
+library(plyr)
+library(ggpubr)
+library(FSA)
+library(ggh4x)
+library(MASS)
+library(ggplot2)+theme_set(theme_bw()+
+                             theme(strip.background = element_rect(fill="white"),
+                                   axis.text.y = element_text(colour = "black", size = 12, face = "bold"),
+                                   axis.text.x = element_text(colour = "black", face = "bold", size = 12),
+                                   legend.text = element_text(size = 8, face ="bold", colour ="black"),
+                                   legend.position = "right", axis.title.y = element_text(face = "bold", size = 14),
+                                   axis.title.x = element_text(face = "bold", size = 14, colour = "black"),
+                                   legend.title = element_text(size = 14, colour = "black", face = "bold"),
+                                   legend.key=element_blank(),
+                                   axis.ticks = element_blank()
+                             ))
+
+dephyloseq = function(phylo_obj){
+  
+  ## get the metadata
+  meta = as.data.frame(as.matrix(phylo_obj@sam_data))
+  
+  ## how many metadata columns you have 
+  metacols = ncol(meta)+1
+  
+  ## get out the otu table 
+  otu = as.data.frame(t(as.matrix(phylo_obj@otu_table)))
+  
+  ## merge the metadata and otu table by the rownames (sample ids from the Illumina sequencing data)
+  mo = merge(meta, otu, by=0)
+  
+  ## get out the taxonomy file 
+  tax = as.data.frame(phylo_obj@tax_table)
+  
+  ## get the ASV ID out. This the matches the placeholder ASV ID in the OTU table
+  tax = tax %>% rownames_to_column(var="asv_id")
+  
+  ## pivot longer to be able to match the ASVs in the OTU table to the taxonomy table 
+  mo = mo %>% pivot_longer(cols = -c(1:metacols), names_to = "asv_id", values_to="asv_abundance")
+  
[…]

### 2024-03-05 kelp_sysrev/dataverse_files/4-SF_alpha_diversity_lab.R

@@ -1,0 +1,146 @@
+##### WORKSPACE SETUP #####
+library(phyloseq)
+library(tidyverse)
+library(data.table)
+library(vegan)
+library(car)
+library(plyr)
+library(ggpubr)
+library(FSA)
+library(ggh4x)
+library(ggplot2)+theme_set(theme_bw()+
+                             theme(strip.background = element_rect(fill="white"),
+                                   axis.text.y = element_text(colour = "black", size = 12, face = "bold"),
+                                   axis.text.x = element_text(colour = "black", face = "bold", size = 12),
+                                   legend.text = element_text(size = 8, face ="bold", colour ="black"),
+                                   legend.position = "right", axis.title.y = element_text(face = "bold", size = 14),
+                                   axis.title.x = element_text(face = "bold", size = 14, colour = "black"),
+                                   legend.title = element_text(size = 14, colour = "black", face = "bold"),
+                                   legend.key=element_blank(),
+                                   axis.ticks = element_blank()
+                             ))
+
+dephyloseq = function(phylo_obj){
+  
+  ## get the metadata
+  meta = as.data.frame(as.matrix(phylo_obj@sam_data))
+  
+  ## how many metadata columns you have 
+  metacols = ncol(meta)+1
+  
+  ## get out the otu table 
+  otu = as.data.frame(t(as.matrix(phylo_obj@otu_table)))
+  
+  ## merge the metadata and otu table by the rownames (sample ids from the Illumina sequencing data)
+  mo = merge(meta, otu, by=0)
+  
+  ## get out the taxonomy file 
+  tax = as.data.frame(phylo_obj@tax_table)
+  
+  ## get the ASV ID out. This the matches the placeholder ASV ID in the OTU table
+  tax = tax %>% rownames_to_column(var="asv_id")
+  
+  ## pivot longer to be able to match the ASVs in the OTU table to the taxonomy table 
+  mo = mo %>% pivot_longer(cols = -c(1:metacols), names_to = "asv_id", values_to="asv_abundance")
+  
[…]

### 2024-03-05 kelp_sysrev/dataverse_files/4-SF_betadiversity_field.R

@@ -1,0 +1,463 @@
+##### WORKSPACE SETUP #####
+## load pacakges
+library(phyloseq)
+library(tidyverse)
+library(vegan)
+library(pairwiseAdonis)
+library(ggpubr)
+library(MASS)
+library(plyr)
+library(microbiome)
+library(car)
+library(ggplot2)+theme_set(theme_bw()+
+                             theme(strip.background = element_rect(fill="white"),
+                                   axis.text.y = element_text(colour = "black", size = 12, face = "bold"),
+                                   axis.text.x = element_text(colour = "black", face = "bold", size = 12),
+                                   legend.text = element_text(size = 8, face ="bold", colour ="black"),
+                                   legend.position = "right", axis.title.y = element_text(face = "bold", size = 14),
+                                   axis.title.x = element_text(face = "bold", size = 14, colour = "black"),
+                                   legend.title = element_text(size = 14, colour = "black", face = "bold"),
+                                   legend.key=element_blank(),
+                                   axis.ticks = element_blank()
+                             ))
+
+dephyloseq = function(phylo_obj){
+  
+  ## get the metadata
+  meta = as.data.frame(as.matrix(phylo_obj@sam_data))
+  
+  ## how many metadata columns you have 
+  metacols = ncol(meta)+1
+  
+  ## get out the otu table 
+  otu = as.data.frame(t(as.matrix(phylo_obj@otu_table)))
+  
+  ## merge the metadata and otu table by the rownames (sample ids from the Illumina sequencing data)
+  mo = merge(meta, otu, by=0)
+  
+  ## get out the taxonomy file 
+  tax = as.data.frame(phylo_obj@tax_table)
+  
+  ## get the ASV ID out. This the matches the placeholder ASV ID in the OTU table
+  tax = tax %>% rownames_to_column(var="asv_id")
+  
+  ## pivot longer to be able to match the ASVs in the OTU table to the taxonomy table 
+  mo = mo %>% pivot_longer(cols = -c(1:metacols), names_to = "asv_id", values_to="asv_abundance")
+  
[…]

### 2024-03-05 kelp_sysrev/dataverse_files/4-SF_betadiversity_lab.R

@@ -1,0 +1,414 @@
+##### WORKSPACE SETUP #####
+## load packages
+library(phyloseq)
+library(tidyverse)
+library(vegan)
+library(pairwiseAdonis)
+library(ggpubr)
+library(car)
+library(plyr)
+library(microbiome)
+library(rstatix)
+library(ggplot2)+theme_set(theme_bw()+
+                             theme(strip.background = element_rect(fill="white"),
+                                   axis.text.y = element_text(colour = "black", size = 12, face = "bold"),
+                                   axis.text.x = element_text(colour = "black", face = "bold", size = 12),
+                                   legend.text = element_text(size = 8, face ="bold", colour ="black"),
+                                   legend.position = "right", axis.title.y = element_text(face = "bold", size = 14),
+                                   axis.title.x = element_text(face = "bold", size = 14, colour = "black"),
+                                   legend.title = element_text(size = 14, colour = "black", face = "bold"),
+                                   legend.key=element_blank(),
+                                   axis.ticks = element_blank()
+                             ))
+
+dephyloseq = function(phylo_obj){
+  
+  ## get the metadata
+  meta = as.data.frame(as.matrix(phylo_obj@sam_data))
+  
+  ## how many metadata columns you have 
+  metacols = ncol(meta)+1
+  
+  ## get out the otu table 
+  otu = as.data.frame(t(as.matrix(phylo_obj@otu_table)))
+  
+  ## merge the metadata and otu table by the rownames (sample ids from the Illumina sequencing data)
+  mo = merge(meta, otu, by=0)
+  
+  ## get out the taxonomy file 
+  tax = as.data.frame(phylo_obj@tax_table)
+  
+  ## get the ASV ID out. This the matches the placeholder ASV ID in the OTU table
+  tax = tax %>% rownames_to_column(var="asv_id")
+  
+  ## pivot longer to be able to match the ASVs in the OTU table to the taxonomy table 
+  mo = mo %>% pivot_longer(cols = -c(1:metacols), names_to = "asv_id", values_to="asv_abundance")
+  
[…]

### 2024-03-05 kelp_sysrev/dataverse_files/4-SF_field_core_over20.R

@@ -1,0 +1,239 @@
+##### SET UP ####
+## load packages
+library(indicspecies)
+library(tidyverse)
+library(data.table)
+library(phyloseq)
+library(plyr)
+library(ggpubr)
+library(ggplot2)+theme_set(theme_bw()+
+                             theme(strip.background = element_rect(fill="white"),
+                                   axis.text.y = element_text(colour = "black", size = 12, face = "bold"),
+                                   axis.text.x = element_text(colour = "black", face = "bold", size = 12),
+                                   legend.text = element_text(size = 8, face ="bold", colour ="black"),
+                                   legend.position = "right", axis.title.y = element_text(face = "bold", size = 14),
+                                   axis.title.x = element_text(face = "bold", size = 14, colour = "black"),
+                                   legend.title = element_text(size = 14, colour = "black", face = "bold"),
+                                   legend.key=element_blank(),
+                                  # axis.ticks = element_blank()
+                             ))
+
+dephyloseq = function(phylo_obj){
+  
+  ## get the metadata
+  meta = as.data.frame(as.matrix(phylo_obj@sam_data))
+  
+  ## how many metadata columns you have 
+  metacols = ncol(meta)+1
+  
+  ## get out the otu table 
+  otu = as.data.frame(t(as.matrix(phylo_obj@otu_table)))
+  
+  ## merge the metadata and otu table by the rownames (sample ids from the Illumina sequencing data)
+  mo = merge(meta, otu, by=0)
+  
+  ## get out the taxonomy file 
+  tax = as.data.frame(phylo_obj@tax_table)
+  
+  ## get the ASV ID out. This the matches the placeholder ASV ID in the OTU table
+  tax = tax %>% rownames_to_column(var="asv_id")
+  
+  ## pivot longer to be able to match the ASVs in the OTU table to the taxonomy table 
+  mo = mo %>% pivot_longer(cols = -c(1:metacols), names_to = "asv_id", values_to="asv_abundance")
+  
+  ## Join the metadata and otu table with the taoxnomy table 
+  mot = full_join(mo, tax)
+  
[…]

### 2024-03-05 kelp_sysrev/dataverse_files/4-SF_lab_condition_day1.R

@@ -1,0 +1,216 @@
+##### WORKSPACE SETUP  ####
+## load packages
+library(tidyverse)
+library(plyr)
+library(stats)
+library(phyloseq)
+library(ggpubr)
+library(ggh4x)
+library(car)
+library(rstatix)
+library(ggplot2); theme_set(theme_bw()+
+                              theme(panel.grid = element_blank(),
+                                    strip.background = element_rect(fill="white"),
+                                    axis.text = element_text(size = 12, colour = "black", face="bold"),
+                                    axis.title = element_text(size=15),
+                                    strip.text = element_text(color="black", size=12),
+                                    legend.text=element_text(size=12),
+                                    axis.line = element_line(colour = "black")))
+
+dephyloseq = function(phylo_obj){
+  
+  ## get the metadata
+  meta = as.data.frame(as.matrix(phylo_obj@sam_data))
+  
+  ## how many metadata columns you have 
+  metacols = ncol(meta)+1
+  
+  ## get out the otu table 
+  otu = as.data.frame(t(as.matrix(phylo_obj@otu_table)))
+  
+  ## merge the metadata and otu table by the rownames (sample ids from the Illumina sequencing data)
+  mo = merge(meta, otu, by=0)
+  
+  ## get out the taxonomy file 
+  tax = as.data.frame(phylo_obj@tax_table)
+  
+  ## get the ASV ID out. This the matches the placeholder ASV ID in the OTU table
+  tax = tax %>% rownames_to_column(var="asv_id")
+  
+  ## pivot longer to be able to match the ASVs in the OTU table to the taxonomy table 
+  mo = mo %>% pivot_longer(cols = -c(1:metacols), names_to = "asv_id", values_to="asv_abundance")
+  
+  ## Join the metadata and otu table with the taoxnomy table 
+  mot = full_join(mo, tax)
+  
+  ## Specify the output for the dephyloseq funciton 
+  output = mot
+}
+
+## set color vector
+salinity.colors = c("cyan","dodgerblue2","navy")
+
+## read in data
+wetweight = read.csv("lab_condition_metadata.csv")
+
[…]

### 2024-03-05 kelp_sysrev/dataverse_files/4-SF_lab_condition_day4.R

@@ -1,0 +1,279 @@
+##### WORKSPACE SETUP  ####
+## load packages
+library(tidyverse)
+library(plyr)
+library(stats)
+library(phyloseq)
+library(ggpubr)
+library(ggh4x)
+library(car)
+library(rstatix)
+library(ggplot2); theme_set(theme_bw()+
+                              theme(panel.grid = element_blank(),
+                                    strip.background = element_rect(fill="white"),
+                                    axis.text = element_text(size = 12, colour = "black", face="bold"),
+                                    axis.title = element_text(size=15),
+                                    strip.text = element_text(color="black", size=12),
+                                    legend.text=element_text(size=12),
+                                    axis.line = element_line(colour = "black")))
+
+dephyloseq = function(phylo_obj){
+  
+  ## get the metadata
+  meta = as.data.frame(as.matrix(phylo_obj@sam_data))
+  
+  ## how many metadata columns you have 
+  metacols = ncol(meta)+1
+  
+  ## get out the otu table 
+  otu = as.data.frame(t(as.matrix(phylo_obj@otu_table)))
+  
+  ## merge the metadata and otu table by the rownames (sample ids from the Illumina sequencing data)
+  mo = merge(meta, otu, by=0)
+  
+  ## get out the taxonomy file 
+  tax = as.data.frame(phylo_obj@tax_table)
+  
+  ## get the ASV ID out. This the matches the placeholder ASV ID in the OTU table
+  tax = tax %>% rownames_to_column(var="asv_id")
+  
+  ## pivot longer to be able to match the ASVs in the OTU table to the taxonomy table 
+  mo = mo %>% pivot_longer(cols = -c(1:metacols), names_to = "asv_id", values_to="asv_abundance")
+  
+  ## Join the metadata and otu table with the taoxnomy table 
+  mot = full_join(mo, tax)
+  
+  ## Specify the output for the dephyloseq funciton 
+  output = mot
+}
+
+## set color vector
+salinity.colors = c("cyan","dodgerblue2","navy")
+
+## read in data
[…]

### 2024-03-05 kelp_sysrev/dataverse_files/4-SF_lab_core_over20.R

@@ -1,0 +1,182 @@
+##### SET UP ####
+## load packages
+library(indicspecies)
+library(tidyverse)
+library(data.table)
+library(phyloseq)
+library(plyr)
+library(ggpubr)
+library(car)
+library(ggplot2)+theme_set(theme_bw()+
+                             theme(strip.background = element_rect(fill="white"),
+                                   axis.text.y = element_text(colour = "black", size = 12, face = "bold"),
+                                   axis.text.x = element_text(colour = "black", face = "bold", size = 12),
+                                   legend.text = element_text(size = 8, face ="bold", colour ="black"),
+                                   legend.position = "right", axis.title.y = element_text(face = "bold", size = 14),
+                                   axis.title.x = element_text(face = "bold", size = 14, colour = "black"),
+                                   legend.title = element_text(size = 14, colour = "black", face = "bold"),
+                                   legend.key=element_blank(),
+                                   # axis.ticks = element_blank()
+                             ))
+
+dephyloseq = function(phylo_obj){
+  
+  ## get the metadata
+  meta = as.data.frame(as.matrix(phylo_obj@sam_data))
+  
+  ## how many metadata columns you have 
+  metacols = ncol(meta)+1
+  
+  ## get out the otu table 
+  otu = as.data.frame(t(as.matrix(phylo_obj@otu_table)))
+  
+  ## merge the metadata and otu table by the rownames (sample ids from the Illumina sequencing data)
+  mo = merge(meta, otu, by=0)
+  
+  ## get out the taxonomy file 
+  tax = as.data.frame(phylo_obj@tax_table)
+  
+  ## get the ASV ID out. This the matches the placeholder ASV ID in the OTU table
+  tax = tax %>% rownames_to_column(var="asv_id")
+  
+  ## pivot longer to be able to match the ASVs in the OTU table to the taxonomy table 
+  mo = mo %>% pivot_longer(cols = -c(1:metacols), names_to = "asv_id", values_to="asv_abundance")
+  
+  ## Join the metadata and otu table with the taoxnomy table 
[…]

### 2024-03-05 kelp_sysrev/dataverse_files/4-SF_size_distribution.R

@@ -1,0 +1,58 @@
+##### WORKSPACE SETUP #####
+## load packages
+library(plyr)
+library(tidyverse)
+library(ggpubr)
+library(car)
+library(ggplot2)+theme_set(theme_bw()+
+                             theme(strip.background = element_rect(fill="white"),
+                                   axis.text.y = element_text(colour = "black", size = 12, face = "bold"),
+                                   axis.text.x = element_text(colour = "black", face = "bold", size = 12),
+                                   legend.text = element_text(size = 8, face ="bold", colour ="black"),
+                                   legend.position = "right", axis.title.y = element_text(face = "bold", size = 14),
+                                   axis.title.x = element_text(face = "bold", size = 14, colour = "black"),
+                                   legend.title = element_text(size = 14, colour = "black", face = "bold"),
+                                   legend.key=element_blank(),
+                                   #axis.ticks = element_blank()
+                                   ))
+
+
+## set the position dodge for ggplot
+pd = position_dodge(0.2) # move them .05 to the left and right
+
+## make color vector for site
+site.colors = c("seagreen1", "#39B54A", "darkgreen", "plum", "#A30262")
+
+## read in data
+metadata = read.csv("field_condition_metadata.csv")
+
+## add julian date 
+metadata$jul.date = paste0(as.numeric(metadata$year), "-", 
+                                  as.numeric(metadata$month), "-", 
+                                  as.numeric(metadata$day))
+metadata$jul.date = as.Date(metadata$jul.date)
+metadata$jul.date = as.POSIXlt(metadata$jul.date)$yday
+
+## order the sites
+metadata$site_id = factor(metadata$site_id, levels=c("lhp", "tb","scp","gw","cp"))
+
+##### PLOT WIDTH BY SITE ######
+ggplot(metadata, aes(y=width_cm, x = site_id, fill=site_id, color=site_id))+
+  geom_boxplot(linewidth=0.5, outlier.size = 5, alpha = 0.3)+
+  scale_colour_manual(values = site.colors)+
+  scale_fill_manual(values=site.colors)+
+  labs(y="Thallus width (cm)", x="Sampling Site")+
+  annotate("text", x=1, y=30, label="a", size=4, color="black")+
+  annotate("text", x=2, y=30, label="b", size=4, color="black")+
+  annotate("text", x=3, y=30, label="b", size=4, color="black")+
[…]

### 2024-03-05 kelp_sysrev/dataverse_files/4-SF_tides_and_temp.R

@@ -1,0 +1,226 @@
+##### SET UP #####
+
+## load packages
+library(plyr)
+library(tidyverse)
+library(lubridate)
+library(sandwich)
+library(car)
+library(plyr)
+library(MASS)
+library(nlme)
+library(ggplot2)+theme_set(theme_bw()+
+                             theme(panel.grid = element_blank(),
+                                   strip.background = element_rect(fill="white"),
+                                   axis.text.y = element_text(colour = "black", size = 12, face = "bold"),
+                                   axis.text.x = element_text(colour = "black", size = 8),
+                                   legend.text = element_text(size = 8, face ="bold", colour ="black"),
+                                   legend.position = "right", axis.title.y = element_text(face = "bold", size = 14),
+                                   axis.title.x = element_text(face = "bold", size = 14, colour = "black"),
+                                   legend.title = element_text(size = 14, colour = "black", face = "bold"),
+                                   legend.key=element_blank()))
+  
+## read in data
+tempdata =read.csv("2022_temp_logger_data.csv")
+logger_elevation = read.csv("logger_elevation.csv")
+tide_obs = read.csv("gov_canada_tide_observations.csv")
+
+## make color vector for site
+site.colors = c("#39B54A", "darkgreen", "plum", "#A30262")
+
+
+##### FORMAT THE TEMPERATURE LOGGER DATA #####
+## remove bad data (before deployment temps and first 12h outside)
+tempdata = subset(tempdata, tempdata$keep=="yes")
+
+## set variable format
+tempdata$datetime.utc = as.POSIXct(tempdata$datetime.utc, format="%Y-%m-%d %H:%M")
+tempdata$datetime.pdt = as.POSIXct(tempdata$datetime.pdt, format="%Y-%m-%d %H:%M")
+tempdata$temp = as.numeric(tempdata$temp)
+
+## split columns
+tempdata = separate(data = tempdata, 
+                         col = site, 
+                         into = c("site","position"), 
+                         sep = "-")
+
+## rename sites
[…]

### 2024-03-05 kelp_sysrev/dataverse_files/4-SF_ysi_data_analysis.R

@@ -1,0 +1,138 @@
+##### WORKSPACE SETUP #####
+## load packages
+library(plyr)
+library(tidyverse)
+library(ggpubr)
+library(ggplot2)+theme_set(theme_bw()+
+                             theme(strip.background = element_rect(fill="white"),
+                                   axis.text.y = element_text(colour = "black", size = 12, face = "bold"),
+                                   axis.text.x = element_text(colour = "black", face = "bold", size = 12),
+                                   legend.text = element_text(size = 8, face ="bold", colour ="black"),
+                                   legend.position = "right", axis.title.y = element_text(face = "bold", size = 14),
+                                   axis.title.x = element_text(face = "bold", size = 14, colour = "black"),
+                                   legend.title = element_text(size = 14, colour = "black", face = "bold"),
+                                   legend.key=element_blank(),
+                                   #axis.ticks = element_blank()
+                                   ))
+
+
+## set the position dodge for ggplot
+pd = position_dodge(0.2) # move them .05 to the left and right
+
+## make color vector for site
+site.colors = c("seagreen1", "#39B54A", "darkgreen", "plum", "#A30262")
+
+## load in data
+abiotic = read.csv("ysi_data_both_years.csv")
+
+lmnotp = subset(abiotic, abiotic$water_type !="tidepool")
+
+
+## group the data by site, sampling date, and water type
+abiotic.grouped = ddply(abiotic, c("site_id", "water_type", "day","month","year"), 
+                       summarise,
+                       N.conduct = length(salinity), ## sample size
+                       mean.conduct = mean(salinity), ## mean
+                       sd.conduct = sd(salinity),
+                      N.temp = length(water_temp_C), ## sample size
+                      mean.temp = mean(water_temp_C), ## mean
+                      sd.temp = sd(water_temp_C))%>% ## standard deviation) %>%
+  mutate(se.conduct = sd.conduct/sqrt(N.conduct), ## standard error
[…]

### 2024-03-05 kelp_sysrev/dataverse_files/4-taxaplots_lab_and_field.R

@@ -1,0 +1,287 @@
+###### SET UP #######
+## load packages
+library(phyloseq)
+library(tidyverse)
+library(plyr)
+library(qualpalr)
+library(ggpubr)
+library(ggh4x)
+library(ggplot2); theme_set(theme_bw()+
+  theme(panel.grid = element_blank(),
+        strip.background = element_rect(fill="white"),
+        axis.text.y = element_text(size = 10, colour = "black"),
+        axis.title = element_text(size=10, face="bold"),
+        strip.text = element_text(color="black", size=10),
+        legend.text=element_text(size=10),
+        axis.line = element_line(colour = "black")))
+
+## load functions
+dephyloseq = function(phylo_obj){
+  
+  ## get the metadata
+  meta = as.data.frame(as.matrix(phylo_obj@sam_data))
+  
+  ## how many metadata columns you have 
+  metacols = ncol(meta)+1
+  
+  ## get out the otu table 
+  ## if your metadta is empty after running this, you need to use 
+  otu = as.data.frame(t(as.matrix(phylo_obj@otu_table)))
+  #otu = as.data.frame(as.matrix(phylo_obj@otu_table))
+  
+  ## merge the metadata and otu table by the rownames (sample ids from the Illumina sequencing data)
+  mo = merge(meta, otu, by=0)
+  
+  ## get out the taxonomy file 
+  tax = as.data.frame(phylo_obj@tax_table)
+  
+  ## get the ASV ID out. This the matches the placeholder ASV ID in the OTU table
+  tax = tax %>% rownames_to_column(var="ASVid")
+  
+  ## pivot longer to be able to match the ASVs in the OTU table to the taxonomy table 
+  mo = mo %>% pivot_longer(cols = -c(1:metacols), names_to = "ASVid", values_to="asv_abundance")
+  
+  ## Join the metadata and otu table with the taoxnomy table 
+  mot = full_join(mo, tax)
+  
+  ## Specify the output for the dephyloseq funciton 
+  output = mot
+}
+
+## tell R whereto get data
+## tell R where to get data
+setwd("")
+path=""
+
+## read in field data
+sf = readRDS("SF_2021_and_2022_filtered001percent_notrarefied.rds")
+
+field = subset_samples(sf, lab_or_field_sampled=="field")
+
+## make field groups
[…]

### 2024-03-05 kelp_sysrev/kelp_bacteria.R

@@ -2,6 +2,7 @@
 
 library(PRISMAstatement)
 library(tidyverse)
+library(clipr)
 
 prisma(found = 1448,
        found_other = 106,
@@ -17,3 +18,8 @@ prisma(found = 1448,
 #files for full-text screening
 inital_screen <- read_csv("./kelp_sysrev/initial_screen.csv")
 
+blank_table <- read_clip_tbl()
+
+blank_table %>% bind_rows(blank_table, blank_table) %>% arrange(title) %>% write_clip()
+
+blank_table%>% write_clip()

### 2024-03-15 bio_dynamics/bio274_hw6.Rmd
bio hw 6

---
title: "BIOL 274 Homework 6"
author: "Ruby Krasnow"
date: "2024-03-15"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1  
Write each of the following systems in matrix/vector notation:

(a)
\begin{align*}
      x' &= 4x-6y \\
      y' &= 0.7x + y
\end{align*}

\begin{equation*}
\begin{bmatrix} x' \\ y' \end{bmatrix} = \begin{bmatrix} 
    4 & -6 \\
    0.7 & 1 \end{bmatrix} \begin{bmatrix} x \\ y \end{bmatrix}, \quad \text{ or }
  \end{equation*}

\begin{equation*}
\dot{X} = \begin{bmatrix} 
    4 & -6 \\
    0.7 & 1 \end{bmatrix} X
  \end{equation*}

(b)
\begin{align*}
    \frac{dx_1}{dt}-16x_2 &=-2x_1 \\
    \frac{dx_2}{dt}+4x_1 & =0
\end{align*}

\begin{equation*}
\begin{bmatrix} x_1' \\ x_2' \end{bmatrix} = \begin{bmatrix} 
    -2 & 16 \\
    -4 & 0 \end{bmatrix} \begin{bmatrix} x_1 \\ x_2 \end{bmatrix}, \quad \text{ or }
  \end{equation*}

\begin{equation*}
\dot{X} = \begin{bmatrix} 
    -2 & 16 \\
    -4 & 0 \end{bmatrix} X
  \end{equation*}

(c)
\begin{align*}
    \dot{v}=v-u+\sin{t} \\
    \dot{u}=2v+u-\cos{t}
\end{align*}

\begin{equation*}
\begin{bmatrix} \dot{v} \\ \dot{u} \end{bmatrix} = \begin{bmatrix} 
    1 & -1 \\
    2 & 1 \end{bmatrix} \begin{bmatrix}v \\ u \end{bmatrix} + \begin{bmatrix} \sin{t} \\ -\cos{t} \end{bmatrix}, \quad \text{ or }
  \end{equation*}

\begin{equation*}
\dot{V} = \begin{bmatrix} 
    1 & -1 \\
    2 & 1 \end{bmatrix} V + \begin{bmatrix} \sin{t} \\ -\cos{t} \end{bmatrix}
  \end{equation*}

## Question 2

Find the general solution for each of the following systems:  
(a)
\begin{align*}
      \dot{x} &= 9x-8y \\
      \dot{y} &= 24x-19y
\end{align*}

\begin{equation*}
A = \begin{bmatrix} 
    9 & -8 \\
    24 & -19 \end{bmatrix}
  \end{equation*}

\begin{equation*}
\det{\begin{bmatrix} 
    9-\lambda & -8 \\
    24 & -19\lambda \end{bmatrix}}=0 \\
\end{equation*}

$$ \left(9-\lambda\right) \left(-19-\lambda\right)+192=0 $$
\begin{equation*}
-171-9\lambda+19\lambda+ \lambda^2+192 = \lambda^2+10\lambda+21 =(\lambda+7)(\lambda+3)
\end{equation*}

So we have found two distinct real eigenvalues, $\lambda_1=-7, \; \lambda_2=-3$. Now we need to solve $AX=\lambda X$, or $(A-\lambda I)X=0$ with these two values of $\lambda$. Let's start with the first eigenvalue, -7.

\begin{equation*}
A+7I=\begin{bmatrix} 
    9+7 & -8 \\
    24 & -19+7 \end{bmatrix}= \begin{bmatrix} 
    16 & -8 \\
    24 & -12 \end{bmatrix} \end{equation*}

Now we use Gauss-Jordan elimination:
 \begin{equation*}   
   \begin{bmatrix} 
  16 & -8 \\  
  24 & -12   \end{bmatrix} \longrightarrow{}
 \begin{bmatrix} 
  -2 & 1 \\  
  -2 & 1
 \end{bmatrix}
\longrightarrow{}
 \begin{bmatrix} 
  -2 & 1 \\  
  0 & 0 \end{bmatrix} \end{equation*}

 \begin{equation*} 
  -2x_1+x_2=0 \quad \rightarrow{} \quad x_2=2x_1 \quad \rightarrow{} \quad 
  v_1=
   \begin{bmatrix} 1 \\ 2  \end{bmatrix}
\end{equation*}

And repeat for the second eigenvalue, -3:
\begin{equation*} A+3I=\begin{bmatrix} 
    9+3 & -8 \\
    24 & -19+3 \end{bmatrix}= \begin{bmatrix} 
    12 & -8 \\
    24 & -16 \end{bmatrix}
\end{equation*}

 \begin{equation*}   
   \begin{bmatrix} 
  12 & -8 \\  
  24 & -16   \end{bmatrix} \longrightarrow{}
 \begin{bmatrix} 
  -3 & 2 \\  
  0 & 0
 \end{bmatrix} \end{equation*}

 \begin{equation*} 
  -3x_1+2x_2=0 \quad \rightarrow{} \quad 2x_2=3x_1 \quad \rightarrow{} \quad v_2=
   \begin{bmatrix} 2 \\ 3  \end{bmatrix}
\end{equation*}

So we have found $\lambda_1=-7$ with eigenvector $\begin{bmatrix} 1 \\ 2  \end{bmatrix}$ and $\lambda_2=-3$ with eigenvector $\begin{bmatrix} 2 \\ 3  \end{bmatrix}$. We can plug these values into our formula for the general solution of a system of linear first-order ODEs with constant coefficients:

$$y=C_1e^{\lambda_1 t}v_1 +C_2e^{\lambda_2 t}v_2$$
\begin{tcolorbox}[colback=white, title=Solution]
$$y=C_1e^{-7 t}\begin{bmatrix} 1 \\ 2  \end{bmatrix} +C_2e^{-3 t}\begin{bmatrix} 2 \\ 3 \end{bmatrix}$$
\end{tcolorbox}

\newpage
Let's also do a quick check using R:
```{r}
A <- matrix(c(9, -8, 24, -19), 2, 2, byrow=TRUE)

lambda1 <- eigen(A)$values[1] # eigenvalue 1
v1<- eigen(A)$vectors[,1] / eigen(A)$vectors[1,1] #eigenvector 1
# divided by the first element to give integer values

lambda2 <- eigen(A)$values[2] # eigenvalue 2
v2 <- eigen(A)$vectors[,2] / eigen(A)$vectors[1,2]*2 #eigenvector 2
# divided by the first element and multiplied by 2 to give integer values
```

```{r}
paste("lambda 1 = ", lambda1, " with eigenvector [", v1[1], ",", v1[2], "]", sep="")
paste("lambda 2 = ", lambda2, " with eigenvector [", v2[1], ",", v2[2], "]", sep="")
```

(b)

$$y=C_1e^{-t} \begin{bmatrix} \sin{2t} \\ \cos{2t} \end{bmatrix}+
C_2e^{-t} \begin{bmatrix} -\cos{2t} \\ \sin{2t} \end{bmatrix} $$

### 2024-03-16 bio_dynamics/bio274_filing_cabinet.Rmd
math hw proiblem 2
finished bio hw6

@@
 ```
+
+\newpage
+Let's first do a quick check of our eigenvalues and eigenvectors using R:
+```{r}
+A <- matrix(c(9, -8, 24, -19), 2, 2, byrow=TRUE)
+
+lambda1 <- eigen(A)$values[1] # eigenvalue 1
+v1<- eigen(A)$vectors[,1] / eigen(A)$vectors[1,1] #eigenvector 1
+# divided by the first element to give integer values
+
+lambda2 <- eigen(A)$values[2] # eigenvalue 2
+v2 <- eigen(A)$vectors[,2] / eigen(A)$vectors[1,2]*2 #eigenvector 2
+# divided by the first element and multiplied by 2 to give integer values
+```
+
+```{r}
+paste("lambda 1 = ", lambda1, " with eigenvector [", v1[1], ",", v1[2], "]", sep="")
+paste("lambda 2 = ", lambda2, " with eigenvector [", v2[1], ",", v2[2], "]", sep="")
+```

### 2024-03-16 bio_dynamics/bio274_hw6.Rmd
finished bio hw6

@@
 author: "Ruby Krasnow"
-date: "2024-03-15"
+date: "2024-03-16"
 output:
@@
 Find the general solution for each of the following systems:
-(a)
+
+### (a)
+
 \begin{align*}
@@
 
-$$y=C_1e^{\lambda_1 t}v_1 +C_2e^{\lambda_2 t}v_2$$
+$$X=C_1e^{\lambda_1 t}v_1 +C_2e^{\lambda_2 t}v_2$$
 \begin{tcolorbox}[colback=white, title=Solution]
-$$y=C_1e^{-7 t}\begin{bmatrix} 1 \\ 2  \end{bmatrix} +C_2e^{-3 t}\begin{bmatrix} 2 \\ 3 \end{bmatrix}$$
+$$X=C_1e^{-7 t}\begin{bmatrix} 1 \\ 2  \end{bmatrix} +C_2e^{-3 t}\begin{bmatrix} 2 \\ 3 \end{bmatrix}$$
 \end{tcolorbox}
 
-\newpage
-Let's also do a quick check using R:
-```{r}
-A <- matrix(c(9, -8, 24, -19), 2, 2, byrow=TRUE)
+We can differentiate our solution and check that it satisfies the original system of differential equations:
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+$$X'=C_1\begin{bmatrix} -7e^{-7 t} \\ -14e^{-7 t}  \end{bmatrix} +C_2\begin{bmatrix} -6e^{-3 t} \\ -9e^{-3 t} \end{bmatrix}$$
 
-lambda1 <- eigen(A)$values[1] # eigenvalue 1
-v1<- eigen(A)$vectors[,1] / eigen(A)$vectors[1,1] #eigenvector 1
-# divided by the first element to give integer values
+$$x'= -7C_1e^{-7 t}-6C_2e^{-3t}=9(C_1e^{-7 t}+2C_2e^{-3 t})-8(2C_1e^{-7 t}+3C_2e^{-3 t})=9x-8y$$
+$$y'= -14C_1e^{-7 t}-9C_2e^{-3t}=24(C_1e^{-7 t}+2C_2e^{-3 t})-19(2C_1e^{-7 t}+3C_2e^{-3 t})=24x-19y$$
+\end{tcolorbox}
 
-lambda2 <- eigen(A)$values[2] # eigenvalue 2
-v2 <- eigen(A)$vectors[,2] / eigen(A)$vectors[1,2]*2 #eigenvector 2
-# divided by the first element and multiplied by 2 to give integer values
-```
+### (b)
 
-```{r}
-paste("lambda 1 = ", lambda1, " with eigenvector [", v1[1], ",", v1[2], "]", sep="")
-paste("lambda 2 = ", lambda2, " with eigenvector [", v2[1], ",", v2[2], "]", sep="")
-```
+\begin{align*}
+      \dot{x} &= 2y-x \\
+      \dot{y} &= -2x-y
+\end{align*}
 
-(b)
+\begin{equation*}
+B = \begin{bmatrix} 
+    -1 & 2 \\
+    -2 & -1 \end{bmatrix}
+  \end{equation*}
 
-$$y=C_1e^{-t} \begin{bmatrix} \sin{2t} \\ \cos{2t} \end{bmatrix}+
+\begin{equation*}
+\det{\begin{bmatrix} 
+    -1-\lambda & 2 \\
+    -2 & -1-\lambda \end{bmatrix}}=0 \\
+\end{equation*}
+
+$$ \left(-1-\lambda\right) \left(-1-\lambda\right)+4=0$$
+$$\lambda^2+2\lambda+5=0$$
+$$\frac{-b\pm\sqrt{b^2-4ac}}{2a} \quad \longrightarrow \quad \frac{-2\pm\sqrt{4-4(5)}}{2} = -1 \pm 2i$$
+So we have complex eigenvalues, $\lambda_1=-1+2i, \; \lambda_2=-1-2i$. Now we need to solve $AX=\lambda X$, or $(A-\lambda I)X=0$ with these two values of $\lambda$.
+
+\begin{equation*}
+B-(-1+2i)I=\begin{bmatrix} 
+    -1+1-2i & 2 \\
+    -2 & -1+1-i \end{bmatrix}= \begin{bmatrix} 
+    -2i & 2 \\
+    -2 & -2i \end{bmatrix} \quad \rightarrow{} \quad  
+    \begin{bmatrix} 
+    i & -1 \\
+    0 & 0 \end{bmatrix}
+\end{equation*}
+
+\begin{equation*} 
+  ix_1-x_2=0 \quad \rightarrow{} \quad ix_1=x_2 \quad \rightarrow{} \quad v_1=
+   \begin{bmatrix} i \\ 1  \end{bmatrix}
+\end{equation*}
+
+Which means the eigenvector for the other eigenvalue is the conjugate of this vector, $v_2=\begin{bmatrix} -i \\ 1  \end{bmatrix}$.
+
+Now we use Euler's formula:
+$$e^{(-1+2i)t} \begin{bmatrix} -i \\ 1  \end{bmatrix} = e^{-t}e^{2it}\begin{bmatrix} -i \\ 1  \end{bmatrix} = e^{-t} \left(\cos{(2t)}+i\sin{(2t)}\right)\begin{bmatrix} -i \\ 1 \end{bmatrix} $$
+$$e^{-t} \begin{bmatrix} -i\cos{(2t)}+\sin{(2t)} \\ \cos{(2t)}+i\sin{(2t)} \end{bmatrix}=e^{-t} \left(  \begin{bmatrix} \sin{(2t)} \\ \cos{(2t)}\end{bmatrix}+i\begin{bmatrix} -\cos{(2t)}\\ \sin{(2t)} \end{bmatrix}\right)$$
+Having found our real and imaginary parts, we can now write our general solution as a linear combination of them:
+\begin{tcolorbox}[colback=white, title=Solution]
+$$X=C_1e^{-t} \begin{bmatrix} \sin{2t} \\ \cos{2t} \end{bmatrix}+
 C_2e^{-t} \begin{bmatrix} -\cos{2t} \\ \sin{2t} \end{bmatrix} $$
+\end{tcolorbox}
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+$$x=C_1e^{-t} \sin{2t}-C_2e^{-t}\cos{2t} $$
+$$y=C_1e^{-t}\cos{2t}+C_2e^{-t}\sin{2t} $$
+
+$$X'= 2 C_1e^{-t} \begin{bmatrix} \cos{2t} \\ -\sin{2t} \end{bmatrix}-
+C_1e^{-t} \begin{bmatrix} \sin{2t}\\ \cos{2t} \end{bmatrix}+ 
+2C_2e^{-t} \begin{bmatrix} \sin{2t} \\ \cos{2t} \end{bmatrix}-
+C_2e^{-t} \begin{bmatrix} -\cos{2t} \\ \sin{2t} \end{bmatrix}$$
+
+\begin{align*}
+\dot{x} &= 2 C_1e^{-t}\cos{2t}-C_1e^{-t}\sin{2t}+2C_2e^{-t}\sin{2t}+ C_2e^{-t}\cos{2t} \\
+        &= 2 \left( C_1e^{-t}\cos{2t}+C_2e^{-t}\sin{2t} \right)- \left( C_1e^{-t} \sin{2t}-C_2e^{-t}\cos{2t}\right) \\
+        &= 2y-x
+\end{align*}
+
+\begin{align*}
+\dot{y} &= -2 C_1e^{-t}\sin{2t}-C_1e^{-t}\cos{2t}+2C_2e^{-t}\cos{2t}-C_2e^{-t}\sin{2t} \\
+        &= -2 \left(C_1e^{-t} \sin{2t}-C_2e^{-t}\cos{2t} \right) - \left(C_1e^{-t}\cos{2t}+C_2e^{-t}\sin{2t} \right) \\
+        &= -2x-y
+\end{align*}
+\end{tcolorbox}
+
+### (c)
+
+\begin{align*}
+      \dot{x} &= x+y \\
+      \dot{y} &= y
+\end{align*}
+
+\begin{equation*}
+C = \begin{bmatrix} 
+    1 & 1 \\
+    0 & 1 \end{bmatrix}
+  \end{equation*}
+
+\begin{equation*}
+\det{\begin{bmatrix} 
+    1-\lambda & 1 \\
+    0 & 1-\lambda \end{bmatrix}}=0 \\
+\end{equation*}
+
+$$ \left(1-\lambda\right) \left(1-\lambda\right)=0$$
+
+So we have a repeated eigenvalue, $\lambda_1=\lambda_2=1$.
+
+If the eigenvalues of A are repeated (i.e.
+$\lambda_1=\lambda_2$), then the general solution to the system $\dot{X}=AX$ is $X=e^{\lambda t} v_0+te^{\lambda t} v_1$, where $v_0=\begin{bmatrix} x_0 \\ y_0 \end{bmatrix}$ and $v_1 = (A-\lambda I)v_0$.
+
+\begin{align*}
+v_1 &= (A-I)v_0 \\
+    &= \begin{bmatrix} 
+    0 & 1 \\
+    0 & 0 \end{bmatrix} \begin{bmatrix} x_0 \\ y_0 \end{bmatrix} \\
+    &= \begin{bmatrix} y_0 \\ 0 \end{bmatrix}
+\end{align*}
+
+Which means the general solution is:
+\begin{tcolorbox}[colback=white, title=Solution]
+$$X=e^{t} \begin{bmatrix} x_0 \\ y_0 \end{bmatrix}+te^{t}\begin{bmatrix} y_0 \\ 0 \end{bmatrix}$$
+\end{tcolorbox}
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+$$x=e^t x_0 + te^t y_0, \quad y=e^t y_0$$
+$$ \dot{x} =e^t x_0 + te^t y_0+ e^t y_0 = x+y $$
+$$\dot{y} =e^t y_0 =y$$
+\end{tcolorbox}
+
+### (d)
+
+\begin{align*}
+      \dot{x} &= 7y-2x \\
+      \dot{y} &= 3y
+\end{align*}
+
+\begin{equation*}
+D = \begin{bmatrix} 
+    -2 & 7 \\
+    0 & 3 \end{bmatrix}
+  \end{equation*}
+
+  \begin{equation*}
+\det{\begin{bmatrix} 
+    -2-\lambda & 7 \\
+    0 & 3-\lambda \end{bmatrix}}=0 \\
+\end{equation*}
+
+$$ \left(-2-\lambda\right) \left(3-\lambda\right)=0$$
+
+So we have found two distinct real eigenvalues, $\lambda_1=-2, \; \lambda_2=3$.
+Now we need to solve $DX=\lambda X$, or $(D-\lambda I)X=0$ with these two values of $\lambda$.
+Let's start with the first eigenvalue, -2.
+
+\begin{equation*}
+D+2I=\begin{bmatrix} 
+    -2+2 & 7 \\
+    0 & 3+2 \end{bmatrix}= \begin{bmatrix} 
+    0 & 7 \\
+    0 & 5 \end{bmatrix} \quad \rightarrow{} \quad
+   \begin{bmatrix} 
+  0 & 1 \\  
+  0 & 0   \end{bmatrix}
+ \end{equation*}
+
+ \begin{equation*} 
+  0x_1+x_2=0 \quad \rightarrow{} \quad x_2=0 \quad \rightarrow{} \quad v_1=\begin{bmatrix} 1 \\ 0 \end{bmatrix}
+\end{equation*}
+
+And repeat for the second eigenvalue, 3:
+\begin{equation*}
+D-3I=\begin{bmatrix} 
+    -2-3 & 7 \\
+    0 & 3-3 \end{bmatrix}= \begin{bmatrix} 
+    -5 & 7 \\
+    0 & 0 \end{bmatrix}
+ \end{equation*}
+
+ \begin{equation*} 
+  -5x_1+7x_2=0 \quad \rightarrow{} \quad 7x_2=5x_1 \quad \rightarrow{} \quad v_2=\begin{bmatrix} 7 \\ 5 \end{bmatrix}
+\end{equation*}
+
+So we have found $\lambda_1=-2$ with eigenvector $\begin{bmatrix} 1 \\ 0  \end{bmatrix}$ and $\lambda_2=3$ with eigenvector $\begin{bmatrix} 7 \\ 5  \end{bmatrix}$.
+We can plug these values into our formula for the general solution of a system of linear first-order ODEs with constant coefficients:
+
+\begin{tcolorbox}[colback=white, title=Solution]
+$$X=C_1e^{-2 t}\begin{bmatrix} 1 \\ 0  \end{bmatrix} +C_2e^{3 t}\begin{bmatrix} 7 \\ 5 \end{bmatrix}$$
+\end{tcolorbox}
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+$$x=C_1e^{-2 t}+7C_2e^{3 t}$$
+$$y=5C_2e^{3 t}$$
+\begin{align*}
+\dot{x} &=-2C_1e^{-2 t}+21C_2e^{3 t} \\
+        &= 7\left(5C_2e^{3 t}\right)-2\left(C_1e^{-2 t}+7C_2e^{3 t}\right) \\
+        &=7y-2x
+\end{align*}
+
+$$\dot{y} =15C_2e^{3 t}=3y$$
+\end{tcolorbox}
+
+### (e)
+
+\begin{align*}
+      \dot{x} &= x-2y \\
+      \dot{y} &= 3x-6y
+\end{align*}
+
+\begin{equation*}
+E = \begin{bmatrix} 
+    1 & -2 \\
+    3 & -6 \end{bmatrix}
+  \end{equation*}
+
+  \begin{equation*}
+\det{\begin{bmatrix} 
+    1-\lambda & -2 \\
+    3 & -6-\lambda \end{bmatrix}}=0 \\
+\end{equation*}
+
+$$ \left(1-\lambda\right) \left(-6-\lambda\right)+6=0$$
+$$-6-\lambda+6\lambda+\lambda^2+6=0$$
+$$\lambda^2+5\lambda=0$$
+Or $\lambda(\lambda+5)=0$, which means $\lambda_1=0$ and $\lambda_2=-5$.
+
+\begin{equation*}
+E-0I=\begin{bmatrix} 
+    1-0 & -2 \\
+    3 & -6-0 \end{bmatrix}= \begin{bmatrix} 
+    1 & -2 \\
+    0 & 0 \end{bmatrix}
+ \end{equation*}
+
+  \begin{equation*} 
+  x_1-2x_2=0 \quad \rightarrow{} \quad 2x_2=x_1 \quad \rightarrow{} \quad v_1=\begin{bmatrix} 2 \\ 1 \end{bmatrix}
+\end{equation*}
+
+\begin{equation*}
+E+5I=\begin{bmatrix} 
+    1+5 & -2 \\
+    3 & -6+5 \end{bmatrix}= \begin{bmatrix} 
+    6 & -2 \\
+    3 &-1 \end{bmatrix}
+ \end{equation*}
+
+ \begin{equation*}
+\begin{bmatrix} 
+    3 & -1 \\
+    0 & 0 \end{bmatrix}
+ \end{equation*}
+
+  \begin{equation*} 
+ 3x_1-x_2=0 \quad \rightarrow{} \quad x_2=3x_1 \quad \rightarrow{} \quad v_2=\begin{bmatrix} 1 \\ 3 \end{bmatrix}
+\end{equation*}
+
+So we have found $\lambda_1=0$ with eigenvector $\begin{bmatrix} 2 \\ 1  \end{bmatrix}$ and $\lambda_2=-5$ with eigenvector $\begin{bmatrix} 1 \\ 3  \end{bmatrix}$.
+We can plug these values into our formula for the general solution of a system of linear first-order ODEs with constant coefficients:
+
+\begin{tcolorbox}[colback=white, title=Solution]
+$$X=C_1\begin{bmatrix} 2 \\ 1  \end{bmatrix} +C_2e^{-5 t}\begin{bmatrix} 1 \\ 3 \end{bmatrix}$$
+\end{tcolorbox}
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+$$x=2C_1+C_2e^{-5 t}$$
+$$y=C_1 + 3C_2e^{-5 t}$$
+$$\dot{x}=-5C_2e^{-5 t}=\left(2C_1+C_2e^{-5 t}\right)-2 \left(C_1 + 3C_2e^{-5 t}\right)=x-2y $$
+$$\dot{y}=-15C_2e^{-5 t}=3 \left(2C_1+C_2e^{-5 t}\right)- 6 \left(C_1 + 3C_2e^{-5 t}\right)=3x-6y  $$
+\end{tcolorbox}

### 2024-03-18 PDEs/math245_hw4.Rmd
started math hw4

---
title: "MATH 245 Homework 4"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-03-18"
output:
  pdf_document:
    extra_dependencies: ["tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1: Find eigenvalues and eigenfunctions

**(a)**

$- \frac{d^2}{dx^2}X(x)=\lambda X(x)$ in $0 < x < l$ with boundary conditions $X'(0) = 0 = X(l)$.

**(b)**

$x^2 X''(x) + xX'(x)+\lambda X(x) = 0$ in $1 < x < e$ with boundary conditions $X(1) = 0 = X(e)$.

**(c)**

On the interval $0 \leq x \leq 1$ of length one, consider the eigenvalue problem
$$-X'' = \lambda X,\quad  X'(0) + X(0) = 0,\quad X(1) = 0$$

(i) Find an eigenfunction with eigenvalue zero. Call it $X_0(x)$.

(ii) Find an equation for the positive eigenvalues $\lambda = \beta^2$.

(iii) Show graphically from part (b) that there are an infinite number of positive eigenvalues.

(iv) Is there a negative eigenvalue?

## Question 2  
Find the Fourier-series of $f(x)$. Does the Fourier-series converge (i) pointwise, or (ii)
uniformly?

**(a)**
\begin{equation*} f(x)=
    \begin{cases}
      1-\lvert x \rvert & \lvert x \rvert \leq 1\\
      1 & 1 < \lvert x \rvert \leq \pi\\
    \end{cases}       
\end{equation*}

**(b)**
\begin{equation*} f(x)=\lvert x \rvert=
    \begin{cases}
      -x & -\pi \leq x \leq 0 \\
      x &  0<x\leq \pi \\
    \end{cases}       
\end{equation*}

**(c)**
$$f(x)=x+x^2, \quad -\pi \leq x \leq \pi $$

## Question 3

**(a)** Find the Fourier-sine-series of
\begin{equation*} f(x)=
    \begin{cases}
      1 & 0 < x < \pi /2 \\
      2 &   \pi /2 < x < \pi \\
    \end{cases}       
\end{equation*}

**(b)** Find the Fourier-cosine-series of $f(x)= \lvert \sin{x} \rvert$. Then find
$$\sum^\infty_{k=1} \frac{1}{4k^2-1} $$

**(c)** The Riemann Zeta function is defined for $s > 1$ by
$$ \zeta(s)= \sum^\infty_{n=1} \frac{1}{n^s}$$

By computing the Fourier series of $x^2$ over $-\pi<x<\pi$ and using Parseval’s identity, compute 
$\zeta(4)$.

**(d)** Use the Fourier series in 2c and the pointwise convergence theorem to find $\zeta(2)$. Then find
$$\sum^\infty_{n=1} \frac{1}{(2n-1)^2} $$

## Question 4
Compute the complex Fourier series of the following functions:

**(a)** Compute the complex Fourier series of $f(x) = e^x$ and show that

$$\coth{\pi}= \frac{1}{\pi}+ \frac{2}{\pi}  \left( \frac{1}{1+1^2} +  \frac{1}{1+2^2} +  \frac{1}{1+3^2} + \;...\right) $$

**(b)** Find the complex Fourier series of $xe^{ix}$. Then use your result to find the Fourier series of $x \cos{x}$ and $x \sin{x}$.

## Question 5.

Find the function represented by the new series which is obtained by termwise integration of the following series from $0$ to $x$.

### 2024-03-20 PDEs/paper_outline.Rmd
started PDE paper

---
title: "Final paper outline"
author: "Ruby Krasnow"
date: "2024-03-19"
output:
  pdf_document:
    extra_dependencies: ["tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Introduction

Reaction-diffusion models are a type of partial differential equation (PDE) that have long been used by ecologists to analyze the migration and population dynamics of invasive species. Recent work has led to innumerable modifications on the original formulations by Fisher (1937) and Skellam (1951). Beyond the fundamental processes of growth and spread, extensions of reaction-diffusion models can be applied to age- and size-structured populations (Neubert and Caswell, 2000), and incorporate environmental heterogeneity, interspecies interactions (competition, predation, mutualism, etc.), Allee effects, invasion control efforts, and long-distance dispersal. For an introduction to the use of PDEs to represent dispersal in a population ecology context, the reader is directed to Holmes et al. (1994) and the seminal papers referenced therein. For an accessible review of recent advancements in modeling invasion using reaction-diffusion equations and integro-difference equations, see Hastings et al. (2005); for a comprehensive treatment, see the textbooks by Shigesada and Kawasaki (1997) or Lewis et al. (2016).

Here, we use a simple one-dimensional reaction-diffusion model to represent the spread of the invasive ascidian *Botrylloides violaceus* in the Gulf of Maine. *B. violaceus* is ... DESCRIPTION HERE
Colonial ascidians generally display indeterminate growth, meaning that colony area tends to increase exponentially with time (Jackson, 1977). 
Thus, our reaction-diffusion model takes the form:

$$\frac{\partial{u}}{\partial{t}}=\text{D}\frac{\partial{u}^2}{\partial{x^2}}+ru $$
where $u(t,x)$ is the population density of *B. violaceus* at time $t$ and spatial coordinate $x$. This is the simplest version of the general RD model, for a ...  (Andow et al., 1990; OTHER SOURCE).

Since *B. violaceus* has overlapping generations and can be considered to be continuously reproducing, it is appropriate to represent the invasion dynamics using a continuous-time PDE rather than a discrete integrodifference equation.

### 2024-03-22 PDEs/math245_hw4.Rmd
started PDE paper
started bio hw7

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-03-18"
+date: "2024-03-22"
 output:
@@
 
-**(a)**
+### (a)
 
-$- \frac{d^2}{dx^2}X(x)=\lambda X(x)$ in $0 < x < l$ with boundary conditions $X'(0) = 0 = X(l)$.
+$- \frac{d^2}{dx^2}X(x)=\lambda X(x)$ in $0 < x < l$ with boundary conditions $X'(0) = 0 = X(l)$
 
-**(b)**
+\underline{Case 1:} Positive eigenvalues, $\lambda = \beta^2 >0$
+
+Re-writing as $X''+\lambda X=0$, we will get the characteristic equation: $r^2+\beta^2=0$. Since our characteristic equation has complex roots $r= \pm i\beta$, our solutions take the form 
+$$X(x)=A\sin{(\beta x)}+B\cos{(\beta x)}$$
+Differentiating, we find that 
+$$X'(x)=\beta A\cos{(\beta x)}-\beta B\sin{(\beta x)}$$
+Now plugging in our initial condition $X'(0) = 0$, we get
+$X'(0)=\beta A=0$.
+And since we are in a case where $\beta \neq 0$, this means $B=0$ so $X(x)=A\sin{(\beta x)}$.
+Now we use our boundary condition $X(l)=0$ to get $X(l)=A\sin{(\beta l)}=0$.
+If $A=0$, then $X(0)=0$ and this contradicts the definition of an eigenfunction.
+Therefore, we must have $\sin{(\beta l)}=0$, which can only occur if $\beta = \frac{(2n+1)\pi}{l}$.
+Therefore, this case gives us eigenvalues
+
+$$\lambda_n= \left(\frac{(2n+1)\pi}{l}\right)^2 \quad n=1,2,3... $$
+with eigenfunctions
+$$X_n(x)=\sin{\left(\frac{(2n+1)\pi x}{l}\right)}$$
+
+\underline{Case 2:} Zero eigenvalues, $\lambda=0$
+$X''=0$ implies that $X(x)$ is of the form $Ax+B$, with derivative $X'(x)=A$.
+Now plugging in our initial condition $X'(0) = 0$, we get $A=0$, which means $X(x)=B$.
+But from the boundary condition $X(l)=0$, we get $B=0$ and so $X(0)=0$.
+Therefore, there are no eigenfunctions $X(x)$ that satisfy $- \frac{d^2}{dx^2}X(x)=\lambda X(x)$ when $\lambda=0$ and hence no zero eigenvalues.
+
+\underline{Case 3:} Negative eigenvalues, $\lambda = -\beta^2 <0$
+This case gives us the characteristic equation: $r^2-\beta^2=0$. Since our characteristic equation has distinct real roots $r= \pm \beta$, our solutions take the form 
+$$X(x)=Ae^{\beta x}+Be^{-\beta x}$$
+Differentiating, we find that
+$$X'(x)=\beta Ae^{\beta x}-\beta Be^{-\beta x}$$
+
+$$X'(0)=\beta A-\beta B=0$$
+Since we are in a case where $\beta \neq 0$, this means $A-B=0$, or $A=B$. Then the boundary condition gives
+$$X(l)=Be^{\beta l}+Be^{-\beta l}=0$$
+
+Since $e^{\beta l}$ and $e^{-\beta l}$ are nonzero for all values of $l$, we must have $B=0$, implying that again $X(x)=0$.
+
+Thus, this problem has only positive eigenvalues and their associated eigenfunctions as found in Case 1.
+
+### (b)
 
@@
 
-**(c)**
+We recognize this equation as having the same form as a second-order Cauchy-Euler equation, a linear homogeneous ODE of the form $ax^2y+bxy'+cy=0$ with the auxiliary equation $ar(r-1)+br+c=0$.
+Here, $a=b=1$ and $c=\lambda$, so we have
+
+$$r(r-1)+r+\lambda=0$$
+$$r^2-r+r+\lambda=0$$
+$$r^2+\lambda=0$$
+\underline{Case 1:} Positive eigenvalues, $\lambda = \beta^2 >0$
+
+$$X(x)=A\sin{(\beta \ln{x})}+B\cos{(\beta \ln{x})}$$
+$$X(1)=A\sin{(0)}+B\cos{(0)}=0 \quad \longrightarrow \quad B=0$$
+$$X(e)=A\sin{(\beta \ln{e})}=0$$
+$$A\sin{(\beta)}=0$$
+If $A=0$, then $X(0)=0$ and this contradicts the definition of an eigenfunction.
+Therefore, we must have $\sin{(\beta)}=0$, which can only occur if $\beta = (2n+1)\pi$.
+Therefore, this case gives us eigenvalues
+$$\lambda_n= (2n+1)^2 \pi^2 \quad n=1,2,3... $$
+with eigenfunctions
+$$X_n(x)=\sin{\left((2n+1)\pi \ln{x}\right)}$$
+
+\underline{Case 2:} Zero eigenvalues, $\lambda=0$
+$X''=0$ implies that $X(x)$ is of the form $Ax+B$.
+Now plugging in our initial condition $X(1) = 0$, we get $A+B=0$, which means $X(x)=Ax-A$.
+But from the boundary condition $X(e)=0$, we get $A(e-1)=0$ which is only possible if $A=0$ and accordingly $X(0)=0$.
+Therefore, there are no eigenfunctions $X(x)$ that satisfy the boundary conditions when $X''=0$ and hence no zero eigenvalues.
+
+\underline{Case 3:} Negative eigenvalues, $\lambda = -\beta^2 <0$
+This case gives us the characteristic equation: $r^2-\beta^2=0$. Since our characteristic equation has distinct real roots $r= \pm \beta$, our solutions take the form 
+$$X(x)=Ax^{\beta }+Bx^{-\beta}$$
+
+$$X(1)=A+B=0 \quad \longrightarrow\quad A=-B$$
+$$X(x)=Ax^{\beta }-Ax^{-\beta}$$
+Then the other boundary condition gives $X(e)=Ae^{\beta}-Ae^{-\beta }=0$.
+Since $e^{\beta}$ and $e^{-\beta}$ are always nonzero, we must have $A=0$, implying that again $X(x)=0$.Thus, this problem has only positive eigenvalues and their associated eigenfunctions as found in Case 1.
+
+### (c)
 
@@
 
+$X''=0$ implies that $X(x)$ is of the form $Ax+B$ and $X'(x)=A$. Then $X(1)=0$ becomes $A+B=0$, or $A=-B$. Now we look at our second condition, 
+$$ X'(0) + X(0) = 0$$
+$X'(0)$ is just $A$ and $X(0)$ is just $B$, so again this gives us $A=-B$. Thus, we have $X(x)=Ax-A=A(x-1)$, so we have found
+$$X_0(x)=x-1 $$
+
 (ii) Find an equation for the positive eigenvalues $\lambda = \beta^2$.
+
+$$X(x)=A\sin{(\beta x)}+B\cos{(\beta x)}$$
+$$X'(x)=\beta A\cos{(\beta x)}-\beta B\sin{(\beta x)}$$
+
+$$X(1)=A\sin{(\beta)}+B\cos{(\beta)}=0$$
+$$A\sin{(\beta)}=-B\cos{(\beta)}$$
+
+\begin{equation}\label{aciii}
+-\frac{B}{A}=\frac{\sin{(\beta)}}{\cos{(\beta)}}=\tan{\beta}
+\end{equation}
+
+$$ X'(0) =\beta A, \quad X(0)=B$$
+$$ X'(0) + X(0) = \beta A+B=0 $$
+\begin{equation}\label{aciv}\beta=-\frac{B}{A}\end{equation}
+
+Combining (\ref{aciii}) and (\ref{aciv}), we get the equation $\beta=\tan{(\beta)}$ for the positive eigenvalues.
 
@@
 
+In part (b), we showed that $\beta$ is a positive eigenvalue if $\beta=\tan{(\beta)}$.
+However, plotting the equations $y=\beta$ and $y=\tan{\beta}$ reveals that these two functions intersect an infinite number of times (Fig.
+1), meaning there are an infinite number of positive eigenvalues.
+
+```{r, echo=FALSE,   fig.height=3, fig.width=5}
+df_1c <- tibble(x = seq(-1*pi, 10*pi, length.out = 1000),
+                y=tan(x)) %>% 
+  mutate(y = ifelse(abs(y) < 30, y, NA))
+
+ggplot(df_1c) + 
+  geom_line(aes(x=x, y=y, color="A"))+
+  geom_line(aes(x=x, y=x, color="B"))+
+  geom_hline(yintercept=0, color="black")+
+ coord_cartesian(ylim=c(-10,10), xlim=c(-1,11))+
+  theme_bw()+
+  labs(y="y", x=expression(beta), color="")+ scale_colour_discrete(labels=c(expression(paste("y=tan(", beta, ")")), expression(paste("y=", beta))))+ggtitle("Figure 1")
+
+```
+
 (iv) Is there a negative eigenvalue?
+
+This case gives us the characteristic equation: $r^2-\beta^2=0$. Since our characteristic equation has distinct real roots $r= \pm \beta$, our solutions take the form 
+$$X(x)=Ae^{\beta x}+Be^{-\beta x}$$
+Differentiating, we find that
+$$X'(x)=\beta Ae^{\beta x}-\beta Be^{-\beta x}$$
+
+$$X'(0)=\beta A-\beta B$$
+$$X(0)=A+B$$
+
+$$ X'(0) + X(0) = \beta A-\beta B+A+B$$
+$$0=A(\beta+1)+B(1-\beta)$$
+$$-A(\beta+1)=B(1-\beta)$$
+$$\frac{\beta+1}{1-\beta}=-\frac{B}{A}$$
+Now turning to our other boundary condition,
+$$X'(1)=\beta Ae^{\beta}-\beta Be^{-\beta}=0$$
+$$\beta Ae^{\beta}=\beta Be^{-\beta}$$
+$$-\frac{B}{A} = -e^{2\beta}$$
+This would give us the following equation:
+$$\frac{\beta+1}{1-\beta}=-e^{2\beta}$$
+But the only time that this equality holds true is when $\beta=0$ (Fig.
+2), but we are in the case when we defined $\beta$ to be nonzero, meaning that we have reached a contradiction and we cannot have any negative eigenvalues.
+
+```{r, echo=FALSE,   fig.height=3, fig.width=5}
+df_1d <- tibble(x = seq(-10, 10, length.out = 1000),
+                y1=exp(2*x),
+                y2=(x+1)/(1-x)) %>% 
+  mutate(y2 = ifelse(abs(y2) < 30, y2, NA))
+
+ggplot(df_1d) + 
+  geom_hline(yintercept=0, color="black")+
+  geom_vline(xintercept=0, color="black")+
+  geom_line(aes(x=x, y=y1, color="A"))+
+  geom_line(aes(x=x, y=y2, color="B"))+
+  annotate(geom = "point",x=0,y=1)+
+  coord_cartesian(ylim=c(-6,8), xlim=c(-4,4))+
+  theme_bw()+
+  labs(y="y", x=expression(beta), color="")+ scale_colour_discrete(labels=c(expression(paste("y=exp(2", beta, ")")), expression(paste("y=(", beta, "+1)/(1-", beta, ")"))))+ggtitle("Figure 2")
+
+```
 
@@
 Find the function represented by the new series which is obtained by termwise integration of the following series from $0$ to $x$.
+
+$$ \sum^\infty_{k=1} (-1)^{k+1} \frac{\cos{kx}}{k}=\log{  \left(2 \cos{\left(\frac{x}{2}\right)}\right) }, \quad -\pi<x<\pi$$

### 2024-03-22 bio_dynamics/bio274_hw7.Rmd
started bio hw7

---
title: "BIOL 274 Homework 7"
author: "Ruby Krasnow"
date: "2024-03-22"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---
# Visualizing Solutions of Linear Systems with Constant Coefficients

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

For each of the following systems, state the general solution and sketch the phase portrait (with at least 4 different solutions / ICs shown):

## Question 1

## Question 2

## Question 3

## Question 4

## Question 5

## Question 6

## Question 7

## Question 8

### 2024-03-24 bio_dynamics/bio274_hw7.Rmd
everything but phase portraits

@@
 author: "Ruby Krasnow"
-date: "2024-03-22"
+date: "2024-03-24"
 output:
@@
 ## Question 1 
+\begin{align*}
+      \dot{x} &= x+2y \\
+      \dot{y} &= 3x+2y
+\end{align*}
+
+\begin{equation*}
+A = \begin{bmatrix} 
+    1 & 2 \\
+    3 & 2 \end{bmatrix}
+  \end{equation*}
+
+\begin{equation*}
+\det{\begin{bmatrix} 
+    1-\lambda & 2 \\
+    3 & 2-\lambda \end{bmatrix}}=0 \\
+\end{equation*}
+
+$$ \left(1-\lambda\right) \left(2-\lambda\right)-6=0 $$
+
+$$2-\lambda-2\lambda+\lambda^2-6=0 $$
+
+$$\lambda^2-3\lambda-4=(\lambda-4)(\lambda+1) $$
+So we have found two distinct real eigenvalues, $\lambda_1=4, \; \lambda_2=-1$.
+Now we need to solve $AX=\lambda X$, or $(A-\lambda I)X=0$ with these two values of $\lambda$.
+Let's start with the first eigenvalue, 4.
+
+\begin{equation*}
+A-4I=\begin{bmatrix} 
+    1-4 & 2 \\
+    3 & 2-4 \end{bmatrix}= \begin{bmatrix} 
+    -3 & 2 \\
+    3 & -2 \end{bmatrix} \longrightarrow{}
+ \begin{bmatrix} 
+  -3 & 2 \\  
+  0 & 0
+ \end{bmatrix} \end{equation*}
+
+ \begin{equation*} 
+  -3x_1+2x_2=0 \quad \rightarrow{} \quad 2x_2=3x_1 \quad \rightarrow{} \quad 
+  v_1=
+   \begin{bmatrix} 2 \\ 3  \end{bmatrix}
+\end{equation*}
+
+And repeat for the second eigenvalue, -1:
+\begin{equation*}
+A-(-I)=\begin{bmatrix} 
+    1+1 & 2 \\
+    3 & 2+1 \end{bmatrix}= \begin{bmatrix} 
+    2 & 2 \\
+    3 & 3 \end{bmatrix} \longrightarrow{}
+ \begin{bmatrix} 
+  1 & 1 \\  
+  0 & 0
+ \end{bmatrix} \end{equation*}
+
+ \begin{equation*} 
+  x_1+x_2=0 \quad \rightarrow{} \quad x_2=-x_1 \quad \rightarrow{} \quad 
+  v_2=
+   \begin{bmatrix} 1 \\ -1  \end{bmatrix}
+\end{equation*}
+
+\begin{tcolorbox}[colback=white, title=Solution]
+$$X=C_1e^{4 t}\begin{bmatrix} 2 \\ 3  \end{bmatrix} +C_2e^{- t}\begin{bmatrix} 1 \\ -1 \end{bmatrix}$$
+\end{tcolorbox}
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+$$X'=C_1\begin{bmatrix} 8e^{4 t} \\ 12e^{4 t}  \end{bmatrix} +C_2\begin{bmatrix} -e^{-t} \\ e^{- t} \end{bmatrix}$$
+$$x'= 8C_1e^{4 t}-C_2e^{-t}=2C_1e^{4 t}+C_2e^{- t}+2(3C_1e^{4 t}-C_2e^{- t})=x+2y$$
+$$y'= 12C_1e^{4 t}+C_2e^{-t}=3(2C_1e^{4 t}+C_2e^{- t})+2(3C_1e^{4 t}-C_2e^{- t})=3x+2y$$
+\end{tcolorbox}
 
@@
 
+\begin{align*}
+      \dot{x} &= 4x+2y \\
+      \dot{y} &= -x+y
+\end{align*}
+
+\begin{equation*}
+A = \begin{bmatrix} 
+    4 & 2 \\
+    -1 & 1 \end{bmatrix}
+  \end{equation*}
+
+\begin{equation*}
+\det{\begin{bmatrix} 
+    4-\lambda & 2 \\
+    -1 & 1-\lambda \end{bmatrix}}=0 \\
+\end{equation*}
+
+$$ \left(4-\lambda\right) \left(1-\lambda\right)+2=0 $$
+$$ \lambda^2-5\lambda+6=0 $$
+$$ \left(\lambda-3\right) \left(\lambda-2\right)=0 $$
+So we have found two distinct real eigenvalues, $\lambda_1=3, \; \lambda_2=2$.
+Now we need to solve $AX=\lambda X$, or $(A-\lambda I)X=0$ with these two values of $\lambda$.
+Let's start with the first eigenvalue, 3.
+
+\begin{equation*}
+A-3I=\begin{bmatrix} 
+    4-3 & 2 \\
+    -1 & 1-3 \end{bmatrix}= \begin{bmatrix} 
+    1 & 2 \\
+    -1 & -2 \end{bmatrix} \longrightarrow{}
+ \begin{bmatrix} 
+  1 & 2 \\  
+  0 & 0
+ \end{bmatrix} \end{equation*}
+
+  \begin{equation*} 
+  x_1+2x_2=0 \quad \rightarrow{} \quad 2x_2=-x_1 \quad \rightarrow{} \quad 
+  v_1=
+   \begin{bmatrix} 2 \\ -1  \end{bmatrix}
+\end{equation*}
+
+ And repeat for the second eigenvalue, 2:
+ \begin{equation*}
+A-2I=\begin{bmatrix} 
+    4-2 & 2 \\
+    -1 & 1-2 \end{bmatrix}= \begin{bmatrix} 
+    2 & 2 \\
+    -1 & -1 \end{bmatrix} \longrightarrow{}
+ \begin{bmatrix} 
+  1 & 1 \\  
+  0 & 0
+ \end{bmatrix} \end{equation*}
+
+  \begin{equation*} 
+  x_1+x_2=0 \quad \rightarrow{} \quad x_2=-x_1 \quad \rightarrow{} \quad 
+  v_2=
+   \begin{bmatrix} 1 \\ -1  \end{bmatrix}
+\end{equation*}
+
+ \begin{tcolorbox}[colback=white, title=Solution]
+$$X=C_1e^{3 t}\begin{bmatrix} 2 \\ -1  \end{bmatrix} +C_2e^{2 t}\begin{bmatrix} 1 \\ -1 \end{bmatrix}$$
+\end{tcolorbox}
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+
+$$x'= 6C_1e^{3 t}+2C_2e^{2t}=4(2C_1e^{3 t}+C_2e^{2 t})+2(-C_1e^{3 t}-C_2e^{2 t})=4x+2y$$
+
+$$y'= -3C_1e^{3 t}-2C_2e^{2t}=-(2C_1e^{3 t}+C_2e^{2 t})+(-C_1e^{3 t}-C_2e^{2 t})=-x+y$$
+\end{tcolorbox}
+
 ## Question 3
+
+\begin{align*}
+      \dot{x} &= 3x-5y \\
+      \dot{y} &= 5x+3y
+\end{align*}
+
+\begin{equation*}
+A = \begin{bmatrix} 
+    3 & -5 \\
+    5 & 3 \end{bmatrix}
+  \end{equation*}
+
+\begin{equation*}
+\det{\begin{bmatrix} 
+    3-\lambda & -5 \\
+    5 & 3-\lambda \end{bmatrix}}=0 \\
+\end{equation*}
+
+$$ \left(3-\lambda\right) \left(3-\lambda\right)+25=0 $$
+$$ \lambda^2-6\lambda+34=0 $$
+$$\frac{-b\pm\sqrt{b^2-4ac}}{2a} \quad \longrightarrow \quad \frac{6\pm\sqrt{36-136}}{2} = \frac{6\pm10i}{2}= 3 \pm 5i$$
+So we have complex eigenvalues, $\lambda_1=3+5i, \; \lambda_2=3-5i$. Now we need to solve $AX=\lambda X$, or $(A-\lambda I)X=0$ with these two values of $\lambda$.
+
+\begin{equation*}
+A-(3+5i)I=\begin{bmatrix} 
+    3-3-5i & -5 \\
+    5 & 3-3-5i \end{bmatrix}= \begin{bmatrix} 
+    -5i & -5 \\
+    5 & -5i \end{bmatrix} \quad \rightarrow{} \quad  
+    \begin{bmatrix} 
+    i & 1 \\
+    0 & 0 \end{bmatrix}
+\end{equation*}
+
+\begin{equation*} 
+  ix_1+x_2=0 \quad \rightarrow{} \quad ix_1=-x_2 \quad \rightarrow{} \quad v_1=
+   \begin{bmatrix} i \\ 1  \end{bmatrix}
+\end{equation*}
+
+Which means the eigenvector for the other eigenvalue is the conjugate of this vector, $v_2=\begin{bmatrix} -i \\ 1  \end{bmatrix}$.
+
+Now we use Euler's formula:
+$$e^{(3+5i)t} \begin{bmatrix} i \\ 1  \end{bmatrix} = e^{3t}e^{5it}\begin{bmatrix} i \\ 1  \end{bmatrix} = e^{3t} \left(\cos{(5t)}+i\sin{(5t)}\right)\begin{bmatrix} i \\ 1 \end{bmatrix} $$
+$$e^{3t} \begin{bmatrix} i\cos{(5t)}-\sin{(5t)} \\ \cos{(5t)}+i\sin{(5t)} \end{bmatrix}=e^{3t} \left(  \begin{bmatrix} -\sin{(5t)} \\ \cos{(5t)}\end{bmatrix}+i\begin{bmatrix} \cos{(5t)}\\ \sin{(5t)} \end{bmatrix}\right)$$
+
+Having found our real and imaginary parts, we can now write our general solution as a linear combination of them:
+\begin{tcolorbox}[colback=white, title=Solution]
+$$X=C_1e^{3t} \begin{bmatrix} -\sin{5t} \\ \cos{5t} \end{bmatrix}+
+C_2e^{3t} \begin{bmatrix} \cos{5t} \\ \sin{5t} \end{bmatrix} $$
+\end{tcolorbox}
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+$$x=-C_1e^{3t}\sin{5t}+C_2e^{3t}\cos{5t} $$
+$$y=C_1e^{3t}\cos{5t}+C_2e^{3t}\sin{5t} $$
+\begin{align*}
+\dot{x} &= -5 C_1e^{3t}\cos{5t}-3C_1e^{3t}\sin{5t}-5C_2e^{3t}\sin{5t}+3 C_2e^{3t}\cos{5t} \\
+        &= 3 \left( -C_1e^{3t}\sin{5t}+C_2e^{3t}\cos{5t} \right)- 5\left( C_1e^{3t}\cos{5t}+C_2e^{3t}\sin{5t}\right) \\
+        &= 3x-5y
+\end{align*}
+
+\begin{align*}
+\dot{y} &= -5 C_1e^{3t}\sin{5t}+3C_1e^{3t}\cos{5t}+5C_2e^{3t}\cos{5t}+3C_2e^{3t}\sin{5t} \\
+        &= 5 \left( -C_1e^{3t}\sin{5t}+C_2e^{3t}\cos{5t} \right)+3\left( C_1e^{3t}\cos{5t}+C_2e^{3t}\sin{5t}\right) \\
+        &= 5x+3y
+\end{align*}
+\end{tcolorbox}
 
@@
 
+\begin{align*}
+      \dot{x} &= 6x-8y \\
+      \dot{y} &= -3x+4y
+\end{align*}
+
+\begin{equation*}
+A = \begin{bmatrix} 
+    6 & -8 \\
+    -3 & 4 \end{bmatrix}
+  \end{equation*}
+
+\begin{equation*}
+\det{\begin{bmatrix} 
+    6-\lambda & -8 \\
+    -3 & 4-\lambda \end{bmatrix}}=0 \\
+\end{equation*}
+
+$$ \left(6-\lambda\right) \left(4-\lambda\right)-24=0 $$
+$$\lambda^2-10\lambda=\lambda(\lambda-10)=0 $$
+So we have found two distinct real eigenvalues, $\lambda_1=0, \; \lambda_2=10$.
+Now we need to solve $AX=\lambda X$, or $(A-\lambda I)X=0$ with these two values of $\lambda$.
+Let's start with the first eigenvalue, 0.
+
+\begin{equation*}
+A-0I=\begin{bmatrix} 
+   6 & -8 \\
+    -3 & 4 \end{bmatrix} \quad \rightarrow{} \quad  
+    \begin{bmatrix} 
+    3 & -4 \\
+    0 & 0 \end{bmatrix} \end{equation*}
+
+\begin{equation*} 
+  3x_1-4x_2=0 \quad \rightarrow{} \quad 3x_1=4x_2 \quad \rightarrow{} \quad v_1=
+   \begin{bmatrix} 4 \\ 3  \end{bmatrix}
+\end{equation*}
+
+ And repeat for the second eigenvalue, 10:
+\begin{equation*}
+A-10I=\begin{bmatrix} 
+   6-10 & -8 \\
+    -3 & 4-10 \end{bmatrix} =
+    \begin{bmatrix} 
+   -4 & -8 \\
+    -3 & -6 \end{bmatrix} \quad \rightarrow{} \quad  
+    \begin{bmatrix} 
+    1 & 2 \\
+    0 & 0 \end{bmatrix} \end{equation*}
+
+\begin{equation*} 
+  x_1+2x_2=0 \quad \rightarrow{} \quad x_1=-2x_2 \quad \rightarrow{} \quad v_2=
+   \begin{bmatrix} -2 \\ 1  \end{bmatrix}
+\end{equation*}
+
+\begin{tcolorbox}[colback=white, title=Solution]
+$$X=C_1\begin{bmatrix} 4 \\ 3  \end{bmatrix} +C_2e^{10 t}\begin{bmatrix} -2 \\ 1 \end{bmatrix}$$
+\end{tcolorbox}
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+
+$$x'= -20C_2e^{10t}=6(4C_1-2C_2e^{10t)}-8(3C_1+C_2e^{10 t})=6x-8y$$
+$$y'= 10C_2e^{10t}=-3(4C_1-2C_2e^{10t)}+4(3C_1+C_2e^{10 t})=-3x+4y$$
+\end{tcolorbox}
+
 ## Question 5
+
+\begin{align*}
+      \dot{x} &= 4x-2y \\
+      \dot{y} &= 2x
+\end{align*}
+
+\begin{equation*}
+A = \begin{bmatrix} 
+    4 & -2 \\
+    2 & 0 \end{bmatrix}
+  \end{equation*}
+
+\begin{equation*}
+\det{\begin{bmatrix} 
+    4-\lambda & -2 \\
+    2 & -\lambda \end{bmatrix}}=0 \\
+\end{equation*}
+
+$$ \left(4-\lambda\right) \left(-\lambda\right)+4=0 $$
+$$\lambda^2-4\lambda+4=(\lambda-2)^2=0 $$
+So we have a repeated eigenvalue, $\lambda_1=\lambda_2=2$.
+
+If the eigenvalues of A are repeated (i.e.
+$\lambda_1=\lambda_2$), then the general solution to the system $\dot{X}=AX$ is $X=e^{\lambda t} v_0+te^{\lambda t} v_1$, where $v_0=\begin{bmatrix} x_0 \\ y_0 \end{bmatrix}$ and $v_1 = (A-\lambda I)v_0$.
+
+\begin{align*}
+v_1 &= (A-2I)v_0 \\
+    &= \begin{bmatrix} 
+    2 & -2 \\
+    2 & -2 \end{bmatrix} \begin{bmatrix} x_0 \\ y_0 \end{bmatrix} \\
+    &= \begin{bmatrix} 2x_0-2y_0 \\ 2x_0-2y_0 \end{bmatrix}
+\end{align*}
+
+Which means the general solution is:
+\begin{tcolorbox}[colback=white, title=Solution]
+$$X=e^{2t} \begin{bmatrix} x_0 \\ y_0 \end{bmatrix}+2te^{2t}\begin{bmatrix} x_0-y_0 \\ x_0-y_0 \end{bmatrix}$$
+\end{tcolorbox}
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+\begin{align*}
+\dot{x}&= 2e^{2t} x_0 + 4te^{2t} (x_0-y_0)+ 2e^{2t} (x_0-y_0) \\
+       &= 4e^{2t} x_0 + 4te^{2t} (x_0-y_0)-2e^{2t} y_0 \\
+       &= 4(e^{2t} x_0 + 2te^{2t} (x_0-y_0))-2(e^{2t} y_0 + 2te^{2t} (x_0-y_0)) \\
+       &= 4x-2y
+\end{align*}
+
+\begin{align*}
+\dot{y}&= 2e^{2t} y_0 + 4te^{2t} (x_0-y_0)+ 2e^{2t} (x_0-y_0) \\
+       &= 2e^{2t} x_0 + 4te^{2t} (x_0-y_0)\\
+       &= 2(e^{2t} x_0 + 2te^{2t} (x_0-y_0)) \\
+       &= 2x
+\end{align*}
+\end{tcolorbox}
 
@@
 
+\begin{align*}
+      \dot{x} &= -4x+y \\
+      \dot{y} &= 3x-2y
+\end{align*}
+
+\begin{equation*}
+A = \begin{bmatrix} 
+    -4 & 1 \\
+    3 & -2 \end{bmatrix}
+  \end{equation*}
+
+\begin{equation*}
+\det{\begin{bmatrix} 
+    -4-\lambda & 1 \\
+    3 & -2-\lambda \end{bmatrix}}=0 \\
+\end{equation*}
+
+$$ \left(-4-\lambda\right) \left(-2-\lambda\right)-3=0 $$
+
+$$\lambda^2+6\lambda+5=(\lambda+5)(\lambda+1)=0 $$
+So we have found two distinct real eigenvalues, $\lambda_1=-5, \; \lambda_2=-1$.
+Now we need to solve $AX=\lambda X$, or $(A-\lambda I)X=0$ with these two values of $\lambda$.
+Let's start with the first eigenvalue, 0.
+
+\begin{equation*}
+A+5I=\begin{bmatrix} 
+    -4+5 & 1 \\
+    3 & -2+5 \end{bmatrix} =
+    \begin{bmatrix} 
+    1 & 1 \\
+    3 & 3 \end{bmatrix} \quad \rightarrow{} \quad  
+    \begin{bmatrix} 
+    1 & 1 \\
+    0 & 0 \end{bmatrix} \end{equation*}
+
+\begin{equation*} 
+  x_1+x_2=0 \quad \rightarrow{} \quad x_1=-x_2 \quad \rightarrow{} \quad v_2=
+   \begin{bmatrix} 1 \\ -1  \end{bmatrix}
+\end{equation*}
+
+\begin{equation*}
+A+I=\begin{bmatrix} 
+    -4+1 & 1 \\
+    3 & -2+1 \end{bmatrix} =
+    \begin{bmatrix} 
+    -3 & 1 \\
+    3 & -1 \end{bmatrix} \quad \rightarrow{} \quad  
+    \begin{bmatrix} 
+    3 & -1 \\
+    0 & 0 \end{bmatrix} \end{equation*}
+
+\begin{equation*} 
+  3x_1-x_2=0 \quad \rightarrow{} \quad 3x_1=x_2 \quad \rightarrow{} \quad v_2=
+   \begin{bmatrix} 1 \\ 3  \end{bmatrix}
+\end{equation*}
+
+ \begin{tcolorbox}[colback=white, title=Solution]
+$$X=C_1e^{-5 t}\begin{bmatrix} 1 \\ -1  \end{bmatrix} +C_2e^{- t}\begin{bmatrix} 1 \\ 3 \end{bmatrix}$$
+\end{tcolorbox}
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+$$x'=-5C_1e^{-5 t}-C_2 e^{- t}=-4(C_1e^{-5 t}+C_2 e^{- t})+(-C_1e^{-5 t}+3C_2 e^{- t})=-4x+y $$
+$$y'=5C_1e^{-5 t}-3C_2 e^{- t}=3(C_1e^{-5 t}+C_2 e^{- t})-2(-C_1e^{-5 t}+3C_2 e^{- t})=3x-2y $$
+\end{tcolorbox}
+
 ## Question 7
 
+\begin{align*}
+      \dot{x} &= -9x-3y \\
+      \dot{y} &= 3x+y
+\end{align*}
+
+\begin{equation*}
+A = \begin{bmatrix} 
+    -9 & -3 \\
+    3 & 1 \end{bmatrix}
+  \end{equation*}
+
+\begin{equation*}
+\det{\begin{bmatrix} 
+    -9-\lambda & -3 \\
+    3 & 1-\lambda \end{bmatrix}}=0 \\
+\end{equation*}
+
+$$ \left(-9-\lambda\right) \left(1-\lambda\right)+9=0 $$
+
+$$\lambda^2+8\lambda=\lambda(\lambda+8)=0 $$
+
+So we have found two distinct real eigenvalues, $\lambda_1=0, \; \lambda_2=-8$.
+Now we need to solve $AX=\lambda X$, or $(A-\lambda I)X=0$ with these two values of $\lambda$.
+Let's start with the first eigenvalue, 0.
+
+\begin{equation*}
+A-0I=\begin{bmatrix} 
+  -9 & -3 \\
+    3 & 1 \end{bmatrix} \quad \rightarrow{} \quad  
+    \begin{bmatrix} 
+    3 & 1 \\
+    0 & 0 \end{bmatrix} \end{equation*}
+
+\begin{equation*} 
+  3x_1+x_2=0 \quad \rightarrow{} \quad 3x_1=-x_2 \quad \rightarrow{} \quad v_1=
+   \begin{bmatrix} 1 \\ -3  \end{bmatrix}
+\end{equation*}
+
+ And repeat for the second eigenvalue, -8:
+\begin{equation*}
+A+8I=\begin{bmatrix} 
+    -9+8 & -3 \\
+    3 & 1+8 \end{bmatrix} =
+    \begin{bmatrix} 
+   -1 & -3 \\
+    3 & 9 \end{bmatrix} \quad \rightarrow{} \quad  
+    \begin{bmatrix} 
+    1 & 3 \\
+    0 & 0 \end{bmatrix} \end{equation*}
+
+\begin{equation*} 
+  x_1+3x_2=0 \quad \rightarrow{} \quad x_1=-3x_2 \quad \rightarrow{} \quad v_2=
+   \begin{bmatrix} 3 \\ -1  \end{bmatrix}
+\end{equation*}
+
+ \begin{tcolorbox}[colback=white, title=Solution]
+$$X=C_1\begin{bmatrix} 1 \\ -3  \end{bmatrix} +C_2e^{-8t}\begin{bmatrix} 3 \\ -1 \end{bmatrix}$$
+\end{tcolorbox}
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+$$x'=-24C_2e^{-8t}= -9(C_1+3C_2e^{-8t})-3(-3C_1-C_2e^{-8t})=-9x-3y$$
+$$y'=8C_2e^{-8t}=3(C_1+3C_2e^{-8t})+(-3C_1-C_2e^{-8t})=3x+y $$
+
+\end{tcolorbox}
+
 ## Question 8
+
+\begin{align*}
+      \dot{x} &= -3x+10y \\
+      \dot{y} &= -x+3y
+\end{align*}
+
+\begin{equation*}
+A = \begin{bmatrix} 
+    -3 & 10 \\
+    -1 & 3 \end{bmatrix}
+  \end{equation*}
+
+\begin{equation*}
+\det{\begin{bmatrix} 
+    -3-\lambda & 10 \\
+    -1 & 3-\lambda \end{bmatrix}}=0 \\
+\end{equation*}
+
+$$ \left(-3-\lambda\right) \left(3-\lambda\right)+10=0 $$
+
+$$\lambda^2+1=0 $$
+
+So we have complex eigenvalues, $\lambda_1=i, \; \lambda_2=-i$. Now we need to solve $AX=\lambda X$, or $(A-\lambda I)X=0$ with these two values of $\lambda$.
+
+\begin{equation*}
+A-iI=\begin{bmatrix} 
+    -3-i & 10 \\
+    -1 & 3-i \end{bmatrix}\quad \rightarrow{} \quad  
+    \begin{bmatrix} 
+    -1 & 3-i \\
+    0 & 0 \end{bmatrix}
+\end{equation*}
+
+\begin{equation*} 
+  -x_1+(3-i)x_2=0 \quad \rightarrow{} \quad x_1=(3-i)x_2 \quad \rightarrow{} \quad v_1=
+   \begin{bmatrix} 3-i \\ 1  \end{bmatrix}
+\end{equation*}
+
+Which means the eigenvector for the other eigenvalue is the conjugate of this vector, $v_2=\begin{bmatrix} 3+i \\ 1  \end{bmatrix}$.
+
+Now we use Euler's formula:
+$$e^{it} \begin{bmatrix} 3-i \\ 1  \end{bmatrix} = \left(\cos{(t)}+i\sin{(t)}\right)\begin{bmatrix} 3-i \\ 1 \end{bmatrix} $$
+
+$$\begin{bmatrix} (3-i)\cos{t}+(3i+1)\sin{t} \\ \cos{t}+i\sin{t} \end{bmatrix}=  
+\begin{bmatrix} 3\cos{t}+\sin{t} \\ \cos{t} \end{bmatrix} +i\begin{bmatrix} 3\sin{t}-\cos{t}\\ \sin{t} \end{bmatrix}$$
+
+Having found our real and imaginary parts, we can now write our general solution as a linear combination of them:
+\begin{tcolorbox}[colback=white, title=Solution]
+$$X=C_1 \begin{bmatrix} 3\cos{t}+\sin{t} \\ \cos{t} \end{bmatrix}+
+C_2 \begin{bmatrix} 3\sin{t}-\cos{t}\\ \sin{t} \end{bmatrix} $$
+\end{tcolorbox}
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
+
+\begin{align*}
+x'&=-3C_1\sin{t}+C_1\cos{t}+3C_2 \cos{t}+C_2\sin{t} \\
+  &= -3(3C_1\cos{t}+C_1\sin{t}+3C_2\sin{t}-C_2\cos{t})+ 10(C_1 \cos{t}+C_2\sin{t}) \\
+  &= -3x+10y
+\end{align*}
+
+\begin{align*}
+y'&=-C_1\sin{t}+C_2 \cos{t} \\
+  &= -(3C_1\cos{t}+C_1\sin{t}+3C_2\sin{t}-C_2\cos{t})+ 3(C_1 \cos{t}+C_2\sin{t}) \\
+  &= -x+3y
+\end{align*}
+\end{tcolorbox}

### 2024-03-25 PDEs/math245_hw4.Rmd
worked on pdes

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-03-22"
+date: "2024-03-24"
 output:
@@
 ## Question 2  
-Find the Fourier-series of $f(x)$. Does the Fourier-series converge (i) pointwise, or (ii)
+Find the Fourier series of $f(x)$. Does the Fourier-series converge (i) pointwise, or (ii)
 uniformly?
 
-**(a)**
+### (a)
 \begin{equation*} f(x)=
@@
 
-**(b)**
+The Fourier series for a function $f(x)$ on an interval $-l<x<l$ is
+$$f(x)=a_0+ \sum_{n=1}^{\infty} \left( a_n \cos{  \left(\frac{n\pi x}{l}\right) }+b_n \sin{ \left(\frac{n\pi x}{l}\right)}\right)$$
+where
+$$a_0=\frac{1}{2l} \int_{-l}^{l}f(x)dx$$ 
+$$ a_n=\frac{1}{l} \int_{-l}^{l}f(x)\cos{  \left(\frac{n\pi x}{l}\right) }dx$$
+$$ b_n=\frac{1}{l} \int_{-l}^{l}f(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx$$
+
+Let's start with $a_0$:
+\begin{align*}
+a_0 &=\frac{1}{2\pi} \int_{-\pi}^{\pi}f(x)dx \\
+    &=\frac{1}{2\pi}\left[ \int_{-\pi}^{-1} 1\;dx + \int_{-1}^{1} (1-\lvert x \rvert) dx+ \int_{1}^{\pi} 1\;dx  \right] \\
+    &= \frac{1}{2\pi}\int_{-\pi}^{\pi}1\;dx - \int_{-1}^{1} \lvert x \rvert\, dx \\
+    &= 1- 2\int_{0}^{1} x \; dx \\
+    &= 1- 1=0
+\end{align*}
+
+where we used the evenness of $\lvert x \rvert$ to rewrite $\int_{-1}^{1} \lvert x \rvert \,dx$ as $2\int_{0}^{1} x \; dx$. Thus, $a_0=0$. Now we perform a similar procedure to find $a_n$:
+
+\begin{align*}
+a_n &= \frac{1}{\pi} \int_{-\pi}^{\pi} f(x) \cos{(nx)}dx \\
+    &= \frac{1}{\pi}\left[ \int_{-\pi}^{-1} \cos{(nx)}\;dx + \int_{-1}^{1} (1-\lvert x \rvert)\cos{(nx)}\; dx+ \int_{1}^{\pi} \cos{(nx)}\;dx  \right] \\
+    &= \frac{1}{\pi} \int_{-\pi}^{\pi} \cos{(nx)}\;dx - \frac{1}{\pi} \int_{-1}^{1} \lvert x \rvert\cos{(nx)}\; dx
+\end{align*}
+
+Since $\sin{(\pi n)}$ and $\sin{(-\pi n)}$ are zero for any integer $n$, the first integral vanishes.
+We use the symmetry of $\lvert x \rvert \cos{(nx)}$ (both functions are even, and the product of two even functions is even) and rewrite the second integral as $$- \frac{2}{\pi} \int_{0}^{1}  x \cos{(nx)} \; dx $$
+
+We then use integration by parts:
+\begin{align*}
+&- \frac{2}{\pi} \int_{0}^{1}  x \cos{(nx)}\; dx \\
+&= - \frac{2}{n\pi} \left[x \sin{(nx)} \Bigr\rvert_0^1 - \int_{0}^{1}\sin{(nx)}\; dx \right] \\
+&= - \frac{2}{n\pi} \left[\sin{(n)}+ \frac{1}{n} \cos{(nx)}\Bigr\rvert_0^1  \right] \\
+&= - \frac{2}{n\pi} \left[\sin{(n)}+ \frac{1}{n} \cos{(n)} - \frac{1}{n} \right]
+\end{align*}
+
+$$ a_n= - \frac{2 \sin{n}}{n\pi} + \frac{2(\cos{n}-1)}{n^2 \pi} $$
+\begin{align*}
+b_n &=
+\frac{1}{\pi}\left[ \int_{-\pi}^{-1} \sin{(nx)}\;dx + \int_{-1}^{1} (1-\lvert x \rvert)\sin{(nx)}\; dx+ \int_{1}^{\pi} \sin{(nx)}\;dx  \right] \\
+    &=  \frac{1}{\pi} \int_{-\pi}^{\pi} \sin{(nx)}\;dx - \frac{1}{\pi} \int_{-1}^{1} \lvert x \rvert\sin{(nx)}\; dx
+\end{align*}
+
+But $\sin{x}$ is an odd function and so is $\lvert x \rvert\sin{(nx)}$, so both of these integrals vanish and we find $b_n=0$.
+
+Now inserting the coefficients $a_n$ we found above, we get the Fourier series:
+$$f(x)=\frac{2}{\pi} \sum_{n=1}^{\infty}  \left(\frac{\sin{n}}{n}+ \frac{\cos{n}+1}{n^2} \right) \cos{ \left( nx \right)}$$
+
+### (b)
 \begin{equation*} f(x)=\lvert x \rvert=
@@
       -x & -\pi \leq x \leq 0 \\
-      x &  0<x\leq \pi \\
+      x &  0<x\leq \pi
     \end{cases}
@@
 
-**(c)**
-$$f(x)=x+x^2, \quad -\pi \leq x \leq \pi $$
+Let's start with $a_0$:
+\begin{align*}
+a_0 &=\frac{1}{2\pi} \int_{-\pi}^{\pi}\lvert x \rvert \, dx \\
+    &=\frac{1}{\pi}\int_{0}^{\pi} x \;dx \\
+    &= \frac{1}{2\pi} x^2 \Bigr\rvert_0^{\pi} = \frac{\pi}{2}
+\end{align*}
+
+\begin{align*}
+a_n &= \frac{1}{\pi} \int_{-\pi}^{\pi}f(x)\cos{\left(nx\right) }dx \\
+    &= \frac{2}{\pi} \int_{0}^{\pi}x\cos{\left(nx\right) }dx \\
+    &= \frac{2}{n \pi} x \sin{(nx)} \Bigr\rvert_0^{\pi} - \frac{2}{n \pi}\int_{0}^{\pi}\sin{\left( nx \right) } dx \\
+    &= 0 + \frac{2}{n^2 \pi} \cos{(nx)}\Bigr\rvert_0^{\pi} \\
+    &= \frac{2}{n^2 \pi} (-1)^{n+1}
+\end{align*}
+
+$$b_n = \frac{1}{\pi} \int_{-\pi}^{\pi} f(x) \sin{(nx)}dx $$
+But since $\lvert x \rvert$ is even and $\sin{(nx)}$ is odd, the product of the two functions is odd and so the integral vanishes over the symmetric interval.
+
+Then we plug in our formulas for $a_0$ and $a_n$:
+$$ f(x) = \frac{\pi}{2} + \frac{2}{\pi} \sum_{n=1}^{\infty} \left[ \frac{1}{n^2} (-1)^{n+1} \cos{(nx)} \right]$$
+
+### (c)
+
+$$f(x)=x+x^2, \quad -\pi \leq x \leq \pi$$
+\begin{align*}
+a_0 &=\frac{1}{2\pi} \int_{-\pi}^{\pi}x + x^2 dx \\
+    &=\frac{1}{2\pi} \left( \frac{x^2}{2}+ \frac{x^3}{3} \right) \Bigr\rvert_{-\pi}^{\pi} \\
+    &= \frac{1}{2\pi} \left( \frac{\pi^2}{2} + \frac{\pi^3}{3} - \frac{\pi^2}{2} + \frac{\pi^3}{3}\right) \\
+    &= \frac{1}{2\pi} \left( \frac{2\pi^3}{3} \right) = \frac{\pi^2}{3}
+\end{align*}
+
+\begin{align*}
+a_n &= \frac{1}{\pi} \int_{-\pi}^{\pi}f(x)\cos{(nx)}dx \\
+    &= \frac{1}{\pi} \int_{-\pi}^{\pi}x\cos{(nx)}+ \frac{1}{\pi} \int_{-\pi}^{\pi}x^2\cos{(nx)}dx 
+\end{align*}
+
+Applying integration by parts to the first integral,
+$$\frac{1}{\pi} \int_{-\pi}^{\pi}x\cos{(nx)}=\frac{1}{n\pi}x\sin{(nx)} \Bigr\rvert_{-\pi}^{\pi}+\frac{1}{n\pi} \int_{-\pi}^{\pi}\sin{(nx)}$$
+$$=\frac{1}{n\pi}x\sin{(nx)} \Bigr\rvert_{-\pi}^{\pi}+\frac{-1}{n^2\pi} \cos{(nx)}\Bigr\rvert_{-\pi}^{\pi}=0+0=0$$
+
+We then use the integration by parts table method demonstrated in class to find that
+
+$$\frac{1}{\pi} \int_{-\pi}^{\pi}x^2\cos{(nx)}= \frac{x^2}{n}\sin{(nx)}\Bigr\rvert_{-\pi}^{\pi} +  \frac{2x}{n^2}\cos{(nx)}\Bigr\rvert_{-\pi}^{\pi}+\frac{2}{n^3}\sin{(nx)}\Bigr\rvert_{-\pi}^{\pi}$$
+The first and last terms (with $\sin{(nx)}$) vanish, while the middle term evaluates to $\frac{4 \cos{(n\pi)}}{n^2}$.
+
+$$\frac{4 \cos{(n\pi)}}{n^2}=\frac{4}{n^2} (-1)^{n}$$
+Thus, we add this to the result of first integral to find that
+$$a_n= \frac{4}{n^2} (-1)^{n} $$
+\begin{align*}
+b_n &= \frac{1}{\pi} \int_{-\pi}^{\pi}f(x)\sin{(nx)}dx \\
+    &= \frac{1}{\pi} \int_{-\pi}^{\pi}x\sin{(nx)}+ \frac{1}{\pi} \int_{-\pi}^{\pi}x^2\sin{(nx)}dx 
+\end{align*}
+
+But since $x^2$ is even and $\sin{}$ is odd, the product under the second integral is odd functions and thus evaluates to zero over the symmetric interval $-\pi<x<\pi$.
+Applying integration by parts to the first integral,
+$$b_n=\frac{-1}{n\pi}x\cos{(nx)} \Bigr\rvert_{-\pi}^{\pi}+\frac{1}{n\pi} \int_{-\pi}^{\pi}\cos{(nx)}$$
+But since $\cos$ is even and $\cos{\pi}=\cos{-\pi}$, both terms are zero. Thus, $b_n=0$ and
+
+$$f(x)=a_0 + \sum_{n=1}^{\infty}  a_n \cos{  \left(\frac{n\pi x}{l}\right) } $$
+$$f(x)=\frac{\pi^2}{3} + 4 \sum_{n=1}^{\infty} \frac{(-1)^{n}\cos{ (nx) }}{n^2}  $$
 
@@
 
-**(a)** Find the Fourier-sine-series of
+### a)  
+Find the Fourier sine series of
 \begin{equation*} f(x)=
@@
 
-**(b)** Find the Fourier-cosine-series of $f(x)= \lvert \sin{x} \rvert$. Then find
+The Fourier sine series of a function $f(x)$ is $$f(x)= \sum_{n=1}^{\infty} A_n \sin{  \left(\frac{n\pi x}{l}\right) }$$
+where 
+$$A_n = \frac{2}{l} \int_{0}^{l}f(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx $$
+$$A_n = \frac{2}{\pi} \int_{0}^{\pi}f(x)\sin{(nx)}dx $$
+\begin{align*}
+A_n &= \frac{2}{\pi}\left[ \int_{0}^{\pi/2}\sin{(nx)}dx + 2 \int_{\pi/2}^{\pi}\sin{(nx)}dx \right] \\
+    &= \frac{2}{\pi}\left[\frac{-1}{n} \cos{(nx)}\Bigr\rvert_{0}^{\pi/2}- \frac{2}{n} \cos{(nx)}\Bigr\rvert_{\pi/2}^{\pi}\right] \\
+    &= \frac{2}{n \pi}\left[\cos{\left(\frac{n\pi}{2}\right)}+1-2\cos{(n\pi)} \right]
+\end{align*}
+
+ $$f(x)=  \frac{2}{\pi} \sum_{n=1}^{\infty} \frac{1}{n}\left[1+\cos{\left(\frac{n\pi}{2}\right)}-2\cos{(n\pi)} \right] \sin{(nx) }$$
+
+### b)
+
+Find the Fourier-cosine-series of $f(x)= \lvert \sin{x} \rvert$. Then find
 $$\sum^\infty_{k=1} \frac{1}{4k^2-1} $$
 
-**(c)** The Riemann Zeta function is defined for $s > 1$ by
+The Fourier cosine series of a function $f(x)$ is $$f(x)=\frac{1}{2}A_0+  \sum_{n=1}^{\infty} A_n \cos{  \left(\frac{n\pi x}{l}\right) }$$
+where 
+$$A_n = \frac{2}{l} \int_{0}^{l}f(x)\cos{  \left(\frac{n\pi x}{l}\right) }dx $$
+Assuming $l=\pi$, we have
+\begin{equation}\label{eqn:3b}
+A_n = \frac{2}{\pi} \int_{0}^{\pi}\lvert \sin{x} \rvert\cos{(nx)}dx
+\end{equation}
+
+But since $\sin$ is non-negative on the interval $0 \leq x \leq \pi$, we can drop the absolute value bars and use the following trigonometric identity to rewrite the integral: $\sin{\alpha}\cos{\beta} = \frac{1}{2} [ \sin{(\alpha + \beta)} + \sin{(\alpha - \beta)}]$
+
+\begin{align*}
+A_n &= \frac{1}{\pi} \int_{0}^{\pi} \sin{((1+n)x)} +\sin{((1-n)x)}dx \\
+    &= \frac{1}{\pi}\left[ \frac{-1}{1+n} \cos{((1+n)x)}\Bigr\rvert_{0}^{\pi} -\frac{1}{1-n}\cos{((1-n)x)}\Bigr\rvert_{0}^{\pi} \right] \\
+    &= \frac{1}{\pi}\left[ \frac{-1}{1+n} [ (-1)^{1+n}-1] -\frac{1}{1-n} [ (-1)^{1-n}-1]\right] \\
+    & = \frac{1}{\pi}  \left[ (-1)^{1+n}-1) \right] \left(\frac{-1}{1+n} -\frac{1}{1-n} \right)
+\end{align*}
+
+Thus, we have found a formula for $A_n$,
+
+$$A_n = \frac{-2}{\pi}  \frac{(-1)^{n}+1}{n^2-1}  $$
+This formula works for all $n \neq 1$
+For $n=0$, 
+$$A_0 = \frac{-2}{\pi}  \frac{(-1)^{0}+1}{0-1}=  \frac{4}{\pi}  $$
+
+For $n=1$, we can go back to (\ref{eqn:3b}) and plug in the appropriate value of $n$:
+$$A_1 = \frac{1}{\pi} \int_{0}^{\pi} \sin{(2x)}\;dx = \frac{1}{2\pi} \cos{(2x)}\Bigr\rvert_{0}^{\pi} \;dx=0 $$
+The Fourier cosine series is therefore
+$$f(x)=\frac{2}{\pi}-\frac{2}{\pi}  \sum_{n=2}^{\infty}   \frac{(-1)^{n}+1}{n^2-1}\cos{(nx) }$$
+
+Now, $(-1)^{n}+1$ will be zero for all odd $n$ and two for all even $n$. So substituting $n=2k$, we gwt
+
+$$\lvert \sin{x} \rvert=\frac{2}{\pi}-\frac{4}{\pi}  \sum_{2k=2}^{\infty}   \frac{1}{4k^2-1}\cos{(2kx) }$$
+
+Starting the index at $2k=2$ is equivalent to starting it at $k=1$, so after making that change and evaluating both sides at $x=0$, we can solve for the sum given in the question:
+$$0=\frac{2}{\pi}-\frac{4}{\pi}  \sum_{k=1}^{\infty}   \frac{1}{4k^2-1}$$
+$$\sum_{k=1}^{\infty}   \frac{1}{4k^2-1}=\frac{1}{2}$$
+
+### c)
+
+The Riemann Zeta function is defined for $s > 1$ by
 $$ \zeta(s)= \sum^\infty_{n=1} \frac{1}{n^s}$$
-
 By computing the Fourier series of $x^2$ over $-\pi<x<\pi$ and using Parseval’s identity, compute
@@
 
-**(d)** Use the Fourier series in 2c and the pointwise convergence theorem to find $\zeta(2)$. Then find
-$$\sum^\infty_{n=1} \frac{1}{(2n-1)^2} $$
+\begin{align*}
+a_0 &=\frac{1}{2\pi} \int_{-\pi}^{\pi}x^2 dx \\
+    &=\frac{1}{2\pi} \left(\frac{x^3}{3}\right) \Bigr\rvert_{-\pi}^{\pi} \\
+    &= \frac{1}{2\pi} \left( \frac{\pi^3}{3} - \frac{-\pi^3}{3}\right) \\
+    &= \frac{1}{2\pi} \left( \frac{2\pi^3}{3} \right) = \frac{\pi^2}{3}
+\end{align*}
 
-## Question 4
-Compute the complex Fourier series of the following functions:
+$$a_n = \frac{1}{\pi} \int_{-\pi}^{\pi}x\cos{(nx)}dx$$
 
-**(a)** Compute the complex Fourier series of $f(x) = e^x$ and show that
+But in part (2b), we found that this evaluates to zero.
 
-$$\coth{\pi}= \frac{1}{\pi}+ \frac{2}{\pi}  \left( \frac{1}{1+1^2} +  \frac{1}{1+2^2} +  \frac{1}{1+3^2} + \;...\right) $$
+<!-- **(d)** Use the Fourier series in 2c and the pointwise convergence theorem to find $\zeta(2)$. Then find -->
+<!-- $$\sum^\infty_{n=1} \frac{1}{(2n-1)^2} $$ -->
 
-**(b)** Find the complex Fourier series of $xe^{ix}$. Then use your result to find the Fourier series of $x \cos{x}$ and $x \sin{x}$.
+<!-- ## Question 4 -->
+<!-- Compute the complex Fourier series of the following functions: -->
 
-## Question 5.
+<!-- **(a)** Compute the complex Fourier series of $f(x) = e^x$ and show that -->
 
-Find the function represented by the new series which is obtained by termwise integration of the following series from $0$ to $x$.
+<!-- $$\coth{\pi}= \frac{1}{\pi}+ \frac{2}{\pi}  \left( \frac{1}{1+1^2} +  \frac{1}{1+2^2} +  \frac{1}{1+3^2} + \;...\right) $$ -->
 
-$$ \sum^\infty_{k=1} (-1)^{k+1} \frac{\cos{kx}}{k}=\log{  \left(2 \cos{\left(\frac{x}{2}\right)}\right) }, \quad -\pi<x<\pi$$
+<!-- **(b)** Find the complex Fourier series of $xe^{ix}$. Then use your result to find the Fourier series of $x \cos{x}$ and $x \sin{x}$. -->
+
+<!-- ## Question 5 -->
+
+<!-- Find the function represented by the new series which is obtained by termwise integration of the following series from $0$ to $x$. -->
+
+<!-- $$ \sum^\infty_{k=1} (-1)^{k+1} \frac{\cos{kx}}{k}=\log{  \left(2 \cos{\left(\frac{x}{2}\right)}\right) }, \quad -\pi<x<\pi$$ -->

### 2024-03-25 PDEs/math245_hw4.Rmd
worked on pdes

@@
 If $A=0$, then $X(0)=0$ and this contradicts the definition of an eigenfunction.
-Therefore, we must have $\sin{(\beta l)}=0$, which can only occur if $\beta = \frac{(2n+1)\pi}{l}$.
+Therefore, we must have $\sin{(\beta l)}=0$, which can only occur if $\beta = \frac{n\pi}{l}$.
 Therefore, this case gives us eigenvalues
 
-$$\lambda_n= \left(\frac{(2n+1)\pi}{l}\right)^2 \quad n=1,2,3... $$
+$$\lambda_n= \left(\frac{n\pi}{l}\right)^2 \quad n=1,2,3... $$
 with eigenfunctions
-$$X_n(x)=\sin{\left(\frac{(2n+1)\pi x}{l}\right)}$$
+$$X_n(x)=\sin{\left( \frac{n \pi x}{l}\right)}$$
 
@@
 If $A=0$, then $X(0)=0$ and this contradicts the definition of an eigenfunction.
-Therefore, we must have $\sin{(\beta)}=0$, which can only occur if $\beta = (2n+1)\pi$.
+Therefore, we must have $\sin{(\beta)}=0$, which can only occur if $\beta = n\pi$.
 Therefore, this case gives us eigenvalues
-$$\lambda_n= (2n+1)^2 \pi^2 \quad n=1,2,3... $$
+$$\lambda_n= n^2 \pi^2 \quad n=1,2,3... $$
 with eigenfunctions
-$$X_n(x)=\sin{\left((2n+1)\pi \ln{x}\right)}$$
+$$X_n(x)=\sin{\left(n\pi \ln{x}\right)}$$
 
@@
 
-Let's start with $a_0$:
+Here, $l=\pi$. Let's start with $a_0$:
 \begin{align*}
@@
     &= 0 + \frac{2}{n^2 \pi} \cos{(nx)}\Bigr\rvert_0^{\pi} \\
+    &= \frac{2}{n^2 \pi} (-1)^{n}-\frac{2}{n^2 \pi} \\
     &= \frac{2}{n^2 \pi} (-1)^{n+1}
@@
 \begin{align*}
-a_n &= \frac{1}{\pi} \int_{-\pi}^{\pi}f(x)\cos{(nx)}dx \\
-    &= \frac{1}{\pi} \int_{-\pi}^{\pi}x\cos{(nx)}+ \frac{1}{\pi} \int_{-\pi}^{\pi}x^2\cos{(nx)}dx 
+a_n &= \frac{1}{\pi} \int_{-\pi}^{\pi}f(x)\cos{(nx)}\;dx \\
+    &= \frac{1}{\pi} \int_{-\pi}^{\pi}x\cos{(nx)}\;dx+ \frac{1}{\pi} \int_{-\pi}^{\pi}x^2\cos{(nx)}\;dx 
 \end{align*}
@@
 Applying integration by parts to the first integral,
-$$\frac{1}{\pi} \int_{-\pi}^{\pi}x\cos{(nx)}=\frac{1}{n\pi}x\sin{(nx)} \Bigr\rvert_{-\pi}^{\pi}+\frac{1}{n\pi} \int_{-\pi}^{\pi}\sin{(nx)}$$
-$$=\frac{1}{n\pi}x\sin{(nx)} \Bigr\rvert_{-\pi}^{\pi}+\frac{-1}{n^2\pi} \cos{(nx)}\Bigr\rvert_{-\pi}^{\pi}=0+0=0$$
+$$\frac{1}{\pi} \int_{-\pi}^{\pi}x\cos{(nx)}\;dx=\frac{1}{n\pi}x\sin{(nx)} \Bigr\rvert_{-\pi}^{\pi}-\frac{1}{n\pi} \int_{-\pi}^{\pi}\sin{(nx)}\;dx$$
+$$=\frac{1}{n\pi}x\sin{(nx)} \Bigr\rvert_{-\pi}^{\pi}-\frac{-1}{n^2\pi} \cos{(nx)}\Bigr\rvert_{-\pi}^{\pi}=0-0=0$$
 
-We then use the integration by parts table method demonstrated in class to find that
+For the second integral, we can use the integration by parts table method demonstrated in class to find that
 
-$$\frac{1}{\pi} \int_{-\pi}^{\pi}x^2\cos{(nx)}= \frac{x^2}{n}\sin{(nx)}\Bigr\rvert_{-\pi}^{\pi} +  \frac{2x}{n^2}\cos{(nx)}\Bigr\rvert_{-\pi}^{\pi}+\frac{2}{n^3}\sin{(nx)}\Bigr\rvert_{-\pi}^{\pi}$$
-The first and last terms (with $\sin{(nx)}$) vanish, while the middle term evaluates to $\frac{4 \cos{(n\pi)}}{n^2}$.
+$$\frac{1}{\pi} \int_{-\pi}^{\pi}x^2\cos{(nx)}\;dx= \frac{x^2}{n\pi}\sin{(nx)}\Bigr\rvert_{-\pi}^{\pi} +  \frac{2x}{n^2\pi}\cos{(nx)}\Bigr\rvert_{-\pi}^{\pi}+
+\frac{2}{n^3\pi}\sin{(nx)}\Bigr\rvert_{-\pi}^{\pi}$$
+The first and last terms (with $\sin{(nx)}$) vanish. Evaluating the middle term, we get:
 
-$$\frac{4 \cos{(n\pi)}}{n^2}=\frac{4}{n^2} (-1)^{n}$$
-Thus, we add this to the result of first integral to find that
+$$\frac{2\pi \cos{(n\pi)}}{n^2 \pi}-\frac{-2\pi \cos{(n\pi)}}{n^2 \pi}=\frac{4 \cos{(n\pi)}}{n^2}=\frac{4}{n^2} (-1)^{n}$$
+Thus, we have found our coefficients $a_n$:
 $$a_n= \frac{4}{n^2} (-1)^{n} $$
+Now for $b_n$:
 \begin{align*}
@@
     &= \frac{2}{\pi}\left[\frac{-1}{n} \cos{(nx)}\Bigr\rvert_{0}^{\pi/2}- \frac{2}{n} \cos{(nx)}\Bigr\rvert_{\pi/2}^{\pi}\right] \\
-    &= \frac{2}{n \pi}\left[\cos{\left(\frac{n\pi}{2}\right)}+1-2\cos{(n\pi)} \right]
+    &= \frac{2}{n \pi}\left[-\cos{\left(\frac{n\pi}{2}\right)}+1-2\cos{(n\pi)}+2\cos{\left(\frac{n\pi}{2}\right)} \right]\\
+    &= \frac{2}{n \pi}\left[1+\cos{\left(\frac{n\pi}{2}\right)}-2\cos{(n\pi)} \right]
 \end{align*}
@@
 
-$$a_n = \frac{1}{\pi} \int_{-\pi}^{\pi}x\cos{(nx)}dx$$
+$$a_n = \frac{1}{\pi} \int_{-\pi}^{\pi}x^2\cos{(nx)}dx$$
 
-But in part (2b), we found that this evaluates to zero.
+But in part (2c), we found that this is equivalent to $\frac{4}{n^2} (-1)^{n}$, so
+$$a_n= \frac{4}{n^2} (-1)^{n} $$
+$$b_n = \frac{1}{\pi} \int_{-\pi}^{\pi}x^2\sin{(nx)}dx$$
+Since $x^2$ is an even function, $b_n=0$.
+
+$$f(x)=\frac{\pi^2}{3}+ 4 \sum_{n=1}^{\infty}  \frac{(-1)^{n}\cos{(nx) } }{n^2} $$
+Now let $x=0$, such that $f(0)=0^2=0$ and $\cos{(nx)}=\cos{(0)}=1$.
+
+$$0=\frac{\pi^2}{3}+ 4 \sum_{n=1}^{\infty}  \frac{(-1)^{n} }{n^2} $$
+$$-\frac{\pi^2}{12}=\sum_{n=1}^{\infty}  \frac{(-1)^{n} }{n^2} $$
 

### 2024-03-29 bio_dynamics/bio274_filing_cabinet.Rmd
biohw

@@
 ```
+
+```{r}
+# Define the rate equation:
+lynx_hare_eq <- c(
+  dHdt ~ r * H - b * H * L,
+  dLdt ~ e * b * H * L - d * L
+)
+
+# Define the parameters (as a named vector):
+lynx_hare_params <- c(r = 2, b = 0.5, e = 0.1, d = 1)
+
+# Define the initial condition (as a named vector):
+lynx_hare_init <- c(H = 1, L = 3)
+
+# Define deltaT and the number of time steps:
+deltaT <- 0.05
+n_steps <- 200
+
+# Compute the solution via Euler's method:
+out_solution <- euler(system_eq = lynx_hare_eq,
+                      parameters = lynx_hare_params,
+                      initial_condition = lynx_hare_init,
+                      deltaT = deltaT,
+                      n_steps = n_steps
+                      )
+
+# Make a plot of the solution,
+# using different colors for lynx or hares:
+ggplot(data = out_solution) +
+  geom_line(aes(x = t, y = H), color = "red") +
+  geom_line(aes(x = t, y = L), color = "blue",linetype='dashed') +
+  labs(
+    x = "Time",
+    y = "Lynx (red) or Hares (blue/dashed)"
+  )
+```
+
+```{r}
+ggplot(data = out_solution) +
+  geom_path(aes(x = L, y = H)) +
+  labs(
+    x = "Lynx",
+    y = "Hares"
+  )
+```
+
+```{r}
+
+lin_eq <- c(
+  dxdt ~ a * x + b * y,
+  dydt ~ c * x + d * y
+)
+
+# Define the parameters (as a named vector):
+lin_params <- c(a = 1, b = 2, c = 3, d = 2)
+
+# Define the initial condition (as a named vector):
+lin_init <- c(x = -0.5, y = 0.5)
+
+# Define deltaT and the number of time steps:
+deltaT <- 0.01
+n_steps <- 200
+
+# Compute the solution via Euler's method:
+out_solution <- euler(system_eq = lin_eq,
+                      parameters = lin_params,
+                      initial_condition = lin_init,
+                      deltaT = deltaT,
+                      n_steps = n_steps
+                      )
+
+ggplot(data = out_solution) +
+  geom_line(aes(x = x, y = y))+
+  xlim(-5,5)+
+  ylim(-5,5)
+```

### 2024-03-29 bio_dynamics/bio274_hw7.Rmd
biohw

@@
 author: "Ruby Krasnow"
-date: "2024-03-24"
+date: "2024-03-28"
 output:
@@
 library(tidyverse)
+library(phaseR)
+library(ggplotify)
+library(demodelr)
 ```
@@
 \end{tcolorbox}
+
+## Phase Portraits
+
+```{r}
+fun <- function(t, y, parameters) {
+  x <- y[1]
+  y <- y[2]
+
+  a <- parameters[1]
+  b <- parameters[2]
+  c <- parameters[3]
+  d <- parameters[4]
+
+  dy <- numeric(2)
+
+  dy[1] <- a*x + b*y
+  dy[2] <- c*x + d*y
+
+  return(list(dy))
+}
+
+```
+
+```{r, echo=FALSE, results='hide'}
+fun2 <- c(dx ~ x+2*y, dy ~3*x+2*y)
+p1 <- phaseplane(fun2,x_var="x",y_var="y", x_window = c(-7, 7), y_window = c(-7, 7))
+
+y0 <- matrix(c(0, 1, 0, 4, -6, 1, 5, 0.5, 0, -3), 5, 2, byrow = TRUE)
+
+y2 <- tibble(x=c(0,0,-6,5,0), y=c(1,4,1,0.5,-3), IC=c("V1", "V2", "V3", "V4", "V5"))%>% mutate(pt = paste0("(",x, ",", y, ")"))
+```
+
+```{r, results='hide', echo=FALSE, message=FALSE, warning=FALSE, fig.keep = "none"}
+traj <- trajectory(fun, y0 = y0,tlim= c(0, 1), 
+                   parameters = c(1, 2, 3, 2), add=FALSE)
+
+t <- c(traj$t)
+
+x <- as_tibble(traj$x) %>% bind_cols("t"=t) %>% 
+  pivot_longer(cols = c(1:5), names_to = "IC", values_to = "x")
+
+y <- as_tibble(traj$y) %>% bind_cols("t"=t) %>% 
+  pivot_longer(cols = c(1:5), names_to = "IC", values_to = "y")
+
+z <- left_join(x,y) %>% filter(x<7 & x>-7 & y<7 & y>-7) %>% left_join(y2 %>% select(IC, pt), by="IC")
+
+```
+
+```{r}
+p1+theme_classic()+
+  geom_abline(slope=-1, color="gray", linewidth=1)+
+  geom_abline(slope=3/2, color="gray", linewidth=1)+
+  geom_path(data=z, aes(x=x, y=y, color=pt), linewidth=1.5)+
+  geom_point(data=y2, aes(x=x,y=y, color=pt), shape=21, fill="white",size=5, stroke=2)+labs(color="IC")
+
+```

### 2024-03-29 bio_dynamics/bio274_hw7.Rmd
phase plane

@@
 fun2 <- c(dx ~ x+2*y, dy ~3*x+2*y)
-p1 <- phaseplane(fun2,x_var="x",y_var="y", x_window = c(-7, 7), y_window = c(-7, 7))
+
+p1 <- phaseplane(fun2, x_var="x" ,y_var="y", x_window = c(-7, 7), y_window = c(-7, 7))
 
@@
 
-y2 <- tibble(x=c(0,0,-6,5,0), y=c(1,4,1,0.5,-3), IC=c("V1", "V2", "V3", "V4", "V5"))%>% mutate(pt = paste0("(",x, ",", y, ")"))
+y0_df <- tibble(x=c(0,0,-6,5,0), y=c(1,4,1,0.5,-3), IC=c("V1", "V2", "V3", "V4", "V5"))%>% mutate(pt = paste0("(",x, ",", y, ")"))
 ```
@@
 
-z <- left_join(x,y) %>% filter(x<7 & x>-7 & y<7 & y>-7) %>% left_join(y2 %>% select(IC, pt), by="IC")
+z <- left_join(x,y) %>% filter(x<7 & x>-7 & y<7 & y>-7) %>% left_join(y0_df %>% select(IC, pt), by="IC")
 
@@
   geom_path(data=z, aes(x=x, y=y, color=pt), linewidth=1.5)+
-  geom_point(data=y2, aes(x=x,y=y, color=pt), shape=21, fill="white",size=5, stroke=2)+labs(color="IC")
+  geom_point(data=y2, aes(x=x,y=y, color=pt), shape=21, fill="white",size=4, stroke=2)+labs(color="IC")
 

### 2024-03-30 bio_dynamics/bio274_hw7.Rmd
finished biohw7

--- 2024-03-24 bio_dynamics/bio274_hw7.Rmd
@@
 author: "Ruby Krasnow"
-date: "2024-03-24"
+date: "2024-03-30"
 output:
@@
 library(tidyverse)
+library(phaseR)
+library(demodelr)
+```
+
+```{r, echo=FALSE}
+# Phase Portrait setup
+y0 <- matrix(c(0, 1, 0, 4, -6, 1, 5, 0.5, 0, -3), 5, 2, byrow = TRUE)
+
+y0_df <- tibble(x=c(0,0,-6,5,0), 
+                y=c(1,4,1,0.5,-3),
+                IC=c("V1", "V2", "V3", "V4", "V5")) %>% 
+  mutate(pt = paste0("(",x, ",", y, ")"))
+
+fun_traj <- function(t, y, parameters) {
+  x <- y[1]
+  y <- y[2]
+  a <- parameters[1]
+  b <- parameters[2]
+  c <- parameters[3]
+  d <- parameters[4]
+  dy <- numeric(2)
+  dy[1] <- a*x + b*y
+  dy[2] <- c*x + d*y  
+  return(list(dy))
+}
+
+fun_plane <- c(dx ~ a*x+b*y, dy ~c*x+d*y)
+
+make_plane <- function(params) {
+  params_unnamed <- unname(params)
+  traj <- trajectory(fun_traj, y0 = y0, tlim= c(0, 1), tstep=0.005,
+                     parameters = params_unnamed, add=FALSE)
+
+  t <- c(traj$t)
+  x <- as_tibble(traj$x) %>% bind_cols("t"=t) %>% 
+    pivot_longer(cols = c(1:5), names_to = "IC", values_to = "x")
+  y <- as_tibble(traj$y) %>% bind_cols("t"=t) %>% 
+    pivot_longer(cols = c(1:5), names_to = "IC", values_to = "y")
+  z <- left_join(x,y, by = join_by(t, IC)) %>% filter(x<7 & x>-7 & y<7 & y>-7) %>% 
+    left_join(y0_df %>% select(IC, pt), by="IC")
+
+  p1 <- phaseplane(fun_plane, x_var="x" ,y_var="y",
+                   x_window = c(-7, 7), y_window = c(-7, 7),
+                   parameters = params)
+
+  A<- matrix(params_unnamed, 2, 2, byrow=TRUE)
+
+  lambda1 <- eigen(A)$values[1] # eigenvalue 1
+  lambda2 <- eigen(A)$values[2] # eigenvalue 2
+
+  if (abs(lambda1) < (1^(-10))) {
+    v1_col <- "black"
+    v1_wid <- 2
+  }
+  else {
+    v1_col <- "gray"
+    v1_wid <- 1
+  }
+
+  if  (abs(lambda2) < (1^(-10))) {
+    v2_col <- "black"
+    v2_wid <- 2
+  }
+  else {
+    v2_col <- "gray"
+    v2_wid <- 1
+  }
+
+  v1_slope <- eigen(A)$vectors[2,1] / eigen(A)$vectors[1,1]
+  v2_slope <- eigen(A)$vectors[2,2] / eigen(A)$vectors[1,2]
+
+  p1+theme_bw()+
+    geom_abline(slope=v1_slope, color=v1_col, linewidth=v1_wid)+
+    geom_abline(slope=v2_slope, color=v2_col, linewidth=v2_wid)+
+    geom_path(data=z, aes(x=x, y=y, color=pt), linewidth=1.5)+
+    geom_point(data=y0_df, aes(x=x,y=y, color=pt), shape=21, fill="white",size=4, stroke=2)+labs(color="IC")
+}
+
+make_plane_complex <- function(params) {
+  params_unnamed <- unname(params)
+  traj <- trajectory(fun_traj, y0 = y0, tlim= c(0, 1), tstep = 0.005,
+                     parameters = params_unnamed, add=FALSE)
+
+  t <- c(traj$t)
+  x <- as_tibble(traj$x) %>% bind_cols("t"=t) %>% 
+    pivot_longer(cols = c(1:5), names_to = "IC", values_to = "x")
+  y <- as_tibble(traj$y) %>% bind_cols("t"=t) %>% 
+    pivot_longer(cols = c(1:5), names_to = "IC", values_to = "y")
+  z <- left_join(x,y, by = join_by(t, IC)) %>% filter(x<7 & x>-7 & y<7 & y>-7) %>% 
+    left_join(y0_df %>% select(IC, pt), by="IC")
+
+  p1 <- phaseplane(fun_plane, x_var="x" ,y_var="y",
+                   x_window = c(-7, 7), y_window = c(-7, 7),
+                   parameters = params)
+
+  p1+theme_bw()+
+    geom_path(data=z, aes(x=x, y=y, color=pt), linewidth=1.5)+
+    geom_point(data=y0_df, aes(x=x,y=y, color=pt), shape=21, fill="white",size=4, stroke=2)+
+    labs(color="IC")
+}
 ```
@@
 
+One positive eigenvalue and one negative eigenvalue gives a saddle:
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE,fig.keep="last"}
+make_plane(c(a=1,b=2,c=3,d=2))
+```
+
 ## Question 2
@@
 
+Two positive eigenvalues gives a source, with solutions going towards the eigenvector $\begin{bmatrix} 2 \\ -1  \end{bmatrix}$ with the eigenvalue $\lambda=3$ of the larger magnitude.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE,fig.keep="last"}
+make_plane(c(a=4,b=2,c=-1,d=1))
+```
+
 ## Question 3
@@
 
+Complex eigenvalues with a positive real part gives a spiral source. Plugging in some test points (e.g., (1,1)) to our original matrix reveals that the spiral rotates counterclockwise.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE,fig.keep="last"}
+make_plane_complex(c(a=3,b=-5,c=5,d=3))
+```
+
 ## Question 4
@@
 
+The zero eigenvalue gives a line of equilibria, emphasized with the black line, with solutions pointing outward (because of the positive eigenvalue 10) parallel to the other eigenvector.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE,fig.keep="last"}
+make_plane(c(a=6,b=-8,c=-3,d=4))
+```
+
 ## Question 5
@@
 \end{tcolorbox}
+
+The positive repeated eigenvalue makes this a source, with solutions going towards the straight-line solution with the same slope (1, because y=x) as $v_1$.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE,fig.keep="last"}
+make_plane(c(a=4,b=-2,c=2,d=0))
+```
 
@@
 \end{tcolorbox}
+Two negative eigenvalues make this a sink, with solutions going towards the eigenvector $\begin{bmatrix} 1 \\ 3 \end{bmatrix}$  with the larger (less negative) eigenvalue.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE,fig.keep="last"}
+make_plane(c(a=-4,b=1,c=3,d=-2))
+```
 
@@
 \end{tcolorbox}
+The zero eigenvalue gives a line of equilibria, with solutions drawn towards the line (because the other eigenvalue is negative) parallel with the other eigenvector.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE,fig.keep="last"}
+make_plane(c(a=-9,b=-3,c=3,d=1))
+```
 
@@
 \end{tcolorbox}
+
+Purely imaginary complex eigenvalues give a center. Testing points with our original matrix reveals a clockwise direction of rotation.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE,fig.keep="last"}
+y1<- matrix(c(0, 1, 
+              10, 4,
+              -6, 10,
+              30, 10,
+              40, 15),
+          5, 2, byrow = TRUE)
+
+y1_df <- tibble(x=c(0,10,-6,30,40), 
+                y=c(1,4,10,10,15),
+                IC=c("V1", "V2", "V3", "V4", "V5")) %>% 
+  mutate(pt = paste0("(",x, ",", y, ")"))
+
+traj8<- trajectory(fun_traj,
+           y0         = y1,
+           tlim       = c(0, 10),
+           parameters = c(-3, 10, -1, 3), add = FALSE)
+
+t8 <- c(traj8$t)
+x8 <- as_tibble(traj8$x) %>% bind_cols("t"=t8) %>% 
+  pivot_longer(cols = c(1:5), names_to = "IC", values_to = "x")
+y8 <- as_tibble(traj8$y) %>% bind_cols("t"=t8) %>% 
+  pivot_longer(cols = c(1:5), names_to = "IC", values_to = "y")
+z8 <- left_join(x8,y8, by = join_by(t, IC)) %>% filter(x<50 & x>-50 & y<40 & y>-40) %>% 
+  left_join(y1_df %>% select(IC, pt), by="IC")
+
+p8 <- phaseplane(fun_plane, x_var="x" ,y_var="y",
+           x_window = c(-50, 50), y_window = c(-40, 40),
+           parameters = c(a=-3, b=10, c=-1, d=3))
+
+p8+theme_bw()+
+  geom_path(data=z8 %>% filter(pt!="(-6,10)"), aes(x=x, y=y, color=pt), linewidth=1.5)+
+  geom_point(data=y1_df%>% filter(pt!="(-6,10)"), aes(x=x,y=y, color=pt), shape=21, fill="white",size=4, stroke=2)+labs(color="IC")
+```

### 2024-03-30 bio_dynamics/test_bio274.R

@@ -1,0 +1,71 @@
+# testing code for bio 274 HW
+
+library(tidyverse)
+
+y0 <- matrix(c(0, 1, 0, 4, -6, 1, 5, 0.5, 0, -3), 5, 2, byrow = TRUE)
+
+y0_df <- tibble(x=c(0,0,-6,5,0), 
+                y=c(1,4,1,0.5,-3),
+                IC=c("V1", "V2", "V3", "V4", "V5")) %>% 
+  mutate(pt = paste0("(",x, ",", y, ")"))
+
+fun_traj <- function(t, y, parameters) {
+  x <- y[1]
+  y <- y[2]
+  a <- parameters[1]
+  b <- parameters[2]
+  c <- parameters[3]
+  d <- parameters[4]
+  dy <- numeric(2)
+  dy[1] <- a*x + b*y
+  dy[2] <- c*x + d*y  
+  return(list(dy))
+}
+
+fun_plane <- c(dx ~ a*x+b*y, dy ~c*x+d*y)
+
+make_plane <- function(params) {
+  params_unnamed <- unname(params)
+  traj <- trajectory(fun_traj, y0 = y0, tlim= c(0, 1), tstep = 0.005,
+                     parameters = params_unnamed, add=FALSE)
+  
+  t <- c(traj$t)
+  x <- as_tibble(traj$x) %>% bind_cols("t"=t) %>% 
+    pivot_longer(cols = c(1:5), names_to = "IC", values_to = "x")
+  y <- as_tibble(traj$y) %>% bind_cols("t"=t) %>% 
+    pivot_longer(cols = c(1:5), names_to = "IC", values_to = "y")
+  z <- left_join(x,y, by = join_by(t, IC)) %>% filter(x<7 & x>-7 & y<7 & y>-7) %>% 
+    left_join(y0_df %>% select(IC, pt), by="IC")
+  
+  p1 <- phaseplane(fun_plane, x_var="x" ,y_var="y",
+                   x_window = c(-7, 7), y_window = c(-7, 7),
+                   parameters = params)
+  
+  A<- matrix(params_unnamed, 2, 2, byrow=TRUE)
+  v1_slope <- eigen(A)$vectors[2,1] / eigen(A)$vectors[1,1]
+  v2_slope <- eigen(A)$vectors[2,2] / eigen(A)$vectors[1,2]
+  
+  p1+
+    theme_bw()+
+    geom_abline(slope=v2_slope, color="gray40", linewidth=1)+
[…]

### 2024-04-01 PDEs/math245_hw4.Rmd
tentative finish pde

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-03-24"
+date: "2024-04-01"
 output:
@@
 
+Since $f(x)$ and $f'(x)$ are both piecewise continuous, the Fourier series converges pointwise. However, since $f(x)$ is not continuous (only piecewise), it does not converge uniformly.
+
 ### (b)
@@
 $$ f(x) = \frac{\pi}{2} + \frac{2}{\pi} \sum_{n=1}^{\infty} \left[ \frac{1}{n^2} (-1)^{n+1} \cos{(nx)} \right]$$
+Since $f(x)$ and $f'(x)$ are both piecewise continuous, the Fourier series converges pointwise.
+Also, $f(x)$ is continuous and $f'(x)$ is piecewise continuous.
+Finally, the periodic boundary conditions are satisfied because $f(-\pi)=\pi=f(\pi)$.
+Therefore, the Fourier series also converges uniformly.
 
@@
 $$f(x)=\frac{\pi^2}{3} + 4 \sum_{n=1}^{\infty} \frac{(-1)^{n}\cos{ (nx) }}{n^2}  $$
+Here, $f(x)$ and $f'(x)$ are both continuous, so the Fourier series converges piecewise.
+However, $f(-\pi)=\pi^2-\pi$ while $f(\pi)=\pi^2+\pi$, so the periodic boundary conditions are not satisfied and the Fourier series does not converge uniformly.
 
@@
 
-The Fourier sine series of a function $f(x)$ is $$f(x)= \sum_{n=1}^{\infty} A_n \sin{  \left(\frac{n\pi x}{l}\right) }$$
+The Fourier sine series of a function $f(x)$ is $$f(x)= \sum_{n=1}^{\infty} B_n \sin{  \left(\frac{n\pi x}{l}\right) }$$
 where 
-$$A_n = \frac{2}{l} \int_{0}^{l}f(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx $$
-$$A_n = \frac{2}{\pi} \int_{0}^{\pi}f(x)\sin{(nx)}dx $$
+$$B_n = \frac{2}{l} \int_{0}^{l}f(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx $$
+$$B_n = \frac{2}{\pi} \int_{0}^{\pi}f(x)\sin{(nx)}dx $$
 \begin{align*}
-A_n &= \frac{2}{\pi}\left[ \int_{0}^{\pi/2}\sin{(nx)}dx + 2 \int_{\pi/2}^{\pi}\sin{(nx)}dx \right] \\
+B_n &= \frac{2}{\pi}\left[ \int_{0}^{\pi/2}\sin{(nx)}dx + 2 \int_{\pi/2}^{\pi}\sin{(nx)}dx \right] \\
     &= \frac{2}{\pi}\left[\frac{-1}{n} \cos{(nx)}\Bigr\rvert_{0}^{\pi/2}- \frac{2}{n} \cos{(nx)}\Bigr\rvert_{\pi/2}^{\pi}\right] \\
@@
  $$f(x)=  \frac{2}{\pi} \sum_{n=1}^{\infty} \frac{1}{n}\left[1+\cos{\left(\frac{n\pi}{2}\right)}-2\cos{(n\pi)} \right] \sin{(nx) }$$
+ $$f(x)=  \frac{2}{\pi} \sum_{n=1}^{\infty} \frac{1}{n}\left[1+\cos{\left(\frac{n\pi}{2}\right)}+2 (-1)^{n+1} \right] \sin{(nx) }$$
+$$f(x)= \sum_{n_{odd}}+\sum_{n_{even}}$$
+\begin{align*}
+\sum_{n_{even}} &=\frac{2}{\pi} \sum_{k=1}^{\infty} \frac{1}{2k}\left[1+\cos{\left(\frac{2k\pi}{2}\right)}+
+ 2 (-1)^{2k+1} \right] \sin{(2kx) } \\
+ &=\frac{2}{\pi} \sum_{k=1}^{\infty} \frac{1}{2k}\left[1+(-1)^n-2 \right] \sin{(2kx) } \\
+ &=\frac{2}{\pi} \sum_{k=1}^{\infty} \frac{(-1)^{n+1}}{2k}\sin{(2kx) }
+ \end{align*}
+
+\begin{align*}
+\sum_{n_{odd}}& =
+ \frac{2}{\pi} \sum_{k=1}^{\infty} \frac{1}{2k-1}\left[1+\cos{\left(\frac{(2k-1)\pi}{2}\right)}+2 (-1)^{2k-1+1} \right] \sin{((2k-1)x) } \\
+&= \frac{2}{\pi} \sum_{k=1}^{\infty} \frac{3}{2k-1}\sin{((2k-1)x) }
+ \end{align*}
+
+ $$f(x)=\frac{2}{\pi} \sum_{k=1}^{\infty} \frac{(-1)^{n+1}}{2k}\sin{(2kx) }+\frac{2}{\pi} \sum_{k=1}^{\infty} \frac{3}{2k-1}\sin{((2k-1)x) }$$
 
@@
 $$f(x)=\frac{\pi^2}{3}+ 4 \sum_{n=1}^{\infty}  \frac{(-1)^{n}\cos{(nx) } }{n^2} $$
-Now let $x=0$, such that $f(0)=0^2=0$ and $\cos{(nx)}=\cos{(0)}=1$.
+Parseval's Identity says that
+$$ \frac{1}{2\pi}\int_{-\pi}^{\pi}(x^2)^2\;dx=a_0^2+\frac{1}{2}\sum_{n=1}^{\infty} (a_n^2+b_n^2)$$
 
-$$0=\frac{\pi^2}{3}+ 4 \sum_{n=1}^{\infty}  \frac{(-1)^{n} }{n^2} $$
-$$-\frac{\pi^2}{12}=\sum_{n=1}^{\infty}  \frac{(-1)^{n} }{n^2} $$
+$$ \frac{1}{2\pi}\int_{-\pi}^{\pi}(x^2)^2\;dx= \left(\frac{\pi^2}{3}\right)^2+ \frac{1}{2}\sum_{n=1}^{\infty}  \left( \frac{4(-1)^{n} }{n^2}\right)^2  $$
 
-<!-- **(d)** Use the Fourier series in 2c and the pointwise convergence theorem to find $\zeta(2)$. Then find -->
-<!-- $$\sum^\infty_{n=1} \frac{1}{(2n-1)^2} $$ -->
+$$ \frac{1}{\pi}\int_{0}^{\pi}x^4\;dx= \frac{\pi^4}{9}+ \sum_{n=1}^{\infty} \frac{8}{n^4} $$
+$$ \frac{1}{\pi} \frac{x^5}{5} \Bigr\rvert_{0}^{\pi}= \frac{\pi^4}{9}+ 8 \sum_{n=1}^{\infty} \frac{1 }{n^4}$$
+$$  \frac{\pi^4}{5}= \frac{\pi^4}{9}+ 8 \sum_{n=1}^{\infty} \frac{1}{n^4}$$
+$$ \frac{4\pi^4}{45}= 8\sum_{n=1}^{\infty} \frac{1}{n^4}$$
+Thus,
+$$\zeta(4)= \sum_{n=1}^{\infty} \frac{1}{n^4}=\frac{\pi^4}{90}$$
 
-<!-- ## Question 4 -->
-<!-- Compute the complex Fourier series of the following functions: -->
+**(d)** Use the Fourier series in 2c and the pointwise convergence theorem to find $\zeta(2)$. Then find
+$$\sum^\infty_{n=1} \frac{1}{(2n-1)^2} $$
 
-<!-- **(a)** Compute the complex Fourier series of $f(x) = e^x$ and show that -->
+Take the following Fourier series and set $x=\pi$.
+$$f(x)=x^2=\frac{\pi^2}{3}+ 4 \sum_{n=1}^{\infty}  \frac{(-1)^{n}\cos{(nx) } }{n^2} $$
 
-<!-- $$\coth{\pi}= \frac{1}{\pi}+ \frac{2}{\pi}  \left( \frac{1}{1+1^2} +  \frac{1}{1+2^2} +  \frac{1}{1+3^2} + \;...\right) $$ -->
+$$\pi^2=\frac{\pi^2}{3}+ 4 \sum_{n=1}^{\infty}  \frac{1 }{n^2} $$
+$$\frac{2\pi^2}{3}=4 \sum_{n=1}^{\infty}  \frac{1 }{n^2} $$
 
-<!-- **(b)** Find the complex Fourier series of $xe^{ix}$. Then use your result to find the Fourier series of $x \cos{x}$ and $x \sin{x}$. -->
+Thus, 
+$$\zeta(2)= \sum_{n=1}^{\infty} \frac{1}{n^2}=\frac{\pi^2}{6}$$
+Now, we want to find 
+$$\sum_{n=1}^{\infty} \frac{1}{(2n-1)^2} = \frac{1}{1}+\frac{1}{9}+\frac{1}{25}+\frac{1}{49}+ \;...$$
+Let's also write out some terms in the series representation of $\zeta(2)$:
+$$\sum_{n=1}^{\infty} \frac{1}{n^2}=\frac{1}{1}+\frac{1}{4}+\frac{1}{9}+\frac{1}{16}+\frac{1}{25}+\frac{1}{36}+\frac{1}{49}+ \;...$$
+We can see that the top series is simply the bottom series without the terms where the denominator is the square of an even integer; which we can write as:
+\begin{align*}
+\sum_{n=1}^{\infty} \frac{1}{(2n-1)^2} 
+&=\sum_{n=1}^{\infty} \frac{1}{n^2}-\sum_{n=1}^{\infty} \frac{1}{(2n)^2} \\
+&=\sum_{n=1}^{\infty} \frac{1}{n^2}-\frac{1}{4}\sum_{n=1}^{\infty} \frac{1}{n^2}\\
+&=\frac{\pi^2}{6}-\frac{1}{4}\frac{\pi^2}{6} \\
+&=\frac{\pi^2}{8}
+\end{align*}
+Therefore, we conclude that $$\sum_{n=1}^{\infty} \frac{1}{(2n-1)^2} =\frac{\pi^2}{8}$$
 
-<!-- ## Question 5 -->
+## Question 4
+Compute the complex Fourier series of the following functions:
 
-<!-- Find the function represented by the new series which is obtained by termwise integration of the following series from $0$ to $x$. -->
+### (a)   
+Compute the complex Fourier series of $f(x) = e^x$ and show that
 
-<!-- $$ \sum^\infty_{k=1} (-1)^{k+1} \frac{\cos{kx}}{k}=\log{  \left(2 \cos{\left(\frac{x}{2}\right)}\right) }, \quad -\pi<x<\pi$$ -->
+$$\coth{\pi}= \frac{1}{\pi}+ \frac{2}{\pi}  \left( \frac{1}{1+1^2} +  \frac{1}{1+2^2} +  \frac{1}{1+3^2} + \;...\right) $$
+
+The complex Fourier series of a function $f(x)$ is
+$$f(x)=\sum_{n=-\infty}^{\infty} c_ne^{in\pi x / l}$$
+where $$c_n= \frac{1}{2l} \int_{-l}^{l}f(x) e^{-in\pi x / l}\;dx $$
+
+\begin{align*}
+c_n&= \frac{1}{2\pi} \int_{-\pi}^{\pi}e^xe^{-in x}\;dx \\
+  &= \frac{1}{2\pi} \int_{-\pi}^{\pi}e^{x(1-in)}\;dx \\
+  &= \frac{1}{2\pi} \frac{1}{1-in}e^{x(1-in)}\Bigr\rvert_{-\pi}^{\pi} \\
+  & = \frac{1}{2\pi} \frac{1}{1-in}\left[e^{\pi}e^{-in\pi}-e^{-\pi}e^{in\pi}\right]
+\end{align*}
+and since $e^{in\pi}=\cos{(n\pi)}+i\sin{(n\pi)}=(-1)^n=\cos{(-n\pi)}+i\sin{(-n\pi)}= e^{-in\pi}$,
+
+\begin{align*}
+c_n& = \frac{1}{2\pi} \frac{(-1)^n}{1-in}\left[e^{\pi}-e^{-\pi}\right]\\
+&= \frac{1}{\pi} \frac{(-1)^n}{1-in} \sinh{\pi}
+\end{align*}
+$$f(x)=e^x=\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{(-1)^n}{1-in} (\sinh{\pi})\,e^{inx}$$
+
+Set $x=\pi$ and $x=-\pi$ and then add the two expressions, noting that $e^{in\pi}=e^{-in\pi}=(-1)^n$ and $(-1)^n(-1)^n=(-1)^{2n}=1$.
+$$e^\pi=\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{1}{1-in} (\sinh{\pi})$$
+
+$$e^{-\pi}=\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{1}{1-in} (\sinh{\pi})$$
+
+$$\frac{e^\pi+e^{-\pi}}{2}=\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{1}{1-in} (\sinh{\pi})$$
+But note that $\frac{e^\pi+e^{-\pi}}{2}=\cosh{\pi}$.
+
+$$\cosh{\pi}=\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{1}{1-in} (\sinh{\pi})$$
+Divide both sides by $\sinh{\pi}$:
+$$\coth{\pi}=\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{1}{1-in}$$
+Multiply the right-hand side by $\frac{1+in}{1+in}$:
+$$\coth{\pi}=\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{1+n}{1+n^2}$$
+
+$$=\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{1}{1+n^2}+\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{n}{1+n^2}$$
+The second sum will disappear because the $n$ term in the numerator means that all negative terms in the series will cancel with the corresponding positive $n$ terms, while the expression evaluates to zero at $n=0$.
+Thus,
+\begin{align*}
+\coth{\pi}&=\sum_{n=-\infty}^{\infty} \frac{1}{\pi} \frac{1}{1+n^2} \\
+&=\frac{1}{\pi}+ \frac{2}{\pi}  \left( \frac{1}{1+1^2} +  \frac{1}{1+2^2} +  \frac{1}{1+3^2} +  \;...\right)
+\end{align*}
+which is what we initially sought to show.
+
+### (b)  
+Find the complex Fourier series of $xe^{ix}$. Then use your result to find the Fourier series of $x \cos{x}$ and $x \sin{x}$.
+
+The complex Fourier series of a function $f(x)$ is
+$$f(x)=\sum_{n=-\infty}^{\infty} c_ne^{in\pi x / l}$$
+where $$c_n= \frac{1}{2l} \int_{-l}^{l}f(x) e^{-in\pi x / l}\;dx $$
+
+ \begin{align*}
+ c_n &= \frac{1}{2\pi} \int_{-\pi}^{\pi}xe^{ix}e^{-in x}\;dx \\
+     &= \frac{1}{2\pi} \int_{-\pi}^{\pi}xe^{ix(1-n)}\;dx
+  \end{align*}
+
+Using integration by parts,
+  $$u=x, \quad du=dx,\quad dv=e^{ix(1-n)}\,dx,\quad v=\frac{1}{i(1-n)}e^{ix(1-n)} $$
+$$\frac{1}{2\pi} \int_{-\pi}^{\pi}xe^{ix(1-n)}\;dx = \frac{1}{2\pi} \left[\frac{x}{i(1-n)}e^{ix(1-n)}\right]^{\pi}_{-\pi}-\frac{1}{2\pi} \int_{-\pi}^{\pi}\frac{1}{i(1-n)}e^{ix(1-n)}\;dx$$
+First, let's evaluate the first term on the right-hand side:
+\begin{align*}
+\frac{1}{2\pi} \left[\frac{x}{i(1-n)}e^{ix(1-n)}\right]^{\pi}_{-\pi} 
+&= \frac{1}{2\pi} \left[ \frac{\pi}{i(1-n)} e^{i \pi (1-n)} -\frac{-\pi}{i(1-n)} e^{-i \pi (1-n)}\right]  \\
+&= \frac{1}{2\pi} \left[ \frac{\pi}{i(1-n)} e^{i \pi}e^{-i \pi n}  -\frac{-\pi}{i(1-n)} e^{-i \pi} e^{i \pi n} \right]  \\
+&= \frac{1}{2\pi} \left[ \frac{\pi}{i(1-n)} (-1)(-1)^n  -\frac{-\pi}{i(1-n)} (-1)(-1)^n \right] \\
+&= \frac{1}{2\pi} \left[ \frac{2\pi}{i(1-n)} (-1)^{n+1}  \right] \\
+&= \frac{(-1)^{n+1}} {i(1-n)} 
+\end{align*}
+
+Now we turn to the integral on the right-hand side:
+\begin{align*}
+-\frac{1}{2\pi} \int_{-\pi}^{\pi}\frac{1}{i(1-n)}e^{ix(1-n)}\;dx &= \frac{1}{2\pi} \left[\frac{e^{ix(1-n)}}{(1-n)^2}\right]^{\pi}_{-\pi}  \\
+&= \frac{1}{2\pi} \left[ \frac{e^{i\pi (1-n)}}{(1-n)^2}  -\frac{e^{-i \pi (1-n)}}{(1-n)^2}\right] \\
+&= \frac{1}{2\pi} \left[ \frac{(-1)^{n+1}}{(1-n)^2}  -\frac{(-1)^{n+1}}{(1-n)^2}\right] \\
+&=0 
+\end{align*}
+
+Thus, $c_n=\frac{(-1)^{n+1}}{i(1-n)}$. Plugging this into the equation for a complex Fourier series, we find that
+
+$$xe^{ix}=\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} e^{inx}$$
+
+Now, we use Euler's identity
+$$e^{ix}=\cos{x}+i\sin{x}$$
+$$\cos{x}=\frac{e^{ix}+e^{-ix}}{2}$$
+\begin{align*}
+x \cos{x} &=  \frac{x\left( e^{ix}+ e^{-ix} \right) } {2} \\
+          &= \frac{x}{2}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} e^{inx} + \frac{x}{2}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} e^{-inx}\\
+          &= \frac{x}{2}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} (\cos{nx}+i\sin{nx}) + \frac{x}{2}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)}(\cos{(-nx)}+i\sin{(-nx)})  \\
+          &=\frac{x}{2}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} (\cos{nx}+i\sin{nx}) + \frac{x}{2}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)}(\cos{(nx)}-i\sin{(nx)}) 
+\end{align*}
+
+The sin terms cancel and the cos terms add, yielding
+$$x \cos{x} =x \sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} \cos{nx} $$
+Similarly,
+
+\begin{align*}
+\sin{x} &=\frac{e^{ix}-e^{-ix}}{2i} \\
+        &= \frac{x}{2i}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} e^{inx} - \frac{x}{2i}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} e^{-inx} \\
+        &= \frac{x}{2i}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} (\cos{nx}+i\sin{nx}) - \frac{x}{2i}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)}(\cos{(-nx)}+i\sin{(-nx)})  \\
+        &=\frac{x}{2i}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)} (\cos{nx}+i\sin{nx}) - \frac{x}{2i}\sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)}(\cos{(nx)}-i\sin{(nx)})
+\end{align*}
+
+The $\cos$ terms cancel and the $\sin$ terms add, yielding
+$$\frac{x}{i} \sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {i(1-n)}(- \sin{nx}) $$
+Multiplying the two $i$s together cancels the negative sign, meaning the Fourier series becomes:
+$$x \sin{x} =x \sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {(1-n)}\sin{nx} $$
+
+## Question 5
+
+Find the function represented by the new series which is obtained by term-wise integration of the following series from $0$ to $x$.
+
+$$ \sum^\infty_{k=1} (-1)^{k+1} \frac{\cos{kx}}{k}=\log{  \left(2 \cos{\left(\frac{x}{2}\right)}\right) }, \quad -\pi<x<\pi$$
+
+$$ \int_{0}^{x} \sum^\infty_{k=1} (-1)^{k+1} \frac{\cos{kx}}{k}= \sum^\infty_{k=1} (-1)^{k+1} \frac{\sin{kx}}{k^2}$$
+
+$$\left[ \sum^\infty_{k=1} (-1)^{k+1} \frac{\sin{kx}}{k^2} \right]\Bigr\rvert_{0}^{x}= \sum^\infty_{k=1} (-1)^{k+1} \frac{\sin{kx}}{k^2}$$
+Therefore,
+$$\int_{0}^{x} \log{  \left(2 \cos{\left(\frac{x}{2}\right)}\right) }= \sum^\infty_{k=1} (-1)^{k+1} \frac{\sin{kx}}{k^2}$$

### 2024-04-03 PDEs/math245_hw4.Rmd
finished pde 4

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-04-01"
+date: "2024-04-02"
 output:
@@
 $X'(0)=\beta A=0$.
-And since we are in a case where $\beta \neq 0$, this means $B=0$ so $X(x)=A\sin{(\beta x)}$.
-Now we use our boundary condition $X(l)=0$ to get $X(l)=A\sin{(\beta l)}=0$.
-If $A=0$, then $X(0)=0$ and this contradicts the definition of an eigenfunction.
-Therefore, we must have $\sin{(\beta l)}=0$, which can only occur if $\beta = \frac{n\pi}{l}$.
+And since we are in a case where $\beta \neq 0$, this means $A=0$ so $X(x)=B\cos{(\beta x)}$.
+Now we use our boundary condition $X(l)=0$ to get $X(l)=B\cos{(\beta l)}=0$.
+If $B=0$, then $X(0)=0$ and this contradicts the definition of an eigenfunction.
+Therefore, we must have $\cos{(\beta l)}=0$, which can only occur if $\beta = \frac{(2n-1)\pi}{2l}$.
 Therefore, this case gives us eigenvalues
 
-$$\lambda_n= \left(\frac{n\pi}{l}\right)^2 \quad n=1,2,3... $$
+$$\lambda_n= \left(\frac{(2n-1)\pi}{2l}\right)^2 \quad n=1,2,3... $$
 with eigenfunctions
-$$X_n(x)=\sin{\left( \frac{n \pi x}{l}\right)}$$
+$$X_n(x)=\cos{\left( \frac{(2n-1) \pi x}{2l}\right)}$$
 
@@
 Now plugging in our initial condition $X(1) = 0$, we get $A+B=0$, which means $X(x)=Ax-A$.
-But from the boundary condition $X(e)=0$, we get $A(e-1)=0$ which is only possible if $A=0$ and accordingly $X(0)=0$.
+But from the boundary condition $X(e)=0$, we get $A(e-1)=0$ which is only possible if $A=0$ and accordingly $B=-A=0$.
 Therefore, there are no eigenfunctions $X(x)$ that satisfy the boundary conditions when $X''=0$ and hence no zero eigenvalues.
@@
 $$X(x)=Ax^{\beta }-Ax^{-\beta}$$
-Then the other boundary condition gives $X(e)=Ae^{\beta}-Ae^{-\beta }=0$.
-Since $e^{\beta}$ and $e^{-\beta}$ are always nonzero, we must have $A=0$, implying that again $X(x)=0$.Thus, this problem has only positive eigenvalues and their associated eigenfunctions as found in Case 1.
+Then the other boundary condition gives $X(e)=Ae^{\beta}-Ae^{-\beta }=0 \quad\Rightarrow A(e^{\beta}-e^{-\beta })=0$.
+If we had $e^{\beta}-e^{-\beta }=0$, that would imply $\beta=-\beta\quad\Rightarrow\beta=0$, but we are in a case when we defined $\beta$ to be nonzero.
+So, we must have $A=0$, implying that again $X(x)=0$.
+Thus, this problem has only positive eigenvalues and their associated eigenfunctions as found in Case 1.
 
@@
 $$A\sin{(\beta)}=-B\cos{(\beta)}$$
+Before dividing by $A$, we need to verify that $A \neq 0$. In this case, if $A = 0$,
+then $X=B \cos{\beta x}$, so $X'(0)+X(0)=0+B=0$ implies $B=0$, a contradiction.
+Similarly, $\cos{\beta}\neq 0$, since if $\cos{\beta}=0$, then $\sin{\beta} \neq 0$ (sin and cos are not zero at the same time), and so
+$A \sin{\beta}=-B\cos{\beta}=0 \quad\Rightarrow A=0$, leading to the same contradiction as before.
 
@@
 Now turning to our other boundary condition,
-$$X'(1)=\beta Ae^{\beta}-\beta Be^{-\beta}=0$$
-$$\beta Ae^{\beta}=\beta Be^{-\beta}$$
-$$-\frac{B}{A} = -e^{2\beta}$$
+$$X(1)=Ae^{\beta}+Be^{-\beta }=0$$
+$$Ae^{\beta}=-Be^{-\beta}$$
+$$-\frac{B}{A} = e^{2\beta}$$
+
+We know $A \neq= 0$, since otherwise $X(1)=Be^{-\beta }=0$ implies $B = 0$, a contradiction.
+Now you also can’t have $1-\beta = 0$, since $\beta = 1$ leads to $-A(1+1) = -2A = 0$. This justifies all the division so far.
+
 This would give us the following equation:
-$$\frac{\beta+1}{1-\beta}=-e^{2\beta}$$
+$$\frac{\beta+1}{1-\beta}=e^{2\beta}$$
 But the only time that this equality holds true is when $\beta=0$ (Fig.
@@
 a_0 &=\frac{1}{2\pi} \int_{-\pi}^{\pi}f(x)dx \\
-    &=\frac{1}{2\pi}\left[ \int_{-\pi}^{-1} 1\;dx + \int_{-1}^{1} (1-\lvert x \rvert) dx+ \int_{1}^{\pi} 1\;dx  \right] \\
-    &= \frac{1}{2\pi}\int_{-\pi}^{\pi}1\;dx - \int_{-1}^{1} \lvert x \rvert\, dx \\
-    &= 1- 2\int_{0}^{1} x \; dx \\
-    &= 1- 1=0
+    &=\frac{1}{\pi} \int_{0}^{\pi}f(x)dx \\
+    &= \frac{1}{\pi} \left(\int_{0}^{1}(1-\lvert x \rvert )\;dx +   \int_{1}^{\pi}\;dx \right) \\
+    &= \frac{1}{\pi} \left(\int_{0}^{1}(1-x )\;dx +  (\pi-1) \right) \\
+    &= \frac{1}{\pi} \left(\int_{0}^{1}(1-x )\;dx +  (\pi-1) \right) \\
+    &= \frac{1}{\pi} \left((x-x^2/2) \Bigr\rvert_{0}^{1} +  (\pi-1) \right) \\
+    &= \frac{1}{\pi} \left( \frac{1}{2} +  (\pi-1) \right) \\
+    &= \frac{1}{\pi} \left( \pi-\frac{1}{2} \right) \\
 \end{align*}
 
-where we used the evenness of $\lvert x \rvert$ to rewrite $\int_{-1}^{1} \lvert x \rvert \,dx$ as $2\int_{0}^{1} x \; dx$. Thus, $a_0=0$. Now we perform a similar procedure to find $a_n$:
+where we used the evenness of $\lvert x \rvert$ to rewrite $\int_{-\pi}^{\pi} f(x)dx \,dx$ as $2\int_{0}^{\pi} f(x)dx \; dx$.
+Thus, $a_0=\frac{1}{\pi} \left( \pi-\frac{1}{2} \right)$.
+Now we perform a similar procedure to find $a_n$:
 
@@
 
-Now inserting the coefficients $a_n$ we found above, we get the Fourier series:
-$$f(x)=\frac{2}{\pi} \sum_{n=1}^{\infty}  \left(\frac{\sin{n}}{n}+ \frac{\cos{n}+1}{n^2} \right) \cos{ \left( nx \right)}$$
+Now inserting the coefficients we found above, we get the Fourier series:
+$$f(x)=\frac{1}{\pi} \left( \pi-\frac{1}{2} \right)+\frac{2}{\pi} \sum_{n=1}^{\infty}  \left(\frac{\sin{n}}{n}+ \frac{\cos{n}+1}{n^2} \right) \cos{ \left( nx \right)}$$
 
@@
 $$x \sin{x} =x \sum_{n=-\infty}^{\infty}  \frac{(-1)^{n+1}} {(1-n)}\sin{nx} $$
+**Note that if $n = 1$, then $c_n = 0$, since you end up with the integral of $x$ over $(-\pi, \pi)$, which is zero. Hence the sum is properly over all terms with $n\neq 1$.
 
@@
 $$\int_{0}^{x} \log{  \left(2 \cos{\left(\frac{x}{2}\right)}\right) }= \sum^\infty_{k=1} (-1)^{k+1} \frac{\sin{kx}}{k^2}$$
+
+If you wanted us to differentiate rather than integrate:
+$$\frac{d}{dx} \sum^\infty_{k=1} (-1)^{k+1} \frac{\cos{kx}}{k}=\frac{d}{dx}\log{  \left(2 \cos{\left(\frac{x}{2}\right)}\right) }$$
+
+\begin{align*}
+\sum^\infty_{k=1} (-1)^{k}\sin{kx} &=\frac{-\sin{\frac{x}{2}}}{2 \cos{\frac{x}{2}}} \\
+&= -\frac{1}{2}\tan{\frac{x}{2}}
+\end{align*}
+
+Thus, 
+$$\sum^\infty_{k=1} (-1)^{k}\sin{kx}=-\frac{1}{2}\tan{\frac{x}{2}}$$
+
+Note that this sum is actually divergent, since the terms don’t go to zero for any $x$ apart from a few special points (like $x = 0$).

### 2024-04-03 bio_dynamics/bio274_hw8.Rmd
finished pde 4

---
title: "BIOL 274 Homework 8"
author: "Ruby Krasnow"
date: "2024-04-02"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---
# Visualizing Solutions of Linear Systems and Runga Kutta

### 2024-04-05 PDEs/math245_hw5.Rmd
started 274 hw8

---
title: "MATH 245 Homework 5"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-04-05"
output:
  pdf_document:
    extra_dependencies: ["tcolorbox", "xcolor"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1:

### 2024-04-05 bio_dynamics/bio274_hw8.Rmd
started 274 hw8

@@
 author: "Ruby Krasnow"
-date: "2024-04-02"
+date: "2024-04-05"
 output:
@@
 ---
-# Visualizing Solutions of Linear Systems and Runga Kutta
+# Visualizing Solutions of Linear Systems and Runge-Kutta
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+library(tidyverse)
+library(phaseR)
+library(demodelr)
+```
+
+## Question 1  
+Consider the epidemic model presented in class.
+
+### a.  
+For the parameters given, we found one single equilibrium point. Determine whether this equilibrium point is stable or unstable and sketch the phase plane near the equilibrium point.
+
+Let $X$ be the number of susceptible individuals at time $t$ and $Y$ be the number of infected individuals (who can transmit the disease).
+The epidemic model presented in class takes the following form:
+$$\frac{dX}{dt}=-\mu XY- \rho $$
+$$\frac{dY}{dt}=\mu XY- \nu Y $$
+where $\mu=0.025,\; \rho=5,$ and $\nu =0.5$. We determined that given these parameters, the model has a single equilibrium point at $\left( \frac{\nu}{\mu}, \frac{\rho}{\nu}\right)=(20,10)$.
+
+We will determine the local stability of this equilibrium point by first computing the Jacobian matrix at $(20,10)$.
+
+Given a system of the type $\dot{x}=f(x,y), \quad \dot{y}=g(x,y)$, the Jacobian is defined as the following matrix:
+$$J(x,y)= \begin{bmatrix} 
+    \frac{\partial{f}}{\partial{x}} &  \frac{\partial{f}}{\partial{y}} \\
+    \frac{\partial{g}}{\partial{x}} & \frac{\partial{g}}{\partial{y}} \end{bmatrix} $$
+For our system, the Jacobian is
+$$J(X,Y)= \begin{bmatrix} 
+    -\mu Y &  -\mu X \\
+    \mu Y & -\mu X-\nu \end{bmatrix} $$
+
+We then evaluate the Jacobian at our equilibrium point:
+$$J(20,10)= \begin{bmatrix} 
+    -0.025(10) &  -0.025(20) \\
+    0.025(10) & 0.025(20)-0.5 \end{bmatrix}= \begin{bmatrix} 
+    -0.25 &  -0.5 \\
+    0.25 & 0 \end{bmatrix} $$
+
+Our next step is to find the eigenvalues of this matrix.
+\begin{equation*}
+\det{\begin{bmatrix} 
+    -0.25-\lambda & -0.5 \\
+    0.25 & 0-\lambda \end{bmatrix}}=0 \\
+\end{equation*}
+
+$$ \left(-0.25-\lambda\right) \left(-\lambda\right)+0.125=0 $$
+$$ \lambda^2+0.25\lambda +0.125=0 $$
+$$\frac{-b\pm\sqrt{b^2-4ac}}{2a} \quad \longrightarrow \quad \frac{-0.25\pm\sqrt{0.0625-4(0.125)}}{2} = \frac{-0.25\pm \sqrt{-4.375}}{2} \approx -0.125 \pm 0.331\,i$$
+We found complex eigenvalues with a negative real part, which means that we have a (stable) spiral sink. Near our equilibrium point, the phase portrait looks like:
+
+```{r, echo=FALSE}
+# Phase Portrait setup
+y0 <- matrix(c(0, 1, 0, 5, -6, 1, 5, 0.5, 0, -3), 5, 2, byrow = TRUE)
+
+y0_df <- tibble(x=c(0,0,-6,5,0), 
+                y=c(1,5,1,0.5,-3),
+                IC=c("V1", "V2", "V3", "V4", "V5")) %>% 
+  mutate(pt = paste0("(",x, ",", y, ")"))
+
+fun_traj <- function(t, y, parameters) {
+  x <- y[1]
+  y <- y[2]
+  a <- parameters[1]
+  b <- parameters[2]
+  c <- parameters[3]
+  d <- parameters[4]
+  dy <- numeric(2)
+  dy[1] <- a*x + b*y
+  dy[2] <- c*x + d*y  
+  return(list(dy))
+}
+
+fun_plane <- c(dx ~ a*x+b*y, dy ~c*x+d*y)
+
+make_plane_complex <- function(params) {
+  params_unnamed <- unname(params)
+  traj <- trajectory(fun_traj, y0 = y0, tlim= c(0, 55), tstep = 0.05,
+                     parameters = params_unnamed, add=FALSE)
+
+  t <- c(traj$t)
+  x <- as_tibble(traj$x) %>% bind_cols("t"=t) %>% 
+    pivot_longer(cols = c(1:5), names_to = "IC", values_to = "x")
+  y <- as_tibble(traj$y) %>% bind_cols("t"=t) %>% 
+    pivot_longer(cols = c(1:5), names_to = "IC", values_to = "y")
+  z <- left_join(x,y, by = join_by(t, IC)) %>% filter(x<7 & x>-7 & y<7 & y>-7) %>% 
+    left_join(y0_df %>% select(IC, pt), by="IC")
+
+  p1 <- phaseplane(fun_plane, x_var="x" ,y_var="y",
+                   x_window = c(-7, 7), y_window = c(-7, 7),
+                   parameters = params)
+
+  p1+theme_bw()+
+    geom_path(data=z, aes(x=x, y=y, group=pt), linewidth=1) +
+  theme(
+    axis.ticks = element_blank(),
+    axis.text = element_blank()
+  )
+}
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE,fig.keep="last",fig.height=4, fig.width=4,fig.align='center'}
+make_plane_complex(c(a=-0.25,b=-0.5,c=0.25,d=0))
+```
+
+### b.   
+Modify the model given in class, assuming that the Susceptible population exhibits exponential growth with an intrinsic growth rate of 0.10 in the absence of any infected individuals.
+
+### c.  
+Find all equilibria of your new Epidemic model from part (b) and determine their stability by sketching the phase plane near each equilibrium.
+
+## Question 2   
+Consider the IVP from class: $$\frac{dy}{dt}=-2ty^2, \quad y(0)=1$$ over the interval $0 \leq t \leq 2$.
+
+### a.  
+Calculate the Runge-Kutta approximation to the solution with $n=4$ steps.
+
+### b.  
+Calculate the total error $e_4$ associated with this approximation.
+
+### c.  
+How many steps are necessary to approximate the solution with an error of less than
+0.0001? Make sure to justify your answer.

### 2024-04-05 bio_dynamics/test_bio274.R

@@ -69,3 +69,45 @@ flowField(fun_traj,  xlim = c(-50, 50),
 
 
 
+# April 5 -----------------------------------------------------------------
+
+# Phase Portrait setup
+y0 <- matrix(c(5, 0), 1, 2, byrow = TRUE)
+# 
+# y0_df <- tibble(x=c(0), 
+#                 y=c(4),
+#                 IC=c("V1")) %>% 
+#   mutate(pt = paste0("(",x, ",", y, ")"))
+
+fun_traj <- function(t, y, parameters) {
+  x <- y[1]
+  y <- y[2]
+  a <- parameters[1]
+  b <- parameters[2]
+  c <- parameters[3]
+  d <- parameters[4]
+  dy <- numeric(2)
+  dy[1] <- a*x + b*y
+  dy[2] <- c*x + d*y  
+  return(list(dy))
+}
+
+fun_plane <- c(dx ~ a*x+b*y, dy ~c*x+d*y)
+
+
+  params_unnamed <- unname(c(a=-0.25,b=-0.5,c=0.25,d=0))
+  traj <- trajectory(fun_traj, y0 = y0, tlim= c(0, 555), tstep = 0.1,
+                     parameters = params_unnamed, add=FALSE)
+  
+  t <- c(traj$t)
+  x <- as_tibble(traj$x) %>% rename("x"="V1") %>% bind_cols("t"=t)
+  y <- as_tibble(traj$y) %>% rename("y"="V1")%>% bind_cols("t"=t)
+  z <- left_join(x,y, by = join_by(t)) %>% filter(x<7 & x>-7 & y<7 & y>-7)
+  
+  p1 <- phaseplane(fun_plane, x_var="x" ,y_var="y",
+                   x_window = c(-7, 7), y_window = c(-7, 7),
+                   parameters = c(a=-0.25,b=-0.5,c=0.25,d=0))
+  
+  p1+theme_bw()+
+    geom_path(data=z, aes(x=x, y=y, group=pt), linewidth=1.5)
+

### 2024-04-05 bio_dynamics/bio274_hw8.Rmd
worked on 274 hw8

@@
 The epidemic model presented in class takes the following form:
-$$\frac{dX}{dt}=-\mu XY- \rho $$
+$$\frac{dX}{dt}=-\mu XY+ \rho $$
 $$\frac{dY}{dt}=\mu XY- \nu Y $$
@@
 
+The differential equation for the infected population will remain the same, so we will still have
+$$\frac{dY}{dt}=\mu XY- \nu Y $$
+However, we need to modify our DE for the susceptible population so that in the absence of infected individuals (i.e., when $Y=0$), the population will exhibit exponential growth.
+We set $r=0.1$ and write our new equation as
+$$\frac{dX}{dt}=-\mu XY+ rX $$
+When $Y=0$, this becomes $\frac{dX}{dt}=rX$.
+We know this simple first-order ODE has the solution $X(t)=X_0 e^{rt}$, meaning the population will grow exponentially over time with an intrinsic growth rate of 0.1, as desired.
+
 ### c.  
 Find all equilibria of your new Epidemic model from part (b) and determine their stability by sketching the phase plane near each equilibrium.
+
+To find the equilibrium points of our new epidemic model, we need to find all of the points where 
+\begin{equation}\label{eqn:dX}
+\frac{dX}{dt}=-\mu XY+ rX=0
+\end{equation}
+
+\begin{equation}\label{eqn:dY}
+\frac{dY}{dt}=\mu XY- \nu Y =0
+\end{equation}
+
+Let's start with the derivative of $Y$, which is the same as when we analyzed it in class. $$\frac{dY}{dt}=Y(\mu X- \nu)=0$$
+This means that either $Y=0$ or $\mu X - \nu = 0$.
+But if $Y=0$, we must have $\frac{dX}{dt}=0+ rX=0$, and since we defined $r=0.1$, $X$ must be zero and we have found the logical but boring equilibrium point $(0,0)$.
+
+The other possibility is $\mu X - \nu = 0$, which we can rearrange to be $X= \frac{\nu}{\mu}$. Plugging this into (\ref{eqn:dX}), we get
+$$\frac{dX}{dt}=-\mu \frac{\nu}{\mu}Y+ r\frac{\nu}{\mu}=0 $$
+$$\frac{dX}{dt}=-\nu Y+ r\frac{\nu}{\mu}=0 $$
+$$\nu Y= r\frac{\nu}{\mu}$$
+$$Y= \frac{r}{\mu}$$
+Therefore, our second equilibrium point is $$\left(\frac{\nu}{\mu},\frac{r}{\mu}\right)=\left(\frac{0.5}{0.025},\frac{0.1}{0.025}\right)=(20,4)$$
+
+Now we look at $\frac{dX}{dt}=-\mu XY+ rX= X(-\mu Y+r)=0$
+Thus, either $X=0$ or $-\mu Y+r=0$.
+If $X=0$, our equation (\ref{eqn:dY}) says that we must have $Y=0$, so we found the same equilibrium point $(0,0)$ as before.
+If $-\mu Y+r=0 \quad\Rightarrow\quad \mu Y=r \quad\Rightarrow\quad Y= \frac{r}{\mu}$
+
+$$\frac{dY}{dt}=\mu XY- \nu Y =0$$
+
+$$\frac{dY}{dt}=\mu X \frac{r}{\mu}- \nu \frac{r}{\mu} =0$$
+$$Xr- \nu \frac{r}{\mu} =0$$
+We can divide by $r$ because it is a non-zero constant:
+$$X- \frac{\nu}{\mu} =0$$
+So we again we have the equilibrium point 
+$$\left(\frac{\nu}{\mu},\frac{r}{\mu}\right)=(20,4)$$
+
+For our new system, the Jacobian is
+$$J(X,Y)= \begin{bmatrix} 
+    -\mu Y + r &  -\mu X \\
+    \mu Y & -\mu X-\nu \end{bmatrix}= \begin{bmatrix} 
+    -\mu Y + r &  -\mu X \\
+    \mu Y & -\mu X-\nu \end{bmatrix} $$
+
+We then evaluate the Jacobian at our equilibrium point:
+$$J(20,4)= \begin{bmatrix} 
+    -0.025(4)+0.1 &  -0.025(20) \\
+    0.025(4) & 0.025(20)-0.5 \end{bmatrix}= \begin{bmatrix} 
+    0 &  -0.5 \\
+    0.1 & 0 \end{bmatrix} $$
+
+Our next step is to find the eigenvalues of this matrix.
+\begin{equation*}
+\det{\begin{bmatrix} 
+    0-\lambda & -0.5 \\
+    0.1 & 0-\lambda \end{bmatrix}}=0 \\
+\end{equation*}
+
+$$ \lambda^2+0.05=0 $$
+
+So $\lambda \approx \pm 0.223 i$. Since we have purely imaginary complex eigenvalues, the phase plane will look like a center, corresponding to periodic solutions:
+```{r, echo=FALSE, message=FALSE, warning=FALSE,fig.keep="last",fig.height=4, fig.width=4,fig.align='center'}
+
+# Phase Portrait setup
+# Phase Portrait setup
+y0 <- matrix(c(0, 1, -6, 1, 5, 0.5, 0, -3), 5, 2, byrow = TRUE)
+
+y0_df <- tibble(x=c(0,-6,5,0), 
+                y=c(1,1,0.5,-3),
+                IC=c("V1", "V2", "V3", "V4")) %>% 
+  mutate(pt = paste0("(",x, ",", y, ")"))
+
+fun_traj <- function(t, y, parameters) {
+  x <- y[1]
+  y <- y[2]
+  a <- parameters[1]
+  b <- parameters[2]
+  c <- parameters[3]
+  d <- parameters[4]
+  dy <- numeric(2)
+  dy[1] <- a*x + b*y
+  dy[2] <- c*x + d*y  
+  return(list(dy))
+}
+
+fun_plane <- c(dx ~ a*x+b*y, dy ~c*x+d*y)
+
+make_plane_complex <- function(params) {
+  params_unnamed <- unname(params)
+  traj <- trajectory(fun_traj, y0 = y0, tlim= c(0, 55), tstep = 0.05,
+                     parameters = params_unnamed, add=FALSE)
+
+  t <- c(traj$t)
+  x <- as_tibble(traj$x) %>% bind_cols("t"=t) %>% 
+    pivot_longer(cols = c(1:4), names_to = "IC", values_to = "x")
+  y <- as_tibble(traj$y) %>% bind_cols("t"=t) %>% 
+    pivot_longer(cols = c(1:4), names_to = "IC", values_to = "y")
+  z <- left_join(x,y, by = join_by(t, IC)) %>% filter(x<7 & x>-7 & y<7 & y>-7) %>% 
+    left_join(y0_df %>% select(IC, pt), by="IC")
+
+  p1 <- phaseplane(fun_plane, x_var="x" ,y_var="y",
+                   x_window = c(-7, 7), y_window = c(-7, 7),
+                   parameters = params)
+
+  p1+theme_bw()+
+    geom_path(data=z, aes(x=x, y=y, group=pt), linewidth=1) +
+  theme(
+    axis.ticks = element_blank(),
+    axis.text = element_blank()
+  )
+}
+
+make_plane_complex(c(a=0,b=-0.5,c=0.1,d=0))
+```
+
+We then evaluate the Jacobian at our equilibrium point:
+$$J(0,0)= \begin{bmatrix} 
+    -0.025(0)+0.1 &  -0.025(0) \\
+    0.025(0) & 0.025(0)-0.5 \end{bmatrix}= \begin{bmatrix} 
+    0.1 &  0 \\
+   0 & -0.5 \end{bmatrix} $$
+
+We can see that our eigenvalues for this matrix are the values along the diagonal, $\lambda_1=0.1, \; \lambda_2=-0.5$.
+Using our knowledge of linear algebra, we know that the eigenvector associated with $\lambda_1$ is just $\begin{bmatrix}1 \\ 0 \end{bmatrix}$ and the eigenvector associated with $\lambda_2$ is $\begin{bmatrix}1 \\ 0 \end{bmatrix}$.
+We have one positive and one negative eigenvalue, which means we will have an (unstable) saddle near the equilibrium at (0,0), with straight-line solutions tending towards the origin on the x-axis (because of the positive eigenvalue) and away from the origin on the y-axis (because of the negative eigenvalue).
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE,fig.keep="last"}
+# Phase Portrait setup
+y0 <- matrix(c(0, 1, 0, 4, -6, 1, 5, 0.5, 0, -3), 5, 2, byrow = TRUE)
+
+y0_df <- tibble(x=c(0,0,-6,5,0), 
+                y=c(1,4,1,0.5,-3),
+                IC=c("V1", "V2", "V3", "V4", "V5")) %>% 
+  mutate(pt = paste0("(",x, ",", y, ")"))
+
+make_plane <- function(params) {
+  params_unnamed <- unname(params)
+  traj <- trajectory(fun_traj, y0 = y0, tlim= c(0, 1), tstep=0.005,
+                     parameters = params_unnamed, add=FALSE)
+
+  t <- c(traj$t)
+  x <- as_tibble(traj$x) %>% bind_cols("t"=t) %>% 
+    pivot_longer(cols = c(1:5), names_to = "IC", values_to = "x")
+  y <- as_tibble(traj$y) %>% bind_cols("t"=t) %>% 
+    pivot_longer(cols = c(1:5), names_to = "IC", values_to = "y")
+  z <- left_join(x,y, by = join_by(t, IC)) %>% filter(x<7 & x>-7 & y<7 & y>-7) %>% 
+    left_join(y0_df %>% select(IC, pt), by="IC")
+
+  p1 <- phaseplane(fun_plane, x_var="x" ,y_var="y",
+                   x_window = c(-7, 7), y_window = c(-7, 7),
+                   parameters = params)
+
+  A<- matrix(params_unnamed, 2, 2, byrow=TRUE)
+
+  lambda1 <- eigen(A)$values[1] # eigenvalue 1
+  lambda2 <- eigen(A)$values[2] # eigenvalue 2
+
+  print(lambda1)
+  print(lambda2)
+
+  v1_col <- "gray"
+  v1_wid <- 1
+  v2_col <- "gray"
+  v2_wid <- 1
+
+  if (abs(lambda1) < (1^(-10))) {
+    v1_col <- "black"
+    v1_wid <- 2
+    print("hi")
+  }
+
+  if (abs(lambda2) < (1^(-10))) {
+    v2_col <- "black"
+    v2_wid <- 2
+    print("hi")
+  }
+
+  p1+theme_bw()+
+    geom_vline(linewidth=v1_wid, xintercept = 0, color=v1_col)+
+    geom_hline(linewidth=v2_wid, yintercept = 0, color=v2_col)+
+    geom_path(data=z, aes(x=x, y=y, color=pt), linewidth=1.5)+
+    geom_point(data=y0_df, aes(x=x,y=y, color=pt), shape=21, fill="white",size=4, stroke=2)+labs(color="IC")
+}
+
+make_plane(c(a=0.1,b=0,c=0,d=-0.5))
+
+```
 

### 2024-04-05 bio_dynamics/test_bio274.R

@@ -53,12 +53,12 @@ make_plane <- function(params) {
     geom_point(data=y0_df, aes(x=x,y=y, color=pt), shape=21, fill="white",size=4, stroke=2)+labs(color="IC")
 }
 
-parameters=c(a=1,b=2,c=3,d=2)
+parameters=c(a=0.1,b=0,c=0,d=-0.5)
 make_plane(parameters)
 
 make_plane_complex(c(a=-3,b=10,c=-1,d=3))
 
-A <- matrix(c(-3, 10, -1, 3), 2, 2, byrow=TRUE)
+
 
 
 flowField(fun_traj,  xlim = c(-50, 50),
@@ -111,3 +111,7 @@ fun_plane <- c(dx ~ a*x+b*y, dy ~c*x+d*y)
   p1+theme_bw()+
     geom_path(data=z, aes(x=x, y=y, group=pt), linewidth=1.5)
 
+  A <- matrix(c(0.1, 0, 0, -0.5), 2, 2, byrow=TRUE)
+  eigen(A)$values  
+  eigen(A)$vectors
+  

### 2024-04-06 bio_dynamics/bio274_hw8.Rmd
finished 274 hw8

@@
 author: "Ruby Krasnow"
-date: "2024-04-05"
+date: "2024-04-06"
 output:
@@
 library(demodelr)
+library(deSolve)
 ```
@@
 
-We then evaluate the Jacobian at our equilibrium point:
+We then evaluate the Jacobian at our other equilibrium point, (0,0):
 $$J(0,0)= \begin{bmatrix} 
@@
 Using our knowledge of linear algebra, we know that the eigenvector associated with $\lambda_1$ is just $\begin{bmatrix}1 \\ 0 \end{bmatrix}$ and the eigenvector associated with $\lambda_2$ is $\begin{bmatrix}1 \\ 0 \end{bmatrix}$.
-We have one positive and one negative eigenvalue, which means we will have an (unstable) saddle near the equilibrium at (0,0), with straight-line solutions tending towards the origin on the x-axis (because of the positive eigenvalue) and away from the origin on the y-axis (because of the negative eigenvalue).
+We have one positive and one negative eigenvalue, which means we will have an (unstable) saddle near the equilibrium at (0,0), with straight-line solutions tending away from the origin on the x-axis (because of the positive eigenvalue) and towards the origin on the y-axis (because of the negative eigenvalue).
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE,fig.keep="last"}
+```{r, echo=FALSE, message=FALSE, warning=FALSE,fig.keep="last",fig.height=4, fig.width=5, fig.align='center'}
 # Phase Portrait setup
-y0 <- matrix(c(0, 1, 0, 4, -6, 1, 5, 0.5, 0, -3), 5, 2, byrow = TRUE)
+y0 <- matrix(c(0.2, 1, 3, 4, -6, -1, -5, 1, 4, -3), 5, 2, byrow = TRUE)
 
-y0_df <- tibble(x=c(0,0,-6,5,0), 
-                y=c(1,4,1,0.5,-3),
+y0_df <- tibble(x=c(0.2,3,-6,-5,4), 
+                y=c(1,  4,-1,1,-3),
                 IC=c("V1", "V2", "V3", "V4", "V5")) %>% 
@@
   params_unnamed <- unname(params)
-  traj <- trajectory(fun_traj, y0 = y0, tlim= c(0, 1), tstep=0.005,
+  traj <- trajectory(fun_traj, y0 = y0, tlim= c(0, 250), tstep=0.01,
                      parameters = params_unnamed, add=FALSE)
@@
 
-  A<- matrix(params_unnamed, 2, 2, byrow=TRUE)
-
-  lambda1 <- eigen(A)$values[1] # eigenvalue 1
-  lambda2 <- eigen(A)$values[2] # eigenvalue 2
-
-  print(lambda1)
-  print(lambda2)
-
-  v1_col <- "gray"
-  v1_wid <- 1
-  v2_col <- "gray"
-  v2_wid <- 1
-
-  if (abs(lambda1) < (1^(-10))) {
-    v1_col <- "black"
-    v1_wid <- 2
-    print("hi")
-  }
-
-  if (abs(lambda2) < (1^(-10))) {
-    v2_col <- "black"
-    v2_wid <- 2
-    print("hi")
-  }
-
   p1+theme_bw()+
-    geom_vline(linewidth=v1_wid, xintercept = 0, color=v1_col)+
-    geom_hline(linewidth=v2_wid, yintercept = 0, color=v2_col)+
-    geom_path(data=z, aes(x=x, y=y, color=pt), linewidth=1.5)+
-    geom_point(data=y0_df, aes(x=x,y=y, color=pt), shape=21, fill="white",size=4, stroke=2)+labs(color="IC")
+    geom_vline(linewidth=1, xintercept = 0, color="gray")+
+    geom_hline(linewidth=1, yintercept = 0, color="gray")+
+    geom_path(data=z, aes(x=x, y=y, group=pt), linewidth=1.5)
 }
@@
 
+Given an initial value problem $\frac{dy}{dt} = f(t, y), \quad y(t_0) = y_0$, the Runge-Kutta method is as follows:
+
+Pick a step-size $h>0$ and define:
+
+\begin{align*}
+y_{n+1} &= y_n + \frac{h}{6}\left(k_1 + 2k_2 + 2k_3 + k_4 \right),\\
+t_{n+1} &= t_n + h \\
+\end{align*}
+
+for $n = 0, 1, 2, 3, ...$ using
+\begin{align*}
+ k_1 &= \ f(t_n, y_n), \\
+ k_2 &= \ f\!\left(t_n + \frac{h}{2}, y_n + h \frac{k_1}{2}\right), \\ 
+ k_3 &= \ f\!\left(t_n + \frac{h}{2}, y_n + h \frac{k_2}{2}\right), \\
+ k_4 &= \ f\!\left(t_n + h, y_n + h k_3\right).
+\end{align*}
+
+We will first compute the result manually using this process, then check our results with two different R packages:
+```{r}
+manual<- tribble(
+  ~val,  ~n_1,        ~n_2,         ~n_3,        ~n_4,
+  "t",  0.5,        1,          1.5,        2,
+  "k1", 0,         -0.63740945,-0.49940322,-0.28490054,
+  "t_n+h/2",0.25,       0.75,       1.25,       1.75,
+  "y_n+h*(k1)/2", 1,          0.6390269,  0.37485072, 0.23694178,
+  "k2",-0.5,       -0.61253307,-0.35128265,-0.19649492,
+  "y_n+h*(k2)/2", 0.875,      0.64524599, 0.41188086, 0.25904318,
+  "k3",-0.3828125, -0.62451359,-0.42411461,-0.2348618,
+  "y_n+h*(k3)", 0.80859375, 0.48612247, 0.28764422, 0.19073601,
+  "k4",-0.65382385,-0.47263011,-0.24821759,-0.14552091,
+  "y",  0.79837926, 0.49970152, 0.30816691, 0.20040567)
+
+# from manual computation of intermediate values
+as.data.frame(manual)
+```
+
+```{r}
+# from demodelr package
+demodelr_out<- as.data.frame(demodelr::rk4(system_eq=c(dy ~ -2*t*(y^2)),
+              initial_condition = c(y=1),
+              deltaT = 0.5,
+              n_steps = 5))
+demodelr_out
+
+# from deSolve paackage
+rk_fun <- function(t, x, parms) {
+  with(as.list(parms), {
+    dx <- -1*a*t*(x[1])^2
+    list(dx) }) }
+
+time <- seq(from=0, to=2, length.out=5)
+deSolve_out <- as.data.frame(deSolve::rk4(c(y=1), time, rk_fun, c(a=2)))
+deSolve_out
+```
+
+In all three cases, we found $y(2) \approx 0.2004057$.
+
 ### b.  
 Calculate the total error $e_4$ associated with this approximation.
+
+We define the total error after $k$ steps to be
+$$e_k = \lvert y(t_k)-y_k\rvert $$
+Since we know the true solution to our ODE is $y(t)=\frac{1}{1+t^2}$, we know that the value of the function at $t=2$ should be $y(2)=\frac{1}{1+4}=0.2$
+
+```{r}
+rk_est <- deSolve_out %>% filter(time==2) %>% pull(y) #our rk4 approx.
+abs(0.2-rk_est) #abs. difference between approx. and real values
+```
+
+Thus, RK4 with $\Delta t=0.5$ gives us $e_4 = 4.057 \times 10^{-4}$.
 
@@
 0.0001? Make sure to justify your answer.
+
+In class, we found that RK4 with $\Delta t=0.2$ gives us $e_{10} = 1.095 \times 10^{-5}$ and with $\Delta t=0.1$ gives us $e_{20} = 6.541 \times 10^{-7}$.
+
+We know that
+\begin{equation}\label{eqn:C}
+e_n \leq C \Delta t
+\end{equation},
+
+so given the error associated with two different time steps $t_1$ and $t_2$, we have the following relationship:
+
+$$e_{n_1} \leq C(\Delta t_1)^4, \quad e_{n_2} \leq C(\Delta t_2)^4 \quad\Longrightarrow\quad 
+\frac{e_{n_1}}{e_{n_2}} \leq \left(\frac{\Delta t_1}{\Delta t_2}\right)^4 
+\quad\Longrightarrow\quad 
+\left(\frac{e_{n_1}}{e_{n_2}}\right)^{1/4} \leq \frac{\Delta t_1}{\Delta t_2} $$
+Which we can rearrange as
+$$\Delta t_1  \geq \left(\frac{e_{n_1}}{e_{n_2}}\right)^{1/4} \Delta t_2 $$
+
+Given an interval $0 \leq t \leq T$ and a known time step $\Delta t_1$, you need $n_1=\frac{T}{\Delta t_1}$ time steps, so if we know the error $e_{n_2}$ associated with a time step $t_2$, we can find the number of steps needed to approximate a target error $e_{n_1}$ with the formula:
+
+$$n_1  \geq \left(\frac{e_{n_1}}{e_{n_2}}\right)^{-1/4} \frac{T}{\Delta t_2} $$
+
+We can write a function to do this in R:
+```{r}
+find_n <- function(e1, e2, t2, T) {
+  out <- ((e1/e2)^(-1/4))*T*(1/t2)
+  out
+}
+```
+
+```{r}
+e4<- 0.0004056722 # The error we calculated in part (b) above
+e10<-1.095e-5     # From class slides
+e20<- 6.541e-7    # From class slides
+
+# Checking that our function works
+test4 <- paste(round(find_n(e1=e4, e2=e10, t2=0.2, T=2),2),
+      round(find_n(e1=e4, e2=e20, t2=0.1, T=2),2), sep=",")
+
+test10<- paste(round(find_n(e1=e10, e2=e4, t2=0.5, T=2),2),
+      round(find_n(e1=e10, e2=e20, t2=0.1, T=2),2), sep=",")
+
+test20 <- paste(round(find_n(e1=e20, e2=e10, t2=0.2, T=2),2),
+     round(find_n(e1=e20, e2=e4, t2=0.5, T=2),2), sep=",")
+
+out <- paste(test4, test10, test20, sep="\n")
+cat(out)
+```
+```{r}
+en <- 0.0001 # Target error
+
+# Applying the function to our target error and all three known error/time step combinations
+out2 <- paste(find_n(e1=en, e2=e10, t2=0.2, T=2), 
+              find_n(e1=en, e2=e20, t2=0.1, T=2),
+              find_n(e1=en, e2=e4, t2=0.5, T=2), sep="\n")
+cat(out2)
+```
+
+Therefore, we would need 6 time steps to approximate our solution with an error of less than 0.001.
+
+We can see that we obtain the same result by finding $C$ directly and plugging in our values to (\ref{eqn:C}):
+
+```{r}
+c1<- 0.0004056722/(0.5)^4 #0.006490755
+c2 <- 1.095e-5/(0.2)^4 #0.00684375
+c3 <- 6.541e-7/(0.1)^4 #0.006541
+c<- mean(c(c1, c2, c3))
+2/((en/c)^(1/4))
+```

### 2024-04-06 bio_dynamics/test_bio274.R

@@ -115,3 +115,47 @@ fun_plane <- c(dx ~ a*x+b*y, dy ~c*x+d*y)
   eigen(A)$values  
   eigen(A)$vectors
   
+  library(clipr)
+
+  A <- read_clip_tbl()
+ 
+  B<- tribble(
+  ~val,  ~n_1,        ~n_2,         ~n_3,        ~n_4,
+  "t",  0.5,        1,          1.5,        2,
+  "k1", 0,         -0.63740945,-0.49940322,-0.28490054,
+  "t_n+h/2",0.25,       0.75,       1.25,       1.75,
+  "y_n+h*(k1)/2", 1,          0.6390269,  0.37485072, 0.23694178,
+  "k2",-0.5,       -0.61253307,-0.35128265,-0.19649492,
+  "y_n+h*(k2)/2", 0.875,      0.64524599, 0.41188086, 0.25904318,
+  "k3",-0.3828125, -0.62451359,-0.42411461,-0.2348618,
+  "y_n+h*(k3)", 0.80859375, 0.48612247, 0.28764422, 0.19073601,
+  "k4",-0.65382385,-0.47263011,-0.24821759,-0.14552091,
+  "y",  0.79837926, 0.49970152, 0.30816691, 0.20040567)
+  
+  
+
+
+
+
+
+
+2/(((e4/e10)^(1/4))*0.2) #0.493
+2/((((e4/e20)^(1/4)))*0.1)  #0.499
+2/(((e10/e20)^(1/4))*0.1)  #0.202
+
+((en/e4)^(1/4))*0.5 #0.352
+(((en/e10)^(1/4)))*0.2  #0.348
+((en/e20)^(1/4))*0.1  #0.3516
+
+#2/n = 0.35
+#2/0.35 = n = 5.7
+
+c1<- 0.0004056722/(0.5)^4 #0.006490755
+c2 <- 1.095e-5/(0.2)^4 #0.00684375
+c3 <- 6.541e-7/(0.1)^4 #0.006541
+c<- mean(c(c1, c2, c3))
+(en/c)^(.25)
+
+find_n(e1=en, e2=e10, t2=0.2)
+find_n(e1=en, e2=e20, t2=0.1)
+find_n(e1=en, e2=e4, t2=0.5)

### 2024-04-07 PDEs/math245_hw5.Rmd
started pde5

@@
 
-## Question 1:
+## Question 5: Radioactive Decay Problem
+
+\begin{equation}\label{eqn:radio} f(x)=
+    \begin{cases}
+     u_t - u_{xx} = Ae^{-ax} \\
+     u(0,x) = \sin{x} \\
+     u(t, 0) = u(t, \pi) = 0
+    \end{cases}       
+\end{equation}
+
+We want to find a separated solution of the form $u(t,x)=X(x)T(t)$.
+Recall that for the analogous homogeneous PDE with homogeneous Dirichlet boundary conditions, we consider the following eigenvalue problem $X'' + \lambda X=0,\quad  X(0) = X(l) = 0$, which we have shown to have the eigenvalues and eigenvectors
+
+$$\lambda_n= \left(\frac{n\pi}{l}\right)^2, \quad X_n(x)=\sin{\left( \frac{n\pi x}{l}\right)},\quad n=1,2,3... $$
+Giving us the Fourier sine series
+\begin{equation}\label{eqn:fss}
+f(x)= \sum_{n=1}^{\infty} b_n \sin{  \left(\frac{n\pi x}{l}\right) }
+\end{equation}
+
+where
+$$ b_n=\frac{2}{l} \int_{0}^{l}f(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx$$
+
+Now, we assume that our solution to (\ref{eqn:radio}) will take a similar form as (\ref{eqn:fss}), where $l=\pi$ and $f(t,x)=Ae^{-ax}$:
+$$u(t,x)= b_0(t)+\sum_{n=1}^{\infty} b_n(t) \sin{(nx)}$$
+
+We can differentiate this and plug it into (\ref{eqn:radio}) as follows:
+$$u_t(t,x)= b_0'(t)+\sum_{n=1}^{\infty} b_n'(t) \sin{(nx)}$$
+$$u_{xx}(t,x)= -\sum_{n=1}^{\infty} b_n(t) \;n^2\sin{(nx)}$$
+
+\begin{equation}\label{eqn:radio2}
+b'_0(t)+\sum_{n=1}^{\infty} b_n'(t) \sin{(nx)}+\sum_{n=1}^{\infty} b_n(t) \;n^2\sin{(nx)}= Ae^{-ax}
+\end{equation}
+
+For each fixed $t$, we write $Ae^{-ax}$ as a Fourier sine series:
+\begin{equation}\label{eqn:radio3}
+Ae^{-ax}= q_0(t)+\sum_{n=1}^{\infty} q_n(t) \sin{(nx)}\end{equation}
+where
+
+\begin{align*}
+q_0 (t) &=\frac{1}{l} \int_{0}^{l}f(t,x)dx \\
+    &=\frac{A}{\pi} \int_{0}^{\pi}e^{-ax}dx \\
+    &= \frac{-A}{a \pi} \left[e^{-ax}\right]^{pi}_{0} \\
+    &= \frac{-A}{a \pi} \left(e^{-a\pi}-1\right)
+\end{align*}
+
+\begin{align*}
+q_n (t) &=\frac{2}{l} \int_{0}^{l}f(t,x)\sin{  \left(\frac{n\pi x}{l}\right) }dx \\
+    &=\frac{2A}{\pi} \int_{0}^{\pi}e^{-ax}\sin{(nx)}dx
+\end{align*}
+
+Now we do integration by parts twice on $\int_{0}^{\pi}e^{-ax}\sin{(nx)}dx$
+
+First with $u=\sin{(nx)}, \;\;du=n \cos{(nx)}\;dx, \;\; dv = e^{-ax}\;dx, \;\; v=\frac{-1}{a}e^{-ax}$,
+
+and the second time with $u=n\cos{(nx)}, \;\;du=-n^2 \sin{(nx)}\;dx, \;\; dv = \frac{1}{a} e^{-ax}\;dx, \;\; v=\frac{-1}{a^2}e^{-ax}$
+
+\begin{align*}
+\int_{0}^{\pi}e^{-ax}\sin{(nx)}dx &= 
+\left[\frac{-\sin{(nx)}}{a}e^{-ax} \right]_{0}^{\pi} +
+\frac{n}{a} \int_{0}^{\pi}e^{-ax}\cos{  \left(\frac{n\pi x}{l}\right) }dx \\
+&= 0-0+\frac{n}{a} \int_{0}^{\pi}e^{-ax}\cos{(nx)}dx \\
+&= \left[\frac{-n\cos{(nx)}}{a^2}e^{-ax} \right]_{0}^{\pi} -\frac{n^2}{a^2}\int_{0}^{\pi}e^{-ax}\sin{(nx)}dx
+\end{align*}
+
+Then moving the integrals to the same side,
+\begin{align*}
+\left(1+\frac{n^2}{a^2}\right)\int_{0}^{\pi}e^{-ax}\sin{(nx)}dx 
+&= \left[\frac{-n\cos{(nx)}}{a^2}e^{-ax} \right]_{0}^{\pi} \\
+&= \frac{-n\cos{(n\pi)}}{a^2}e^{-a\pi} +\frac{n}{a^2} \\
+(n^2+a^2) \int_{0}^{\pi}e^{-ax}\sin{(nx)}dx &= n(-1)^{n+1} e^{-a\pi}+n
+\end{align*}
+
+Thus, $$\int_{0}^{\pi}e^{-ax}\sin{(nx)}dx= \frac{n(-1)^{n+1} e^{-a\pi}+n}{n^2+a^2}$$
+
+Which means that 
+$$q_n=\frac{2A}{\pi} \frac{n(-1)^{n+1} e^{-a\pi}+n}{n^2+a^2}$$
+Now, by (\ref{eqn:radio2}) and  (\ref{eqn:radio3}), we get the following equations:
+
+\begin{equation*}
+\begin{cases}
+b'_0(t)=q_0(t) \\
+b_n'(t)+b_n(t) n^2= q_n(t)
+\end{cases}
+\end{equation*}
+
+From $b'_0(t)=q_0(t)$ we get $b_0(t)=\int_{0}^{t} q_0(s)\;ds$ Since $q_0=\frac{-A}{a \pi} \left(e^{-a\pi}-1\right)$, this means $b_0(t)=\frac{-At}{a \pi} \left(e^{-a\pi}-1\right)$
+
+On the other hand, we have $b_n'(t)+b_n(t) n^2= q_n(t)$, which we solve as follows:
+
+$$\mu(t)=\exp{ \left(\int_{0}^{t}n^2\;ds \right)}=\exp{(n^2t)}$$
+$$b_n(t)=\frac{1}{\mu(t)}\left[\int_{0}^{t} \mu(s)q_n(s)\;ds+b_n(0)\right]$$
+$$b_n(t)=b_n(0)\,\mu(t)^{-1}+\int_{0}^{t} \frac{\mu(s)}{\mu(t)}q_n(s)\;ds$$
+
+$$b_n(t)=e^{-n^2}b_n(0)+\int_{0}^{t} \frac{\exp{(n^2s)}}{\exp{(n^2t)}}q_n(s)\;ds$$
+$$b_n(t)=e^{-n^2}b_n(0)+\int_{0}^{t} e^{n^2(s-t)}q_n(s)\;ds$$
+
+Therefore, our solution to (\ref{eqn:radio}) is:
+$$u(t,x)= b_0(t)+\sum_{n=1}^{\infty} b_n(t) \sin{(nx)}, \quad \text{where} \\ b_n(t)=e^{-n^2}b_n(0)+\int_{0}^{t} e^{n^2(s-t)}q_n\;ds,\\
+q_n=\frac{2A}{\pi} \frac{n(-1)^{n+1} e^{-a\pi}+n}{n^2+a^2}, \\ 
+b_0(t)=\frac{-At}{a \pi} \left(e^{-a\pi}-1\right)$$
+
+$$u(0,x)=b_0(0)+\sum_{n=1}^{\infty} \left[e^{-n^2}b_n(0)+\int_{0}^{0} e^{n^2(s-0)}q_n(s)\;ds   \right] \sin{(nx)}$$
+
+$$u(0,x)=b_0(0)+\sum_{n=1}^{\infty}e^{-n^2}b_n(0) \sin{(nx)}$$

### 2024-04-07 PDEs/paper_outline.Rmd
started pde5

@@
 author: "Ruby Krasnow"
-date: "2024-03-19"
+date: "2024-04-06"
 output:
@@
 Here, we use a simple one-dimensional reaction-diffusion model to represent the spread of the invasive ascidian *Botrylloides violaceus* in the Gulf of Maine. *B. violaceus* is ... DESCRIPTION HERE
-Colonial ascidians generally display indeterminate growth, meaning that colony area tends to increase exponentially with time (Jackson, 1977). 
-Thus, our reaction-diffusion model takes the form:
+
+Colonial ascidians generally display indeterminate growth, meaning that colony area tends to increase exponentially with time (Jackson, 1977). Thus, our reaction-diffusion model takes the form:
 

### 2024-04-07 PDEs/math245_hw5.Rmd
pde hw 5 prob 5 and 6

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-04-05"
+date: "2024-04-07"
 output:
@@
 
-## Question 5: Radioactive Decay Problem
+## Problem 1: Inhomogeneous Heat Equation
 
-\begin{equation}\label{eqn:radio} f(x)=
+Using the method of separation of variables, solve the inhomogeneous heat equation:
+\begin{equation}\label{eqn:inhomoheat}
+    \begin{cases}
+     u_{t} -k u_{xx}=x \cos{t}          & 0<x<\pi, \quad t>0 \\
+     u(0,x) = \sin{(\pi x)}             & 0<x<\pi\\
+     u(t,0)=t^2, \;\; u(t,\pi)=2t       & t>0
+    \end{cases}       
+\end{equation}
+
+## Problem 2: Inhomogeneous Wave Equation
+
+Using the method of separation of variables, solve the inhomogeneous wave equation:
+\begin{equation}\label{eqn:inhomowave}
+    \begin{cases}
+     u_{tt} -c^2 u_{xx}=F(t,x)    & 0<x<L, \quad t>0 \\
+     u(0,x) = \phi(x)             & 0<x<L\\
+     u_t(0,x) = \psi(x)           & 0<x<L\\
+     u_x(t,0)=h(t), \;\; u_x(t,L)=g(t) & t>0
+    \end{cases}       
+\end{equation}
+
+for constants $a$ and $k>0$
+
+## Problem 3: Damped Heat Equation
+
+Using the method of separation of variables, solve the damped heat equation:
+\begin{equation}\label{eqn:dampheat}
+    \begin{cases}
+     u_{t} +au=ku_{xx}    & -\pi<x<\pi, \quad t>0 \\
+     u(0,x) = \phi(x)     & -\pi<x<\pi\\
+     u(t,\pi)=u(t,-\pi)   & t>0 \\
+     u_x(t,\pi)=u_x(t,-\pi) & t>0
+    \end{cases}       
+\end{equation}
+
+for constants $a$ and $k>0$
+
+## Problem 4: Beam Equation
+
+Using the method of separation of variables, solve the beam equation:
+\begin{equation}\label{eqn:beam}
+    \begin{cases}
+     u_{tt} =c^2 u_{xxxx} & 0<x<L, \quad t>0 \\
+     u(0,x) = \phi(x)     & 0<x<L\\
+     u_t(0,x) = \psi(x)   & 0<x<L\\
+     u(t, 0) = u(t, L) = 0 & t>0 \\
+     u_{xx}(t,0)=u_{xx}(t,L)=0
+    \end{cases}       
+\end{equation}
+
+## Problem 5: Radioactive Decay Problem
+
+Using the method of separation of variables, solve the radioactive decay problem, for constants $A, a>0$.
+\begin{equation}\label{eqn:radio}
     \begin{cases}
@@
 
-From $b'_0(t)=q_0(t)$ we get $b_0(t)=\int_{0}^{t} q_0(s)\;ds$ Since $q_0=\frac{-A}{a \pi} \left(e^{-a\pi}-1\right)$, this means $b_0(t)=\frac{-At}{a \pi} \left(e^{-a\pi}-1\right)$
+From $b'_0(t)=q_0(t)$ we get $$b_0(t)=\int_{0}^{t} q_0(s)\;ds$$
+
+Since $q_0=\frac{-A}{a \pi} \left(e^{-a\pi}-1\right)$, this means
+$$b_0(t)=\frac{-At}{a \pi} \left(e^{-a\pi}-1\right)+b_0(0)$$
 
@@
 
+$$u(0,x)= \sin{x} = b_0(0)+\sum_{n=1}^{\infty} b_n(0) \sin{(nx)}$$
+
+so using our equations to find the  coefficients of a Fourier sine series, 
+$$b_0(0)=\frac{2}{\pi} \int_{0}^{\pi}\sin{(nx)}dx$$
+$$=\frac{-2}{n\pi} \cos{(nx)}\Bigr\rvert^{\pi}_{0}\;dx\\
+=  \frac{2}{n\pi}(1+(-1)^{n+1})$$
+
+$$b_n(0)=\frac{1}{\pi} \int_{0}^{\pi}\sin{(x)} \sin{(nx)}dx$$
+
+We evaluate this integral using the following trigonometric identity: $$\sin{\alpha}\sin{\beta}=\frac{\cos{(\alpha-\beta)}-\cos{(\alpha+\beta)}}{2}
+\quad \Longrightarrow\quad \sin{(x)} \sin{(nx)}=\frac{\cos{(-x)}-\cos{(3x)}}{2}$$
+
+$$b_n(0)=\frac{1}{2\pi} \int_{0}^{\pi} \cos{(x)}-\cos{(3x)}dx=0$$
+Because $\sin{0}, \sin{\pi}$, and $\sin{3\pi}$ all equal zero.
+
 Therefore, our solution to (\ref{eqn:radio}) is:
-$$u(t,x)= b_0(t)+\sum_{n=1}^{\infty} b_n(t) \sin{(nx)}, \quad \text{where} \\ b_n(t)=e^{-n^2}b_n(0)+\int_{0}^{t} e^{n^2(s-t)}q_n\;ds,\\
-q_n=\frac{2A}{\pi} \frac{n(-1)^{n+1} e^{-a\pi}+n}{n^2+a^2}, \\ 
-b_0(t)=\frac{-At}{a \pi} \left(e^{-a\pi}-1\right)$$
+\begin{tcolorbox}[colback=white, title=Solution]
+\begin{align*}
+& u(t,x)= b_0(t)+\sum_{n=1}^{\infty} b_n(t) \sin{(nx)}, \quad \text{where} \\ 
+& b_n(t)=\int_{0}^{t} e^{n^2(s-t)}q_n\;ds,\\
+& q_n=\frac{2A}{\pi} \frac{n(-1)^{n+1} e^{-a\pi}+n}{n^2+a^2}, \\ 
+& b_0(t)=\frac{-At}{a \pi} \left(e^{-a\pi}-1\right)+b_0(0), \\
+& b_0(0)=\frac{2}{n\pi}(1+(-1)^{n+1})
+\end{align*}
+\end{tcolorbox}
 
-$$u(0,x)=b_0(0)+\sum_{n=1}^{\infty} \left[e^{-n^2}b_n(0)+\int_{0}^{0} e^{n^2(s-0)}q_n(s)\;ds   \right] \sin{(nx)}$$
+## Problem 6: Telegraph Equation
 
-$$u(0,x)=b_0(0)+\sum_{n=1}^{\infty}e^{-n^2}b_n(0) \sin{(nx)}$$
+Using the method of separation of variables, solve the telegraph equation:
+\begin{equation}\label{eqn:tele}
+    \begin{cases}
+     u_{tt} + au_t + bu = c^2 u_{xx}   & 0<x<l, \quad t>0 \\
+     u(0,x) = \phi(x)             & 0 \leq x \leq l\\
+     u_t(0,x) = \psi(x)           & 0 \leq x \leq l\\
+     u(t,0)=u(t,l)=0 & t>0
+    \end{cases}       
+\end{equation}
+
+for constants $a, b >0$. Only find the solution when the characteristic equation of the time problem has real roots. Define the following energy:
+
+$$E(t)=\frac{1}{2} \int_{0}^{l}(u_t^2+c^2u_x^2+bu^2)\;dx $$
+Show that $E(t) \leq E(0)$ for all $t>0$. Then prove that the telegraph equation has a unique solution.
+
+- - - -
+
+Suppose we have a separated solution of the form $u(t,x)=X(x)T(t)$.
+
+If we plug this into the homogeneous Dirichlet BCs, we find $u(t,0)=T(t)X(0)=0, u(t,l)=T(t)X(l)=0$. In order for both of these to be true, we must have either $T(t)=0$ or $X(0)=X(l)=0$.
+
+But if $T(t)=0$, then $u(t,x)=0$ for all $x$, which contradicts our ICs. Thus, we must have $X(0)=X(l)=0$.
+
+Plugging our expression $u(t,x)=X(x)T(t)$ into the PDE in (\ref{eqn:tele}), it becomes $$T''(t)X(x)+aT'(t)X(x)+bT(t)X(x)=c^2T(t)X''(x)$$
+$$- \frac{T''(t)+aT'(t)+bT(t)}{c^2 T(t)}=-\frac{X''(x)}{X(x)}=\lambda$$
+
+Based on our assumptions thus far and the IBVP (\ref{eqn:tele}), we have three problems:
+
+**Spatial problem**
+\begin{align*} \begin{cases}
+X''+\lambda X=0 \\
+X(0)=X(l)=0 \end{cases} \end{align*}
+
+**Time problem**
+\begin{equation}\label{eqn:time6}T''+aT'+(b+\lambda c^2) T=0\end{equation}
+
+**IVP**
+$$u(0,x) = T(0)X(x)=\phi(x)$$
+
+The spatial problem is an eigenvalue problem with homogenous Dirichlet BCs, which we already know has the eigenvalues and eigenvectors
+$$\lambda_n= \left(\frac{n\pi}{l}\right)^2, \quad X_n(x)=\sin{\left( \frac{n\pi x}{l}\right)},\quad n=1,2,3... $$
+
+Now moving to the time problem, the equation (\ref{eqn:time6}) becomes the characteristic equation $r^2 + ar + (b+\lambda c^2)=0$.
+From the Pythagorean Theorem, the roots of this equation are $$\frac{-a\pm\sqrt{a^2-4(b+\lambda c^2)}}{2}$$
+
+In order for the characteristic equation of the time problem to have only real roots, we must have $a^2-4(b+\lambda c^2) \geq 0$, or $$\lambda \leq \frac{a^2-4b}{4c^2}$$
+
+For each $\lambda_n$ associated with the spatial problem, we get a solution to the time problem:
+$$T_n(t)=C_n \exp{\left[\frac{t}{2}(-a-a^2-4b-4\lambda_n c^2)\right]}+D_n \exp{\left[\frac{t}{2}(-a+a^2-4b-4\lambda_n c^2)\right]} $$
+
+where $C_n, D_n \in \mathbb{R}$
+
+Therefore, the following are solutions of the PDE in (\ref{eqn:tele}):
+$u_n(t,x)=X_n(x)\;T_n(t), \quad n=1,2,3...$
+
+By the principle of superposition,
+\begin{align*}
+u(t,x) &= \sum^{\infty}_{n=1} c'_n \;u_n(t,x) \\
+       &=\sum^{\infty}_{n=1} c'_n \;T_n(t)\;X_n(x) \\
+       &=\sum^{\infty}_{n=1} A_n \;\exp{\left[\frac{t}{2}(-a-a^2-4b-4\lambda_n c^2)\right]}\;\sin{\left( \frac{n\pi x}{l}\right)}+ \\
+       &\sum^{\infty}_{n=1} B_n \;\exp{\left[\frac{t}{2}(-a+a^2-4b-4\lambda_n c^2)\right]}\;\sin{\left( \frac{n\pi x}{l}\right)}
+\end{align*}
+
+where $A_n = c'_n C_n$ and $B_n = c'_n D_n$, with $A_n, B_n \in \mathbb{R}$.
+
+We can get rid of these arbitrary constants by using our initial conditions,  and .
+First, let us simplify the notation by defining $\gamma=\frac{t}{2}(-a-a^2-4b-4\lambda_n c^2)$ and $\zeta=\frac{t}{2}(-a-a^2-4b-4\lambda_n c^2)$.
+
+Using our first initial condition, $u(0,x) = \phi(x)$:
+$$u(t,x)= \sum^{\infty}_{n=1} A_n \,e^{\gamma t} \sin{\left( \frac{n\pi x}{l}\right)}+ \\
+\sum^{\infty}_{n=1} B_n \,e^{\zeta t} \sin{\left( \frac{n\pi x}{l}\right)}$$
+
+$$u(0,x) = \phi(x) = \sum^{\infty}_{n=1} \left(A_n+B_n\right) \sin{\left( \frac{n\pi x}{l}\right)}$$
+
+We can use the equation for the coefficients inside a Fourier sine series to find $A_n+B_n$
+$$ A_n+B_n=\frac{2}{l} \int_{0}^{l}\phi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx$$
+
+Now repeating for our second initial condition, $u_t(0,x) = \psi(x)$:
+$$u_t(t,x)= \sum^{\infty}_{n=1} \gamma\, A_n \,e^{\gamma t} \;\sin{\left( \frac{n\pi x}{l}\right)}+  \sum^{\infty}_{n=1} \zeta\,B_n \,e^{\zeta t} \;\sin{\left( \frac{n\pi x}{l}\right)}$$
+
+$$u_t(0,x)= \psi(x) = \sum^{\infty}_{n=1} \left(\gamma A_n+\zeta B_n\right) \sin{\left( \frac{n\pi x}{l}\right)}$$
+
+$$\left(\gamma A_n+\zeta B_n\right)=\frac{2}{l} \int_{0}^{l}\psi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx$$
+Which can be rearranged as
+$$A_n= \frac{1}{\gamma}\left[-\zeta B_n +\frac{2}{l} \int_{0}^{l}\psi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx \right]$$
+
+Now we can solve for $B_n$:
+
+$$\frac{-\zeta B_n}{\gamma}+\frac{2}{\gamma l} \int_{0}^{l}\psi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx= -B_n+\frac{2}{l} \int_{0}^{l}\phi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx$$
+
+$$B_n-\frac{\zeta B_n}{\gamma}= \left(1-\frac{\zeta}{\gamma}\right)B_n=
+\frac{2}{l} \int_{0}^{l}\phi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx-\frac{2}{\gamma l} \int_{0}^{l}\psi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx$$
+
+Therefore, our solution is:
+\begin{tcolorbox}[colback=white, title=Solution]
+$$u(t,x)= \sum^{\infty}_{n=1} A_n \,e^{\gamma t} \sin{\left( \frac{n\pi x}{l}\right)}+ \\
+\sum^{\infty}_{n=1} B_n \,e^{\zeta t} \sin{\left( \frac{n\pi x}{l}\right)}, \quad \text{where}$$
+
+$$ A_n=\frac{2}{l} \int_{0}^{l}\phi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx-B_n$$
+$$B_n=\frac{2\gamma}{(\gamma-\zeta)l} \int_{0}^{l}\phi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx-\frac{2(\gamma-\zeta)}{l} \int_{0}^{l}\psi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx,$$
+$$\gamma=\frac{t}{2}(-a-a^2-4b-4\lambda_n c^2)$$
+$$\zeta=\frac{t}{2}(-a-a^2-4b-4\lambda_n c^2)$$
+$$\lambda_n= \left(\frac{n\pi}{l}\right)^2, \quad n=1,2,3...$$
+\end{tcolorbox}

### 2024-04-09 PDEs/math245_hw5.Rmd
finally finished beam problem

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-04-07"
+date: "2024-04-08"
 output:
@@
 \end{equation}
+
+Suppose we have a separated solution of the form $u(t,x)=X(x)T(t)$.
+
+If we plug this into the homogeneous Dirichlet BCs for $u$, we find $u(t,0)=T(t)X(0)=0, u(t,L)=T(t)X(L)=0$.
+In order for both of these to be true, we must have either $T(t)=0$ or $X(0)=X(L)=0$.
+But if $T(t)=0$, then $u(t,x)=0$ for all $x$, which contradicts our ICs.
+Thus, we must have $X(0)=X(L)=0$.
+Similarly, from our other BCs we find that $$X''(0)=X''(L)=0$$.
+
+Plugging our expression $u(t,x)=X(x)T(t)$ into the PDE in (\ref{eqn:beam}), it becomes $$T''(t)X(x)=c^2T(t)X^{(4)}(x)$$
+$$\frac{T''(t)}{c^2 T(t)}=\frac{X^{(4)}(x)}{X(x)}=\lambda$$
+$$X^{(4)}-\lambda X=0 $$
+$$r^4-\lambda =0 $$
+
+### Case 1: Zero eigenvalues, $\lambda=0$
+
+By the PDE, $X^{(4)}(x)=0$ implies that $u_{xxxx}=0$, and thus that $X(x)$ is of the form $Ax^3+Bx^2+Cx+D$. Now plugging in our initial conditions:
+
+$$X(0) = 0 \quad\Longrightarrow\quad D=0$$
+
+$$X''(x)=6Ax+2B \quad\Longrightarrow\quad X''(0)=2B=0 \quad\Longrightarrow\quad B=0$$
+
+$$X''(l)=6Al=0 \quad\Longrightarrow\quad A=0$$
+
+$$X(l)=Cl=0 \quad\Longrightarrow\quad C=0$$
+
+Therefore, we would have $X(x)=0$, so there are no eigenfunctions $X(x)$ that satisfy $X^{(4)}+\lambda X=0$ when $\lambda=0$ and hence no zero eigenvalues.
+
+### Case 2: Negative eigenvalues, $\lambda = -\beta^4 <0$
+
+We will use Green's second identity to show that we cannot have any negative eigenvalues. Start with $X^{(4)}=\lambda X$, multiply by $\bar{X}$, and integrate both sides from 0 to $l$:
+
+\begin{align*}
+X^{(4)} &=\lambda X \\
+\int_0^l X^{(4)} \bar{X}\,dx &=\lambda \int_0^l  X \bar{X}\,dx \\
+&=\lambda \int_0^l  \lvert X \rvert ^2\,dx \end{align*}
+
+Now Green's second identity states that for any $u,v \in C^2 [a,b]$, we have
+$$\int_a^bvu'' dx=\int_a^buv'' dx + [vu'-uv']^{x=b}_{x=a}$$
+
+Letting $u= X^{(4)}$ and $v=\bar{X}$,
+
+$$\int_0^l X^{(4)} \bar{X}\; dx= \int_0^l \lvert X^{''} \rvert ^2 \;dx \;+ \left[ \bar{X}X^{(5)}-X^{(4)}\bar{X}' \right]^{x=l}_{x=0}$$
+But since $X(0)=X(L)=0$, the boundary terms disappear and we have
+
+$$\int_0^l \lvert X^{''} \rvert ^2 \;dx =\lambda \int_0^l  \lvert X \rvert ^2\,dx$$
+
+$$\lambda = \frac{\int_0^l \lvert X^{''} \rvert ^2 \;dx }{\int_0^l  \lvert X \rvert ^2\,dx} \quad\Longrightarrow\quad \lambda \geq 0$$
+Therefore, we have only positive eigenvalues.
+
+### Case 3: Positive eigenvalues, $\lambda = \beta^4 >0$
+This case gives us the characteristic equation: $r^4-\beta^4=0$, or $r^2= \pm \beta^2, r=\pm \beta, \pm \beta i$
+
+$$X(x)=Ae^{\beta x}+Be^{-\beta x}+C\sin{(\beta x)}+D\cos{(\beta x)}$$
+\begin{equation}\label{eqn:x0}
+X(0)=A+B+D=0
+\end{equation}
+
+$$X''(x)=\beta^2Ae^{\beta x}+\beta^2Be^{-\beta x}-\beta^2C\sin{(\beta x)}-\beta^2D\cos{(\beta x)}$$
+Since we are in a case where we defined $\beta$ to be non-zero, we can divide by it to find that
+$$X''(0)=\beta^2A+\beta^2B-\beta^2D=0 \quad\Longrightarrow\quad A+B=D$$
+Combining this with (\ref{eqn:x0}), we find that $D=0$.
+
+$$X(L)=Ae^{\beta L}+Be^{-\beta L}+C\sin{(\beta L)}=0$$
+
+$$X''(L)=\beta^2Ae^{\beta L}+\beta^2Be^{-\beta L}-\beta^2C\sin{(\beta L)}=0$$
+
+$$C\sin{(\beta L)}=-C\sin{(\beta L)}, \quad 2C\sin{(\beta L)}=0$$
+Either $C=0$ or $\sin{(\beta L)}=0$. Say $C=0$, then $X(0)=A+B=0$ implies $A=-B$. Then the boundary condition gives
+$$X(L)=Be^{\beta L}-Be^{-\beta L}=0, \quad B(e^{\beta L}-e^{-\beta L})=0$$
+
+Since $e^{\beta L}$ and $e^{-\beta L}$ are only equal when $L=0$ and we defined $L>0$, we must have $B=A=0$, which would imply that $X(x)=0$, the trivial solution.
+Therefore, we must have $\sin{(\beta L)}=0$, such that $\beta = \frac{n\pi}{L}$.
+
+Returning to our boundary conditions,
+
+$$X(0)=Ae^0+Be^0+C\sin{0}=0$$
+
+$$A+B=0, \quad A=-B$$
+
+$$X(L)=Ae^{\left(\frac{n\pi}{L}\right) L}-Ae^{-\left(\frac{n\pi}{L}\right) L}+C\sin{\left(\left(\frac{n\pi}{L}\right) L\right)}=0$$
+
+$$X(L)=Ae^{n\pi}-Ae^{-n\pi}+C\sin{(n\pi)}=0$$
+
+$$Ae^{n\pi}=Ae^{-n\pi}$$
+
+But for $n=1,2,3...$, this is only true if $A=0$, which means $B$ is also zero.
+
+Thus, we have eigenvalues of $\lambda_n= \left(\frac{n\pi}{L}\right)^4$ with eigenfunctions 
+$$\quad X_n(x)=\sin{\left( \frac{n\pi x}{L}\right)},\quad n=1,2,3...$$
+
+Now we return to the time equation,
+
+$$T''(t)-\lambda c^2 T(t)=0$$
+
+$$T''-\left(\frac{n\pi}{L}\right)^4 c^2 T=0$$
+
+$$r^2-\left(\frac{n\pi}{L}\right)^4 c^2 =0$$
+
+$$T_n(t)=Ae^{\left(\frac{n\pi}{L}\right)^2 ct}+Be^{-\left(\frac{n\pi}{L}\right)^2 ct} $$
+
+$$u_n(t,x)=\left[Ae^{\left(\frac{n\pi}{L}\right)^2 ct}+Be^{-\left(\frac{n\pi}{L}\right)^2 ct} \right]\sin{\left( \frac{n\pi x}{L}\right)},\quad n=1,2,3... $$
+
+Check if this actually satisfies our PDE:
+$$u_{tt} =c^2 u_{xxxx} $$
+
+$$u\left(\frac{n\pi}{L}\right)^4c^2 =c^2 \left(\frac{n\pi}{L}\right)^4 u $$
+
+Now we can use the principle of superposition to find our general solution:
+
+$$u(t,x)= \sum_{n=1}^{\infty} \left[Ae^{\left(\frac{n\pi}{L}\right)^2 ct}+Be^{-\left(\frac{n\pi}{L}\right)^2 ct} \right]\sin{\left( \frac{n\pi x}{L}\right)}$$
+
+$$u(0,x)= \phi(x)=\sum_{n=1}^{\infty} \left[A+B \right]\sin{\left( \frac{n\pi x}{L}\right)}$$
+
+$$A+B = \frac{2}{L} \int_{0}^{L}\phi(x)\sin{  \left(\frac{n\pi x}{L}\right) }dx$$
+
+$$u_t(0,x)= \psi(x)=\sum_{n=1}^{\infty} \left(\frac{n\pi}{L}\right)^2 c\left[A-B \right]\sin{\left( \frac{n\pi x}{L}\right)}$$
+
+$$A-B =  \left(\frac{n\pi}{L}\right)^2 \frac{2c}{L} \int_{0}^{L}\psi(x)\sin{  \left(\frac{n\pi x}{L}\right) }dx$$
+$$-2B +\frac{2}{L} \int_{0}^{L}\phi(x)\sin{  \left(\frac{n\pi x}{L}\right) }dx=  \left(\frac{n\pi}{L}\right)^2 \frac{2c}{L} \int_{0}^{L}\psi(x)\sin{  \left(\frac{n\pi x}{L}\right) }dx$$
+\begin{align*}
+B & =\frac{1}{L} \int_{0}^{L}\phi(x)\sin{  \left(\frac{n\pi x}{L}\right) }dx-  \left(\frac{n\pi}{L}\right)^2 \frac{c}{L} \int_{0}^{L}\psi(x)\sin{  \left(\frac{n\pi x}{L}\right) }dx \\
+&=\frac{1}{L} \int_{0}^{L} \left[\phi(x)- c\left(\frac{n\pi}{L}\right)^2  \psi(x)\right]\sin{  \left(\frac{n\pi x}{L}\right) }dx \end{align*}
+
+$$A=-B + \frac{2}{L} \int_{0}^{L}\phi(x)\sin{  \left(\frac{n\pi x}{L}\right) }dx$$
+
+Therefore, our solution to (\ref{eqn:beam}) is:
+\begin{tcolorbox}[colback=white, title=Solution]
+\begin{align*}
+& u(t,x)= \sum_{n=1}^{\infty} \left[Ae^{\left(\frac{n\pi}{L}\right)^2 ct}+Be^{-\left(\frac{n\pi}{L}\right)^2 ct} \right]\sin{\left( \frac{n\pi x}{L}\right)}, \quad \text{where} \\ 
+& B=\frac{1}{L} \int_{0}^{L} \left[\phi(x)- c\left(\frac{n\pi}{L}\right)^2  \psi(x)\right]\sin{  \left(\frac{n\pi x}{L}\right) }dx\\
+& A=\frac{1}{L} \int_{0}^{L} \left[\phi(x)+ c\left(\frac{n\pi}{L}\right)^2  \psi(x)\right]\sin{  \left(\frac{n\pi x}{L}\right) }dx
+\end{align*}
+\end{tcolorbox}
 
@@
 \begin{tcolorbox}[colback=white, title=Solution]
-$$u(t,x)= \sum^{\infty}_{n=1} A_n \,e^{\gamma t} \sin{\left( \frac{n\pi x}{l}\right)}+ \\
-\sum^{\infty}_{n=1} B_n \,e^{\zeta t} \sin{\left( \frac{n\pi x}{l}\right)}, \quad \text{where}$$
+\begin{equation}\label{eqn:teleSol}
+u(t,x)= \sum^{\infty}_{n=1} A_n \,e^{\gamma t} \sin{\left( \frac{n\pi x}{l}\right)}+ \\
+\sum^{\infty}_{n=1} B_n \,e^{\zeta t} \sin{\left( \frac{n\pi x}{l}\right)},\end{equation}
 
-$$ A_n=\frac{2}{l} \int_{0}^{l}\phi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx-B_n$$
+$$\text{where}  \quad A_n=\frac{2}{l} \int_{0}^{l}\phi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx-B_n$$
 $$B_n=\frac{2\gamma}{(\gamma-\zeta)l} \int_{0}^{l}\phi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx-\frac{2(\gamma-\zeta)}{l} \int_{0}^{l}\psi(x)\sin{  \left(\frac{n\pi x}{l}\right) }dx,$$
@@
 \end{tcolorbox}
+
+Now we use an energy argument to show that the telegraph equation has a unique solution. We start with the given energy equation and differentiate:
+\begin{align*}
+\frac{d}{dt} E(t) &=\frac{1}{2} \int_{0}^{l} \frac{d}{dt}(u_t^2+c^2u_x^2+bu^2)\;dx \\
+                  &=\frac{1}{2} \int_{0}^{l} 2u_tu_{tt}+2c^2u_xu_{xt}+2buu_t)\;dx \\
+                  &=\int_{0}^{l} u_tu_{tt}+\int_{0}^{l}c^2u_xu_{xt}+\int_{0}^{l}buu_t \\
+                  &= \int_{0}^{l}c^2u_xu_{xt}\; dx = c^2u_xu_t \Bigr\rvert^{l}_0- \int_{0}^{l}c^2u_tu_{xx}\;dx
+\end{align*}
+
+The term $c^2u_xu_t \Bigr\rvert^{l}_0$ disappears because of our boundary conditions, so we now have
+$$\frac{d}{dt} E(t)=\int_{0}^{l} u_tu_{tt}+buu_t-c^2u_tu_{xx}\;dx = \int_{0}^{l}u_t(u_{tt}+bu-c^2u_{xx})\;dx $$
+
+But our PDE, $u_{tt} + au_t + bu = c^2 u_{xx}$, can be re-written as $u_{tt} + bu - c^2 u_{xx}=-au_t$, which means
+
+$$\frac{d}{dt} E(t)=-\int_{0}^{l}a(u_t^2) \;dx $$
+
+Since $a$ and $u_t^2$ are necessarily non-negative, $\int_{0}^{l}a(u_t^2) \;dx \geq 0$ which means that 
+$$\frac{d}{dt} E(t)=-\int_{0}^{l}a(u_t^2) \;dx \leq 0$$.
+
+Since we just showed that the derivative of $E(t)$ is less than or equal to zero for all $t$, we have that $E(t) \leq E(0)$ for all $t>0$.
+
+Let $u$ and $v$ be two solutions to (\ref{eqn:tele}) and define $w=u-v$. Then $w$ satisfies the problem
+
+\begin{equation}\label{eqn:teleW}
+    \begin{cases}
+     w_{tt} + aw_t + bw = c^2 w_{xx}   & 0<x<l, \quad t>0 \\
+     w(0,x) = 0             & 0 \leq x \leq l\\
+     w_t(0,x) = 0           & 0 \leq x \leq l\\
+     w(t,0)=w(t,l)=0 & t>0
+    \end{cases}       
+\end{equation}
+
+$$E(0)=\frac{1}{2} \int_{0}^{l}(w_t(0,x))^2+c^2(w_x(0,x))^2+b(w(0,x)^2)\;dx = \frac{1}{2} \int_{0}^{l}(0+0+0)\;dx=0$$
+
+So $E(t) \leq E(0)$ for all $t>0$ but $E(0)=0$ for our solution $w(t,x)$, which means that
+
+$$E(t)=\frac{1}{2} \int_{0}^{l}(w_t^2+c^2w_x^2+bw^2)\;dx \leq 0$$
+
+Since all of the terms under the integrand are non-negative, this is only possible if $w_t=w_x=w=0$ for all $t>0$.
+Thus, $w=u-v=0 \quad\Rightarrow\quad u=v$, meaning any solution to (\ref{eqn:tele}) is in fact the unique solution.

### 2024-04-09 PDEs/math245_old.Rmd
pde hw 5 prob 5 and 6
finally finished beam problem

@@
 <!-- paste0("(", "x[0]","','","t[0]", ")")) -->
+
+### HW5
+
+Then the time problem has repeated real roots $r=\frac{-a}{2}$, giving us $T(t)=C_1 e^{-a/2}+C_2te^{-a/2}$.
+
+Now we can return to our solution (\ref{eqn:teleSol}).
+Since our new IBVP (\ref{eqn:teleW}) has $\phi(x)=\psi(x)=0$, our coefficients $A_n$ and $B_n$ will be zero everywhere, which means $w(t,x)=0$ and therefore $u=v$, so any solution to (\ref{eqn:tele}) is in fact the unique solution.
+
+$$T''+\lambda T=0 $$
+
+$$ r^2+\lambda=0$$
+
+$(r^2-\beta^2)(r^2+\beta^2)=(r+\beta)(r-\beta)(r+\beta i)(r-\beta i)$.
+
+\underline{Case 3:} Negative eigenvalues, $\lambda = -\beta^4 <0$
+This case gives us the characteristic equation: $r^4+\beta^4=0$, or $r^2= \pm \beta^2i$.
+
+From $e^{i\theta}=\cos{\theta}+i\sin{\theta}$, we get that $e^{\frac{i\pi}{2}}=i$. Thus, $$r^2=\beta^2e^{\frac{i\pi}{2}}, r^2=\beta^2e^{\frac{-i\pi}{2}}$$
+$$r=\beta e^{\frac{i\pi}{4}},\;-\beta e^{\frac{i\pi}{4}},\;\beta e^{\frac{-i\pi}{4}},\;-\beta e^{\frac{-i\pi}{4}} $$
+$e^{\frac{\pi}{4}i}=\cos{\frac{\pi}{4}}+i\sin{\frac{\pi}{4}}= \frac{\sqrt{2}}{2}+i\frac{\sqrt{2}}{2}$
+
+$e^{ \frac{-\pi}{4}i}=\cos{\frac{-\pi}{4}}+i\sin{\frac{-\pi}{4}}= \frac{\sqrt{2}}{2}-i\frac{\sqrt{2}}{2}$
+
+$$r = \frac{\beta\sqrt{2}}{2} \pm i\frac{\beta\sqrt{2}}{2}, \frac{-\beta\sqrt{2}}{2} \pm i\frac{\beta\sqrt{2}}{2} $$
+
+Our solutions would then take the form
+$$X(x)=e^{\frac{-\beta\sqrt{2}}{2}}\left[A\cos{\left(\frac{\beta\sqrt{2}}{2} x\right)}+B\sin{\left(\frac{\beta\sqrt{2}}{2} x\right)}\right] + e^{\frac{\beta\sqrt{2}}{2}}\left[C\cos{\left(\frac{\beta\sqrt{2}}{2} x\right)}+D\sin{\left(\frac{\beta\sqrt{2}}{2} x\right)}\right]$$
+
+$$X(0)=e^{\frac{-\beta\sqrt{2}}{2}}A+e^{\frac{\beta\sqrt{2}}{2}}C=0 $$
+$$A=-e^{\beta\sqrt{2}}C $$
+Let $\frac{\beta\sqrt{2}}{2}=\alpha$
+
+$$X(x)=e^{-\alpha}\left[-Ce^{2\alpha}\cos{\left(\alpha x\right)}+B\sin{\left(\alpha x\right)}\right] + e^{\alpha}\left[C\cos{\left(\alpha x\right)}+D\sin{\left(\alpha x\right)}\right]$$
+
+$$X(x)=e^{-\alpha}B\sin{\left(\alpha x\right)}-e^{-\alpha} e^{2\alpha}C \cos{\left(\alpha x\right)} + e^{\alpha}C\cos{\left(\alpha x\right)}+e^{\alpha}D\sin{\left(\alpha x\right)}$$
+$$X(x)=e^{-\alpha}B\sin{\left(\alpha x\right)}+e^{\alpha}D\sin{\left(\alpha x\right)}$$
+
+$$X(L)=e^{-\alpha}B\sin{(\alpha L)}+e^{\alpha}D\sin{(\alpha L)}=0$$
+$$X''(L)=-\alpha^2e^{-\alpha}B\sin{(\alpha L)}-\alpha^2e^{\alpha}D\sin{(\alpha L)}=0$$
+
+Assume $\sin{(\alpha L)} \neq 0$.
+
+$$e^{-\alpha}B+e^{\alpha}D=0 \quad \Longrightarrow\quad B=-De^{2\alpha}$$
+$$X(x)=-De^{2\alpha}e^{-\alpha}\sin{\left(\alpha x\right)}+e^{\alpha}D\sin{\left(\alpha x\right)}$$
+
+Then $X(x)=0$ for all $x$, which is just the trivial solution that doesn't satisfy our ICs.
+So we must have $\sin{(\alpha L)}=0, \alpha=\frac{\beta\sqrt{2}}{2}=\frac{n \pi}{L}, \beta=\frac{2 n \pi}{\sqrt{2}L}$
+
+<!-- \underline{Case 3:} Positive eigenvalues, $\lambda = \beta^4 >0$ -->
+
+<!-- This case gives us the characteristic equation: $r^4+\beta^4=0$.  -->
+<!-- $$(r^2+\beta^2i)(r^2-\beta^2i)=(r+\beta\sqrt{i})(r-\beta\sqrt{i})(r+\beta i^{3/2})(r-\beta i^{3/2})$$ -->
+<!-- $$X(x) = A e^{(-1)^{3/4} \beta x} + B e^{-(-1)^{3/4} \beta x} + C e^{(-1)^{1/4} \beta x} + D e^{-(-1)^{1/4} \beta x} $$ -->
+
+<!-- $$X(L) = A e^{(-1)^{3/4} \beta L} + B e^{-(-1)^{3/4} \beta L} + C e^{(-1)^{1/4} \beta L} + D e^{-(-1)^{1/4} \beta L}=0 $$ -->
+<!-- $$A+B+C+D=0 $$ -->
+
+<!-- $$X''(x) = -iA\beta^2 e^{(-1)^{3/4} \beta x} -iB\beta^2 e^{-(-1)^{3/4} \beta x} + iC\beta^2 e^{(-1)^{1/4} \beta x} + iD\beta^2 e^{-(-1)^{1/4} \beta x}=0 $$ -->
+<!-- $$i\beta^2(-A-B+C+D)=0 $$ -->
+<!-- $A+B=C+D$ -->
+
+<!-- $$-A e^{(-1)^{3/4} \beta L} -Be^{-(-1)^{3/4} \beta L} + C e^{(-1)^{1/4} \beta L} + D e^{-(-1)^{1/4} \beta L}=0 $$ -->
+
+\underline{Case 2:} Negative eigenvalues, $\lambda = -\beta^2 <0$
+This case gives us the characteristic equation: $r^2-\beta^2=0$. Since our characteristic equation has distinct real roots $r= \pm \beta$, our solutions take the form 
+$$T(t)=Ae^{\beta t}+Be^{-\beta t}$$
+
+$\beta A - \beta B = \psi(x)$
+$A+B=\phi(x)$

### 2024-04-09 PDEs/math245_hw5.Rmd
pde hw5 prob 2

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-04-08"
+date: "2024-04-09"
 output:
@@
 ```
+
+## Done: 2, 4, 5, 6
+## Not started: 1, 3
 
@@
 
-for constants $a$ and $k>0$
+### Step 1
+
+We will use the principle of superposition and write $u(t,x)=v(t,x)+w(t,x)$, so $v=u-w$. Then for $v(t,x)$, we have
+
+\begin{equation}\label{eqn:inhomowaveV1}
+    \begin{cases}
+     \square v =F(t,x) - \square w = F(t,x)-w_{tt}+c^2 w_{xx} =H(t,x)    & 0<x<L, \quad t>0 \\
+     v(0,x) = u(0,x)-w(0,x) = \phi(x) -w(0,x)  = \Phi(x)          & 0<x<L\\
+     v_t(0,x) = u_t(0,x)-w_t(0,x)= \psi(x)  -  w_t(0,x)  = \Psi(x)      & 0<x<L\\
+     v_x(t,0)=u_x(t,0)-w_x(t,0) = h(t)-h(t)=0 \\
+     v_x(t,L)=u_x(t,L)-w_x(t,L) = g(t)-g(t)=0 & t>0
+    \end{cases}       
+\end{equation}
+
+\begin{equation}\label{eqn:inhomowaveV}
+    \begin{cases}
+     \square v =H(t,x)   & 0<x<L, \quad t>0 \\
+     v(0,x) =  \Phi(x)          & 0<x<L\\
+     v_t(0,x) =\Psi(x)      & 0<x<L\\
+     v_x(t,0)=0 \\
+     v_x(t,L)=0 & t>0
+    \end{cases}       
+\end{equation}
+
+and for $w$, we have
+\begin{equation}\label{eqn:inhomowaveW}
+    \begin{cases}
+     w_{xx}=0 \\
+     w_x(t,0)=h(t) \\
+     w_x(t,L)=g(t)
+    \end{cases}       
+\end{equation}
+
+For $w(t,x)$, we will use a function of the form 
+$$w(t,x)= \left(x-\frac{x^2}{2L}\right)h(t)+\frac{x^2}{2L}g(t)$$
+$$ w_{x}=h(t)- \frac{x}{L} h(t) + \frac{x}{L}g(t)$$
+$$ w_{xx}=\frac{1}{L} \Bigl( g(t)- h(t) \Bigr) $$
+So that we will have $w_x(t,0)= h(t), w_x(t,L)=g(t)$.
+
+### Step 2:  Solve (\ref{eqn:inhomowaveV})
+
+Assume (\ref{eqn:inhomowaveV}) has the separated solution $v(t,x)=X(x)T(t)$ and consider the following eigenvalue problem
+
+\begin{equation*}
+    \begin{cases}
+     x''+\lambda X = 0 \\
+     X'(0)=X'(L)=0
+    \end{cases}       
+\end{equation*}
+
+W know that the solution to this eigenvalue problem is
+$$\lambda_n= \left(\frac{n\pi}{L}\right)^2, \quad X_n(x)=\cos{\left( \frac{n \pi x}{L}\right)},  \quad n=0,1,2,3... $$
+
+### Step 3
+
+Now, let's write $v(t,x)$ as
+
+$$v(t,x)= a_0(t)+\sum_{n=1}^{\infty}  a_n(t) \cos{  \left(\frac{n\pi x}{L}\right) } $$
+and substitute this into the PDE of (\ref{eqn:inhomowaveV}).
+
+$$v_{tt}(t,x)= a_0''(t)+\sum_{n=1}^{\infty}  a_n''(t) \cos{  \left(\frac{n\pi x}{L}\right) } $$
+$$c^2 v_{xx}(t,x)= -\sum_{n=1}^{\infty} \left(\frac{n\pi c}{L}\right)^2 a_n(t) \cos{  \left(\frac{n\pi x}{L}\right) } $$
+
+\begin{equation}\label{eqn:step3}
+H(t,x)=a_0''(t)+ \sum_{n=1}^{\infty}\left[  a_n''(t) +  \left(\frac{n\pi c}{L}\right)^2 a_n(t)    \right] \cos{  \left(\frac{n\pi x}{L}\right) }
+\end{equation}
+
+### Step 4
+
+For a fixed $t$, we write $H(t,x)$ as the Fourier cosine series
+\begin{equation}\label{eqn:step4}
+H(t,x)=q_0(t)+ \sum_{n=1}^{\infty}q_n(t) \cos{  \left(\frac{n\pi x}{l}\right) } \end{equation}
+
+Such that
+$$q_n (t)= \frac{2}{L} \int_{0}^{L}H(t,x)\cos{  \left(\frac{n\pi x}{L}\right) }dx $$
+$$q_0(t)=\frac{1}{L} \int_{0}^{L}H(t,x)dx$$
+
+### Step 5
+
+\begin{tcolorbox}[colback=white, title=Variation of Parameters]
+Consider an ODE $$(\ast)\quad y''+a(x)y'+b(x)y=c(x) $$
+and assume $\{ y_1(x), y_2(x)\}$ are fundamental solutions of the associated homogeneous equation $$(\ast \ast)\quad y''+a(x)y'+b(x)y=0 $$
+Then the general solution of $(\ast \ast)$ is 
+$y_h(x)=c_1y_1(x)+c_2y_2(x)$ for $c_1, c_2 \in \mathbb{R}$ and the solution of $(\ast)$ is the sum of the general and particular solutions, $y=y_h+y_p$, where
+$$y_p = -y_1(x) \int \frac{y_2(s)c(s)}{W(y_1,y_2)}ds+ y_2(x) \int \frac{y_1(s)c(s)}{W(y_1,y_2)}ds$$
+where $$W(y_1,y_2)= \begin{vmatrix}
+y_1 & y_2\\
+y_1' & y_2'
+\end{vmatrix}$$ \end{tcolorbox}
+
+Combine (\ref{eqn:step3}) and (\ref{eqn:step4}):
+
+$$q_0(t)=a_0''(t)$$
+$$q_n (t)=  a_n''(t) +  \left(\frac{n\pi c}{L}\right)^2 a_n(t)   $$
+First we find the fundamental solution of 
+$$ a_0''(t) =0$$
+$$a_0^H=A_0t+B_0$$
+for $A_0, B_0 \in \mathbb{R}$. Call $y_1=1, y_2=t$.
+
+$$W(1,t)= \begin{vmatrix}
+1 & t\\
+0 & 1 \end{vmatrix}= 1 $$
+
+$$a_0^P(t) = -\int_0^t s \,q_n(s)\,ds+ t \int_0^t q_n(s)\,ds=\int_0^t(t-s)q_n(s)\,ds$$
+
+$$a_0(t)=A_0t+B_0+\int_0^t(t-s)q_n(s)\,ds$$
+
+Now we find the fundamental solutions of
+
+$$ a_n''(t) +  \left(\frac{n\pi c}{L}\right)^2 a_n(t) =0$$
+This has the characteristic equation
+$r^2+  \left(\frac{n\pi c}{L}\right)^2 =0$, giving us $\{y_1,y_2\}=\left\{ \cos{\left(\frac{n\pi ct}{L}\right)}, \sin{\left(\frac{n\pi ct}{L}\right)} \right\}$.
+Thus, the general solution to the homogeneous problem is
+
+$$a_n^H(t)=A_n\cos{\left(\frac{n\pi ct}{L}\right)}+B_n\sin{\left(\frac{n\pi ct}{L}\right)}$$
+for $A_n, B_n \in \mathbb{R}, n \in \mathbb{N}$.
+
+$$W(y_1,y_2)(s)= \begin{vmatrix}
+\cos{\left(\frac{n\pi cs}{L}\right)} & \sin{\left(\frac{n\pi cs}{L}\right)}\\
+-\frac{n\pi c}{L}\sin{\left(\frac{n\pi cs}{L}\right)} & \frac{n\pi c}{L}\cos{\left(\frac{n\pi cs}{L}\right)}
+\end{vmatrix}= \frac{n\pi c}{L}$$
+
+\begin{align*}
+a_n^P(t)& = -\cos{\left(\frac{n\pi ct}{L}\right)} \int_0^t \frac{\sin{\left(\frac{n\pi cs}{L}\right)}q_n(s)}{W(y_1,y_2)}ds + \sin{\left(\frac{n\pi ct}{L}\right)} \int_0^t \frac{\cos{\left(\frac{n\pi cs}{L}\right)}q_n(s)}{W(y_1,y_2)}ds \\
+& =\int_0^t\frac{\sin{\left(\frac{n\pi c}{L} (t-s)\right)}q_n(s)}{W(y_1,y_2)}ds \\ &=\int_0^t\frac{\sin{\left(\frac{n\pi c}{L} (t-s)\right)}q_n(s)}{\frac{n\pi c}{L}}ds
+\end{align*}
+
+Therefore, $a_n(t)= a_n^H(t)+a_n^P(t)$, but the $A_n, B_n$ in $a_n^H(t)$ and the $A_0, B_0$ in $a_0^H(t)$ are still unknown.
+
+### Step 6
+
+To find $A_0, B_0, A_n, B_n$, we use the initial conditions from (\ref{eqn:inhomowaveV}).
+
+$$\Phi(x)=v(0,x)= a_0(0)+\sum_{n=1}^{\infty}  a_n(0) \cos{  \left(\frac{n\pi x}{L}\right) } =B_0+\sum_{n=1}^{\infty}  A_n \cos{  \left(\frac{n\pi x}{L}\right) }$$
+
+$$\Psi(x)=v_t(0,x)= A_0+\sum_{n=1}^{\infty} \frac{n\pi c}{L} B_n \cos{  \left(\frac{n\pi x}{L}\right) }$$
+
+$$A_0= \frac{1}{L} \int_{0}^{L}\Psi(x)dx$$
+$$B_0= \frac{1}{L} \int_{0}^{L}\Phi(x)dx$$
+$$A_n= \frac{2}{L} \int_{0}^{L}\Phi(x)\cos{  \left(\frac{n\pi x}{L}\right) }dx$$
+$$ \frac{n\pi c}{L}B_n= \frac{2}{L} \int_{0}^{L}\Psi(x)\cos{  \left(\frac{n\pi x}{L}\right) }dx$$
+
+### Step 7: Solution
+
+The solution of (\ref{eqn:inhomowave}) is then
+$$u(t,x)= a_0(t)+\sum_{n=1}^{\infty}  a_n(t) \cos{  \left(\frac{n\pi x}{L}\right) }$$
+where $$a_0(t)=\frac{1}{L} \int_{0}^{L} \left[t\,\Psi(x)+\Phi(x) \right]dx+\int_0^t(t-s)q_n(s)\,ds$$
+
+$$a_n(t)=A_n\cos{\left(\frac{n\pi ct}{L}\right)}+B_n\sin{\left(\frac{n\pi ct}{L}\right)}+\int_0^t\frac{\sin{\left(\frac{n\pi c}{L} (t-s)\right)}q_n(s)}{\frac{n\pi c}{L}}ds$$
+$$A_n= \frac{2}{L} \int_{0}^{L}\Phi(x)\cos{  \left(\frac{n\pi x}{L}\right) }dx, \quad B_n=\frac{2}{n\pi c}\int_{0}^{L}\Psi(x)\cos{  \left(\frac{n\pi x}{L}\right) }dx$$
+$$q_n (t)= \frac{2}{L} \int_{0}^{L}H(t,x)\cos{  \left(\frac{n\pi x}{L}\right) }dx$$
+\begin{align*}
+H(t,x)&=F(t,x)-w_{tt}+c^2w_{xx} \\
+&=F(t,x)-\left[\left(x-\frac{x^2}{2L}\right)h''(t)+\frac{x^2}{2L}g''(t)\right]
++\frac{c^2}{L} \Bigl( g(t)- h(t) \Bigr) \\
+\Phi(x)  &= u(0,x)-w(0,x) \\
+&= \phi(x) -w(0,x)  \\
+&= \phi(x) -\left(x-\frac{x^2}{2L}\right)h(0)-\frac{x^2}{2L}g(0) \\
+\Psi(x) &=u_t(0,x)-w_t(0,x) \\
+&= \psi(x)  -  w_t(0,x)  \\
+&= \psi(x)- \left(x-\frac{x^2}{2L}\right)h'(0)-\frac{x^2}{2L}g'(0)
+\end{align*}
 

### 2024-04-09 PDEs/math245_hw5.Rmd
pde hw5 prob 1

@@
 
-## Done: 2, 4, 5, 6
-## Not started: 1, 3
+## Done: 1, 2, 4, 5, 6
+## Not started: 3
 
@@
 \end{equation}
+
+### Step 1
+
+We will use the principle of superposition and write $u(t,x)=v(t,x)+w(t,x)$, so $v=u-w$. Then for $v(t,x)$, we have
+
+\begin{equation}\label{eqn:inhomoheatV1}
+    \begin{cases}
+     v_t-kv_{xx} = x \cos{t}-w_t+kw_{xx}                  & 0<x<\pi, \quad t>0 \\
+     v(0,x) = u(0,x)-w(0,x) = \sin{(\pi x)}-w(0,x)  & 0<x<\pi\\
+     v(t,0)=u(t,0)-w(t,0) = t^2-t^2=0                         & t>0       \\
+     v(t,\pi)=u(t,\pi)-w(t,\pi) = 2t-2t=0                     & t>0
+    \end{cases}       
+\end{equation}
+
+So we want a function $w(t,x)$ such that $w(t,0)=t^2$ and $w(t,\pi)=2t$.
+The following function satisfies these requirements, and also has the convenient properties that $w_{xx}(t,x)=0$ and $w(0,x)=0$.
+
+$$w(t,x)=t^2+ \frac{2t-t^2}{\pi}x $$
+$$w_t(t,x)=2t+ \frac{2-2t}{\pi}x $$
+
+Thus, set $H(t,x)=x \cos{t}-2t-\frac{2-2t}{\pi}x$, and our problem simplifies to
+\begin{equation}\label{eqn:inhomoheatV}
+    \begin{cases}
+     v_t-kv_{xx} = H(t,x) & 0<x<\pi, \quad t>0 \\
+     v(0,x) = \sin{(\pi x)}  & 0<x<\pi\\
+     v(t,0)=0                & t>0       \\
+     v(t,\pi)=0              & t>0
+    \end{cases}       
+\end{equation}
+
+### Step 2
+
+Assume (\ref{eqn:inhomoheatV}) has a separated solution $v(t,x)=X(x)T(t)$. Based on the Dirichlet BCs, we will assume the solution takes the form
+
+$$v(t,x)= \sum_{n=1}^{\infty}  b_n(t) \sin{(nx)} $$
+Where the $\pi$ in the numerator cancels with $L=\pi$ in the denominator of $\sin{  \left(\frac{n\pi x}{L}\right) }$. Now we substitute this into the PDE in (\ref{eqn:inhomoheatV}):
+
+\begin{equation}\label{eqn:step2a}
+H(t,x)=\sum_{n=1}^{\infty}\left[b_n''(t) +  kn^2 b_n(t)\right] \sin{(nx))}
+\end{equation}
+
+### Step 3
+
+For each fixed $t$, we write $H(t,x)$ as the Fourier sine series
+\begin{equation}\label{eqn:step3a}
+H(t,x)=\sum_{n=1}^{\infty}q_n(t) \sin{(nx)} \end{equation}
+
+Such that
+\begin{align*}
+q_n (t)&= \frac{2}{L} \int_{0}^{L}H(t,x)\sin{  \left(\frac{n\pi x}{L}\right) }dx \\
+      &= \frac{2}{\pi} \int_{0}^{\pi}H(t,x)\sin{ (nx)}dx
+\end{align*}
+
+### Step 4
+
+By (\ref{eqn:step2a}) and (\ref{eqn:step3a}), we get
+$$b_n''(t) +  kn^2 b_n(t)=q_n(t) $$
+
+$$\mu(t) = \exp\left(\int_0^t kn^2ds\right)=e^{kn^2t} $$
+
+$$b_n(t)= \frac{1}{\mu (t)} \left[\int_0^t \mu (s) q_n (s)+b_n(0)\right] $$
+
+$$b_n(t)= e^{-kn^2t}\;b_n(0)+ \int_0^t e^{kn^2(s-t)} q_n (s)$$
+
+### Step 5
+
+To find $b_n(0)$, we use the IC from (\ref{eqn:inhomoheatV}):
+
+$$v(0,x)= \sin{(\pi x)}=\sum_{n=1}^{\infty}  b_n(0) \sin{(nx)} $$
+
+\begin{align*}
+b_n(0)& =\frac{2}{\pi} \int_{0}^{\pi} \sin{(\pi x)} \cos{(nx)}dx \\ 
+      &=\frac{1}{\pi} \int_{0}^{\pi} \sin{(x(\pi+n))} +\sin{(x(\pi-n))}dx \\                   &=\left[\frac{-\cos{(x(\pi+n))}}{\pi(\pi+n)}\right]^\pi_0+\left[\frac{-\cos{(x(\pi-n))}}{\pi(\pi-n)}\right]^\pi_0 \\
+      &=\left[\frac{1-\cos{(\pi(\pi+n))}}{\pi(\pi+n)}\right]+\left[\frac{1-\cos{(\pi(\pi-n))}}{\pi(\pi-n)}\right]\\
+      &=2\left[\frac{1+\cos(\pi^2)(-1)^{n+1}}{\pi(\pi+n)}\right]
+\end{align*}
+
+Where we used $\cos(\alpha+\beta)=\cos\alpha\cos\beta-\sin\alpha\sin\beta$ to find that
+$$\cos{(\pi^2+n\pi)}=\cos(\pi^2)\cos(n\pi)-\sin(\pi^2)\sin(n\pi)=\cos(\pi^2)(-1)^n$$
+$$\cos{(\pi^2-n\pi)}=\cos(\pi^2)\cos(n\pi)+\sin(\pi^2)\sin(n\pi)=\cos(\pi^2)(-1)^n$$
+
+### Step 6
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Solution]
+The solution of (\ref{eqn:inhomoheat}) is then
+$$u(t,x)= \sum_{n=1}^{\infty} \Bigl[ b_n(t) \sin{(nx)}\Bigr]+ t^2+ \frac{2t-t^2}{\pi}x$$
+where 
+$$b_n(t)= 2e^{-kn^2t}\;\left[\frac{1+\cos(\pi^2)(-1)^{n+1}}{\pi(\pi+n)}\right]+ \int_0^t e^{kn^2(s-t)} q_n (s)$$
+$$q_n (t)= \frac{2}{\pi} \int_{0}^{\pi}H(t,x)\sin{ (nx)}dx, \quad \quad H(t,x)=x \cos{t}-2t-\frac{2-2t}{\pi}x$$
+\end{tcolorbox}
 
@@
 
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Solution]
 The solution of (\ref{eqn:inhomowave}) is then
-$$u(t,x)= a_0(t)+\sum_{n=1}^{\infty}  a_n(t) \cos{  \left(\frac{n\pi x}{L}\right) }$$
+$$u(t,x)= a_0(t)+\sum_{n=1}^{\infty}  a_n(t) \cos{  \left(\frac{n\pi x}{L}\right) }+\left(x-\frac{x^2}{2L}\right)h(t)+\frac{x^2}{2L}g(t)$$
 where $$a_0(t)=\frac{1}{L} \int_{0}^{L} \left[t\,\Psi(x)+\Phi(x) \right]dx+\int_0^t(t-s)q_n(s)\,ds$$
@@
 \end{align*}
+\end{tcolorbox}
 
@@
 Therefore, our solution to (\ref{eqn:beam}) is:
-\begin{tcolorbox}[colback=white, title=Solution]
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Solution]
 \begin{align*}
@@
 Therefore, our solution to (\ref{eqn:radio}) is:
-\begin{tcolorbox}[colback=white, title=Solution]
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Solution]
 \begin{align*}
@@
 
-Therefore, our solution is:
-\begin{tcolorbox}[colback=white, title=Solution]
+Therefore, our solution to (\ref{eqn:tele}) is:
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Solution]
 \begin{equation}\label{eqn:teleSol}

### 2024-04-11 PDEs/math245_hw5.Rmd
PDE proj

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-04-09"
+date: "2024-04-10"
 output:
@@
 ```
-
-## Done: 1, 2, 4, 5, 6
-## Not started: 3
 
@@
 for constants $a$ and $k>0$
+
+Assume (\ref{eqn:dampheat}) has a separated solution $u(t,x)=X(x)T(t)$.
+
+From our boundary conditions, $X(\pi)T(t)=X(-\pi)T(t)$ and $X'(\pi)T(t)=X'(-\pi)T(t)$.
+Since we do not want the trivial solution $T(t)=0 \rightarrow u(t,x)=0$, we must have $X(\pi)=X(-\pi), \; X'(\pi)=X'(-\pi)$.
+
+Plugging $u(t,x)=X(x)T(t)$ into the PDE in (\ref{eqn:dampheat}), we get $$X(x)T'(t)+aX(x)T(t)=kX''(x)T(t)$$
+
+$$\frac{T'}{kT} +\frac{a}{k} =\frac{X''}{X}=-\lambda  $$
+
+This gives us the following (eigenvalue) time problem:
+\begin{equation*}
+    \begin{cases}
+     X''+\lambda X = 0 \\
+     X(\pi)=X(-\pi)\\
+     X'(\pi)=X'(-\pi)\\
+    \end{cases}       
+\end{equation*}
+
+### Case 1: Zero eigenvalues, $\lambda=0$
+
+$X''=0$ implies that $X(x)$ is of the form $Ax+B$. Now plugging in our initial conditions: $X(\pi)=A\pi+B= -A\pi+B=X(-\pi)$ means we must have $A=0$, so that $X(x)=B$.
+
+From our time equation, we would get $\frac{T'}{T}=-a$, meaning that $T(t)=e^{-at}$.
+
+$u(t,x)=Be^{-at}$.
+$u(0,x)=B=\phi(x)$.
+$u(t,x)=\phi e^{-at}$.
+$-a\phi e^{-at}+a\phi e^{-at}=0$ so this satisfies our PDE and BCs, but I don't think it satisfies the IC because $B$ is not a function of $x$.
+
+### Case 2: Negative eigenvalues, $-\lambda^2<0$
+
+This case gives us the characteristic equation: $r^2-\beta^2=0$. Since our characteristic equation has distinct real roots $r= \pm \beta$, our solutions take the form 
+$$X(x)=Ae^{\beta x}+Be^{-\beta x}$$
+From our BCs,
+$X(\pi)=Ae^{\beta \pi}+Be^{-\beta \pi}$ must equal $X(-\pi)=Ae^{-\beta \pi}+Be^{\beta \pi}$ and $X'(\pi)=\beta Ae^{\beta \pi}-\beta Be^{-\beta \pi}$ must equal $X'(-\pi)=\beta Ae^{\beta \pi}-\beta Be^{\beta \pi}$.
+Since we have assumed $\beta \neq 0$, we can divide by it to give the following system of equations:
+
+\begin{equation*}
+    \begin{cases}
+     Ae^{\beta \pi}+Be^{-\beta \pi} = Ae^{-\beta \pi}+Be^{\beta \pi} \\
+     Ae^{\beta \pi}-Be^{-\beta \pi} = Ae^{-\beta \pi}-Be^{\beta \pi}
+    \end{cases}       
+\end{equation*}
+
+Solve for $Ae^{\beta \pi}$ in the first equation:
+$$Ae^{\beta \pi} = Ae^{-\beta \pi}+Be^{\beta \pi}-Be^{-\beta \pi} $$
+Plug it into the second equation:
+$$Ae^{-\beta \pi}+Be^{\beta \pi}-Be^{-\beta \pi}-Be^{-\beta \pi} = Ae^{-\beta \pi}-Be^{\beta \pi} $$
+This simplifies to $Be^{\beta \pi}=Be^{-\beta \pi}$, which for non-zero $\beta$ only holds when $B=0$.
+If $B=0$, our boundary conditions then indicate $Ae^{\beta \pi} = Ae^{-\beta \pi}$ and so $A$ is similarly zero, meaning we have only the trivial solution and we do not have any negative eigenvalues.
+
+### Case 3: Positive eigenvalues, $\lambda^2>0$
+
+We will get the characteristic equation $r^2+\beta^2=0$. Since our characteristic equation has complex roots $r= \pm i\beta$, our solutions take the form 
+$$X(x)=A\sin{(\beta x)}+B\cos{(\beta x)}$$
+Differentiating, we find that 
+$$X'(x)=\beta A\cos{(\beta x)}-\beta B\sin{(\beta x)}$$
+
+$$X(\pi)=A\sin{(\beta \pi)}+B\cos{(\beta \pi)}$$
+$$X(-\pi)=A\sin{(-\beta \pi)}+B\cos{(-\beta \pi)}$$
+$$=-A\sin{(\beta \pi)}+B\cos{(\beta \pi)}$$
+So we have
+$$A\sin{(\beta \pi)}+B\cos{(\beta \pi)}=-A\sin{(\beta \pi)}+B\cos{(\beta \pi)}$$
+$$2A\sin{(\beta \pi)}=0 $$
+
+This is only true if $A=0$ or $\beta=n$. Let $A=0$. Then
+
+$$X(x)=B\cos{(\beta x)}, \quad X'(x)=-B\sin{(\beta x)}$$
+$$\quad X'(\pi)=-B\sin{(\beta \pi)}, \quad X'(-\pi)=B\sin{(\beta \pi)} $$
+$$2B\sin{(\beta \pi)}=0 $$
+Since if $\sin{(\beta \pi)}=0$ we would not be able to determine anything about $A$, if $A=0$ we must also have $B=0$, giving us just the trivial solution. Thus, we have found
+$\sin{(\beta \pi)}=0$ so $\beta=n$.
+$$\lambda_n = \beta^2=n^2, \quad X_n(x)=A\cos{(nx)}+B\sin{(nx)} $$
+
+Now we return to our time problem.
+
+$$\frac{T'}{kT}+\frac{a}{k}=-\lambda $$
+
+$$\int \frac{T'}{T} dt= \int (-\lambda k-a)\; dt$$
+$$T_n(t)=e^{-(kn+a)t} $$
+$$u_n(t,x)= e^{-(kn+a)t}\Bigl[A_n\cos{(nx)}+B_n\sin{(nx)}\Bigr], \quad n=0,1,2,3...$$
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Solution]
+$$u(t,x)= A_0+\sum^{\infty}_{n=1}e^{-(kn+a)t}\Bigl[A_n\cos{(nx)}+B_n\sin{(nx)}\Bigr] $$
+
+From from our initial condition $u(0,x)= \phi(x)$ and the equations for the coefficients of Fourier series, we find that:
+
+$$A_0=\frac{1}{2\pi} \int_{-\pi}^{\pi}\phi(x)dx$$ 
+$$ A_n=\frac{1}{\pi} \int_{-\pi}^{\pi}\phi(x)\cos{(nx)}dx$$
+$$ B_n=\frac{1}{\pi} \int_{-\pi}^{\pi}\phi(x)\sin{(nx)}dx$$
+\end{tcolorbox}
 

### 2024-04-11 PDEs/math245_old.Rmd
PDE proj

@@
 $A+B=\phi(x)$
+
+### Paper
+
+We assume that *B violaceus* occupies a section of the New England coastline of length $L$, so that $0<X<L$, and that environmental conditions outside this domain are unfavorable to *B.
+violaceus*; for instance, temperatures are above or below the tunicate's thermal tolerance.
+This gives us the homogeneous Dirichlet boundary conditions, $u(0,t)=u(0,L)=0$.
+
+"Our central thesis has been that the principal objective of any modeling effort is to abstract those details that are relevant to patterns on particular
+scales of investigation, separating signal from noise
+by suppressing unnecessary detail. Thus, the adequacy of a particular model can be evaluated only
+relevant to the purposes at hand, and to particular
+scales of investigation.
+When this philosophy is applied to models of
+spread, it is clear that the most useful models are
+not necessarily those that retain great detail at the
+level of the individual behavior.
+
+```{r}
+library(tidyverse)
+library(spocc)
+
+#tunicate<-st_as_sf(df_cleaned, coords=c("longitude", "latitude"), crs=4326)
+# ggplot()+
+#   geom_sf(data=poly)+
+#   geom_polygon(data=map_data("usa"), aes(x=long, y = lat, group = group))+
+#   geom_sf(data=tunicate_inat, aes(color=date))
+```
+
+<!-- # iNaturalist map -->
+<!-- df_inat <- occ(query = 'Botrylloides violaceus', from = 'inat', has_coords = TRUE, geometry = poly, limit = 1000) -->
+<!-- df_inat_cleaned <- occ2df(df_inat) %>%  -->
+<!--   mutate(lon=as.numeric(longitude), lat=as.numeric(latitude)) %>% -->
+<!--   filter(!is.na(date)) -->
+
+<!-- tunicate_inat <- df_inat_cleaned %>%  -->
+<!--   mutate(lon=as.numeric(longitude), lat=as.numeric(latitude), date=ymd(date), .keep="unused") -->
+
+<!-- bbox <- make_bbox(lon, lat, data = df_inat_cleaned) -->
+<!-- map <- get_stadiamap(bbox = bbox, maptype = "stamen_toner_lite", zoom=6) -->
+
+<!-- ggmap(map) + -->
+<!--   geom_point(data = df_inat_cleaned, aes(color=year(date))) -->

### 2024-04-11 PDEs/paper_outline.Rmd
PDE proj

@@
 author: "Ruby Krasnow"
-date: "2024-04-06"
+date: "2024-04-11"
 output:
@@
 
-Reaction-diffusion models are a type of partial differential equation (PDE) that have long been used by ecologists to analyze the migration and population dynamics of invasive species.
+Reaction-diffusion models are a type of partial differential equation (PDE) that have long been used by ecologists to analyze the migration and population dynamics of invasive species (Cosner, 2008).
 Recent work has led to innumerable modifications on the original formulations by Fisher (1937) and Skellam (1951).
@@
 
-Here, we use a simple one-dimensional reaction-diffusion model to represent the spread of the invasive ascidian *Botrylloides violaceus* in the Gulf of Maine. *B. violaceus* is ... DESCRIPTION HERE
+Here, we use a simple two-dimensional reaction-diffusion model to represent the spread of the invasive ascidian *Botrylloides violaceus* in the Gulf of Maine.
+Colonial tunicates (or ascidians) are compound organisms formed by a mass of small zooids organized within a common gelatinous matrix.
+These colonies often form visible crusts on rocks, vessel hulls, and other artificial floating structures and submerged objects.
+Colonial tunicates also commonly grow on macroalgae and bivalves such as shellfish and oysters, making them a nuisance for the low-trophic aquaculture industry.
 
-Colonial ascidians generally display indeterminate growth, meaning that colony area tends to increase exponentially with time (Jackson, 1977). Thus, our reaction-diffusion model takes the form:
+In New England, economic impacts from biofouling are estimated to be NUMBER. Add additional background about impact of *Botrylloides violaceus* on kelp and oyster aquaculture in the GOM here.
 
-$$\frac{\partial{u}}{\partial{t}}=\text{D}\frac{\partial{u}^2}{\partial{x^2}}+ru $$
-where $u(t,x)$ is the population density of *B. violaceus* at time $t$ and spatial coordinate $x$. This is the simplest version of the general RD model, for a ...  (Andow et al., 1990; OTHER SOURCE).
+Most common fouling tunicate at shellfish aquaculture sites along the East coast (Carman et al., 2010)
+
+*Botrylloides violaceus* has a complex life cycle, including a cyclical growth pattern dependent on both sexual and asexual reproduction.
+The growth and reproduction of colonial tunicates can be highly variable based on a number of environmental factors, including food availability, temperature, salinity, sediment load, and water pollution (Carver et al., 2006; Nelson, 2014).
+The microscale mechanisms behind tunicate dispersion are therefore very complicated, and a fully-resolved spatiotemporal model of *B.
+violaceus* spread would need also to incorporate local hydrodynamic conditions, human-mediated transport, competition, and predation.
+
+However, simple reaction-diffusion models can accurately capture population spread without replying on many assumptions and often unknown information about organism-level behavior.
+The value of these models is derived from their tremendous utility in ecological applications, despite abstracting many biological details (Andow et al., 1990).
+By eliminating unnecessary complexity and noise, simple reaction-diffusion models present a tangible method of connecting the mathematical theory of partial differential equations to observed data on invasive species, with implications for real-world ecological and economic problems.
+
+# Model Development & Theory
+
+Here, we follow the derivation method of Lewis et al. (2016) to justify our use of the following equation to represent the rate of spread of *B. violaceus*:
+
+\begin{equation}\label{eqn:V} V=\sqrt{4rD} \end{equation}
+
+where $V>0$ is the velocity (distance/time) of the advancing front of the invasive population and $D$ and $r$ are positive constants representing the diffusion coefficient and maximum per capita growth rate, respectively.
 
@@
 violaceus* has overlapping generations and can be considered to be continuously reproducing, it is appropriate to represent the invasion dynamics using a continuous-time PDE rather than a discrete integrodifference equation.
+We represent the the population density of *B.
+violaceus* at time $t$ and spatial coordinate $x$ as $u(t,x)$.
+
+Assume that the population initially occupies a finite section of the New England coastline of length $L$ and that the initial population distribution is compact.
+
+\begin{align}\label{eqn:IC_1} u(0,x) =
+\begin{cases}
+\phi(x)>0 & 0<x<L \\
+0 & \text{otherwise}
+\end{cases}
+\end{align}
+
+The reaction-diffusion model describing the population dynamics takes the form
+\begin{equation}\label{eqn:PDE}
+\frac{\partial{u}}{\partial{t}}=\text{D}\frac{\partial{u}^2}{\partial{x^2}}+F(u)
+\end{equation}
+
+Determining the correct form of the population growth term in (\ref{eqn:PDE}) is challenging.
+Colonial ascidians generally display indeterminate growth, meaning that colony area tends to increase exponentially with time (Jackson, 1977).
+If we assumed simple linear population growth, our PDE would take the form
+
+\begin{equation}\label{eqn:lin_PDE}
+\frac{\partial{U}}{\partial{t}}=\text{D}\frac{\partial{U}^2}{\partial{x^2}}+rU
+\end{equation}
+
+This is the simplest version of the general reaction-diffusion model, for a homogeneous, unstructured population growing exponentially and dispersing in a uniform, one-dimensional environment (Andow et al., 1990; Skellam, 1951).
+However, it is unlikely that colony-scale dynamics would apply over much larger spatial scales.
+Leaving $F(u)$ as an unspecified function allows for the possibility of logistic growth, as in the Fisher-KPP equation:
+
+\begin{equation}\label{eqn:Fisher}
+\frac{\partial{u}}{\partial{t}}=
+\text{D}\frac{\partial{u}^2}{\partial{x^2}}+r(1-\frac{u}{K})
+\end{equation}
+
+The use of nonlinear reaction diffusion models in a biological context is well-established (see Cherniha and Davydovych, 2017).
+However, note that for biologically reasonable choices of $F(u)$, the growth function will be bounded by the linear function, such that $F(u) \leq ru$.
+Thus,
+
+\begin{equation}\label{eqn:comp1}
+\text{D}\frac{\partial{u}^2}{\partial{x^2}}+F(u) \leq \text{D}\frac{\partial{U}^2}{\partial{x^2}}+rU
+\end{equation}
+
+Now, note that for compact initial conditions like (\ref{eqn:IC_1}), there exist positive constants $G$ and $\sigma$ such that we bound the initial distribution from above by a Gaussian distribution:
+\begin{equation}\label{eqn:initGaus}
+U(x,0)=\frac{G}{\sqrt{4\pi\sigma^2}}\exp{\left(-\frac{x^2}{4\sigma^2}\right)}
+\end{equation}
+
+\begin{equation}\label{eqn:comp2} 
+\text{such that}\quad u(0,x) \leq U(0,x)\quad \text{for all }x
+\end{equation}
+
+From (\ref{eqn:comp1}) and (\ref{eqn:comp2}), we use the comparison theorem for PDEs to find that the solution to the linear equation in (\ref{eqn:comp1}) provides an upper bound to the solution of the potentially nonlinear equation (\ref{eqn:PDE}).
+
+A detailed derivation of the solution of (\ref{eqn:lin_PDE}) can be found in Lewis et al.
+(2016).
+In brief, we utilize the change of variables $U(t,x)= W(t,x)e^{rt}$, reducing (\ref{eqn:lin_PDE}) to the standard diffusion equation for $W$.
+We then apply Fourier and inverse Fourier transforms to the new PDE and the initial condition (\ref{eqn:initGaus}) to find that
+
+$$W(t,x)=\frac{G}{\sqrt{4\pi(\sigma^2+Dt)}}\exp{\left(-\frac{x^2}{4(\sigma^2+Dt)}\right)}  $$
+
+Converting back to our original variable, we conclude that the solution to (\ref{eqn:lin_PDE}) with the initial condition (\ref{eqn:initGaus}) is a normal distribution:
+
+\begin{equation}\label{eqn:PDEsol}
+U(t,x)=\frac{G}{\sqrt{4\pi(\sigma^2+Dt)}}\exp{\left(-\frac{x^2}{4(\sigma^2+Dt)}+rt\right)}\end{equation}
+
+Now, we can quantify the spatial spread of the population by considering how quickly a level set $U(x,t)=U_c$ moves in space, where $U_c$ could be considered a threshold of detection.
+In other words, below $U_c$ it is so difficult to find the invasive organisms that for practical purposes they are not present.
+By rearranging (\ref{eqn:PDEsol}), we can get the following expression
+$$\left(\frac{x}{t}\right) ^2=4rD+\frac{4r\sigma^2}{t}-\frac{4(\sigma^2+Dt)}{t^2}\ln{  \left(\frac{U_{cr}\sqrt{4\pi(\sigma^2+Dt)}}{G}\right) }$$
+
+As $t$ becomes large, this formula simplifies to $\frac{\lvert x \rvert}{t}\rightarrow \sqrt{4rD}$. Thus, the asymptotic rate of spread is $V=\sqrt{4rD}$, as we initially sought to show.
+
+This relationship between $V, r$, and $D$ holds true for nonlinear forms of $F(u)$, can be extended to higher dimensions, and does not need to rely on the idea of a threshold of detection— in fact, true advancing fronts with asymptotic speeds given by (\ref{eqn:V}) form (Skellam, 1951, Okubo and Levin, 2001).
+For invasion of a two-dimensional homogeneous habitat described by the reaction-diffusion equation
+\begin{equation}\label{eqn:PDE2D}
+\frac{\partial{u(x,y)}}{\partial{t}}=\text{D}\left[\frac{\partial{u}^2}{\partial{x^2}}+\frac{\partial{u}^2}{\partial{y^2}}\right]+F(u),
+\end{equation}
+the radius of the species range should increase linearly with time, with the slope of the increase given by $\sqrt{4rD}$.
+This theory has been tested for numerous ecological systems and often agrees well with the empirical relationship between measured spread rates, per capita growth rates, and diffusion coefficients.
+Microscale observations can be used to parameterize (\ref{eqn:V}) and predict the spreading rate for a given species at a geographic scale.
+Such spreading-speed analyses have been conducted for a number of terrestrial animal species, including the muskrat, cereal leaf beetle, small cabbage white butterfly, grey squirrel (Andow et al., 1990; Gorsholz, 1996)
+
+Spreading speed theory has also been applied to numerous aquatic species, including the brown trout, green crab, hermit crab, periwinkle, barnacle, mussels, and molluscs (MacDonald et al., 2022; Grosholz, 1996; Marculis and Lui, 2016; Gharouni et al., 2015).
+Possibly the only application of this method to a tunicate was performed by Grozholz (1996) for *B.
+leachi*.
+His estimates of $r=1.8$ and $D=1,800$ yielded a predicted rate of spread of $114$ km/year, much faster than the observed rate of $16$ km/year.
+In addition to providing insight on the spread of a specific invasive species of commercial concern, comparing the accuracy of our estimation for *B.
+violaceus* to the previous results for *B.
+leachi* could allow for broader conclusions to be drawn regarding the applicability of this approach for colonial tunicates in general.
+
+# Application
+
+Population growth rate was estimated from Grey (2011) to range from 0.8-1.9.
+
+```{r}
+colony_day <- 0.225 #0.225 km per colony/day
+raft_prob <- (2/3) # probability that a given colony will raft
+hab_prob <- 0.5 # probability that a given rafting colony will reach a suitable habitat
+km_year_raft <- colony_day*raft_prob*hab_prob*365
+D_raft <- 2*(km_year_raft^2)/(pi)
+
+# allen_2017 <- 2*(10*52/(10^5))^2/pi # 10 cm/week
+# v_raft_allen <- 2*sqrt(1.9*allen_2017) #0.01 km/yr
+
+v_raft_low_r <- 2*sqrt(0.8*D_raft) # ~40 km/year
+v_raft_high_r <- 2*sqrt(1.9*D_raft) # ~60 km/year
+
+# 2 km/year observed by Hewitt (1993)
+
+#0.6m/day*180 days repro/yr
+2*sqrt(1.9*(2*((0.6*180)^2)/(pi))) #237 km/yr
+2*sqrt(0.8*(2*((0.6*180)^2)/(pi))) #154 km/yr
+
+#Grosholz for B. leachi predicted 114 km/year and observed 16
+```
+
+# Discussion
+
+This model attempted to predict the macro-scale dispersal of an invasive tunicate using a simple reaction-diffusion model.
+
+Multiple avenues are apparent for future work modeling the spread of *B.
+violaceus* using diffusion-type equations.
+First, there exist many other possible formulations for the term representing population growth.
+The Fisher-KPP equation, which assumes logistic rather than exponential growth, has been well-studied from a mathematical perspective and has seen widespread application across ecological contexts (SOURCES).
+Other diffusion models incorporate Allee effects, an ecological phenomenon characterized by decreased individual fitness (lower per capita growth rates) at small population sizes.
+Accounting for Allee effects in continuous reaction-diffusion models or analogous discrete integrodifference models can alter predictions about the potential success or failure of an invasion, rates of spread, and plausible ranges of an invasive species (Keitt et al., 2001; Kot et al., 1996; Liu and Kot, 2019).
+
+The model from which (\ref{eqn:V}) was derived assumes a uniform environment and isotropic dispersal.
+In reality, invasive species are more likely to disperse into more favorable habitats, so assuming that the rate of spread does not vary based on the direction of movement is not realistic.
+For example, the growth and reproduction of *B.
+violaceus* generally increase with warmer temperatures, so in a relatively cold climate, southward dispersal may be more likely than northward dispersal.
+
+*B.
+violaceus* can disperse naturally by rafting on eelgrass and other floating vegetation or debris (Carver et al., 2006; Worcester, 1994).
+Human-mediated introduction via hull fouling on recreational or commercial watercraft and aquaculture activities are also major pathways by which invasive tunicates are introduced to new locations (Carver et al., 2006).
+Although *B.
+violaceus* has a relatively low attachment strength, the small drag coefficient associated with their low profile means that its displacement velocity is above the top speeds of sailboats, barges, and other slower-moving vessels (Clarke Murray et al., 2012).
+Clarke Murray et al.
+(2011) found that *B.
+violaceus* was attached to the hulls of 9.8% of recreational boats at nine distinct marinas in British Columbia.
+
+Future work could utilize discrete integrodifference equations with leptokurtic dispersion kernels to account for transport via hull fouling and other potential vectors of long-distance dispersion.
+The potential for long-distance dispersal events via natural or human-mediated pathways often prompts the use of integrodifference equations for diffusion models of invasive populations (Lutscher, 2019; Lewis et al., 2016; Neubert and Caswell, 2000; Kot and Neubert, 2008).
+These models assume that growth and dispersal occur in distinct and asynchronous stages.
+The ability of *B.
+violaceus* to utilize multiple forms of reproduction and the fact that both larvae and post-larvae can contribute to the colonization of new habitat potentially violates the assumption of temporally separated growth and dispersion.
+However, stratified models that separate spatially continuous short-distance dispersal from discrete long-distance dispersal can be valuable tools to reconstruct realistic invasion dynamics (e.g., Gilioli et al., 2013) and their potential for modeling the spread of *B.
+violaceus* merits further investigation.
+
+Recent work coupling deterministic models with a stochastic component has led to major improvements in the biological realism of PDE-based invasion models.
+These stochastic models can account for both process and observation error, incorporating the randomness inherent in natural systems.
+For example, Roques et al.
+(2011) used a "statistical-reaction–diffusion approach" to analyze the expansion of the pine processionary moth based on moth nest presence/absence data.
+Using a "mechanistic-statistical" diffusion model assuming logistical growth and imperfect species observation data, Louvrier et al.
+(2020) was able to forecast the expanding distribution of the grey wolf in France.
+Applying similar methods to data on *B.
+violaceus* would be valuable for developing useful predictions of potential future invasions.
+
+Finally, competition-diffusion models that represent the population growth and spread of two competing species in the same habitat admit similar traveling front solutions to those described above for the simple reaction-diffusion equation (Carrère, 2018).
+It is often interesting to investigate the stability and long-term behavior of a system of competition-diffusion models for different parameter values, including varying the propagation speeds of the two organisms or the differential impacts of the competitive interactions on each species (Carrère, 2018; Ducrot et al., 2014).
+Experimental evidence suggests that *B.
+violaceus* colonies can outcompete *Botryllus schlosseri*, a different invasive tunicate that has been equally or more problematic for the New England aquaculture industry (Gittenberger and Moons, 2011; Dijkstra et al., 2007).
+However, these studies have been limited to small-scale experiments conducted on a single substrate (e.g., an oyster cage, SOURCE).
+Our single-species reaction diffusion model could be extended to a system of competition-diffusion equations to examine the potential impacts of competition between *B.
+violaceus* and *B.
+schlosseri* on the spread of these two species on a broader scale.

### 2024-04-11 PDEs/pde_mod.R

@@ -1,0 +1,29 @@
+# Ruby Krasnow
+# Code for PDE final project
+
+#load packages
+library(spocc)
+library(sf)
+library(ggmap)
+
+#define spatial boundary for observations
+polygon_list = list(rbind(c(-72, 41), c(-59.52, 41), c(-59.52, 50), c (-72, 50), c(-72, 41)))
+poly<- st_polygon(polygon_list)
+poly<-st_sfc(poly, crs=4326)
+
+
+# GBIF map
+df_gbif <- occ(query = 'Botrylloides violaceus', from = 'gbif', has_coords = TRUE, geometry = poly, limit = 1000)
+df_gbif_cleaned <- occ2df(df_gbif) %>% mutate(lon=as.numeric(longitude), lat=as.numeric(latitude)) %>% filter(!is.na(date))
+
+tunicate_gbif <- df_gbif_cleaned %>% 
+  mutate(lon=as.numeric(longitude), lat=as.numeric(latitude), date=ymd(date), .keep="unused")
+
+bbox_gbif <- make_bbox(lon, lat, data = df_gbif_cleaned)
+map_gbif <- get_stadiamap(bbox = bbox_gbif, maptype = "stamen_toner_lite", zoom=6)
+
+ggmap(map_gbif) +
+  geom_point(data = df_gbif_cleaned, aes(color=as.factor(year(date))))
+
+# create circles with expanding radii by year
+# mean inc. in radius/year ~ spreading speed

### 2024-04-12 PDEs/math245_hw6.qmd
pde hw6

---
title: "MATH 245 Homework 6"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-04-12"
format: 
  pdf:
    fig-width: 4
    fig-height: 3
execute: 
  echo: false
---

```{r}
#| label: load_packages
#| warning: false
#| message: false

library(tidyverse)
```

## Done: 1,2
## Not started: 3-9

## Problem 1

The function $u: D \subset \mathbb{R^2} \rightarrow \mathbb{R}$ is subharmonic if $\Delta u = u_{xx}+u_{yy} \geq 0$ and superharmonic if $\Delta u = u_{xx}+u_{yy} \leq 0$. Does the maximum and minimum principle hold for subharmonic functions or superharmonic on a connected bounded domain $D \subset \mathbb{R^2}$? If yes, state and prove it, otherwise give a counterexample.

### Subharmonic

**Minimum - No**  
Take $u(x,y)=x^2$ on the connected bounded domain $D=\{(x,y):-1\leq x \leq 1, -1\leq y \leq 1\}$. Then $\Delta u = 2 >0$, so $u$ is subharmonic. However, the minimum of $u$ on $D$ occurs at $x=0$ where $u(0,y)=0$ and not on the boundary of $D$, where $u=1$.

```{r}
#| label: counterexample1
#| fig-align: "center"

x<- seq(-1,1, by=0.1)
y<-x^2
ggplot()+
  geom_line(aes(x=x,y=y))+
  theme_light()
```

**Maximum - Yes**

Let $D$ be a bounded domain in $\mathbb{R^2}$ and $u: D \subset \mathbb{R^2} \rightarrow \mathbb{R}$. Assume $u$ is continuous on $\bar{D}$, where $\bar{D}=D \cup \partial D$, and is a continuous subharmonic function such that $\Delta u = u_{xx}+u_{yy} \geq 0$.

Then because $D$ is bounded and $u$ is continuous on $\bar{D}$, $u$ must attain its maximum and minimum in $\bar{D}$. Define $M=\max_{\partial D} u=u(x_0)$ on $\partial D$. Our goal is to show that $\max_{D}u \leq \max_{\partial D}u=M$.

For every $\varepsilon >0$, we define $v(x)=u(x)+\varepsilon \lvert x \rvert^2$. Then
\begin{align*}
\Delta v& = \Delta u + \varepsilon \Delta \lvert x \rvert^2 \\
        &= \Delta u+\varepsilon \sum_{i=1}^{n} \frac{\partial^2 \Delta \lvert x\rvert^2}{\partial x_{i}^2} \\
        &= \Delta u+ \varepsilon(2+2+...+2) \\
        & = \Delta u+ 2n\varepsilon
\end{align*}

And since $u$ is subharmonic, we have that
\begin{equation}\label{eqn:max_princ}
\Delta v = \Delta u+ 2n\varepsilon > 0
\end{equation}

Now assume the max of $v$ is inside $D$, at some $x_1 \in D$. Then by (\ref{eqn:max_princ}), we have $\Delta v(x_1)>0$. By generalization of the second derivative theorem, we know that at the maximum of a function $v:\mathbb{R^n} \rightarrow \mathbb{R}$, the Hessian of $v$, $\nabla^2 v(x_1)$, should be negative semi-definite, meaning that all eigenvalues $\lambda_i$ are less than or equal to zero. But $\lambda_i \leq 0$ implies that
$$\Delta v(x_1)=\text{tr}\left(\nabla^2 v(x_1)\right)=(\lambda_1+\lambda_2+...+\lambda_n)(x_1) \leq 0$$
which is a contradiction of (\ref{eqn:max_princ}). Therefore, the maximum of $v$ is at the boundary,
$$v(x) \leq \max_{\partial D}v \quad \forall \;x \in D $$
So, $\forall \;x \in D,$
\begin{align*}
u(x)&=v(x)-\varepsilon\lvert x\rvert^2 \\
&\leq \max_{\partial D}v-\varepsilon\lvert x\rvert^2
\end{align*}
But since we can write
\begin{align*}
\max_{\partial D}v-\varepsilon\lvert x\rvert^2& = \max_{\partial D} \left(u(x)+\varepsilon\lvert x\rvert^2  \right)-\varepsilon\lvert x\rvert^2 \\
&= \max_{\partial D} u+\varepsilon\left(R-\lvert x\rvert^2 \right),
\end{align*}
where $R$ is the largest radius of the domain, we must have
$$u(x)\leq \max_{\partial D} u+\varepsilon\left(R-\lvert x\rvert^2 \right). $$
Since this is true for any $\varepsilon$, we take $\varepsilon \rightarrow 0$ to find that 
$u(x) \leq \max_{\partial D} u.\quad \square$

### Superharmonic

**Maximum - No**  
Take $u(x,y)=-x^2$ on the connected bounded domain $D=\{(x,y):-1\leq x \leq 1, -1\leq y \leq 1\}$. Then $\Delta u = -2 < 0$, so $u$ is superharmonic. However, the maximum of $u$ on $D$ occurs at $x=0$ where $u(0,y)=0$ and not on the boundary of $D$, where $u=-1$.

```{r}
#| label: counterexample2
#| fig-align: "center"

y<-(-1)*(x^2)
ggplot()+
  geom_line(aes(x=x,y=y))+
  theme_light()
```

**Minimum - Yes**

Let $D$ be a bounded domain in $\mathbb{R^2}$ and $u: D \subset \mathbb{R^2} \rightarrow \mathbb{R}$. Assume $u$ is continuous on $\bar{D}$, where $\bar{D}=D \cup \partial D$, and is a continuous superharmonic function such that $\Delta u = u_{xx}+u_{yy} \leq 0$.

Then because $D$ is bounded and $u$ is continuous on $\bar{D}$, $u$ must attain its maximum and minimum in $\bar{D}$. Define $m=\min_{\partial D} u=u(x_0)$ on $\partial D$. Our goal is to show that $\min_{D}u \geq \min_{\partial D}u=m$. For every $\varepsilon >0$, we define $v(x)=u(x)-\varepsilon \lvert x \rvert^2$. Then
\begin{align*}
\Delta v& = \Delta u - \varepsilon \Delta \lvert x \rvert^2 \\
        &= \Delta u-\varepsilon \sum_{i=1}^{n} \frac{\partial^2 \Delta \lvert x\rvert^2}{\partial x_{i}^2} \\
        &= \Delta u- \varepsilon(2+2+...+2) \\
        & = \Delta u- 2n\varepsilon
\end{align*}
And since $u$ is superharmonic, we have that
\begin{equation}\label{eqn:min_princ}
\Delta v = \Delta u- 2n\varepsilon <0
\end{equation}

Now assume the min of $v$ is inside $D$, at some $x_1 \in D$. Then by (\ref{eqn:min_princ}), we have $\Delta v(x_1)<0$. By generalization of the second derivative theorem, we know that at the minimum of a function $v:\mathbb{R^n} \rightarrow \mathbb{R}$, the Hessian of $v$, $\nabla^2 v(x_1)$, should be positive semi-definite, meaning that all eigenvalues $\lambda_i$ are greater than or equal to zero. But $\lambda_i \geq 0$ implies that
$$\Delta v(x_1)=\text{tr}\left(\nabla^2 v(x_1)\right)=(\lambda_1+\lambda_2+...+\lambda_n)(x_1) \geq 0$$
which is a contradiction of (\ref{eqn:min_princ}). Therefore, the minimum of $v$ is at the boundary,
$$v(x) \geq \min_{\partial D}v \quad \forall \;x \in D $$
So, $\forall \;x \in D,$
\begin{align*}
u(x)&=v(x)+\varepsilon\lvert x\rvert^2 \\
&\leq \min_{\partial D}v+\varepsilon\lvert x\rvert^2
\end{align*}
But since we can write
\begin{align*}
\min_{\partial D}v+\varepsilon\lvert x\rvert^2& = \min_{\partial D} \left(u(x)-\varepsilon\lvert x\rvert^2  \right)+\varepsilon\lvert x\rvert^2 \\
&= \min_{\partial D} u+\varepsilon\left(\lvert x\rvert^2 -R\right),
\end{align*}

where $R$ is the largest radius of the domain, we must have
$$u(x) \leq \min_{\partial D} u+\varepsilon\left(\lvert x\rvert^2-R \right). $$
Since this is true for any $\varepsilon$, we take $\varepsilon \rightarrow 0$ to find that 
$u(x) \geq \min_{\partial D} u.\quad \square$

## Problem 2

Suppose $u$ is harmonic on the disk $D=\{(x, y):x^2+y^2< 4\}$ and $u = \sin{(\theta)} + 1$ on
$\partial D$. Without finding the solution $u$, find the maximum value of $u$ in $D \cup \partial D$ and value of $u$ at the origin.

By the maximum principle, $u$ obtains its maximum in $D \cup \partial D$ on the boundary $\partial D$ and not inside. The maximum of $u=\sin{\theta}+1$ on $\partial D$ is at $\theta=\frac{\pi}{2}$, where $\sin{\theta}$ is at its maximum. Therefore, the maximum value of $u$ in $D \cup \partial D$ is $u=\sin{\left(\frac{\pi}{2}\right)}+1=1+1=2$.

The mean value property says that for a harmonic function $u$ in a disk $D$, continuous in its closure $\bar{D}$, the value of $u$ at the center of $D$ equals the average of $u$ on its circumference. The average value of a continuous function $f(x)$ over the interval $[a,b]$ is given by 
$$f_{avg}(x)=\frac{1}{b-a}\int_a^b f(x)\;dx$$
Here,
\begin{align*}
u_{avg}(\theta)& =\frac{1}{2\pi}\int_0^{2\pi} (\sin\theta+1)\;d\theta \\
&=\frac{1}{2\pi}\Bigl[(\cos\theta+\theta) \Bigr]_0^{2\pi} \\
&=\frac{1}{2\pi}\Bigl[(\cos{2\pi}+2\pi)-\cos{0} \Bigr] \\
&= \frac{1}{2\pi}(1+2\pi-1)=1
\end{align*}
Therefore, the value of $u$ at the origin is 1.

## Problem 3

### 2024-04-15 PDEs/math245_hw6.qmd
circles for pdes

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-04-12"
+date: "2024-04-14"
 format: 
@@
 
-## Done: 1,2
-## Not started: 3-9
+## Done: 1-4
+## Not started: 5-9
 
@@
 ## Problem 3
+
+Use the maximum principle to show that the solution of the Dirichlet problem for $\Delta u=0$ depends continuously on the boundary data.
+
+Define $u_1$ and $v$ to be two solutions to the Dirichlet problems
+\begin{equation*}
+    \begin{cases}
+     \Delta u_1 =0         & \text{in }D \\
+      u_1=h(x)       & \text{on }\partial D
+    \end{cases}       
+\end{equation*}
+
+\begin{equation*}
+    \begin{cases}
+     \Delta u_2 =0         & \text{in }D \\
+      u_2=h(x)+\varepsilon(x)      & \text{on }\partial D
+    \end{cases}       
+\end{equation*}
+
+where $\varepsilon(x)$ is small $\forall x$ on $\partial D$. Now let $w=u_1-u_2$. Then $w$ solves the following Dirichlet problem:
+\begin{equation*}
+    \begin{cases}
+     \Delta w =0         & \text{in }D \\
+      w=\varepsilon(x)       & \text{on }\partial D
+    \end{cases}       
+\end{equation*}
+
+By the maximum and minimum principles,
+$$\min_{\partial D}\varepsilon(x)\leq w\leq \max_{\partial D}\varepsilon(x)$$
+If $\varepsilon \rightarrow 0$ for all $x$ on $\partial D$, then $w=u_1-u_2$ approaches zero for all $x \in D$.
+
+## Problem 4
+
+Solve the following problem on semi-annulus domain:
+\begin{equation}\label{eqn:semi}
+    \begin{cases}
+     u_{rr}+\frac{1}{r}u_r+ \frac{1}{r^2}u_{\theta\theta}=0  & 0<a<r<b, 0<\theta<\pi \\
+     u(r,0)=u(r,\pi)=0       & 0<a<r<b \\
+    u_r(a, \theta)=\theta       &  0<\theta<\pi\\
+     u_r(b, \theta)=0      &  0<\theta<\pi\\
+    \end{cases}       
+\end{equation}
+
+We look for a separated solution of the form $u=\Theta(\theta)R(r)$. First, we rewrite $u_{rr}+\frac{1}{r}u_r+ \frac{1}{r^2}u_{\theta\theta}=0$ as 
+$$R''\Theta + \frac{1}{r}R'\Theta+\frac{1}{r^2}R\Theta''=0$$
+
+Separating variables, we get
+$$\frac{R''}{R}r^2 + \frac{R'}{R}r=-\frac{\Theta''}{\Theta}$$
+Since the left-hand side is a function of only $R$ and the right-hand side is a function of only $\Theta$, we can set both sides equal to a constant $\lambda$ and separate this into two problems, one in $\Theta$ and one in $R$.
+Start with $\theta$:
+$$\Theta''+\Theta\lambda=0$$
+From our boundary conditions, 
+$$R(r)\Theta(0)=R(r)\Theta(\pi)=0 \quad\Longrightarrow\quad  \Theta(\pi)=\Theta(0)=0$$
+So we have an eigenvalue problem with homogenous Dirichlet BCs, which we have shown to have the eigenvalues and eigenvectors
+$$\lambda_n= \left(\frac{n\pi}{l}\right)^2, \quad X_n(x)=\sin{\left(\frac{n\pi x}{l}\right)},\quad n=1,2,3...$$
+Here, that means
+$$\lambda_n=n^2, \quad X_n(x)=\sin{(n\theta)},\quad n=1,2,3... $$
+
+Now onto the $R$ problem:
+$$\frac{R''}{R}r^2 + \frac{R'}{R}r=\lambda$$
+$$R''r^2 + R'r-\lambda R=0$$
+This is an Euler ODE of the form $ax^2y''+bxy'+cy=0$, which has the characteristic equation $am(m-1)+bm+c$ and, if $m_1, m_2$ are distinct real roots, the general solution $y=.
+(Note that I have used $m$ instead of the standard $r$ to avoid confusion with the variable $r$ from (\ref{eqn:semi}).
+
+Here, $a=1, b=1, c=-\lambda$, so the characteristic equation becomes $m(m-1)+m-\lambda$, which simplifies to $m^2-\lambda$. Thus, our roots are $\pm \sqrt{\lambda}=\pm n$. Thus, $R(r)=Cr^n+Dr^{-n}$.
+
+Our separated solutions thus take the form
+$$u_n(r,\theta)=\left(Cr^n+Dr^{-n}\right)\sin{(n\theta)}, \quad n=1,2,3... $$
+$$u(r,\theta)=\sum_{n=1}^{\infty} \left(Cr^n+Dr^{-n}\right)\sin{(n\theta)} $$
+
+Now we use our boundary conditions $u_r(a, \theta)=\theta$ and $u_r(b, \theta)=0$:
+$$u_r(r,\theta)=\sum_{n=1}^{\infty} \left(nCr^{n-1}-nDr^{-n-1}\right)\sin{(n\theta)} $$
+$$u_r(b,\theta)=0=\sum_{n=1}^{\infty} \left(nCb^{n-1}-nDb^{-n-1}\right)\sin{(n\theta)} $$
+$$Cb^{n}=Db^{-n} \rightarrow C=Db^{-2n}$$
+$$u_r(a,\theta)=\theta=\sum_{n=1}^{\infty} \left(nDb^{-2n}a^{n-1}-nDa^{-n-1}\right)\sin{(n\theta)} $$
+$$\frac{nD}{a}\left(b^{-2n}a^{n}-a^{-n}\right)=\int_0^\pi \theta\sin{(n\theta)}d\theta$$
+$$\int_0^\pi \theta\sin{(n\theta)}d\theta=\theta(-1/n\cos{n\theta})\Bigr\rvert^\pi_0=\pi(-1/n)(-1)^n=\frac{\pi(-1)^{n+1}}{n}$$
+$$\frac{nD}{a}\left(b^{-2n}a^{n}-a^{-n}\right)=\frac{\pi(-1)^{n+1}}{n}$$
+
+Therefore, our solution to (\ref{eqn:semi}) is
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Solution]
+$$u(r,\theta)=\sum_{n=1}^{\infty} \left(Cr^n+Dr^{-n}\right)\sin{(n\theta)}$$
+where 
+$$D=\frac{a\pi(-1)^{n+1}}{n^2\left(b^{-2n}a^{n}-a^{-n}\right)}$$
+$$C=Db^{-2n}$$
+\end{tcolorbox}
+
+## Problem 5
+
+Consider the Neumann problem with periodic BC:
+\begin{equation}\label{eqn:prob5}
+    \begin{cases}
+     \Delta u=0 & r^2>a^2 \\
+     u_r(a,\theta)=g(\theta)     & r^2=a^2
+    \end{cases}       
+\end{equation}
+
+Find the necessary condition for the existence of a solution and then solve the problem. Hint: Use Green's identity
+$$\iint\limits_{D}(u\Delta v-v\Delta u)\,dA=\int\limits_{\partial D} \left(u\frac{\partial v}{\partial n} - v\frac{\partial u}{\partial n}\right)\,ds $$
+where $D$ is a domain and $n$ is a unit normal vector. Note that for a 2-dimensional domain $D$, $dA=dx\,dy$ and $ds$ is arc length.
+
+## Problem 6.
+
+Prove the uniqueness of the Neumann problem on $D \subset \mathbb{R}$
+\begin{equation}\label{eqn:unique}
+    \begin{cases}
+     \Delta u=f  & \text{in }D \\
+     \frac{\partial u}{\partial n}=g    & \text{on }\partial D
+    \end{cases}       
+\end{equation}
+
+up to a constant. Use the following energy and Green's identity:
+$$E(u)=\frac{1}{2}\int_D\lvert \nabla u\rvert^2\,dA$$
+
+Suppose we have two solutions $u$ and $v$ to (\ref{eqn:unique}) and define $w=u-v$. Then we have a new Neumann problem
+\begin{equation}\label{eqn:unique2}
+    \begin{cases}
+     \Delta w=0  & \text{in }D \\
+     \frac{\partial w}{\partial n}=0   & \text{on }\partial D
+    \end{cases}       
+\end{equation}

### 2024-04-15 PDEs/pde_mod.R

@@ -1,10 +1,13 @@
 # Ruby Krasnow
 # Code for PDE final project
+# Last modified: April 15, 2024
 
 #load packages
+library(tidyverse)
 library(spocc)
 library(sf)
-library(ggmap)
+library(lwgeom)
+library(leaflet)
 
 #define spatial boundary for observations
 polygon_list = list(rbind(c(-72, 41), c(-59.52, 41), c(-59.52, 50), c (-72, 50), c(-72, 41)))
@@ -13,17 +16,54 @@ poly<-st_sfc(poly, crs=4326)
 
 
 # GBIF map
-df_gbif <- occ(query = 'Botrylloides violaceus', from = 'gbif', has_coords = TRUE, geometry = poly, limit = 1000)
-df_gbif_cleaned <- occ2df(df_gbif) %>% mutate(lon=as.numeric(longitude), lat=as.numeric(latitude)) %>% filter(!is.na(date))
+df_gbif <- occ(query = 'Botrylloides violaceus',
+               from = 'gbif', has_coords = TRUE,
+               geometry = c(-72, 41, -59.52, 50), limit = 1000)
 
-tunicate_gbif <- df_gbif_cleaned %>% 
-  mutate(lon=as.numeric(longitude), lat=as.numeric(latitude), date=ymd(date), .keep="unused")
+df_gbif_cleaned <- occ2df(df_gbif) %>% 
+  mutate(lng=as.numeric(longitude), lat=as.numeric(latitude), .keep="unused") %>% 
+  filter(!is.na(date))
 
-bbox_gbif <- make_bbox(lon, lat, data = df_gbif_cleaned)
-map_gbif <- get_stadiamap(bbox = bbox_gbif, maptype = "stamen_toner_lite", zoom=6)
+tunicate <- st_as_sf(df_gbif_cleaned %>% 
+                       select(date, lng, lat) %>% 
+                       mutate(year = year(date)),  
+                     coords = c("lng","lat"))
 
-ggmap(map_gbif) +
-  geom_point(data = df_gbif_cleaned, aes(color=as.factor(year(date))))
+ years <- tunicate %>% group_by(year) %>% summarise(year=mean(year))
+ 
 
-# create circles with expanding radii by year
-# mean inc. in radius/year ~ spreading speed
+ get_yrs <- function(yr) {
+   out <- years %>% filter(year <= yr) %>% st_union()
+   }
+ 
+ 
+years_vec <- c(2005:2024)
[…]

### 2024-04-16 PDEs/math245_old.Rmd
circles for pdes
lots of PDE progress

@@
 <!--   geom_point(data = df_inat_cleaned, aes(color=year(date))) -->
+
+<!-- # tunicate_gbif <- df_gbif_cleaned %>%  -->
+<!-- #   mutate(lon=as.numeric(longitude),  -->
+<!-- #          lat=as.numeric(latitude), -->
+<!-- #          date=ymd(date), .keep="unused") -->
+
+# m %>% 
+#   addProviderTiles(providers$Stadia.StamenTonerLite) %>% 
+#   addCircles(data=tunicate %>% slice_min(date), color="black") %>% 
+#   addCircles(lat=c(43.934945),lng= c(-69.580869), color="red")
+
+# allen_2017 <- 2*(10*52/(10^5))^2/pi # 10 cm/week
+# v_raft_allen <- 2*sqrt(1.9*allen_2017) #0.01 km/yr
+
+v_raft_low_r <- 2*sqrt(0.8*D_raft) # ~40 km/year
+v_raft_high_r <- 2*sqrt(1.9*D_raft) # ~60 km/year
+
+# 2 km/year observed by Hewitt (1993)
+
+#0.6m/day*180 days repro/yr
+2*sqrt(1.9*(2*((0.6*180)^2)/(pi))) #237 km/yr
+2*sqrt(0.8*(2*((0.6*180)^2)/(pi))) #154 km/yr
+
+#Grosholz for B. leachi predicted 114 km/year and observed 16
+
+We calculated $V$ to be between 39-60 km/year, depending on the population growth rate, with a mean of 49 km/year.
+Species occurrence data suggests that over the past two decades, *B.
+violaceus* has spread at rates around 27-84 km/year, with a mean of 46 km/year.
+While such accurate prediction

### 2024-04-16 PDEs/paper_outline.qmd
lots of PDE progress

--- 2024-04-11 PDEs/paper_outline.Rmd
@@
 author: "Ruby Krasnow"
-date: "2024-04-11"
-output:
-  pdf_document:
-    extra_dependencies: ["tcolorbox", "xcolor"]
+date: "2024-04-16"
+format: pdf
+execute:
+  echo: false
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, include=FALSE}
+#| label: load_packages
+
 library(tidyverse)
+library(gt)
+library(sf)
+library(spocc)
+library(PNWColors)
 ```
@@
 Possibly the only application of this method to a tunicate was performed by Grozholz (1996) for *B.
-leachi*.
-His estimates of $r=1.8$ and $D=1,800$ yielded a predicted rate of spread of $114$ km/year, much faster than the observed rate of $16$ km/year.
+leachi*, but his estimates of $r=1.8$ and $D=1,800$ yielded a predicted rate of spread an order of magnitude larger than observed rates.
 In addition to providing insight on the spread of a specific invasive species of commercial concern, comparing the accuracy of our estimation for *B.
@@
 
-Population growth rate was estimated from Grey (2011) to range from 0.8-1.9.
+## Predicting invasion speed
+The diffusion coefficient, $D$, was calculated using the formula
+$$D=\frac{2M_D(t)^2}{\pi t} $$
+where $M_D (t)$ is the mean displacement of tracked organisms $t$ time units after they begin moving (Andow et al., 1990; Skellam, 1973).
+
+We estimated $M_D$ using information on the dispersal potential of post-larval *Botrylloides sp.* colonies collected by Worcester (1994) (Table 1).
+We used the dispersal distance/speed of rafting colonies rather than swimming larvae to better reflect the likely mechanisms by which *B.
+violaceus* has spread along the coastline.
+Additionally, calculating dispersal potential from larval swimming speeds would be very difficult for a species with a complex reproduction cycle in which the timing and magnitude of spawning events varies between locations and can be highly affected by environmental conditions (SOURCES).
 
 ```{r}
+#| label: calc_D
 colony_day <- 0.225 #0.225 km per colony/day
@@
 
-# allen_2017 <- 2*(10*52/(10^5))^2/pi # 10 cm/week
-# v_raft_allen <- 2*sqrt(1.9*allen_2017) #0.01 km/yr
+obs_r <- c(1.887,1.842,1.665,1.512,0.972,0.984, 0.867,0.778)
+test_v <- function(x) {
+  2*sqrt(x*D_raft)
+}
+preds <- map_vec(obs_r, test_v)
+```
 
-v_raft_low_r <- 2*sqrt(0.8*D_raft) # ~40 km/year
-v_raft_high_r <- 2*sqrt(1.9*D_raft) # ~60 km/year
+```{r}
+#| label: table1_setup
 
-# 2 km/year observed by Hewitt (1993)
+params <- tribble(
+  ~Parameter, ~Value,
+  "Distance traveled per colony per day (km)", 0.225,
+  "Probability that a given colony will raft", 2/3,
+  "Probability that a given rafting colony will reach a suitable habitat", 0.5,
+  "Mean displacement (km/year)", 27.38,
+  "D (km\U00B2/year)", 477.1
+)
+table1 <- gt(params)  %>% 
+  fmt_number(
+    columns = Value,
+    decimals = 3,
+    use_seps = FALSE,
+   drop_trailing_zeros = TRUE
+  ) %>% 
+  cols_width(Parameter ~ pct(78),
+    Value ~pct(22)) %>% 
+  tab_header(title = "Table 1: Calculating the diffusion coefficient")  %>%  opt_align_table_header(align = "left") %>% 
+  tab_options(table.border.top.width = 0,table.font.size =12,
+              heading.title.font.size = 14)
 
-#0.6m/day*180 days repro/yr
-2*sqrt(1.9*(2*((0.6*180)^2)/(pi))) #237 km/yr
-2*sqrt(0.8*(2*((0.6*180)^2)/(pi))) #154 km/yr
+table1 %>% as_latex()
+```
 
-#Grosholz for B. leachi predicted 114 km/year and observed 16
+Eight estimates of population growth rate ($r$) were taken from Grey (2011), ranging from 0.87-1.89.
+Each $r$ value was combined with our estimate of $D$ to find a predicted invasion speed, $V$, in km/year.
+
+## Model validation
+
+We validated our invasion speed estimates by comparing
+
+We retrieved species occurrence data from the Global Biodiversity Information Facility using the package 'spocc' (SOURCE).
+
+We arbitrarily defined the region of interest to be
+
+```{r}
+#| label: query_data
+
+# Define spatial boundary for observations
+polygon_list = list(rbind(c(-72, 41), c(-59.52, 41), c(-59.52, 50), c (-72, 50), c(-72, 41)))
+poly<- st_polygon(polygon_list)
+poly<-st_sfc(poly, crs=4326)
+
+# Query GBIF database
+df_gbif <- occ(query = 'Botrylloides violaceus',
+               from = 'gbif', has_coords = TRUE,
+               geometry = c(-72, 41, -59.52, 50), limit = 1000)
+
+# Convert lon and lat to numbers, remove observations with no date
+df_gbif_cleaned <- occ2df(df_gbif) %>% 
+  mutate(lng=as.double(longitude), lat=as.double(latitude), .keep="unused") %>% 
+  filter(!is.na(date))
 ```
+
+```{r}
+#| label: calc_V
+
+# Convert df to sf
+tunicate <- st_as_sf(df_gbif_cleaned %>% 
+                    select(date, lng, lat) %>% # Select relevant columns
+                    mutate(year = year(date)),  # Add year column
+                    coords = c("lng","lat")) # Specify which cols are coordinates
+
+years <- tunicate %>% group_by(year) %>% summarise(year=mean(year)) # Group by year of obs
+
+# Define function to get and merge all obs from previous years
+get_yrs <- function(yr) { 
+   out <- years %>% filter(year <= yr) %>% st_union()
+   }
+
+years_vec <- c(2005:2024) # Years with observations in the GBIF dataset
+
+years2 <- years %>% mutate(years_agg = map_vec(year, get_yrs))
+
+init_pt <- tunicate %>% 
+  st_set_crs(4326) %>% # Define CRS
+  filter(year==2005) %>% # Select only the year of introduction
+  pull(geometry) # Pull the coordinates
+
+dist <- tunicate %>% 
+  filter(year>2005) %>% # Remove year of introduction
+  st_set_crs(4326) %>% # Define CRS
+  ungroup() %>% 
+  mutate(dist = st_distance(geometry, init_pt)) %>% # Find the distance b/w the initial point and a given observation
+  group_by(year) %>% 
+  st_drop_geometry() %>% 
+  select(year, dist) %>%  # Select only year and distance columns
+  summarize(max_dist = max(dist)) %>% # Find the maximum distance b/w observations and initial point in each year
+  mutate(max_dist=cummax(max_dist)) # Maximum distance the species has been observed from the initial point before and during a given year
+
+speed <- dist %>% mutate(km=as.double(max_dist/1000), # Convert from m to km
+                         speed=km/(year-2005)) %>% # Speed = distance (km)/time (yr)
+                         na.omit()
+
+```
+
+All calculations and data visualization were conducted in R version 4.3.2 (R Core Team 2023). Packages central to the analysis included 'spocc', 'sf', and the tidyverse collection (SOURCES).
+
+# Results
+
+```{r}
+# summary(speed$speed)
+# summary(preds)
+```
+
+```{r}
+#| label: histogram
+#| warning: false
+#| fig-align: "center"
+
+ggplot()+
+  geom_histogram(aes(x=speed %>% filter(year>2006) %>% pull(speed)), bins=22,
+                 color=pnw_palette("Starfish")[2], fill=pnw_palette("Starfish")[2], alpha=0.5)+
+  geom_vline(aes(xintercept = min(preds)),lty="dashed", color="black")+
+  geom_vline(aes(xintercept = max(preds)),lty="dashed", color="black")+
+  labs(x="Observed invasion speed (km/year)", y="Count", lty=NULL)+lims(x=c(0,NA))+
+  theme_classic()+
+ scale_linetype(guide=FALSE)+
+ # scale_linetype_manual(values=c("pred"="dashed"), labels=c("Predicted range"))+
+  theme(axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
+        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
+        text = element_text(size = 13))
+```
+**Figure 1:** Histogram of observed invasion speeds for *B.
+violaceus* in the Northeast Atlantic, calculated from GBIF species occurrence data from 2006-2024.
+Dashed lines represent the minimum and maximum predicted invasion speeds when varying population growth rates were used as input values.
 
@@
 This model attempted to predict the macro-scale dispersal of an invasive tunicate using a simple reaction-diffusion model.
+Given the low biological realism of our simple diffusion model, the very strong agreement between observed and predicted invasion speeds is quite remarkable.
+Most studies that have applied this method of predicting invasion speed characterized their results as successful if estimates were within the same order of magnitude as observed spreading speeds.
+For example, Grosholz (1994) predicted the spreading speed of the tunicate *B.
+leachi* (a member of the same genus as to be *B.
+violaceus*) to be $114$ km/year, which was much faster than the observed rate of $16$ km/year.
+
+In contrast, our predicted speed of 49 km/year was extremely close to the mean of 46 km/year derived from observations of *B.
+violaceus* made in the Gulf of Maine over the past two decades.
+While such accurate prediction is unlikely to be replicable with other data sets, it does suggest that estimating spreading speed from empirical data on local diffusion potential and experimentally-derived population growth rates could be an underutilized approach in invasion ecology.
 

### 2024-04-16 PDEs/pde_mod.R

@@ -1,6 +1,6 @@
 # Ruby Krasnow
 # Code for PDE final project
-# Last modified: April 15, 2024
+# Last modified: April 16, 2024
 
 #load packages
 library(tidyverse)
@@ -8,42 +8,94 @@ library(spocc)
 library(sf)
 library(lwgeom)
 library(leaflet)
+library(PNWColors)
+library(gt)
 
-#define spatial boundary for observations
+# Define spatial boundary for observations
 polygon_list = list(rbind(c(-72, 41), c(-59.52, 41), c(-59.52, 50), c (-72, 50), c(-72, 41)))
 poly<- st_polygon(polygon_list)
 poly<-st_sfc(poly, crs=4326)
 
-
-# GBIF map
+# Query GBIF database
 df_gbif <- occ(query = 'Botrylloides violaceus',
                from = 'gbif', has_coords = TRUE,
                geometry = c(-72, 41, -59.52, 50), limit = 1000)
 
+# Convert lon and lat to numbers, remove observations with no date
 df_gbif_cleaned <- occ2df(df_gbif) %>% 
-  mutate(lng=as.numeric(longitude), lat=as.numeric(latitude), .keep="unused") %>% 
+  mutate(lng=as.doublelongitude, lat=as.double(latitude), .keep="unused") %>% 
   filter(!is.na(date))
 
+# Convert df to sf
 tunicate <- st_as_sf(df_gbif_cleaned %>% 
-                       select(date, lng, lat) %>% 
-                       mutate(year = year(date)),  
-                     coords = c("lng","lat"))
+                    select(date, lng, lat) %>% # Select relevant columns
+                    mutate(year = year(date)),  # Add year column
+                    coords = c("lng","lat")) # Specify which cols are coordinates
 
- years <- tunicate %>% group_by(year) %>% summarise(year=mean(year))
+years <- tunicate %>% group_by(year) %>% summarise(year=mean(year)) # Group by year of obs
  
-
- get_yrs <- function(yr) {
+# Define function to get and merge all obs from previous years
+get_yrs <- function(yr) { 
    out <- years %>% filter(year <= yr) %>% st_union()
    }
  
- 
-years_vec <- c(2005:2024)
+years_vec <- c(2005:2024) # Years with observations in the GBIF dataset
 
-years2<- years %>% mutate(years_agg = map_vec(year, get_yrs))
[…]

### 2024-04-17 PDEs/paper_outline.qmd
pde proj

--- 2024-04-11 PDEs/paper_outline.Rmd
@@
 ---
-title: "Final paper outline"
+title: "Application of a reaction-diffusion equation to the spread of the invasive tunicate *Botrylloides violaceus*"
 author: "Ruby Krasnow"
-date: "2024-04-11"
-output:
-  pdf_document:
-    extra_dependencies: ["tcolorbox", "xcolor"]
+date: "2024-04-17"
+format:
+  pdf:
+    documentclass: article
+    include-in-header:
+      - text: \usepackage{booktabs, caption, longtable, colortbl, array}
+execute:
+  echo: false
+bibliography: references.bib
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, include=FALSE}
+#| label: load_packages
+
 library(tidyverse)
+library(gt)
+library(sf)
+library(spocc)
+library(PNWColors)
+library(leaflet)
+library(webshot2)
+library(ggmap)
+library(rstatix)
 ```
@@
 
-Reaction-diffusion models are a type of partial differential equation (PDE) that have long been used by ecologists to analyze the migration and population dynamics of invasive species (Cosner, 2008).
-Recent work has led to innumerable modifications on the original formulations by Fisher (1937) and Skellam (1951).
-Beyond the fundamental processes of growth and spread, extensions of reaction-diffusion models can be applied to age- and size-structured populations (Neubert and Caswell, 2000), and incorporate environmental heterogeneity, interspecies interactions (competition, predation, mutualism, etc.), Allee effects, invasion control efforts, and long-distance dispersal.
+Reaction-diffusion models are a type of partial differential equation (PDE) that have long been used by ecologists to analyze the migration and population dynamics of invasive species [@cosner2008].
+Recent work has led to innumerable modifications on the original formulations by Fisher [-@fisher1937] and Skellam [-@skellam1951].
+Beyond the fundamental processes of growth and spread, extensions of reaction-diffusion models can be applied to age- and size-structured populations [@neubert2000], and incorporate environmental heterogeneity, interspecies interactions (competition, predation, mutualism, etc.), Allee effects, invasion control efforts, and long-distance dispersal.
 For an introduction to the use of PDEs to represent dispersal in a population ecology context, the reader is directed to Holmes et al.
-(1994) and the seminal papers referenced therein.
+[-@holmes1994] and the seminal papers referenced therein.
 For an accessible review of recent advancements in modeling invasion using reaction-diffusion equations and integro-difference equations, see Hastings et al.
-(2005); for a comprehensive treatment, see the textbooks by Shigesada and Kawasaki (1997) or Lewis et al.
-(2016).
+[-@hastings2005]; for a comprehensive treatment, see the textbooks by Shigesada and Kawasaki [-@shigesada1997] or Lewis et al.
+[-@lewis2016].
 
@@
 *Botrylloides violaceus* has a complex life cycle, including a cyclical growth pattern dependent on both sexual and asexual reproduction.
-The growth and reproduction of colonial tunicates can be highly variable based on a number of environmental factors, including food availability, temperature, salinity, sediment load, and water pollution (Carver et al., 2006; Nelson, 2014).
+The growth and reproduction of colonial tunicates can be highly variable based on a number of environmental factors, including food availability, temperature, salinity, sediment load, and water pollution [@carver2006; @nelson2014].
 The microscale mechanisms behind tunicate dispersion are therefore very complicated, and a fully-resolved spatiotemporal model of *B.
@@
 However, simple reaction-diffusion models can accurately capture population spread without replying on many assumptions and often unknown information about organism-level behavior.
-The value of these models is derived from their tremendous utility in ecological applications, despite abstracting many biological details (Andow et al., 1990).
+The value of these models is derived from their tremendous utility in ecological applications, despite abstracting many biological details [@andow1990].
 By eliminating unnecessary complexity and noise, simple reaction-diffusion models present a tangible method of connecting the mathematical theory of partial differential equations to observed data on invasive species, with implications for real-world ecological and economic problems.
@@
 
+```{=tex}
 \begin{equation}\label{eqn:V} V=\sqrt{4rD} \end{equation}
-
+```
 where $V>0$ is the velocity (distance/time) of the advancing front of the invasive population and $D$ and $r$ are positive constants representing the diffusion coefficient and maximum per capita growth rate, respectively.
@@
 
+```{=tex}
 \begin{align}\label{eqn:IC_1} u(0,x) =
@@
 \end{align}
-
-The reaction-diffusion model describing the population dynamics takes the form
-\begin{equation}\label{eqn:PDE}
+```
+The reaction-diffusion model describing the population dynamics takes the form \begin{equation}\label{eqn:PDE}
 \frac{\partial{u}}{\partial{t}}=\text{D}\frac{\partial{u}^2}{\partial{x^2}}+F(u)
@@
 
+```{=tex}
 \begin{equation}\label{eqn:lin_PDE}
@@
 \end{equation}
-
-This is the simplest version of the general reaction-diffusion model, for a homogeneous, unstructured population growing exponentially and dispersing in a uniform, one-dimensional environment (Andow et al., 1990; Skellam, 1951).
+```
+This is the simplest version of the general reaction-diffusion model, for a homogeneous, unstructured population growing exponentially and dispersing in a uniform, one-dimensional environment [@andow1990; @skellam1951].
 However, it is unlikely that colony-scale dynamics would apply over much larger spatial scales.
@@
 
+```{=tex}
 \begin{equation}\label{eqn:Fisher}
@@
 \end{equation}
-
+```
 The use of nonlinear reaction diffusion models in a biological context is well-established (see Cherniha and Davydovych, 2017).
@@
 
+```{=tex}
 \begin{equation}\label{eqn:comp1}
@@
 \end{equation}
-
-Now, note that for compact initial conditions like (\ref{eqn:IC_1}), there exist positive constants $G$ and $\sigma$ such that we bound the initial distribution from above by a Gaussian distribution:
-\begin{equation}\label{eqn:initGaus}
+```
+Now, note that for compact initial conditions like (\ref{eqn:IC_1}), there exist positive constants $G$ and $\sigma$ such that we bound the initial distribution from above by a Gaussian distribution: \begin{equation}\label{eqn:initGaus}
 U(x,0)=\frac{G}{\sqrt{4\pi\sigma^2}}\exp{\left(-\frac{x^2}{4\sigma^2}\right)}
@@
 
+```{=tex}
 \begin{equation}\label{eqn:comp2} 
@@
 \end{equation}
-
+```
 From (\ref{eqn:comp1}) and (\ref{eqn:comp2}), we use the comparison theorem for PDEs to find that the solution to the linear equation in (\ref{eqn:comp1}) provides an upper bound to the solution of the potentially nonlinear equation (\ref{eqn:PDE}).
@@
 A detailed derivation of the solution of (\ref{eqn:lin_PDE}) can be found in Lewis et al.
-(2016).
+[-@lewis2016].
 In brief, we utilize the change of variables $U(t,x)= W(t,x)e^{rt}$, reducing (\ref{eqn:lin_PDE}) to the standard diffusion equation for $W$.
@@
 
+```{=tex}
 \begin{equation}\label{eqn:PDEsol}
 U(t,x)=\frac{G}{\sqrt{4\pi(\sigma^2+Dt)}}\exp{\left(-\frac{x^2}{4(\sigma^2+Dt)}+rt\right)}\end{equation}
-
+```
 Now, we can quantify the spatial spread of the population by considering how quickly a level set $U(x,t)=U_c$ moves in space, where $U_c$ could be considered a threshold of detection.
 In other words, below $U_c$ it is so difficult to find the invasive organisms that for practical purposes they are not present.
-By rearranging (\ref{eqn:PDEsol}), we can get the following expression
-$$\left(\frac{x}{t}\right) ^2=4rD+\frac{4r\sigma^2}{t}-\frac{4(\sigma^2+Dt)}{t^2}\ln{  \left(\frac{U_{cr}\sqrt{4\pi(\sigma^2+Dt)}}{G}\right) }$$
+By rearranging (\ref{eqn:PDEsol}), we can get the following expression $$\left(\frac{x}{t}\right) ^2=4rD+\frac{4r\sigma^2}{t}-\frac{4(\sigma^2+Dt)}{t^2}\ln{  \left(\frac{U_{cr}\sqrt{4\pi(\sigma^2+Dt)}}{G}\right) }$$
 
@@
 
-This relationship between $V, r$, and $D$ holds true for nonlinear forms of $F(u)$, can be extended to higher dimensions, and does not need to rely on the idea of a threshold of detection— in fact, true advancing fronts with asymptotic speeds given by (\ref{eqn:V}) form (Skellam, 1951, Okubo and Levin, 2001).
-For invasion of a two-dimensional homogeneous habitat described by the reaction-diffusion equation
-\begin{equation}\label{eqn:PDE2D}
+This relationship between $V, r$, and $D$ holds true for nonlinear forms of $F(u)$, can be extended to higher dimensions, and does not need to rely on the idea of a threshold of detection— in fact, true advancing fronts with asymptotic speeds given by (\ref{eqn:V}) form ([@skellam1951; @okubo2001], Okubo and Levin, 2001).
+For invasion of a two-dimensional homogeneous habitat described by the reaction-diffusion equation \begin{equation}\label{eqn:PDE2D}
 \frac{\partial{u(x,y)}}{\partial{t}}=\text{D}\left[\frac{\partial{u}^2}{\partial{x^2}}+\frac{\partial{u}^2}{\partial{y^2}}\right]+F(u),
-\end{equation}
-the radius of the species range should increase linearly with time, with the slope of the increase given by $\sqrt{4rD}$.
+\end{equation} the radius of the species range should increase linearly with time, with the slope of the increase given by $\sqrt{4rD}$.
+
 This theory has been tested for numerous ecological systems and often agrees well with the empirical relationship between measured spread rates, per capita growth rates, and diffusion coefficients.
 Microscale observations can be used to parameterize (\ref{eqn:V}) and predict the spreading rate for a given species at a geographic scale.
-Such spreading-speed analyses have been conducted for a number of terrestrial animal species, including the muskrat, cereal leaf beetle, small cabbage white butterfly, grey squirrel (Andow et al., 1990; Gorsholz, 1996)
+Such spreading-speed analyses have been conducted for a number of terrestrial animal species, including the muskrat, cereal leaf beetle, small cabbage white butterfly, and grey squirrel [@andow1990; @grosholz1996].
 
-Spreading speed theory has also been applied to numerous aquatic species, including the brown trout, green crab, hermit crab, periwinkle, barnacle, mussels, and molluscs (MacDonald et al., 2022; Grosholz, 1996; Marculis and Lui, 2016; Gharouni et al., 2015).
-Possibly the only application of this method to a tunicate was performed by Grozholz (1996) for *B.
-leachi*.
-His estimates of $r=1.8$ and $D=1,800$ yielded a predicted rate of spread of $114$ km/year, much faster than the observed rate of $16$ km/year.
+Spreading speed theory has also been applied to numerous aquatic species, including the brown trout, green crab, hermit crab, periwinkle, barnacle, mussels, and molluscs [@macdonald2022; @grosholz1996; @marculis2016; @gharouni2015].
+Possibly the only application of this method to a tunicate was performed by Grozholz [-@grosholz1996] for *B.
+leachi*, but his estimates of $r=1.8$ and $D=1800$ yielded a predicted rate of spread an order of magnitude larger than observed rates.
 In addition to providing insight on the spread of a specific invasive species of commercial concern, comparing the accuracy of our estimation for *B.
@@
 
-Population growth rate was estimated from Grey (2011) to range from 0.8-1.9.
+## Predicting invasion speed
+
+The diffusion coefficient, $D$, was calculated using the formula $$D=\frac{2M_D(t)^2}{\pi t} $$ where $M_D (t)$ is the mean displacement of tracked organisms $t$ time units after they begin moving [@andow1990]; Skellam, 1973).
+
+We estimated $M_D$ using information on the dispersal potential of post-larval *Botrylloides sp.* colonies collected by Worcester [-@worcester1994] (@tbl-params).
+*B.
+violaceus* can disperse naturally by rafting on eelgrass and other floating vegetation or debris [@carver2006; @worcester1994].
+The dispersal distance/speed of rafting colonies was used rather than swimming larvae to better reflect the likely mechanisms by which *B.
+violaceus* has spread along the coastline.
+Additionally, calculating dispersal potential from larval swimming speeds would be very difficult for a species with a complex reproduction cycle in which the timing and magnitude of spawning events varies between locations and can be highly affected by environmental conditions (SOURCES).
 
 ```{r}
+#| label: calc_D
+
 colony_day <- 0.225 #0.225 km per colony/day
@@
 
-# allen_2017 <- 2*(10*52/(10^5))^2/pi # 10 cm/week
-# v_raft_allen <- 2*sqrt(1.9*allen_2017) #0.01 km/yr
+obs_r <- c(1.887,1.842,1.665,1.512,0.972,0.984, 0.867,0.778)
+test_v <- function(x) {
+  2*sqrt(x*D_raft)
+}
+preds <- map_vec(obs_r, test_v)
+```
 
-v_raft_low_r <- 2*sqrt(0.8*D_raft) # ~40 km/year
-v_raft_high_r <- 2*sqrt(1.9*D_raft) # ~60 km/year
+```{r}
+#| label: tbl-params
+#| tbl-cap: "Calculating the diffusion coefficient"
 
-# 2 km/year observed by Hewitt (1993)
+params <- tribble(
+  ~Parameter, ~Value,
+  "Distance traveled per colony per day (km)", 0.225,
+  "Probability that a given colony will raft", 2/3,
+  "Probability that a given rafting colony will reach a suitable habitat", 0.5,
+  "Mean displacement (km/year)", 27.38,
+  "D (km\U00B2/year)", 477.1
+)
+table1 <- gt(params)  %>% 
+  fmt_number(
+    columns = Value,
+    decimals = 3,
+    use_seps = FALSE,
+   drop_trailing_zeros = TRUE
+  ) %>% 
+  cols_width(Parameter ~ pct(78),
+    Value ~pct(22)) %>% 
+  tab_options(table.font.size =12)
 
-#0.6m/day*180 days repro/yr
-2*sqrt(1.9*(2*((0.6*180)^2)/(pi))) #237 km/yr
-2*sqrt(0.8*(2*((0.6*180)^2)/(pi))) #154 km/yr
+table1
+```
 
-#Grosholz for B. leachi predicted 114 km/year and observed 16
+Eight estimates of population growth rate ($r$) were taken from Grey [-@grey2011], ranging from 0.87-1.89.
+Each $r$ value was combined with our estimate of $D$ to find a predicted invasion speed, $V$, in km/year.
+
+## Model validation
+
+We validated our invasion speed estimates against the rate of boundary displacement for the region in which *B.
+violaceus* has been observed.
+We retrieved species occurrence data from the Global Biodiversity Information Facility (GBIF, https://www.gbif.org/) using the package 'spocc' [@spocc].
+GBIF is an international database of biological data that has been used in over 10,000 peer-reviewed scientific articles and incorporates data from a variety of sources, primarily the citizen science platform iNaturalist.
+Although citizen science data can contain spatial and temporal bias [@dicecco2021], it is a valuable and widely-used tool for monitoring the spread of invasive species [@encarnação2021].
+Furthermore, iNaturalist observations must meet certain quality standards to be included in GBIF (https://www.inaturalist.org/pages/help#GBIFdata), and tunicates like *B.
+violaceus* have a relatively low risk of misidentification by citizen volunteers [@lawson2022].
+
+The region of interest for the GBIF query was arbitrarily defined to be the rectangular region bounded by the coordinates (72°W, 41°N, 59°W, 50°N).
+We only included observations labeled as "Human Observation" in order to exclude preserved specimens located in museum collections.
+Although *B.
+violaceus* may have been introduced to the east coast of the United States as early as the 1970s (SOURCE), the earliest record in the GBIF database is from 2005.
+
+For each year ($t$) since the year of the initial observation ($t_0=2005$), we calculated $V$ as $$V=\frac{\max{\lvert x_{i}}-x_0\rvert}{t-t_0}$$ where $x_0$ was the coordinate location of the 2005 observation and $x_i$ is a single observation within the vector $\bf{x}$ of locations where *B.
+violaceus* had been observed over the period $t_0$ to $t$.
+While there are various methods of quantifying macroscale patterns of species' range expansions, similar boundary displacement approaches are common in invasion ecology [@clark2018; @tobin2007].
+We discarded the very low speed estimated for 2006 (1.1 km/year) as a clear outlier, but all other data was retained.
+
+```{r}
+#| label: query_data
+##| cache: true
+
+# Define spatial boundary for observations
+polygon_list = list(rbind(c(-72, 41), c(-59, 41), c(-59, 50), c (-72, 50), c(-72, 41)))
+poly<- st_polygon(polygon_list)
+poly<-st_sfc(poly, crs=4326)
+
+# Query GBIF database
+df_gbif <- occ(query = 'Botrylloides violaceus',
+               from = 'gbif',
+               has_coords = TRUE,
+               geometry = c(-72, 41, -59, 50),
+               limit = 1000,
+               gbifopts = list(basisOfRecord ="HUMAN_OBSERVATION"))
+
+# Convert lon and lat to numbers, remove observations with no date
+df_gbif_cleaned <- occ2df(df_gbif) %>% 
+  mutate(lng=as.double(longitude), lat=as.double(latitude), .keep="unused") %>% 
+  filter(!is.na(date))
+```
+
+```{r}
+#| label: calc_V
+
+# Convert df to sf
+tunicate <- st_as_sf(df_gbif_cleaned %>% 
+                    select(date, lng, lat) %>% # Select relevant columns
+                    mutate(year = year(date)),  # Add year column
+                    coords = c("lng","lat")) # Specify which cols are coordinates
+
+years <- tunicate %>% group_by(year) %>% summarise(year=mean(year)) # Group by year of obs
+
+# Define function to get and merge all obs from previous years
+get_yrs <- function(yr) { 
+   out <- years %>% filter(year <= yr) %>% st_union()
+   }
+
+years_vec <- c(2005:2024) # Years with observations in the GBIF dataset
+
+years2 <- years %>% mutate(years_agg = map_vec(year, get_yrs))
+
+init_pt <- tunicate %>% 
+  st_set_crs(4326) %>% # Define CRS
+  filter(year==2005) %>% # Select only the year of introduction
+  pull(geometry) # Pull the coordinates
+
+dist <- tunicate %>% 
+  filter(year>2005) %>% # Remove year of introduction
+  st_set_crs(4326) %>% # Define CRS
+  ungroup() %>% 
+  mutate(dist = st_distance(geometry, init_pt)) %>% # Find the distance b/w the initial point and a given observation
+  group_by(year) %>% 
+  st_drop_geometry() %>% 
+  select(year, dist) %>%  # Select only year and distance columns
+  summarize(max_dist = max(dist)) %>% # Find the maximum distance b/w observations and initial point in each year
+  mutate(max_dist=cummax(max_dist)) # Maximum distance the species has been observed from the initial point before and during a given year
+
+speed <- dist %>% mutate(km=as.double(max_dist/1000), # Convert from m to km
+                         speed=km/(year-2005)) %>% # Speed = distance (km)/time (yr)
+                         na.omit() %>% 
+  mutate(outlier = is_outlier(speed), extreme=is_extreme(speed))
+
+speed2 <- speed %>% 
+  filter(year>2006)
+```
+
+All calculations were performed in R version 4.3.2 [@rcoreteam2023].
+R packages central to data analysis and visualization included 'spocc', 'sf', 'leaflet', and the tidyverse collection [@sf; @tidyverse; @spocc; @leaflet].
+
+# Results
+
+![A map of the spread of *B.
+violaceus* in the Gulf of Maine.
+Black dots represent confirmed observations of *B.
+violaceus* from the GBIF database.
+Colored circles (centered at the location of the first observation) illustrate the expanding range.](map_radii.png){fig-align="center"}
+
+```{r}
+#| label: tbl-summaries
+#| tbl-cap: "Summary statistics for observed and predicted *B. violaceus* invasion speed (km/year)."
+
+test<-data.frame("Observed"=c(summary(speed2$speed)),
+           "Predicted"=c(summary(preds))) %>% rownames_to_column() %>% gt() %>% 
+  fmt_number(
+    columns = c(Observed, Predicted),
+    decimals = 1,
+    use_seps = FALSE
+  )
+test
+```
+
+```{r}
+#| label: fig-hist
+#| warning: false
+#| fig-align: "center"
+#| fig-cap: "Histogram of observed invasion speeds for *B.
+violaceus* in the Gulf of Maine, calculated from GBIF species occurrence data from 2006-2024.
+Dashed lines represent the minimum and maximum predicted invasion speeds when varying population growth rates were used as input values."
+#| fig-pos: "H"
+
+ggplot()+
+  geom_histogram(aes(x=speed2 %>% pull(speed)), bins=23,
+                 color=pnw_palette("Starfish")[2], fill=pnw_palette("Starfish")[2], alpha=0.5)+
+  geom_vline(aes(xintercept = min(preds)),lty="dashed", color="black")+
+  geom_vline(aes(xintercept = max(preds)),lty="dashed", color="black")+
+  labs(x="Observed invasion speed (km/year)", y="Count", lty=NULL)+
+  lims(x=c(0,NA))+
+  theme_classic()+
+ scale_linetype(guide=FALSE)+
+  theme(axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
+        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
+        text = element_text(size = 13))
 ```
@@
 This model attempted to predict the macro-scale dispersal of an invasive tunicate using a simple reaction-diffusion model.
+Given the low biological realism of our simple diffusion model, the very strong agreement between observed and predicted invasion speeds is quite remarkable.
+Most studies that have applied this method of predicting invasion speed characterized their results as successful if estimates were within the same order of magnitude as observed spreading speeds.
+For example, Grosholz [-@grosholz1996] predicted the spreading speed of the tunicate *B.
+leachi* (a member of the same genus as to be *B.
+violaceus*) to be $114$ km/year, which was much faster than the observed rate of $16$ km/year.
+
+In contrast, our predicted speed of 49 km/year was extremely close to the mean of 41 km/year derived from observations of *B.
+violaceus* made in the Gulf of Maine over the past two decades (@tbl-summaries).
+Both the observed and predicted invasion speeds were generally in the range of 40-60 km/year ([Fig @fig-hist]).
+While such accurate prediction is unlikely to be replicable with other data sets, it does suggest that estimating spreading speed from empirical data on local diffusion potential and experimentally-derived population growth rates could be an underutilized approach in invasion ecology.
 
@@
 Other diffusion models incorporate Allee effects, an ecological phenomenon characterized by decreased individual fitness (lower per capita growth rates) at small population sizes.
-Accounting for Allee effects in continuous reaction-diffusion models or analogous discrete integrodifference models can alter predictions about the potential success or failure of an invasion, rates of spread, and plausible ranges of an invasive species (Keitt et al., 2001; Kot et al., 1996; Liu and Kot, 2019).
+Accounting for Allee effects in continuous reaction-diffusion models or analogous discrete integrodifference models can alter predictions about the potential success or failure of an invasion, rates of spread, and plausible ranges of an invasive species [@keitt2001; @kot1996; @liu2019].
 
@@
 
-*B.
-violaceus* can disperse naturally by rafting on eelgrass and other floating vegetation or debris (Carver et al., 2006; Worcester, 1994).
-Human-mediated introduction via hull fouling on recreational or commercial watercraft and aquaculture activities are also major pathways by which invasive tunicates are introduced to new locations (Carver et al., 2006).
+Human-mediated introduction via hull fouling on recreational or commercial watercraft and aquaculture activities are also major pathways by which invasive tunicates are introduced to new locations [@carver2006].
 Although *B.
-violaceus* has a relatively low attachment strength, the small drag coefficient associated with their low profile means that its displacement velocity is above the top speeds of sailboats, barges, and other slower-moving vessels (Clarke Murray et al., 2012).
+violaceus* has a relatively low attachment strength, the small drag coefficient associated with their low profile means that its displacement velocity is above the top speeds of sailboats, barges, and other slower-moving vessels [@clarkemurray2012].
 Clarke Murray et al.
-(2011) found that *B.
+[-@clarkemurray2011] found that *B.
 violaceus* was attached to the hulls of 9.8% of recreational boats at nine distinct marinas in British Columbia.
@@
 Future work could utilize discrete integrodifference equations with leptokurtic dispersion kernels to account for transport via hull fouling and other potential vectors of long-distance dispersion.
-The potential for long-distance dispersal events via natural or human-mediated pathways often prompts the use of integrodifference equations for diffusion models of invasive populations (Lutscher, 2019; Lewis et al., 2016; Neubert and Caswell, 2000; Kot and Neubert, 2008).
+The potential for long-distance dispersal events via natural or human-mediated pathways often prompts the use of integrodifference equations for diffusion models of invasive populations [@lewis2016; @lutscher2019; @kot2008; @neubert2000].
 These models assume that growth and dispersal occur in distinct and asynchronous stages.
@@
 violaceus* to utilize multiple forms of reproduction and the fact that both larvae and post-larvae can contribute to the colonization of new habitat potentially violates the assumption of temporally separated growth and dispersion.
-However, stratified models that separate spatially continuous short-distance dispersal from discrete long-distance dispersal can be valuable tools to reconstruct realistic invasion dynamics (e.g., Gilioli et al., 2013) and their potential for modeling the spread of *B.
+However, stratified models that separate spatially continuous short-distance dispersal from discrete long-distance dispersal can be valuable tools to reconstruct realistic invasion dynamics [e.g., @gilioli2013] and their potential for modeling the spread of *B.
 violaceus* merits further investigation.
@@
 For example, Roques et al.
-(2011) used a "statistical-reaction–diffusion approach" to analyze the expansion of the pine processionary moth based on moth nest presence/absence data.
+[-@roques2011] used a "statistical-reaction–diffusion approach" to analyze the expansion of the pine processionary moth based on moth nest presence/absence data.
 Using a "mechanistic-statistical" diffusion model assuming logistical growth and imperfect species observation data, Louvrier et al.
-(2020) was able to forecast the expanding distribution of the grey wolf in France.
+[-@louvrier2020] were able to forecast the expanding distribution of the grey wolf in France.
 Applying similar methods to data on *B.
@@
 
-Finally, competition-diffusion models that represent the population growth and spread of two competing species in the same habitat admit similar traveling front solutions to those described above for the simple reaction-diffusion equation (Carrère, 2018).
-It is often interesting to investigate the stability and long-term behavior of a system of competition-diffusion models for different parameter values, including varying the propagation speeds of the two organisms or the differential impacts of the competitive interactions on each species (Carrère, 2018; Ducrot et al., 2014).
+Finally, competition-diffusion models that represent the population growth and spread of two competing species in the same habitat admit similar traveling front solutions to those described above for the simple reaction-diffusion equation [@carrère2018].
+It is often interesting to investigate the stability and long-term behavior of a system of competition-diffusion models for different parameter values, including varying the propagation speeds of the two organisms or the differential impacts of the competitive interactions on each species [@carrère2018; @ducrot2014].
 Experimental evidence suggests that *B.
-violaceus* colonies can outcompete *Botryllus schlosseri*, a different invasive tunicate that has been equally or more problematic for the New England aquaculture industry (Gittenberger and Moons, 2011; Dijkstra et al., 2007).
+violaceus* colonies can outcompete *Botryllus schlosseri*, a different invasive tunicate that has been equally or more problematic for the New England aquaculture industry [@dijkstra2007; @gittenberger2011].
 However, these studies have been limited to small-scale experiments conducted on a single substrate (e.g., an oyster cage, SOURCE).

### 2024-04-17 PDEs/pde_mod.R

@@ -10,20 +10,21 @@ library(lwgeom)
 library(leaflet)
 library(PNWColors)
 library(gt)
+library(mapview)
 
 # Define spatial boundary for observations
-polygon_list = list(rbind(c(-72, 41), c(-59.52, 41), c(-59.52, 50), c (-72, 50), c(-72, 41)))
+polygon_list = list(rbind(c(-72, 41), c(-59, 41), c(-59, 50), c (-72, 50), c(-72, 41)))
 poly<- st_polygon(polygon_list)
 poly<-st_sfc(poly, crs=4326)
 
 # Query GBIF database
 df_gbif <- occ(query = 'Botrylloides violaceus',
                from = 'gbif', has_coords = TRUE,
-               geometry = c(-72, 41, -59.52, 50), limit = 1000)
+               geometry = c(-72, 41, -59, 50), limit = 1000)
 
 # Convert lon and lat to numbers, remove observations with no date
 df_gbif_cleaned <- occ2df(df_gbif) %>% 
-  mutate(lng=as.doublelongitude, lat=as.double(latitude), .keep="unused") %>% 
+  mutate(lng=as.double(longitude), lat=as.double(latitude), .keep="unused") %>% 
   filter(!is.na(date))
 
 # Convert df to sf
@@ -107,9 +108,9 @@ ggplot()+geom_sf(data=circles3)
 # Create a continuous palette function
 pal <- colorNumeric(
   palette = "viridis",
-  domain = circles3$year)
+  domain = years_vec)
 
-m <- leaflet() %>% fitBounds(-72,41,-59.52,50) %>% 
+m <- leaflet() %>% fitBounds(-72,41,-59,50) %>% 
   addProviderTiles(providers$Stadia.StamenTonerLite)
 
 m  %>% 
@@ -124,7 +125,7 @@ radii <- speed %>% select(year, max_dist, km) %>% arrange(-year)
 
 map <- leaflet(options = leafletOptions(zoomControl = FALSE)) %>% 
   fitBounds(-72,41,-59.52,50) %>% 
-  addProviderTiles(providers$Stadia.StamenTonerLite) %>% 
+  addProviderTiles(providers$CartoDB.Positron) %>% 
   addCircles(data=tunicate, color="black") %>% 
   addCircles(data=radii, 
              lng=-70.03979, 
[…]

### 2024-04-17 PDEs/math245_old.Rmd
close to done with pde proj

@@
 While such accurate prediction
+
+table1 <- gt(params)  %>% 
+  fmt_number(
+    columns = Value,
+    decimals = 3,
+    use_seps = FALSE,
+    drop_trailing_zeros = TRUE
+  ) %>% 
+  cols_width(Parameter ~ pct(78),
+             Value ~pct(22)) %>% 
+  tab_header(title = "Table 1: Calculating the diffusion coefficient")  %>%  opt_align_table_header(align = "left") %>% 
+  tab_options(table.border.top.width = 0,table.font.size =12,
+              heading.title.font.size = 14)
+
+table1 %>% gt::as_latex()
+
+# circles <- years2 %>%
+#   rowwise() %>%
+#   mutate(circle = st_minimum_bounding_circle(years_agg), .keep="unused") %>% ungroup()
+# 
+# circles2 <- circles %>% st_drop_geometry() %>% 
+#   select(year, circle)
+# 
+# circles3 <- st_as_sf(circles2) %>% arrange(-year)
+# 
+# ggplot()+geom_sf(data=circles3)
+
+m <- leaflet() %>% fitBounds(-72,41,-59,50) %>% 
+  addProviderTiles(providers$Stadia.StamenTonerLite)
+
+m  %>% 
+  addCircles(data=tunicate %>% slice_min(date), color="black") %>% 
+  addPolygons(data=circles3, fillOpacity = 0.2, color = ~pal(year))
+
+m  %>% 
+  addCircles(data=tunicate, color="black") %>% 
+  addPolygons(data=circles3, fillOpacity = 0.2, color = ~pal(year))

### 2024-04-17 PDEs/paper_outline.qmd
close to done with pde proj

@@
 ---
-title: "Application of a reaction-diffusion equation to the spread of the invasive tunicate *Botrylloides violaceus*"
+title: "Application of a reaction-diffusion equation to the spread of the invasive ascidian *Botrylloides violaceus*"
 author: "Ruby Krasnow"
@@
   pdf:
-    documentclass: article
+    documentclass: scrartcl
     include-in-header:
-      - text: \usepackage{booktabs, caption, longtable, colortbl, array}
+      - text: \usepackage{booktabs, caption, longtable, colortbl, array, graphicx}
 execute:
@@
 bibliography: references.bib
+csl: molecular-ecology.csl
+link-external-newwindow: true 
 ---
 
-```{r, include=FALSE}
-#| label: load_packages
+```{r}
+#| echo: false
+#| message: false
+#| results: "hide"
 
-library(tidyverse)
-library(gt)
-library(sf)
-library(spocc)
-library(PNWColors)
-library(leaflet)
-library(webshot2)
-library(ggmap)
-library(rstatix)
+source("pde_mod.R")
 ```
@@
 Here, we use a simple two-dimensional reaction-diffusion model to represent the spread of the invasive ascidian *Botrylloides violaceus* in the Gulf of Maine.
-Colonial tunicates (or ascidians) are compound organisms formed by a mass of small zooids organized within a common gelatinous matrix.
-These colonies often form visible crusts on rocks, vessel hulls, and other artificial floating structures and submerged objects.
-Colonial tunicates also commonly grow on macroalgae and bivalves such as shellfish and oysters, making them a nuisance for the low-trophic aquaculture industry.
+Colonial ascidians (or tunicates) are compound organisms formed by a mass of small zooids organized within a common gelatinous matrix [@wagstaff2017].
+These colonies often form visible crusts on rocks, vessel hulls, and other artificial floating structures and submerged objects [@wagstaff2017].
+Colonial ascidians also commonly grow on macroalgae and bivalves such as mussels and oysters.
+Ascidian growth on cultivation equipment and the cultivated organisms themselves can have many harmful effects, including smothering the bivalves, reducing local water circulation, consuming nutrients and dissolved oxygen, and causing structural damage to cultivation equipment [@cahill2012; @fitridge2012; @aldred2014].
+The detrimental impacts of ascidians on farm production and efficiency present significant economic costs for the North American aquaculture industry [@fitridge2012].
 
-In New England, economic impacts from biofouling are estimated to be NUMBER. Add additional background about impact of *Botrylloides violaceus* on kelp and oyster aquaculture in the GOM here.
+In New England and Atlantic Canada, aquaculturists have had little success controlling the spread or impact of invasive ascidians, particularly species like *B.
+violaceus* that have few natural predators and can outcompete other fouling organisms [@carman2016; @carver2006; @gittenberger2011; @bullard2013].
+*B.
+violaceus* is the most common fouling ascidian at shellfish aquaculture sites along the East coast of North America and the most widely distributed nonindigenous ascidian in British Columbia.
+[@simkanin2013; @carman2010a].
 
-Most common fouling tunicate at shellfish aquaculture sites along the East coast (Carman et al., 2010)
-
-*Botrylloides violaceus* has a complex life cycle, including a cyclical growth pattern dependent on both sexual and asexual reproduction.
+*B.* *violaceus* has a complex life cycle, including a cyclical growth pattern dependent on both sexual and asexual reproduction.
 The growth and reproduction of colonial tunicates can be highly variable based on a number of environmental factors, including food availability, temperature, salinity, sediment load, and water pollution [@carver2006; @nelson2014].
@@
 
-Here, we follow the derivation method of Lewis et al. (2016) to justify our use of the following equation to represent the rate of spread of *B. violaceus*:
+Here, we follow the derivation method of Lewis et al. [-@lewis2016] to justify our use of the following equation to represent the rate of spread of *B. violaceus*:
 
-```{=tex}
 \begin{equation}\label{eqn:V} V=\sqrt{4rD} \end{equation}
-```
+
 where $V>0$ is the velocity (distance/time) of the advancing front of the invasive population and $D$ and $r$ are positive constants representing the diffusion coefficient and maximum per capita growth rate, respectively.
@@
 
-```{=tex}
 \begin{align}\label{eqn:IC_1} u(0,x) =
@@
 \end{align}
-```
+
 The reaction-diffusion model describing the population dynamics takes the form \begin{equation}\label{eqn:PDE}
@@
 Determining the correct form of the population growth term in (\ref{eqn:PDE}) is challenging.
-Colonial ascidians generally display indeterminate growth, meaning that colony area tends to increase exponentially with time (Jackson, 1977).
+Colonial ascidians generally display indeterminate growth, meaning that colony area tends to increase exponentially with time [@jackson1977].
 If we assumed simple linear population growth, our PDE would take the form
 
-```{=tex}
 \begin{equation}\label{eqn:lin_PDE}
@@
 \end{equation}
-```
+
 This is the simplest version of the general reaction-diffusion model, for a homogeneous, unstructured population growing exponentially and dispersing in a uniform, one-dimensional environment [@andow1990; @skellam1951].
@@
 
-```{=tex}
 \begin{equation}\label{eqn:Fisher}
@@
 \end{equation}
-```
-The use of nonlinear reaction diffusion models in a biological context is well-established (see Cherniha and Davydovych, 2017).
+
+We can generalize our section of the coastline of length $L$ to an $n$-dimensional domain $\Omega$, with boundary $\partial\Omega$.
+If the function $u(t,x)$ and $\partial\Omega$ are smooth and $\phi(x)$ is continuous on the $\Omega \cup \partial\Omega$, the problem presented by (\ref{eqn:IC_1}) and (\ref{eqn:PDE}) will be well posed in the sense that it will have a unique solution $u(x,t)$ on $\Omega \times (0,T]$ for some $T > 0$, with $u(x,t)$ depending continuously on $\phi(x)$ [@cosner2008].
+Further discussion of the existence, uniqueness, and smoothness of solutions to reaction-diffusion equations can be found in @cosner2008, along with a review of the relevant maximum principle and comparison theorem and a discussion of the stability and propagation of solutions.
+
+The use of nonlinear reaction diffusion models in a biological context is well-established [see @cherniha2017].
 However, note that for biologically reasonable choices of $F(u)$, the growth function will be bounded by the linear function, such that $F(u) \leq ru$.
@@
 
-```{=tex}
 \begin{equation}\label{eqn:comp1}
@@
 \end{equation}
-```
+
 Now, note that for compact initial conditions like (\ref{eqn:IC_1}), there exist positive constants $G$ and $\sigma$ such that we bound the initial distribution from above by a Gaussian distribution: \begin{equation}\label{eqn:initGaus}
@@
 
-```{=tex}
 \begin{equation}\label{eqn:comp2} 
@@
 \end{equation}
-```
+
 From (\ref{eqn:comp1}) and (\ref{eqn:comp2}), we use the comparison theorem for PDEs to find that the solution to the linear equation in (\ref{eqn:comp1}) provides an upper bound to the solution of the potentially nonlinear equation (\ref{eqn:PDE}).
@@
 
-```{=tex}
 \begin{equation}\label{eqn:PDEsol}
 U(t,x)=\frac{G}{\sqrt{4\pi(\sigma^2+Dt)}}\exp{\left(-\frac{x^2}{4(\sigma^2+Dt)}+rt\right)}\end{equation}
-```
+
 Now, we can quantify the spatial spread of the population by considering how quickly a level set $U(x,t)=U_c$ moves in space, where $U_c$ could be considered a threshold of detection.
@@
 
-This relationship between $V, r$, and $D$ holds true for nonlinear forms of $F(u)$, can be extended to higher dimensions, and does not need to rely on the idea of a threshold of detection— in fact, true advancing fronts with asymptotic speeds given by (\ref{eqn:V}) form ([@skellam1951; @okubo2001], Okubo and Levin, 2001).
+This relationship between $V, r$, and $D$ holds true for nonlinear forms of $F(u)$, can be extended to higher dimensions, and does not need to rely on the idea of a threshold of detection— in fact, true advancing fronts with asymptotic speeds given by (\ref{eqn:V}) form [@skellam1951; @okubo2001; @holmes1994].
 For invasion of a two-dimensional homogeneous habitat described by the reaction-diffusion equation \begin{equation}\label{eqn:PDE2D}
@@
 Microscale observations can be used to parameterize (\ref{eqn:V}) and predict the spreading rate for a given species at a geographic scale.
-Such spreading-speed analyses have been conducted for a number of terrestrial animal species, including the muskrat, cereal leaf beetle, small cabbage white butterfly, and grey squirrel [@andow1990; @grosholz1996].
+Such spreading-speed analyses have been conducted for a number of terrestrial animal species, including the muskrat, cereal leaf beetle, European starling, small cabbage white butterfly, and grey squirrel [@andow1990; @grosholz1996; @holmes1993].
 
@@
 
-The diffusion coefficient, $D$, was calculated using the formula $$D=\frac{2M_D(t)^2}{\pi t} $$ where $M_D (t)$ is the mean displacement of tracked organisms $t$ time units after they begin moving [@andow1990]; Skellam, 1973).
+The diffusion coefficient, $D$, was calculated using the formula $$D=\frac{2M_D(t)^2}{\pi t} $$ where $M_D (t)$ is the mean displacement of tracked organisms $t$ time units after they begin moving [@andow1990; @othmer1988].
 
-We estimated $M_D$ using information on the dispersal potential of post-larval *Botrylloides sp.* colonies collected by Worcester [-@worcester1994] (@tbl-params).
+We estimated $M_D$ using information on the dispersal potential of post-larval *Botrylloides sp.* colonies collected by Worcester [-@worcester1994] (@tbl-parameters).
 *B.
-violaceus* can disperse naturally by rafting on eelgrass and other floating vegetation or debris [@carver2006; @worcester1994].
+violaceus* can disperse naturally by rafting on eelgrass and other floating vegetation or debris [@carver2006; @worcester1994; @carman2010].
 The dispersal distance/speed of rafting colonies was used rather than swimming larvae to better reflect the likely mechanisms by which *B.
 violaceus* has spread along the coastline.
-Additionally, calculating dispersal potential from larval swimming speeds would be very difficult for a species with a complex reproduction cycle in which the timing and magnitude of spawning events varies between locations and can be highly affected by environmental conditions (SOURCES).
 
 ```{r}
-#| label: calc_D
-
-colony_day <- 0.225 #0.225 km per colony/day
-raft_prob <- (2/3) # probability that a given colony will raft
-hab_prob <- 0.5 # probability that a given rafting colony will reach a suitable habitat
-km_year_raft <- colony_day*raft_prob*hab_prob*365
-D_raft <- 2*(km_year_raft^2)/(pi)
-
-obs_r <- c(1.887,1.842,1.665,1.512,0.972,0.984, 0.867,0.778)
-test_v <- function(x) {
-  2*sqrt(x*D_raft)
-}
-preds <- map_vec(obs_r, test_v)
-```
-
-```{r}
-#| label: tbl-params
+#| label: tbl-parameters
 #| tbl-cap: "Calculating the diffusion coefficient"
-
-params <- tribble(
-  ~Parameter, ~Value,
-  "Distance traveled per colony per day (km)", 0.225,
-  "Probability that a given colony will raft", 2/3,
-  "Probability that a given rafting colony will reach a suitable habitat", 0.5,
-  "Mean displacement (km/year)", 27.38,
-  "D (km\U00B2/year)", 477.1
-)
-table1 <- gt(params)  %>% 
-  fmt_number(
-    columns = Value,
-    decimals = 3,
-    use_seps = FALSE,
-   drop_trailing_zeros = TRUE
-  ) %>% 
-  cols_width(Parameter ~ pct(78),
-    Value ~pct(22)) %>% 
-  tab_options(table.font.size =12)
 
@@
 
-Eight estimates of population growth rate ($r$) were taken from Grey [-@grey2011], ranging from 0.87-1.89.
+Eight estimates of population growth rate ($r$) were taken from Grey [-@grey2011], ranging from 0.78-1.89.
 Each $r$ value was combined with our estimate of $D$ to find a predicted invasion speed, $V$, in km/year.
@@
 violaceus* has been observed.
-We retrieved species occurrence data from the Global Biodiversity Information Facility (GBIF, https://www.gbif.org/) using the package 'spocc' [@spocc].
+We retrieved species occurrence data from the Global Biodiversity Information Facility (GBIF, \url{https://www.gbif.org/}) using the package 'spocc' [@spocc].
 GBIF is an international database of biological data that has been used in over 10,000 peer-reviewed scientific articles and incorporates data from a variety of sources, primarily the citizen science platform iNaturalist.
+
 Although citizen science data can contain spatial and temporal bias [@dicecco2021], it is a valuable and widely-used tool for monitoring the spread of invasive species [@encarnação2021].
-Furthermore, iNaturalist observations must meet certain quality standards to be included in GBIF (https://www.inaturalist.org/pages/help#GBIFdata), and tunicates like *B.
+Furthermore, iNaturalist observations must meet certain quality standards to be included in GBIF (\url{https://www.inaturalist.org/pages/help#GBIFdata}), and tunicates like *B.
 violaceus* have a relatively low risk of misidentification by citizen volunteers [@lawson2022].
@@
 Although *B.
-violaceus* may have been introduced to the east coast of the United States as early as the 1970s (SOURCE), the earliest record in the GBIF database is from 2005.
+violaceus* may have been introduced to the east coast of the United States as early as the 1970s [@bock2011; @dijkstra2007], the earliest record in the GBIF database that did not have any issues or flags is from 2005.
 
-For each year ($t$) since the year of the initial observation ($t_0=2005$), we calculated $V$ as $$V=\frac{\max{\lvert x_{i}}-x_0\rvert}{t-t_0}$$ where $x_0$ was the coordinate location of the 2005 observation and $x_i$ is a single observation within the vector $\bf{x}$ of locations where *B.
+For each year ($t$) since the year of the initial observation ($t_0=2005$), we calculated $V$ as $$V=\frac{\max{\lvert x_{i}}-x_0\rvert}{t-t_0}$$ where $x_0$ was the coordinate location of the 2005 observation and $x_i$ is a single observation within the vector ${x}$ of locations where *B.
 violaceus* had been observed over the period $t_0$ to $t$.
@@
 We discarded the very low speed estimated for 2006 (1.1 km/year) as a clear outlier, but all other data was retained.
-
-```{r}
-#| label: query_data
-##| cache: true
-
-# Define spatial boundary for observations
-polygon_list = list(rbind(c(-72, 41), c(-59, 41), c(-59, 50), c (-72, 50), c(-72, 41)))
-poly<- st_polygon(polygon_list)
-poly<-st_sfc(poly, crs=4326)
-
-# Query GBIF database
-df_gbif <- occ(query = 'Botrylloides violaceus',
-               from = 'gbif',
-               has_coords = TRUE,
-               geometry = c(-72, 41, -59, 50),
-               limit = 1000,
-               gbifopts = list(basisOfRecord ="HUMAN_OBSERVATION"))
-
-# Convert lon and lat to numbers, remove observations with no date
-df_gbif_cleaned <- occ2df(df_gbif) %>% 
-  mutate(lng=as.double(longitude), lat=as.double(latitude), .keep="unused") %>% 
-  filter(!is.na(date))
-```
-
-```{r}
-#| label: calc_V
-
-# Convert df to sf
-tunicate <- st_as_sf(df_gbif_cleaned %>% 
-                    select(date, lng, lat) %>% # Select relevant columns
-                    mutate(year = year(date)),  # Add year column
-                    coords = c("lng","lat")) # Specify which cols are coordinates
-
-years <- tunicate %>% group_by(year) %>% summarise(year=mean(year)) # Group by year of obs
-
-# Define function to get and merge all obs from previous years
-get_yrs <- function(yr) { 
-   out <- years %>% filter(year <= yr) %>% st_union()
-   }
-
-years_vec <- c(2005:2024) # Years with observations in the GBIF dataset
-
-years2 <- years %>% mutate(years_agg = map_vec(year, get_yrs))
-
-init_pt <- tunicate %>% 
-  st_set_crs(4326) %>% # Define CRS
-  filter(year==2005) %>% # Select only the year of introduction
-  pull(geometry) # Pull the coordinates
-
-dist <- tunicate %>% 
-  filter(year>2005) %>% # Remove year of introduction
-  st_set_crs(4326) %>% # Define CRS
-  ungroup() %>% 
-  mutate(dist = st_distance(geometry, init_pt)) %>% # Find the distance b/w the initial point and a given observation
-  group_by(year) %>% 
-  st_drop_geometry() %>% 
-  select(year, dist) %>%  # Select only year and distance columns
-  summarize(max_dist = max(dist)) %>% # Find the maximum distance b/w observations and initial point in each year
-  mutate(max_dist=cummax(max_dist)) # Maximum distance the species has been observed from the initial point before and during a given year
-
-speed <- dist %>% mutate(km=as.double(max_dist/1000), # Convert from m to km
-                         speed=km/(year-2005)) %>% # Speed = distance (km)/time (yr)
-                         na.omit() %>% 
-  mutate(outlier = is_outlier(speed), extreme=is_extreme(speed))
-
-speed2 <- speed %>% 
-  filter(year>2006)
-```
 
@@
 
-![A map of the spread of *B.
-violaceus* in the Gulf of Maine.
-Black dots represent confirmed observations of *B.
-violaceus* from the GBIF database.
-Colored circles (centered at the location of the first observation) illustrate the expanding range.](map_radii.png){fig-align="center"}
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.75\textwidth]{map_radii.png}
+\caption{A map of the spread of \textit{B.
+violaceus} in the Gulf of Maine.
+Black dots represent confirmed observations of \textit{B.
+violaceus} from the GBIF database.
+Colored circles (centered at the location of the first observation) illustrate the expanding range.}
+\end{figure}
+
+Predicted invasion speeds using (\ref{eqn:V}) varied from 38.5 to 60.0 km/year, corresponding to the lowest ($r=0.78$) and highest ($r=1.89$) population growth rates assumed for *B.
+violaceus* (@tbl-summaries).
 
@@
 
-test<-data.frame("Observed"=c(summary(speed2$speed)),
-           "Predicted"=c(summary(preds))) %>% rownames_to_column() %>% gt() %>% 
-  fmt_number(
-    columns = c(Observed, Predicted),
-    decimals = 1,
-    use_seps = FALSE
-  )
-test
+summaries_df
 ```
+
+There was significant overlap between the range of predicted speeds and the distribution of "true" invasion speeds, estimated by dividing the distance between the introduction point and furthest observation of *B.
+violaceus* in each year by the time elapsed since introduction (@fig-hist).
+The mean observed and predicted invasion speeds differed by fewer than 9 km/year, a remarkably small disparity (@tbl-summaries).
 
@@
 #| fig-pos: "H"
+#| fig-height: 2.5
+#| fig-width: 4.5
 
-ggplot()+
-  geom_histogram(aes(x=speed2 %>% pull(speed)), bins=23,
-                 color=pnw_palette("Starfish")[2], fill=pnw_palette("Starfish")[2], alpha=0.5)+
-  geom_vline(aes(xintercept = min(preds)),lty="dashed", color="black")+
-  geom_vline(aes(xintercept = max(preds)),lty="dashed", color="black")+
-  labs(x="Observed invasion speed (km/year)", y="Count", lty=NULL)+
-  lims(x=c(0,NA))+
-  theme_classic()+
- scale_linetype(guide=FALSE)+
-  theme(axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
-        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
-        text = element_text(size = 13))
+fig2
 ```
@@
 
-In contrast, our predicted speed of 49 km/year was extremely close to the mean of 41 km/year derived from observations of *B.
+In contrast, our predicted speed of ~49 km/year was extremely close to the mean of ~40 km/year derived from observations of *B.
 violaceus* made in the Gulf of Maine over the past two decades (@tbl-summaries).
@@
 First, there exist many other possible formulations for the term representing population growth.
-The Fisher-KPP equation, which assumes logistic rather than exponential growth, has been well-studied from a mathematical perspective and has seen widespread application across ecological contexts (SOURCES).
+The Fisher equation, which assumes logistic rather than exponential growth, has been well-studied from a mathematical perspective and has seen widespread application across ecological contexts [e.g., @holmes1994; @berestycki2009].
 Other diffusion models incorporate Allee effects, an ecological phenomenon characterized by decreased individual fitness (lower per capita growth rates) at small population sizes.
@@
 
-Finally, competition-diffusion models that represent the population growth and spread of two competing species in the same habitat admit similar traveling front solutions to those described above for the simple reaction-diffusion equation [@carrère2018].
+Finally, competition-diffusion models that represent the population growth and spread of two competing species in the same habitat admit similar wavefront solutions to those described above, and may result in periodic spatial patterns driven by the dynamics of the interspecific interactions [@carrère2018; @holmes1994].
 It is often interesting to investigate the stability and long-term behavior of a system of competition-diffusion models for different parameter values, including varying the propagation speeds of the two organisms or the differential impacts of the competitive interactions on each species [@carrère2018; @ducrot2014].
@@
 violaceus* colonies can outcompete *Botryllus schlosseri*, a different invasive tunicate that has been equally or more problematic for the New England aquaculture industry [@dijkstra2007; @gittenberger2011].
-However, these studies have been limited to small-scale experiments conducted on a single substrate (e.g., an oyster cage, SOURCE).
+However, these studies have been limited to small-scale experiments conducted on a single substrate like a PVC panel [@wagstaff2017].
 Our single-species reaction diffusion model could be extended to a system of competition-diffusion equations to examine the potential impacts of competition between *B.
@@
 schlosseri* on the spread of these two species on a broader scale.
+
+## References

### 2024-04-17 PDEs/pde_mod.R

@@ -1,16 +1,34 @@
 # Ruby Krasnow
 # Code for PDE final project
-# Last modified: April 16, 2024
+# Last modified: April 17, 2024
 
 #load packages
 library(tidyverse)
-library(spocc)
-library(sf)
-library(lwgeom)
-library(leaflet)
-library(PNWColors)
 library(gt)
-library(mapview)
+library(sf)
+library(spocc)
+library(PNWColors)
+library(leaflet)
+library(webshot2)
+library(ggmap)
+library(rstatix)
+
+
+# Predictions -------------------------------------------------------------
+colony_day <- 0.225 #0.225 km per colony/day
+raft_prob <- (2/3) # probability that a given colony will raft
+hab_prob <- 0.5 # probability that a given rafting colony will reach a suitable habitat
+km_year_raft <- colony_day*raft_prob*hab_prob*365
+D_raft <- 2*(km_year_raft^2)/(pi)
+
+r <- c(1.887,1.842,1.665,1.512,0.972,0.984, 0.867,0.778)
+test_v <- function(x) {
+  2*sqrt(x*D_raft)
+}
+preds <- map_vec(r, test_v)
+
+
+# Observations ------------------------------------------------------------
 
 # Define spatial boundary for observations
 polygon_list = list(rbind(c(-72, 41), c(-59, 41), c(-59, 50), c (-72, 50), c(-72, 41)))
@@ -19,8 +37,11 @@ poly<-st_sfc(poly, crs=4326)
 
 # Query GBIF database
 df_gbif <- occ(query = 'Botrylloides violaceus',
-               from = 'gbif', has_coords = TRUE,
-               geometry = c(-72, 41, -59, 50), limit = 1000)
+               from = 'gbif',
+               has_coords = TRUE,
+               geometry = c(-72, 41, -59, 50),
+               limit = 1000,
+               gbifopts = list(basisOfRecord ="HUMAN_OBSERVATION"))
 
 # Convert lon and lat to numbers, remove observations with no date
 df_gbif_cleaned <- occ2df(df_gbif) %>% 
@@ -62,65 +83,59 @@ dist <- tunicate %>%
 
 speed <- dist %>% mutate(km=as.double(max_dist/1000), # Convert from m to km
                          speed=km/(year-2005)) %>% # Speed = distance (km)/time (yr)
-                         na.omit()
+                         na.omit() %>% 
[…]

### 2024-04-17 PDEs/testing.R

@@ -1,0 +1,115 @@
+
+
+gbif_orig <- df_gbif[["gbif"]][["data"]][["Botrylloides_violaceus"]]
+
+
+years_na <- occ2df(df_gbif) %>% 
+  mutate(lng=as.double(longitude), lat=as.double(latitude), .keep="unused") %>% 
+  mutate(year=year(date))
+
+
+with_old <- years_na %>% 
+  left_join(gbif_orig %>% select(key, year), by="key") %>% 
+  rename(year=year.y) %>% 
+  select(-year.x) %>% mutate(year = as.numeric(year))
+
+tunicate_old <- st_as_sf(with_old %>% 
+                       select(year, lng, lat),coords = c("lng","lat")) %>% filter(year>1980)
+
+years_old <- tunicate_old %>% group_by(year) %>% summarise(year=mean(year)) # Group by year of obs
+
+years_vec_old <- c(1988:2024) # Years with observations in the GBIF dataset
+
+get_yrs_old <- function(yr) { 
+  out <- years_old %>% filter(year <= yr) %>% st_union()
+}
+years2_old <- years_old %>% mutate(years_agg = map_vec(year, get_yrs_old))
+
+init_old <- tunicate_old %>% 
+  st_set_crs(4326) %>% # Define CRS
+  filter(year==1988) %>% # Select only the year of introduction
+  pull(geometry) # Pull the coordinates
+
+dist_old <- tunicate_old %>% 
+  filter(year>1988) %>% # Remove year of introduction
+  st_set_crs(4326) %>% # Define CRS
+  ungroup() %>% 
+  mutate(dist = st_distance(geometry, init_pt)) %>% # Find the distance b/w the initial point and a given observation
+  group_by(year) %>% 
+  st_drop_geometry() %>% 
+  select(year, dist) %>%  # Select only year and distance columns
+  summarize(max_dist = max(dist)) %>% # Find the maximum distance b/w observations and initial point in each year
+  mutate(max_dist=cummax(max_dist)) # Maximum distance the species has been observed from the initial point before and during a given year
+
+
+speed_old <- dist_old %>% mutate(km=as.double(max_dist/1000), # Convert from m to km
[…]

### 2024-04-23 PDEs/math245_hw6.qmd
pde hw6

@@
 author: "Ruby Krasnow and Tommy Thach"
-date: "2024-04-14"
+date: "2024-04-23"
 format: 
@@
 ```
-
-## Done: 1-4
-## Not started: 5-9
 
@@
 $$R''r^2 + R'r-\lambda R=0$$
-This is an Euler ODE of the form $ax^2y''+bxy'+cy=0$, which has the characteristic equation $am(m-1)+bm+c$ and, if $m_1, m_2$ are distinct real roots, the general solution $y=.
+This is an Euler ODE of the form $ax^2y''+bxy'+cy=0$, which has the characteristic equation $am(m-1)+bm+c$ and, if $m_1, m_2$ are distinct real roots, the general solution $y=C_1 x^{m_1}+C_2x^{m_2}$.
 (Note that I have used $m$ instead of the standard $r$ to avoid confusion with the variable $r$ from (\ref{eqn:semi}).
@@
 Find the necessary condition for the existence of a solution and then solve the problem. Hint: Use Green's identity
-$$\iint\limits_{D}(u\Delta v-v\Delta u)\,dA=\int\limits_{\partial D} \left(u\frac{\partial v}{\partial n} - v\frac{\partial u}{\partial n}\right)\,ds $$
+\begin{equation}\label{eqn:Green}
+\iint\limits_{D}(u\Delta v-v\Delta u)\,dA=\int\limits_{\partial D} \left(u\frac{\partial v}{\partial n} - v\frac{\partial u}{\partial n}\right)\,ds 
+\end{equation}
 where $D$ is a domain and $n$ is a unit normal vector. Note that for a 2-dimensional domain $D$, $dA=dx\,dy$ and $ds$ is arc length.
+
+- - - -
+
+Since (\ref{eqn:Green}) must hold for any $(u,v)$, take $v=1$. Then $\Delta v = 0 = \frac{\partial v}{\partial n}$, so
+\begin{align*}
+\iint\limits_{D}(u\Delta v-v\Delta u)\,dA &=\int\limits_{\partial D} \left(u\frac{\partial v}{\partial n} - v\frac{\partial u}{\partial n}\right)\,ds \quad \text{becomes} \\
+\iint\limits_{D}(0-\Delta u)\,dA &=\int\limits_{\partial D} \left(0 - \frac{\partial u}{\partial n}\right)\,ds \\
+0 &=-\int\limits_{\partial D} g\left( \theta \right)\,ds
+\end{align*}
+
+Where we used the fact that by (\ref{eqn:prob5}), $\Delta u = 0$ and $\frac{\partial u}{\partial n}=u_r=g(\theta)$.
+Thus, the necessary condition for the existence of a solution is $$\int\limits_{\partial D} g\left( \theta \right)\,ds=0$$
+
+To solve the problem, we now look for a separated solution of the form $u=\Theta(\theta)R(r)$.
+In polar coordinates, $\Delta u=0$ becomes $u_{rr}+\frac{1}{r}u_r+ \frac{1}{r^2}u_{\theta\theta}=0$, giving us $R''\Theta + \frac{1}{r}R'\Theta+\frac{1}{r^2}R\Theta''=0$.
+From the previous problem, we know that this yields two problems, one in $\Theta$ and one in $R$.
+
+For $\Theta$, we have  $\Theta''+\Theta\lambda=0$ with periodic BCs. Using the table provided and setting $l=\pi$, this means we have eigenvalues
+\begin{equation*}
+    \begin{cases}
+     \lambda_0=0 \\
+     \lambda_n = \left( \frac{n \pi}{l} \right)^2=n^2 \quad n=1,2,3,...
+    \end{cases}       
+\end{equation*}
+
+with eigenfunctions
+
+\begin{equation*}
+    \begin{cases}
+     \Theta_0(\theta)=C_0 \\
+     \Theta_n(\theta) = \Bigl\{ \cos{ \left( n\theta \right)},\sin{\left( n\theta \right)} \Bigr\}
+    \end{cases}       
+\end{equation*}
+
+Now we return to the $R$ problem. As before, we get an Euler equation with $a=1, b=1, c=-\lambda$, so the characteristic equation becomes $m(m-1)+m-\lambda$, which simplifies to $m^2-\lambda$.
+
+For $\lambda_0$, we get the repeated root $m_1=m_2=0$, giving us the solution $R_0(r)=C_1+C_2 \ln{r}$.
+For $\lambda_n=n^2$, our roots are $m=\pm n$. Thus, $R_n(r)=C_3r^n+C_4r^{-n}$.
+
+We are only explicitly provided one BC, so in order to determine the unknown coefficients, we will need to assume that we should disregard the solutions that go to infinity as $r$ approaches infinity, since our domain is the region *outside* the disk with radius $a$.
+
+Thus, we exclude $\ln{r}$ and $r^n$, which will both go to $+\infty$ as $r\rightarrow+\infty$, and sum the remaining solutions to find that
+
+$$u(r,\theta)=A_0+\sum_{n=1}^{\infty}r^{-n} \left(A_n\cos{(n\theta)}+B_n\sin{(n\theta)}\right) $$
+$$u_r(r,\theta)=A_0+\sum_{n=1}^{\infty}-nr^{-n-1} \left(A_n\cos{(n\theta)}+B_n\sin{(n\theta)}\right) $$
+
+Finally, we use the inhomogeneous BC at $r=a$. Setting $r=a$ in the series above, we require that
+$$g(\theta)=A_0+\sum_{n=1}^{\infty}-na^{-n-1} \left(A_n\cos{(n\theta)}+B_n\sin{(n\theta)}\right) $$
+
+Since this is a Fourier series for $g(\theta)$, we can now solve for the unknown coefficients:
+
+$$A_0=\frac{1}{2\pi} \int_{-\pi}^{\pi}g(\theta)\,d\theta $$
+But from the necessary condition for the existence of a solution, this integral must be zero.
+Therefore, our solution to (\ref{eqn:prob5}) is
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Solution]
+
+$$u(r,\theta)=\sum_{n=1}^{\infty}r^{-n} \left(A_n\cos{(n\theta)}+B_n\sin{(n\theta)}\right) $$
+where
+$$A_n=\frac{-1}{n\pi(a^{n+1})} \int_{-\pi}^{\pi}g(\theta) \cos{(n\theta)}\,d\theta $$
+$$B_n=\frac{-1}{n\pi(a^{n+1})} \int_{-\pi}^{\pi}g(\theta) \sin{(n\theta)}\,d\theta $$
+\end{tcolorbox}
 
@@
 
-Prove the uniqueness of the Neumann problem on $D \subset \mathbb{R}$
+Prove the uniqueness of the Neumann problem on $D \subset \mathbb{R^2}$
 \begin{equation}\label{eqn:unique}
@@
 \end{equation}
+
+Using the version of Green's first identity given in the textbook but in two dimensions, we have
+
+$$\int\limits_{\partial D}  v\frac{\partial u}{\partial n} \,ds = \iint \limits_{D} \nabla v \cdot \nabla u \;d \mathbf{x}+ \iint \limits_{D}v\Delta u \;d \mathbf{x}$$
+
+Now set $u=v=w$:
+
+$$\int\limits_{\partial D}  w\frac{\partial w}{\partial n} \,ds = \iint \limits_{D} \lvert \nabla w\rvert^2 \;d \mathbf{x}+ \iint \limits_{D}w\Delta w \;d \mathbf{x}$$
+
+But from (\ref{eqn:unique2}), $\Delta w=\frac{\partial w}{\partial n}=0$, so
+
+$$0 = \iint \limits_{D} \lvert\nabla w\rvert^2\;d \mathbf{x}$$
+
+By the vanishing theorem, it follows that $\lvert\nabla w\rvert^2 \equiv 0$ in $D$, so $\nabla w=0$.
+But since a function with a vanishing gradient must be constant, provided that $D$ is connected, this means that $w=c$.
+Thus, $c=u-v$ and the two solutions to the Neumann problem are unique up to a constant.
+$\square$

### 2024-04-23 PDEs/paper_outline.qmd
pde hw6

@@
 author: "Ruby Krasnow"
-date: "2024-04-17"
+date: "2024-04-22"
 format:

### 2024-04-23 PDEs/testing.R

@@ -60,56 +60,29 @@ speed_old <- dist_old %>% mutate(km=as.double(max_dist/1000), # Convert from m t
 
 # 44°56'47.6"N 66°54'19.7"W in 2009, from Martin et al. (2011): doi: 10.3391/ai.2011.6.4.05  
 
-# http://www.aquaticinvasions.net/2011/AI_2011_6_4_Sephton_etal_Supplement.pdf 
-# 2006: 43.71370 65.96947
-# 2006: 43.72360 65.84057
-# 2006: 43.56273 65.36245
-# 2006: 44.36113 64.33362
-# 2006: 44.37525 64.33069
-# 2006: 45.51055 61.01768
-# 2006: 45.50252 60.96400
-# 2006: 46.62680 61.01597
-# 2006: 46.28117 60.42500 
-# 2006: 43.54773 65.43217
-# 2006: 43.56273 65.36245 
+canada <- read_csv("./PDEs/sephton2011.csv",
+                   col_names=TRUE,
+                   col_types="cddllll")
+canada <- canada %>% mutate(lng = -1*lon) %>% mutate(across(where(is.logical),~replace_na(.x, FALSE)))
 
-# 2007
-# 44.19365 66.16688
-# 43.71370 65.96947
-# 43.72360 65.84057
-# 43.44470 65.63510
-# 43.56273 65.36245
-# 43.75780 65.32200
-# 43.69950 65.10740
-# 44.37525 64.33069
-# 44.45598 64.30683 
-# 44.53780 64.23840
-# 45.33450 60.98610
-# 46.20680 60.24900
-# 46.62680 61.01597 
+canada <- canada %>% 
+  pivot_longer(cols=where(is.logical), 
+               values_to = "present", names_to = "year") %>%  filter(present==TRUE)%>%
+  mutate(year=as.numeric(year)) %>% group_by(site) %>% slice_min(year) %>% ungroup()
 
-# 2008
-# 44.19365 66.16688
-# 43.81648 66.14785
-# 43.71370 65.96947
-# 43.72360 65.84057
-# 43.44470 65.63510
-# 43.56273 65.36245
-# 43.69950 65.10740 
-# 44.37525 64.33069
-# 44.45598 64.30683
-# 44.44782 64.37437
-# 44.53780 64.23840
-# 45.50710 60.96050
-# 45.58870 60.96180
-# 45.58333 60.74067 
-# 45.66115 60.87440
[…]

### 2024-04-29 PDEs/math245_hw6.qmd
presentation

@@
 ```
+
+# (Wednesday morning correction to Problem 5)
 
@@
 \begin{align*}
-u_{avg}(\theta)& =\frac{1}{2\pi}\int_0^{2\pi} (\sin\theta+1)\;d\theta \\
+u(0) &=\text{average of u over } \partial{B}(0,2) \\
+&=\frac{1}{2\pi R} \int_{\partial{B}(0,R)} u \; ds \\
+&=\frac{1}{4\pi} \int_{\partial{B}(0,R)} u \; ds \\
+&=\frac{1}{4\pi} \int_{\theta=0}^{\theta=2\pi} u \cdot R \; d\theta \\
+& =\frac{1}{2\pi}\int_0^{2\pi} u\;d\theta \\
+& =\frac{1}{2\pi}\int_0^{2\pi} (\sin\theta+1)\;d\theta \\
 &=\frac{1}{2\pi}\Bigl[(\cos\theta+\theta) \Bigr]_0^{2\pi} \\
@@
 \iint\limits_{D}(0-\Delta u)\,dA &=\int\limits_{\partial D} \left(0 - \frac{\partial u}{\partial n}\right)\,ds \\
-0 &=-\int\limits_{\partial D} g\left( \theta \right)\,ds
+0 &=-\int\limits_{\partial D} g\left( \theta \right)\,d\theta
 \end{align*}
@@
 Where we used the fact that by (\ref{eqn:prob5}), $\Delta u = 0$ and $\frac{\partial u}{\partial n}=u_r=g(\theta)$.
-Thus, the necessary condition for the existence of a solution is $$\int\limits_{\partial D} g\left( \theta \right)\,ds=0$$
+Thus, the necessary condition for the existence of a solution is $$\int\limits_{\partial D} g\left( \theta \right)\,d\theta=0$$
 
@@
 $$u(r,\theta)=A_0+\sum_{n=1}^{\infty}r^{-n} \left(A_n\cos{(n\theta)}+B_n\sin{(n\theta)}\right) $$
-$$u_r(r,\theta)=A_0+\sum_{n=1}^{\infty}-nr^{-n-1} \left(A_n\cos{(n\theta)}+B_n\sin{(n\theta)}\right) $$
+$$u_r(r,\theta)=\sum_{n=1}^{\infty}-nr^{-n-1} \left(A_n\cos{(n\theta)}+B_n\sin{(n\theta)}\right) $$
 
 Finally, we use the inhomogeneous BC at $r=a$. Setting $r=a$ in the series above, we require that
-$$g(\theta)=A_0+\sum_{n=1}^{\infty}-na^{-n-1} \left(A_n\cos{(n\theta)}+B_n\sin{(n\theta)}\right) $$
+$$g(\theta)=\sum_{n=1}^{\infty}-na^{-n-1} \left(A_n\cos{(n\theta)}+B_n\sin{(n\theta)}\right) $$
 
@@
 
-$$A_0=\frac{1}{2\pi} \int_{-\pi}^{\pi}g(\theta)\,d\theta $$
-But from the necessary condition for the existence of a solution, this integral must be zero.
+$$-nA_na^{-n-1}=\frac{1}{\pi} \int_{-\pi}^{\pi}g(\theta)\cos(n\theta)\,d\theta $$
+$$-nB_na^{-n-1}=\frac{1}{\pi} \int_{-\pi}^{\pi}g(\theta)\sin(n\theta)\,d\theta $$
+
 Therefore, our solution to (\ref{eqn:prob5}) is
@@
 
-$$u(r,\theta)=\sum_{n=1}^{\infty}r^{-n} \left(A_n\cos{(n\theta)}+B_n\sin{(n\theta)}\right) $$
+$$u(r,\theta)=A_0+\sum_{n=1}^{\infty}r^{-n} \left(A_n\cos{(n\theta)}+B_n\sin{(n\theta)}\right) $$
 where
-$$A_n=\frac{-1}{n\pi(a^{n+1})} \int_{-\pi}^{\pi}g(\theta) \cos{(n\theta)}\,d\theta $$
-$$B_n=\frac{-1}{n\pi(a^{n+1})} \int_{-\pi}^{\pi}g(\theta) \sin{(n\theta)}\,d\theta $$
+$$A_n=-\frac{a^{n+1}}{n\pi} \int_{-\pi}^{\pi}g(\theta) \cos{(n\theta)}\,d\theta $$
+$$B_n=-\frac{a^{n+1}}{n\pi} \int_{-\pi}^{\pi}g(\theta) \sin{(n\theta)}\,d\theta $$
 \end{tcolorbox}
@@
 
-Now set $u=v=w$:
+Now replace $u$ and $v$ in this formula with $w$:
 
@@
 
-But from (\ref{eqn:unique2}), $\Delta w=\frac{\partial w}{\partial n}=0$, so
+But from (\ref{eqn:unique2}), $\Delta w=0$ in $D$ and $\frac{\partial w}{\partial n}=0$ on $\partial{D}$, so
 

### 2024-04-29 PDEs/paper_outline.qmd
presentation

@@
 author: "Ruby Krasnow"
-date: "2024-04-22"
+date: "2024-04-29"
 format:
@@
 ```
+
+# Abstract
+Modeling is a critical tool for minimizing the harmful effects of invasive species on natural ecosystems.
+Despite the recent popularity of correlative environmental niche models, mechanistic reaction-diffusion models remain powerful tools for predicting and controlling the spread of invasive species.
+The colonial ascidian *Botrylloides violaceus* has rapidly spread across the Gulf of Maine, outcompeting and smothering native species and presenting a major nuisance for shellfish aquaculture.Here, we apply arelationship between the rate of spread of an invasive species and select biological parameters(namely, the population growth rate and diffusion coefficient) derived from the generalized reaction-diffusion equation.
+We validated the predicted rate of spread for *B.
+violaceus* against rates calculated from species occurrence data from the citizen science platform iNaturalist.
+The model predicted a mean speed of ~49 km/year, which was extremely close to the observed mean of ~40 km/year.
+Although simple reaction-diffusion models have limited biological realism, the accuracy of our results suggest that estimating spreading speed based on empirical data on local
+diffusion potentials and population growth rates could be an underutilized approach in invasion ecology.
 
@@
 \frac{\partial{u}}{\partial{t}}=
-\text{D}\frac{\partial{u}^2}{\partial{x^2}}+r(1-\frac{u}{K})
+\text{D}\frac{\partial{u}^2}{\partial{x^2}}+ru\left(1-\frac{u}{K}\right)
 \end{equation}
@@
 In other words, below $U_c$ it is so difficult to find the invasive organisms that for practical purposes they are not present.
-By rearranging (\ref{eqn:PDEsol}), we can get the following expression $$\left(\frac{x}{t}\right) ^2=4rD+\frac{4r\sigma^2}{t}-\frac{4(\sigma^2+Dt)}{t^2}\ln{  \left(\frac{U_{cr}\sqrt{4\pi(\sigma^2+Dt)}}{G}\right) }$$
+By rearranging (\ref{eqn:PDEsol}), we can get the following expression $$\left(\frac{x}{t}\right) ^2=4rD+\frac{4r\sigma^2}{t}-\frac{4(\sigma^2+Dt)}{t^2}\ln{  \left(\frac{U_{c}\sqrt{4\pi(\sigma^2+Dt)}}{G}\right) }$$
 

### 2024-04-29 PDEs/pres.Rmd
presentation

---
title: "PDE pres"
author: "Ruby Krasnow"
date: "2024-04-27"
output:
  pdf_document:
    extra_dependencies: ["tcolorbox", "xcolor"]

bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

We represent the the population density of *B. violaceus* at time $t$ and spatial coordinate $x$ as $u(t,x)$.

Assume that the population occupies a domain of fixed length $L$, so that $0 < x < L$.
The environment outside the domain is hostile, so there is no population there: if
$x\leq 0$ or $x \geq L$, then $u(x,t)= 0$ for all $t$.

If we assumed simple linear population growth, our PDE would take the form

\begin{equation}\label{eqn:lin_PDE}
\frac{\partial{u}}{\partial{t}}=\text{D}\frac{\partial{u}^2}{\partial{x^2}}+ru, \quad 0<x<L
\end{equation}

with the Dirichlet BCs $u(0,t)=u(0,L)=0$.

This is the simplest version of the general reaction-diffusion model, for a homogeneous, unstructured population growing exponentially and dispersing in a uniform, one-dimensional environment [@andow1990; @skellam1951].

We know how to solve this! The solution is
$$u(t,x)= \sum_{n=1}^{\infty} C_n e^{(r-\lambda_n)t} \sin{\left(\frac{n\pi x}{L}\right)}, $$
where

$$\lambda_n= \left(\frac{n\pi}{L}\right)^2D,\quad n=1,2,3... $$
and the coefficients $C_n$ are determined by the initial density $u(x,0)$.

$$\lambda_1= \left(\frac{\pi}{L}\right)^2D \quad \Longrightarrow \quad r= \left(\frac{\pi}{L}\right)^2D \quad \Longrightarrow \quad L_{cr}=\pi \sqrt{\frac{D}{r}}$$

A more realistic model considers the evolution of a compact initial
population distribution in a large open space. Rather than imposing the restriction that the population always occupy a fixed region, we instead allow the population to grow freely. Only the
initial condition is bounded, assumed for simplicity to be connected:
\begin{align}\label{eqn:IC_1} u(0,x) =
\begin{cases}
\phi(x)>0 & 0<x<L \\
0 & \text{otherwise}
\end{cases}
\end{align}

Here, $L$ is the size of the initially occupied area and the function $\phi$ describes the
initial population distribution. The equation describing the population dynamics,
that is,

\begin{equation}\label{eqn:PDE}
\frac{\partial{u}}{\partial{t}}=\text{D}\frac{\partial{u}^2}{\partial{x^2}}+F(u)
\end{equation}

should now be defined on a much larger domain than $L$; here, we assume $-\infty<x<\infty$. Note that $F(u)$ can now be a nonlinear function (like logistic growth), which is much more biologically realistic.

For biologically reasonable choices of $F(u)$, the growth function will be bounded by the linear function, such that $F(u) \leq ru$.

\begin{equation}\label{eqn:comp1}
\text{D}\frac{\partial{u}^2}{\partial{x^2}}+F(u) \leq \text{D}\frac{\partial{U}^2}{\partial{x^2}}+rU
\end{equation}

Now, note that for compact initial conditions like (\ref{eqn:IC_1}), there exist positive constants $G$ and $\sigma$ such that we bound the initial distribution from above by a Gaussian distribution: \begin{equation}\label{eqn:initGaus}
U(0,x)=\frac{G}{\sqrt{4\pi\sigma^2}}\exp{\left(-\frac{x^2}{4\sigma^2}\right)}
\end{equation}

\begin{equation}\label{eqn:comp2} 
\text{such that}\quad u(0,x) \leq U(0,x)\quad \text{for all }x
\end{equation}

\begin{equation}
\frac{\partial{U}}{\partial{t}}=\text{D}\frac{\partial{U}^2}{\partial{x^2}}+rU
\end{equation}

The linear equation in (1) can be readily solved, since the change of variables $U(t,x)= W(t,x)e^{rt}$ reduces it to the standard diffusion equation for $W$. Using forward and inverse Fourier transforms, we find that
$$W(t,x)=\frac{G}{\sqrt{4\pi(\sigma^2+Dt)}}\exp{\left(-\frac{x^2}{4(\sigma^2+Dt)}\right)}  $$

Converting back to our original variable, we conclude that the solution to (1) with the initial condition (5a) is a normal distribution, specifically
$$U(t,x)=\frac{G}{\sqrt{4\pi(\sigma^2+Dt)}}\exp{\left(-\frac{x^2}{4(\sigma^2+Dt)}+rt\right)}$$

Now, we can quantify the spatial spread of the population by considering how quickly a level set $U(x,t)=U_c$ moves in space, where $U_c$ could be considered a threshold of detection. Rearranging the above equation, we can get the following expression
$$\left(\frac{x}{t}\right) ^2=4rD+\frac{4r\sigma^2}{t}-\frac{4(\sigma^2+Dt)}{t^2}\ln{  \left(\frac{U_{c}\sqrt{4\pi(\sigma^2+Dt)}}{G}\right) }$$

As $t$ becomes large, this formula simplifies to $\frac{\lvert x \rvert}{t}\rightarrow \sqrt{4rD}$. Thus, the asymptotic rate of spread is $V=\sqrt{4rD}$, as we initially sought to show.

## General reaction-diffusion equation
$$\frac{\partial{u}}{\partial{t}}=
\text{D}\left(\frac{\partial{u}^2}{\partial{x^2}}\right)+f(u)$$

## Linear growth
$$\frac{\partial{u}}{\partial{t}}=
\text{D}\left(\frac{\partial{u}^2}{\partial{x^2}}\right)+ru$$

## Fisher's equation
$$\frac{\partial{u}}{\partial{t}}=
\text{D}\left(\frac{\partial{u}^2}{\partial{x^2}}\right)+ru\left(1-\frac{u}{K}\right)$$

## Harvest/fishing mortality
$$\frac{\partial{u}}{\partial{t}}=
\text{D}\left(\frac{\partial{u}^2}{\partial{x^2}}\right)+ru\left(1-\frac{u}{K}\right)-qE(x)u$$

## With advection
$$\frac{\partial{u}}{\partial{t}}=
\text{D}\left(\frac{\partial{u}^2}{\partial{x^2}}\right)-a\frac{\partial{u}}{\partial{x}}+f(u)$$

# References

### 2024-04-29 PDEs/pde_mod.R

@@ -1,6 +1,6 @@
 # Ruby Krasnow
 # Code for PDE final project
-# Last modified: April 17, 2024
+# Last modified: April 29, 2024
 
 #load packages
 library(tidyverse)

### 2024-07-23 PDEs/paper_outline.qmd
reorganizing files

@@
 Despite the recent popularity of correlative environmental niche models, mechanistic reaction-diffusion models remain powerful tools for predicting and controlling the spread of invasive species.
-The colonial ascidian *Botrylloides violaceus* has rapidly spread across the Gulf of Maine, outcompeting and smothering native species and presenting a major nuisance for shellfish aquaculture.Here, we apply arelationship between the rate of spread of an invasive species and select biological parameters(namely, the population growth rate and diffusion coefficient) derived from the generalized reaction-diffusion equation.
+The colonial ascidian *Botrylloides violaceus* has rapidly spread across the Gulf of Maine, outcompeting and smothering native species and presenting a major nuisance for shellfish aquaculture.
+Here, we apply a relationship between the rate of spread of an invasive species and select biological parameters (namely, the population growth rate and diffusion coefficient) derived from the generalized reaction-diffusion equation.
 We validated the predicted rate of spread for *B.
@@
 The model predicted a mean speed of ~49 km/year, which was extremely close to the observed mean of ~40 km/year.
-Although simple reaction-diffusion models have limited biological realism, the accuracy of our results suggest that estimating spreading speed based on empirical data on local
-diffusion potentials and population growth rates could be an underutilized approach in invasion ecology.
+Although simple reaction-diffusion models have limited biological realism, the accuracy of our results suggest that estimating spreading speed based on empirical data on local diffusion potentials and population growth rates could be an underutilized approach in invasion ecology.
 

### 2024-07-23 PDEs/pde_mod.R

@@ -63,6 +63,7 @@ get_yrs <- function(yr) {
  
 years_vec <- c(2005:2024) # Years with observations in the GBIF dataset
 
+# Apply function to all years in the years_vec vector
 years2 <- years %>% mutate(years_agg = map_vec(year, get_yrs))
 
 init_pt <- tunicate %>% 

### 2024-07-23 PDEs/testing.R

@@ -1,88 +1,116 @@
 
 
-gbif_orig <- df_gbif[["gbif"]][["data"]][["Botrylloides_violaceus"]]
+library(clipr)
+library(tidyverse)
+library(ggpubr)
+
+mom<-read_clip_tbl()
+
+mom2 <- tribble(~Date,  ~Time,
+ #"11/25/21", "34:00",
+   "8/6/23", "27:05",
+  "9/10/23", "26:50",
+  "10/1/23", "26:03",
+ #"11/23/23", "29:44",
+   "1/1/24", "26:26",
+   "3/3/24", "25:29",
+   "4/6/24", "26:55",
+  "5/12/24", "26:29")
+
+mom2 <- mom2 %>% mutate(Date=mdy(Date), Time=ms(Time))
+
+ggplot(data=mom2, aes(x=Date, y=Time))+
+  geom_point()+
+  theme_classic()+
+  geom_smooth(method="lm")+
+  stat_cor(aes(label = after_stat(r.label)),  r.accuracy = 0.001, label.x.npc=0.1, label.y.npc = "bottom")+
+  scale_y_time(breaks = c(ms("27:30","27:00", "26:30", "26:00", "25:30")), labels = c("27:30","27:00", "26:30", "26:00" ,"25:30"))+theme(axis.title.y = element_text(margin = margin(r = 10)))
 
 
-years_na <- occ2df(df_gbif) %>% 
-  mutate(lng=as.double(longitude), lat=as.double(latitude), .keep="unused") %>% 
-  mutate(year=year(date))
 
-
-with_old <- years_na %>% 
-  left_join(gbif_orig %>% select(key, year), by="key") %>% 
-  rename(year=year.y) %>% 
-  select(-year.x) %>% mutate(year = as.numeric(year))
-
-tunicate_old <- st_as_sf(with_old %>% 
-                       select(year, lng, lat),coords = c("lng","lat")) %>% filter(year>1980)
-
-years_old <- tunicate_old %>% group_by(year) %>% summarise(year=mean(year)) # Group by year of obs
-
-years_vec_old <- c(1988:2024) # Years with observations in the GBIF dataset
-
-get_yrs_old <- function(yr) { 
-  out <- years_old %>% filter(year <= yr) %>% st_union()
-}
-years2_old <- years_old %>% mutate(years_agg = map_vec(year, get_yrs_old))
-
-init_old <- tunicate_old %>% 
-  st_set_crs(4326) %>% # Define CRS
[…]
